### 2016-06-15 testbed_nfs.tex
Started on nfs setup

\label{sec:testbed_nfs}

\subsection{NFS}

% Define variable]
\begin{lstlisting}[]
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
$ export DEV="/dev/mmcblk0"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

% Check out the image   % The dollar hack was to fix vim syntax
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

### 2016-06-15 setup_scripts/nfs_http_sdcard.sh
Added script to setup sdcard for nfs

@@ -1,0 +1,55 @@
+#!/usr/bin/env bash
+
+set -e -u
+
+# Target device
+dev=/dev/mmcblk0
+
+# Print error
+echoerr() { cat <<< "$@" 1>&2; }
+
+# Make sure script is run with root permissions
+if [ "$(id -u)" != "0" ]; then
+   echoerr "This script must be run as root"; exit 1
+fi
+
+# Make sure we write to a block device
+if [ ! -b ${dev} ]; then
+    echoerr "Error: ${dev} is not a block device"; exit 1
+fi
+
+# Make sure the user knows what he is doing
+read -r -p "WARNING: Everything on ${dev} will be erased. Continue [y/N]? " response
+[[ ! ${response,,} =~ ^(yes|y) ]] && exit 1
+
+
+
+# Erase partition table (4M because: rather too much than too little)
+dd if=/dev/zero of=${dev} bs=4M count=1
+
+(echo o;                                        # Create DOS partition table
+    echo n; echo p; echo 1; echo ; echo +64M;   # Create boot partition
+    echo t; echo b;                             # Set boot partition type to FAT
+    echo n; echo p; echo 2; echo ; echo ;       # Create home partition (default partition type: Linux)
+    echo w                                      # Write table to disk
+    ) | fdisk ${dev}
+
+# Print new partition table
+fdisk -lu ${dev}
+
+# Format partitions
+mkfs.vfat -n BOOT ${dev}p1
+mkfs.ext4 -L HOME ${dev}p2
+
+# SHOULD THE BELOW BE COPIED FROM THE ROOT IMAGE?
+
+# Create temp directory that is erased when script exits
+mnt=$(mktemp -d)
+
+# Create home directory for user: pi (uid=1000) in group pi (gid=1000). (gid=100 if group users are preferred)
+# mode=700 means: drwx------ (i.e. only the user and root has access to his directory)
+mount ${dev}p2 ${mnt}
+install -d -o 1000 -g 1000 -m 700 ${mnt}/pi && sync
+
+# Cleanup
+umount ${mnt} && rm -r ${mnt}

### 2016-06-21 testbed_http.tex
more on http

\label{sec:testbed_http}

\subsection{HTTP}

\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Extract root partition from image
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
# dd if=${DEV}p1 of=${HTTPBOOTIMAGE}.img bs=4M
# dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Mount http root and home directories
\begin{lstlisting}[]
$ export HTTPROOTDIR="${WORKDIR}/httproot"
$ mkdir -p ${HTTPROOTDIR}
# mount ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
# mount ${HTTPBOOTIMAGE}.img ${HTTPROOTDIR}/boot
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create the following file (THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT)
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/initramfs-tools/scripts/http}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
# Fetch, cache, and mount root image using HTTP

# Create mount point directory
[ -d /imagedir ] || mkdir /imagedir

# Parse additional command line options
for x in ${cat /proc/cmdline}; do
    case $x in
    image=*)
        IMAGE=${x#image=}
        # Extract device and path
        # E.g. /dev/mmcblkp2:/root.img --> /dev/mmcblkp2 and /imagedir/root.img
        IMAGEDEV=${IMAGE%:*}
        IMAGEPATH="/imagedir/${${IMAGE#*:}#/}"
        ;;
    url=*)
        URL=${${x#url=}%/}  # If url has trailing slash, then remove it
        ;;
done

# Check if variables are declared
[ -z "${IMAGE}" ] && echo "Error: Please add 'image' in cmdline.txt" && exit 1
[ -z "${URL}" ] && echo "Error: Please add 'url' in cmdline.txt" && exit 1

# Fetch and cache root image
http_fetch_image()
{
    configure_networking

    # Fetch list of root image names
    if ! wget -O /tmp/image.csv "${URL}/image.csv"; then
        echo "WARNING failed to fetch image list"
        return
    fi

    # Extract name of image to download based on eth0 MAC
    mac=$(cat /sys/class/net/eth0/address)
    img=$(grep $mac /tmp/image.csv | cut -f2 -d' ')

    # If the image is not cached, then get it
    if [ ! -f ${img} ] && wget --spider -q "${URL}/${img}"; then
        # Delete old image
        rm -f $(readlink -f ${IMAGEPATH})

        # Fetch new image
        if wget -O "${IMAGEPATH%/*}/${img}" "${URL}/${img}"; then
            # If success, create symbolic link to new image
            ln -f -s "${IMAGEPATH%/*}/${img}" ${IMAGEPATH}
        fi
    fi

    # Cleanup
    #rm /tmp/image.csv
}

http_mount_root()
{
    modprobe ext4

    # Mount imagedir
    #mount -n -t ext4 ${IMAGEDEV} /${${IMAGEPATH#/}%%/*}
    mount -n -t ext4 ${IMAGEDEV} /imagedir

    # Fetch and cache root image
    http_fetch_image

    # Mount root as readonly
    losetup /dev/loop0 $(readlink -f ${IMAGEPATH})
    mount -r -n -t ext4 -o nodiratime,noatime /dev/loop0 /root
}

mountroot()
{
    http_mount_root
}

\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Make the script executable
\begin{lstlisting}[]
# chmod +x ${HTTPROOTDIR}/etc/initramfs-tools/scripts/http
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Add the below to \$\{HTTPROOTDIR\}/boot/cmdline.txt (CHANGE URL HERE) (remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
\begin{lstlisting}[]
init=http ro ip=dhcp image=/dev/mmcblk0p2:/root.img root=/dev/loop0 url=http://192.168.87.106/raspi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% ip=dhcp root=/dev/http http-options=hard,intr,ro

Edit fstab to look like this (exclude the last line if persistent
home storage directory is not desired):
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{HTTPROOTDIR\}/etc/fstab}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
proc            /proc           proc    defaults          0       0
/dev/mmcblk0p1  /boot           vfat    defaults          0       2
/dev/loop0      /               ext4    defaults,noatime  0       1
/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       1
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Chroot into the image
% CHROOT to OS image
\begin{lstlisting}[]
$ cd $HTTPROOTDIR
# mount -t proc proc proc/
# mount --bind /sys sys/
# mount --bind /dev dev/
# mount --bind /dev/pts dev/pts
# proot -q qemu-arm-static -S ${HTTPROOTDIR}
# export PS1="(chroot) $PS1"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create new initramfs (remember that your kernel might be different)
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.11+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Exit chroot
\begin{lstlisting}[]
(chroot) # exit
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?

% Define variable
% export IMAGE="2016-05-27-raspbian-jessie-lite"
% export WORKDIR="${HOME}/Raspbian"
\begin{lstlisting}[]
$ export HTTPDEV="/dev/mmcblk0"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\begin{lstlisting}[]
<@\Suppressnumber @># (echo o;                                      # Create DOS partition table
    echo n; echo p; echo 1; echo ; echo +64M;   # Create boot partition
    echo t; echo b;                             # Set boot partition type to FAT
    echo n; echo p; echo 2; echo ; echo ;       # Create home partition
    echo w                                      # Write table to disk
    ) | sudo fdisk ${HTTPDEV} <@\Reactivatenumber @>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Print partition table (SHOULD THE BOOT FLAG HAVE BEEN SET TO PARTITION P1?)
\begin{lstlisting}[]
# fdisk -u sectors -l ${HTTPDEV}
Disk /dev/mmcblk0: 3.7 GiB, 3965190144 bytes, 7744512 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x00e61b04

Device         Boot  Start     End Sectors  Size Id Type
/dev/mmcblk0p1        2048  133119  131072   64M  b W95 FAT32
/dev/mmcblk0p2      133120 7744511 7611392  3.6G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Format partitions
\begin{lstlisting}[]
# mkfs.vfat -n BOOT ${HTTPDEV}p1
# mkfs.ext4 -L HOME ${HTTPDEV}p2
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%Mount image:
%\begin{lstlisting}[]
%# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
%# mount ${DEV}p2 ${ROOTDIR}
%# mount ${DEV}p1 ${ROOTDIR}/boot
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

Copy boot directory:
\begin{lstlisting}[]
# cp -r ${ROOTDIR}/boot/* ${HTTPBOOTDIR}/ 
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Prepare home directory (Maybe it is better just to copy the content in home from the
image's root partition):
\begin{lstlisting}[]
# install -d -o 1000 -g 1000 -m 700 ${HOMEDIR}/pi && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\begin{lstlisting}[]
# mount ${HTTPDEV}p1 ${HTTPROOTDIR}/boot
# mount ${HTTPDEV}p2 ${HTTPROOTDIR}/home
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Unmount directories:
\begin{lstlisting}[]
# umount ${HTTPBOOTDIR} ${HTTPHOMEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

### 2016-06-21 testbed_nfs.tex
more on http

@@
 
-% Define variable]
 \begin{lstlisting}[]
-$ export IMAGE="2016-05-27-raspbian-jessie-lite"
-$ export WORKDIR="${HOME}/Raspbian"
-$ export DEV="/dev/mmcblk0"
+# export NFSBOOTIMAGE="${WORKDIR}/nfsboot"
+# export NFSROOTIMAGE="${WORKDIR}/nfsroot"
 \end{lstlisting}
@@
 
-% Check out the image   % The dollar hack was to fix vim syntax
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${DEV}
+Extract root partition from image
+\begin{lstlisting}[]
+# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
+# dd if=${DEV}p1 of=${NFSBOOTIMAGE}.img bs=4M
+# dd if=${DEV}p2 of=${NFSROOTIMAGE}.img bs=4M
 \end{lstlisting}
@@
 \vspace{-5mm}
+
+Mount nfs root, boot, and home directories
+\begin{lstlisting}[]
+$ export NFSROOTDIR="${WORKDIR}/nfsroot"
+$ mkdir -p ${NFSROOTDIR}
+# mount ${NFSROOTIMAGE}.img ${NFSROOTDIR}
+# mount ${NFSBOOTIMAGE}.img ${NFSROOTDIR}/boot
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Add the below to NFSROOTDIR/boot/cmdline.txt
+\begin{lstlisting}[]
+ip=dhcp root=/dev/nfs nfs-options=hard,intr,ro
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{NFSROOTDIR\}/etc/initramfs-tools/hooks/hooks-rootimg}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+#!/bin/sh
+
+. /usr/share/initramfs-tools/scripts/functions
+. /usr/share/initramfs-tools/hook-functions
+
+copy_exec /sbin/losetup
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{NFSROOTDIR\}/etc/initramfs-tools/hooks/init-premount-rootimg}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+#!/bin/sh
+
+PREREQ=""
+prereqs()
+{
+   echo "$PREREQ"
+}
+
+case $1 in
+prereqs)
+   prereqs
+   exit 0
+   ;;
+esac
+
+. /scripts/functions
+log_begin_msg "Starting rootimg"
+log_end_msg
+
+# Setup loop device
+losetup /dev/loop0 /imgdir/root.img
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?
+
+% Define variable
+% export IMAGE="2016-05-27-raspbian-jessie-lite"
+% export WORKDIR="${HOME}/Raspbian"
+\begin{lstlisting}[]
+$ export NFSDEV="/dev/mmcblk0"
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\begin{lstlisting}[]
+<@\Suppressnumber @># (echo o;                                      # Create DOS partition table
+    echo n; echo p; echo 1; echo ; echo +64M;   # Create boot partition
+    echo t; echo b;                             # Set boot partition type to FAT
+    echo n; echo p; echo 2; echo ; echo ;       # Create home partition
+    echo w                                      # Write table to disk
+    ) | sudo fdisk ${NFSDEV} <@\Reactivatenumber @>
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Print partition table (SHOULD THE BOOT FLAG HAVE BEEN SET TO PARTITION P1?)
+\begin{lstlisting}[]
+# fdisk -u sectors -l ${NFSDEV}
+Disk /dev/mmcblk0: 3.7 GiB, 3965190144 bytes, 7744512 sectors
+Units: sectors of 1 * 512 = 512 bytes
+Sector size (logical/physical): 512 bytes / 512 bytes
+I/O size (minimum/optimal): 512 bytes / 512 bytes
+Disklabel type: dos
+Disk identifier: 0x00e61b04
+
+Device         Boot  Start     End Sectors  Size Id Type
+/dev/mmcblk0p1        2048  133119  131072   64M  b W95 FAT32
+/dev/mmcblk0p2      133120 7744511 7611392  3.6G 83 Linux
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Format partitions
+\begin{lstlisting}[]
+# mkfs.vfat -n BOOT ${NFSDEV}p1
+# mkfs.ext4 -L HOME ${NFSDEV}p2
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+%Mount image:
+%\begin{lstlisting}[]
+%# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
+%# mount ${DEV}p2 ${ROOTDIR}
+%# mount ${DEV}p1 ${ROOTDIR}/boot
+%\end{lstlisting}
+%\FloatBarrier
+%\vspace{-5mm}
+
+Copy boot directory:
+\begin{lstlisting}[]
+# cp -r ${ROOTDIR}/boot/* ${NFSBOOTDIR}/ 
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Prepare home directory (Maybe it is better just to copy the content in home from the
+image's root partition):
+\begin{lstlisting}[]
+# install -d -o 1000 -g 1000 -m 700 ${HOMEDIR}/pi && sync
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\begin{lstlisting}[]
+# mount ${NFSDEV}p1 ${NFSROOTDIR}/boot
+# mount ${NFSDEV}p2 ${NFSROOTDIR}/home
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Unmount directories:
+\begin{lstlisting}[]
+# umount ${NFSBOOTDIR} ${NFSHOMEDIR}
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}

### 2016-06-23 testbed.tex
Major push on http image and image fixes

\label{sec:testbed}

We proceed to define the testbed environment by each of its conforming
parts. We later indicate a procedure to set up the testbed. In this procedure,
we summarize all the details regarding on setting the connectivity and
system files for a set of \ac{Raspi}s in a decentralized fashion.
Afterwards, we indicate how to cross-compile C++ code and the Kodo
network coding library in an easy way.
%Finally, we provide further information about Kodo itself in terms
%of testing, other platforms supported and source code documentation
%for further references.

\subsection{System Overview}

\begin{figure}[ht!]
\centering
\includegraphics[width=0.85\textwidth]{images/testbed_setup2.eps}
\caption{Testbed setup}
\label{fig:testbed_setup}
\end{figure}

%This section will present a testbed of thirty \ac{Raspi}s and a \ac{PC}
The testbed consists of thirty \ac{Raspi}s that will be setup and controlled
by a testbed administrator from a \ac{PC}.
%Twenty \ac{Raspi} 1 model B revision 2 and ten \ac{Raspi} 2 model B V1.1.
The \ac{Raspi}s are each equipped with a memory card and a wired and wireless
network interface. They are all connected to the
university network by wire using their ethernet interface (eth0).

To easily discover the \ac{Raspi}s within the university network,
%by a testbed administrator from a \ac{PC}. The \ac{Raspi}s are each equipped
%with a memory card and connected to the
%university network by wire using their ethernet interface (eth0). To easily
%discover the \ac{Raspi}s within the university network,
they have all been assigned a unique name under a common domain.
This eliminates the need to remember their individual \ac{IP} addresses.
A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
We have two types of \ac{Raspi}s in our setup.
Table~\ref{tab:rasp_naming} shows the naming convention.

\begin{table}[ht!]
  \centering
  \begin{tabular}{|c|c|}
    \hline
      \textbf{Hostname} & \textbf{Model} \\ \hline
    rasp00 - rasp19 & Raspberry Pi 1 model B rev. 2 \\ \hline
    rasp20 - rasp29 & Rasperry Pi 2 model B V1.1 \\ \hline
  \end{tabular}
  \caption{Raspberry Pi naming convention}
  \label{tab:rasp_naming}
\end{table}

%Accessing them is easily done using \ac{SSH}.
%To address them within the university network,
%The \ac{Raspi}s

%The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
%connected by wire to the university's \ac{LAN}. The university network
%administration was provided
%the \ac{MAC} addresses of the ethernet interfaces to automatically assign
%the devices the desired \ac{IP} addresses as shown in the figure. In a personal
%\ac{LAN}, another option could be to manually provide the devices with static
%\ac{IP}s.

%This section will be used to present a \ac{Raspi} testbed and the steps to
%install and configure its devices.

%The testbed is depicted in figure~\ref{fig:testbed_setup}. It consists of twenty
%Raspberry Pi 1 model B rev 2 (Rasp01 to Rasp20) and
%ten Raspberry Pi 2 model b V1.1 (Rasp2\_01 to Rasp2\_10).

The testbed is available for students to perform simulations and
measurements as part of their semester projects. The \ac{Raspi}s should
be configured identically, but allow the students to alter the
configurations and how each \ac{Raspi} behaves.
%It is therefore desired
%that the \ac{Raspi}s are configured identically, but still allows the
%students to alter the configurations.

When working on a testbed with many devices, it is not always desired
to make persistent
changes to devices. We present how the concept of stacked
filesystems is used to enable the means of both persistent and
%The testbed should be usable by students to perform simulations and
%measurements as part of their semester projects. It is therefore desired
%that the \ac{Raspi}s are configured identically, but still allows the
%students to alter the configurations. When working on a testbed, it is
%not always desired to make persistent
%changes to the devices. We also present how the concept of stacked
%filesystems can be used to enable the means of both persistent and
temporary storage.

%It is therefore
%mandatory that the \ac{Raspi}s are easily configurable but also
%desired to

%The testbed should be usable by students that should be able to configure
%the \ac{Raspi}'s as desired. It is therefore

%The testbed will be used by students that might need to change how the
%devices are work either temporary or permanently. We will illustrate how this
%can be accomplished using an overlaying filesystem.

%The goal is to configure all the \ac{Raspi}s identically with the same software
%and configurations. There are however a few differences between them that we
%need to take into account. E.g. hostname and network addresses.

%Finally, it should be possible to store data that are generated from
%simulations on the memory cards that are in the individual \ac{Raspi}.
%We using a overlaying filesystem that
%We will illustrate
%how that can be done. In the case of temporary changes, it is important that
%all changes can be easily reverted. We put a \ac{RAM} filesystem on top of
%the the root filesystem to .
%the configure the
%devices to behave according to their test specifications. We
%will also present how that might be done.

%their linkings. That could be to change network config

%The goal is to configure all the devices to be as identical as possible. This
%means that it is desired to have the same software on every device, but

%The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
%connected by wire to the university's \ac{LAN}. The university network
%administration was provided
%the \ac{MAC} addresses of the ethernet interfaces to automatically assign
%the devices the desired \ac{IP} addresses as shown in the figure. In a personal
%\ac{LAN}, another option could be to manually provide the devices with static
%\ac{IP}s.
%that are
%assigned the \ac{IP} addresses in the figure and connected wired to the
%university's \ac{LAN}.

%We use standard usb power adapters to power the \ac{Raspi}s.

\subsubsection{Installing Raspbian}

To get started, we first need to install an \ac{OS} on the
\ac{Raspi}s. We will use the popular Debian-based Raspbian
Linux~\cite{raspbian}

We will be using a \ac{PC} running a Debian-based Linux distribution
%\ac{Raspi}s. We will use Raspbian Linux~\cite{raspbian}. We will be
%using a \ac{PC} running a Debian-based Linux distribution
to setup a customized Raspbian Linux image that can be written to
a memory card and work in any \ac{Raspi} version 1 or 2.
%We will use a desktop that has linux installed to download,
%configure, and write the image to the memory cards.
%We have decided to use a linux desktop machine to create a
%common image of Raspbian Linux that can be used in any of the Raspberry Pis.

%for all devices. After the image has been configured it can be written to
%a memory card and used in any of the Raspberry Pis. This means that each
%device needs a memory card, but it does not matter which card is put into
%which device.

%We have decided to go with Raspbian linux~\cite{raspbian} and
%to install it on memory cards.

%There exists a number of ways to setup a testbed in a centralized manner.
%E.g. using e.g. \ac{NFS}.

%There exists a number of ways to setup a testbed
%
%with multiple devices. We
%have decided on a very simple approach of simply instal

%-
%- Download image (wget)
%- Update/Modify image (chroot)
%- Burn image
%- Bang! linux is running
%-

%Open a terminal, and then...

%The first thing we need to do is to download the linux installation file.
%For Raspbian, this is available at \url{http://downloads.raspberrypi.org}.

Raspbian is made available in two bundles: Raspbian and Raspbian lite.
The difference between the two is that Raspbian contains a pre-installed desktop
environment, and Raspbian lite is by default only accessible through a 
command shell. Both Raspbian and Raspbian lite will both by default be remotely
accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
connect to another machine through a command shell.
%In fact it is the successor
%of \ac{Telnet} that has been designed with network security in mind.

The \ac{Raspi}s in our testbed is not connected to monitors, so we
have decided to go with Raspbian lite.
%Our testbed does not have monitors so we will install Raspbian lite.
A desktop environment can be installed using the package manager
(apt-get) in case a desktop environment will be desired at a later
time.

Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
%connect to another machine through a command shell. In fact it is the successor
%of \ac{Telnet} that has been designed with network security in mind.
%
%Our testbed does not have monitors so we will install Raspbian lite.
%A desktop environment can be installed using Raspbian's package manager
%(apt-get) in case a desktop environment will be desired sometime in the future.
%Below is the procedure we did to setup the \ac{Raspi}s in our testbed. We
have performed a simple setup where each \ac{Raspi} is equipped with a
memory card that contains a slightly customized Raspbian lite image. The image
of the \ac{OS} in all \ac{Raspi}s will be identical.
%with identical settings.

%The reason this can be done easily is because the official Raspbian lite image
%that we will download simply conntains a complete linux installation that are
%packaged in an image rather than written to a harddrive, memory card, or usb
%pen.

The overall procedure to customize the official Raspbian lite image
are:
\begin{enumerate}
    \item Download Raspbian lite
    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
        in the official Raspbian lite image
    \item Change root, i.e. change root filesystem into the Raspbian lite
        image to update and install software packages
    \item Write image to memory cards
\end{enumerate}

Hashmark (\#) and dollar (\$) will be used throughout the code blocks in
the paper to indicate if an operation needs to be run with root permissions or
user permissions.
%In the following, the hashmark (\#) and dollar (\$) signs will be used
%throughout the procedure to indicate if an operation should be run with
%root permissions or user permissions.
%\textbf{hashmark (\#) means to perform operation as root, dollar (\$) means
%to perform operation as user}

\subsection{Download Raspbian lite}

The latest Raspbian lite image can be downloaded at~\cite{raspbian}.
%To download the latest Raspbian lite, we go to the url
%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
At the time of this writing, the latest bundle was
2016-05-27-raspbian-jessie-lite.zip.

we want to proceed downloading the image in a shell instead of the browser.
to do this, first open a command shell on a linux machine. this guide is
assuming a debian-based distribution, but there should only be minor differences
if you have another package manager. i.e. the package managers and name of software packages.

After the terminal has been opened, we start by declaring a few variables to reduce
repeated typing. The first variables we declare is the Raspbian image name.
Notice that the extension has been omitted. This is because the image has been
packed into a zip file. The other variable we declare is a working directory.
This is where we will download the image to and work on it. In other words, it
will be stored in \texttt{/home/<username>/Raspbian}

%To download Raspbian lite, we go to the \url{http://downloads.raspberrypi.org}.
%There should be a folder called raspbian\lite and
%to find the download page at which Raspbian lite should be located. Instead of
%downloading with the browser, we just extract the download url to the latest
%image. At the time of this writing, that was

% Define variable]
\begin{lstlisting}[]
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

A final thing to notice is the way \$ and \# are used in the code blocks. \$
means that the command can be run with normal user permissions where \# indicates that the
command needs root permissions.
Running a command with root permissions can on most systems be done by putting
the command "sudo" in front of each line. This is illustrated in the code below
with the command "whoami" that prints the username of the caller. Lines without
leading \$ or \# is the terminal output.

% Root example
\begin{lstlisting}[]
$ whoami
<USERNAME>
$ sudo whoami
root
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Next, we make the working directory and change directory (cd) into the working
directory. We can then use wget to download Raspbian lite. Unpack the image
when the download is complete.

% Download and unpack image
\begin{lstlisting}[]
$ mkdir -p ${WORKDIR}
$ cd ${WORKDIR}
$ wget http://kom.aau.dk/project/TuneSCode/raspi/${IMAGE}.zip
$ unzip ${IMAGE}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

A more recent version of Raspbian lite may be available at~\cite{raspbian}.
%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.

\subsection{The image}

After Raspbian lite has been unpacked, there should be an .img file in the directory.
"fdisk" can be used to display the content of the image. Pass the arguments "-u sectors" to
display the sizes in sectors and "-l" to display partitions and automatically
exit fdisk again.

% Check out the image   % The dollar hack was to fix vim syntax
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
\end{lstlisting}
\FloatBarrier

The image is in total 2709504 sectors (1.3GiB) in size and contains two
partitions. The first partition starts at sector 8192 and
the other partition starts at sector 137216. The partitions is of type FAT32
and Linux and have sizes 63M and 1.2G. This indicates that the first partition
is most likely a boot partition and the second partition is a traditional Linux file
system. Actually, it is the Linux root filesystem (i.e. /).

\subsection{Resize image}
1.2G is not very much to hold the root filesystem given that it is desired to
in later steps alter the files and install more packages. It may therefore be
necessary to increase the partition size slightly.
The following steps illustrates how the image and its root filesystem
can be expanded by one \ac{GB}.

First, expand the image one \ac{GB} and wait for the data to be completely written
to the disk:
\begin{lstlisting}[]
$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Use \texttt{fdisk} to check that the image is one \ac{GB} bigger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, to expand the root filesystem it is required to first remove the Linux partition
and then add it again one \ac{GB} bigger.
We use \texttt{fdisk} to alter the partition table. The echo's within the parenthesis
will function as
key-presses in \texttt{fdisk} and hence delete last partition, create a new
primary partition, print new partition table, and write new partition table
to the image:

\begin{lstlisting}[]
$ (echo d; echo ; echo n; echo p; echo ; echo 137216; echo ; echo p; echo w) | fdisk ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Use a loopback device to make the Raspbian image available as a block device: 
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
/dev/loop0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

It is seen that the image is now available as /dev/loop0. Use \texttt{lsblk}
to view the partitions:

\begin{lstlisting}[]
# lsblk
NAME      MAJ:MIN RM  SIZE RO TYPE MOUNTPOINT
...
loop0       7:0    0  2.3G  0 loop 
|-loop0p1 259:2    0   63M  0 loop 
|-loop0p2 259:3    0  2.2G  0 loop
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Check and resize the root partition:
\begin{lstlisting}[]
# e2fsck -f ${DEV}p2
# resize2fs ${DEV}p2
\end{lstlisting}
\FloatBarrier

%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
%We also notice that each sector is 512 bytes.
%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"
%Mount root partition:

\subsection{Mount image}

Browsing and altering the files in the image is easy. Simply mount the partitions.
%To browser and alter the files in the image, it is possible to mount the partitions.

Lets first mount the root partition. This is done by creating an empty directory
that we will be used as a mountpoint. We name it \texttt{root} and create it in the
working directory.
After this, simply mount the partitions:
%After this,
%we mount the partition starting at sector 137216. To provide this offset information
%to mount, we need to find the offset in bytes. From the fdisk command we could also
%see that each sector is 512 bytes. Thus, we multiply the sector size and the sector
%start to get the offset in bytes.

\begin{lstlisting}[]
$ export ROOTDIR="${WORKDIR}/root"
$ mkdir -p ${ROOTDIR}
# mount ${DEV}p2 ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%# mount -o loop,offset=$((137216*512)) ${IMAGE}.img ${ROOTDIR}

After this, we can also mount the boot partition inside the newly mounted
root filesystem. This is convenient because that partition will be mounted
on this directory when a \ac{Raspi} starts up with a memory card containing
the image. %It is therefore the natural place to mount it for later purposes.

Mount boot partition:
\begin{lstlisting}[]
# mount ${DEV}p1 ${ROOTDIR}/boot
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%# mount -o loop,offset=$((8192*512)) ${IMAGE}.img ${ROOTDIR}/boot

We can now change all the files in the disk image as desired.

\subsection{Configuring OS files and scripts}

The \ac{Raspi}s should be setup as similar as possible. One of those
few settings that may be different in the devices is their hostname that
helps the user to distinguish the devices from each other.
%The hostname helps distinguish them from each other.
One approach to create unique hostnames would be to rename the hostname
in each memory card. Another way is to create a start up script to automatically
configure the hostname. We decided to define the hostnames based on the \ac{MAC}
addresses of the \ac{Raspi}s wired ethernet interface.

%The \ac{Raspi}s should be setup to be as similar as possible. One of those
%things that is preferred to be different on the devices is their hostname.
%This helps distinguish them from each other. One solution to this in our
%approach would be to uniquely change the hostname in each memory card.
%This is however not desired because that would either require modifying
%the customized Raspbian lite image before writing the content to the
%memory card or to manually change it in the \ac{Raspi}s after the image has
%been written to a memory card.
%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
%in our testbed. Those had to be used to put them in a domain name anyway.
%
%Below is a part of the file where we keep track of the MAC addresses and assign
%hostnames. This file will be available online and copied to the Raspbian lite
%image in a later step under the listed filename. For different \ac{Raspi}s
%the MAC addresses has to be changed. The MAC address of a network card can be found
%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.

%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
%in our testbed. Those had to be used to put them in a domain name anyway.

Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:

%Below is a part of the file where we keep track of the MAC addresses and assign
%hostnames.
%This file will be available online and copied to the Raspbian lite
%image in a later step under the listed filename. For different \ac{Raspi}s
%the MAC addresses has to be changed. The MAC address of a network card can be found
%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.

% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Hostname
b8:27:eb:5b:da:20 rasp00
b8:27:eb:7b:c3:91 rasp01
b8:27:eb:54:9c:64 rasp02
b8:27:eb:95:bd:11 rasp03
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
%the testbed, it still needs to know which name to give itself.
%We do this using a small Bash script:
We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
hostname in \texttt{nodes.csv}.

% Set hostname
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

script_path="$(dirname $(realpath $0))"
config_file=${script_path}/nodes.csv 
mac=$(cat /sys/class/net/eth0/address)
old_hostname=$(hostname)
new_hostname=$(grep $mac $config_file | cut -f2 -d' ')

# Assign hostname found in nodes.csv
if [ ! -z ${new_hostname} ]; then
    echo ${new_hostname} > /etc/hostname
    hostname ${new_hostname}
    sed -i.old -e "s:${old_hostname}:${new_hostname}:g" /etc/hosts
fi
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Line 1 tells the system to interpret the script using Bash. Line 3 gets the
path of the script itself. Line 4 gets the MAC address. This script will be
located in all \ac{Raspi}s and thus return the MAC address of the node itself
when executed.
Line 4 searches line by line for this mac address.
Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
then it will keep the default hostname of the system.

It is likely that the testbed administrator needs to either change or add
scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
all \ac{Raspi}s. One way around this is to make the \ac{Raspi}s fetch the
configuration scripts during start up before running the configuration scripts.

The below script attempts to fetch the scripts:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

url="http://kom.aau.dk/project/TuneSCode/raspi/"
config_file="rasp_config.zip"

# Attempt to fetch new configuration files
if ! wget -O /tmp/${config_file} ${url%/}/${config_file}; then
    echo "Warning: Unable to update rasp_config files"
    exit 1
fi

# Unzip and overwrite configurationn files to root's home directory
unzip -q -o ${config_file} -d /home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We have packed the files and uploaded them to a webserver. These can be
retrieved with \texttt{wget} and unzipped into the Raspbian lite image that
we are customizing:

% Get the files
\begin{lstlisting}[]
$ wget http://kom.aau.dk/project/TuneSCode/raspi/rasp_config.zip
$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For the above to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
the http server from which the \ac{Raspi} retrieves the files from.
%"nodes.csv" can be changed using a text editor. For vim, this could be done
%the following way:
%\begin{lstlisting}[]
%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

Finally, to actually make the \ac{Raspi}s change hostname, we have to make
each \ac{Raspi} call the above script when it starts up. Optionally,
also run the update script before running other scripts.
Do this with root permissions by inserting the line below before
\texttt{"exit 0"} using your favorite editor:
%"\$\{ROOTDIR\}/etc/rc.local" using
%your favorite text editor..
%The file
%can be opened with your favorite text editor.

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/set_hostname
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Configuring \texttt{rc.local} makes it really cumbersome for the testbed
administrator to add new scripts since a new script first needs to be added
in \texttt{rasp\_config.zip} and then run by adding an additional line of code
to \texttt{rc.local} in the root image or in all memory cards depending on
the final testbed setup.

A solution to this is to have yet another script in \texttt{rasp\_config.zip}
that calls all the other scripts in the desired order. This would look like
below:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

bash /home/pi/rasp_config/set_hostname
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, to use the above script, it is required to change \texttt{rc.local}
slightly:
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Notice that \texttt{set\_hostname} is now called through the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at system startup. Had the update script
been called in \texttt{main} instead would mean that changes to the 
\texttt{main} script would also first have been available at next startup.

%This can be done by inserting a call to the script
%set\_hostname in rc.local before it exits.
%
%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
%\begin{lstlisting}[]
%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
%\end{lstlisting}
%\FloatBarrier
%$ line="bash /home/pi/.rasp_config/set_hostname"
%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>

WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.

All other files could be updated as well.

\subsection{Installation and Updating the image}

It may be desired to pre-install some programs in the image before it is
written to all the memory cards that goes into the \ac{Raspi}s.
This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
(change root).
%\url{https://wiki.debian.org/QemuUserEmulation}
%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}

Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
install some additional software:
%Because \ac{Raspi}s are equipped with an \ac{ARM} processor 
%%it is not as
%straightforward as performing the task for two Linux systems of the same
%architecture.

% CHROOT to OS image
\begin{lstlisting}[]
# apt-get install binfmt-support qemu qemu-user-static
# update-binfmts --display qemu-arm
qemu-arm (enabled):
     package = qemu-user-static
        type = magic
      offset = 0
       magic = \x7fELF\x01\x01\x01\x00\x00\x00\x00\x00\x00\x00\x00\x00\x02\x00\x28\x00
        mask = \xff\xff\xff\xff\xff\xff\xff\x00\xff\xff\xff\xff\xff\xff\xff\xff\xfe\xff\xff\xff
 interpreter = /usr/bin/qemu-arm-static
    detector = 
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% This was neccessary on arch
%# update-binfmts --importdir /var/lib/binfmts/ --import

Make sure the second command writes "enabled" as the output above. If that 
is not the case, then try enabling it:

\begin{lstlisting}[]
# update-binfmts --enable qemu-arm
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Provided that qemu-arm is enabled, we should now be able to change root (chroot)
into our Raspbian lite image. Change root is a method i linux that enables
the root (/) to be changed. Thus, enabling a linux installation within a
linux installation.

There are a few commands to be performed before we can change root.
First, to get internet access from within the Raspbian lite image it is needed
to copy our resolv.conf into the filesystem. Line 1 steps into Raspbian lite's
root filesystem. Then line 2 copies resolv.conf.

% CHROOT to OS image
\begin{lstlisting}[]
$ cd $ROOTDIR
# cp /etc/resolv.conf etc/resolv.conf
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, because of the ARM architecture, it is also required to copy
/usr/bin/qemu-arm-static into the image before we can continue:

% CHROOT to OS image
\begin{lstlisting}[]
# cp /usr/bin/qemu-arm-static usr/bin
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The final preparation before chaning root is to populate the directories proc,
sys, and dev:

% CHROOT to OS image
\begin{lstlisting}[]
# mount -t proc proc proc/
# mount --bind /sys sys/
# mount --bind /dev dev/
# mount --bind /dev/pts dev/pts
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, the filesystem is ready for us to change root. This can be done with
"proot". It may be required to install proot:

% CHROOT to OS image
\begin{lstlisting}[]
# proot -q qemu-arm-static -S ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Most available online material uses the more known chroot command as written in
the code block below. This did not work correctly in our machines, but we show
it as an alternative in case proot may not work on other systems.

% CHROOT to OS image
\begin{lstlisting}[]
# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If everything went well, we should now be standing in the Raspbian lite filesystem
under its root. Optionally, we can change the prompt title to indicate that we
are chrooted:

% Optionally, we may create a unique prompt to indicate we have changed root
\begin{lstlisting}[]
# export PS1="(chroot) $PS1"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The Raspbian lite system should now be possible to use as if it had been booted
in a \ac{Raspi}. The only difference is that the laptop that we are using is
significantly faster than a \ac{Raspi}. Thus enabling us to for example update,
upgrade, and install software packages:

% Update system
\begin{lstlisting}[]
(chroot) # apt-get update
(chroot) # apt-get upgrade
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Lets install some useful applications:
% Install packages
\begin{lstlisting}[]
(chroot) # apt-get install vim git screen
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

All the changes that are made here will exists in all \ac{Raspi}s when the
image is written to a memory card.

\subsection{Installation an overlaying filesystem}

Our goal was to enable the users of the testbed to change however the
\ac{Raspi} are configured while still being able to revert those changes easily.

We do this by storing a users work in the \ac{Raspi}s in \ac{RAM}. Thus a
\ac{Raspi}, when a \ac{Raspi} is rebooted, it will forget all changes.

A way to do this is using stacked filesystem. Stacking filesystem enables
a filesystem to overlay another filesystem. Whenever a file is accessed,
the upper filesystem will forward the request to the lower filesystem in
case it does not have it itself. If it has it, it will simpy return the file.

This idea can be used in our setup to mount the root filesystem (i.e. Raspbian
lite) in the \ac{Raspi}s duing startup as read-only. Thus, the files will not
be changable. Some programs needs to change files and the testbed admin might
want to change configurations. To enable this, it is possible to
overlay a filesystem that is mounted in \ac{RAM} as read/writable on top of
the regular root. Thus, reading a file may return a file from the lower filesytem,
but if it is stored, it will be saved in the upper filesystem. Accessing this file
again will return the stored file from the upper layer.

%We do this using an overlay filesystem. The idea is to let the \ac{Raspi}s
%mount their root filesystem as 
%
%We do this using an overlay filesystem (overlayfs).
%Overlayfs enables a filesystem to be stacked on top of another filesystem. Thus,
%if the \ac{Raspi}s mount their root filesystem (/) as read-onlyif we mount our root (/) filesystem as 
%
%
%Overlayfs enables filesystems
%to be stacked on top of each other. We will be using this to make a lower
%filesystem and an upper filesystem. The lower filesystem will be our
%traditional filesystem and the upper filesystem will be mounted in tmpfs
%(i.e. mounted in \ac{RAM}).
%
%Thus, when a file is requsted from stacked filesystem, it will first attempt
%to fetch the file in the upper filesystem if it does not exists there, it will
%request it from the lower filessytem.

Assuming that you are still \texttt{chrooted} into the Raspbian lite filesystem, we
can setup the overlay filesystem (overlayfs).

There already exists implementations overlaying the root filesystem. We will
use an existing implementation
available at github: \cite{overlayroot}.

We installed git in a previous step, so lets clone the repository. The below
code stores it in /tmp which is really mounted in \ac{RAM}. Thus, the files
will disappear when the machine is rebooted.

% Get files
\begin{lstlisting}[]
(chroot) # OVERLAYROOTDIR="/tmp/overlayroot"
(chroot) # git clone https://github.com/chesty/overlayroot.git $OVERLAYROOTDIR
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To enable the overlaying filesystem it is required to generate an initial \ac{RAM}
file system (initramfs). Initramfs is an initial filesystem that gets loaded into
\ac{RAM} during the startup process of a linux machine to prepare the real
filesystem.

For this purpose, we need BuxyBox:

% Install required packages
\begin{lstlisting}[]
(chroot) # apt-get install busybox
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Add and activate the overlaying filesystem, we need to add the scripts and tell initramfs
to load the overlaying filesytem:

% Setup initramfs
\begin{lstlisting}[]
(chroot) # cp ${OVERLAYROOTDIR}/hooks-overlay /etc/initramfs-tools/hooks/
(chroot) # cp ${OVERLAYROOTDIR}/init-bottom-overlay /etc/initramfs-tools/scripts/init-bottom/
(chroot) # echo "overlay" > /etc/initramfs-tools/modules
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The bootloader should be made aware of the initramfs that we will generate:
%Raspbian lite uses "u-boot" to boot, so it should be made aware of the
%initramfs that we will generate.

% Tell uboot to load initramfs
\begin{lstlisting}[]
(chroot) # echo "initramfs init.gz" > /boot/config.txt
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, lets generate the image. This is done using "mkinitramfs", that will
by default look for the kernel modules of the current system. Remember we are
chrooted, so we need to specify the correct kernel modules to look for. These
can be found in "/lib/modules".

% Create an initramfs
\begin{lstlisting}[]
(chroot) # ls /lib/modules/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For \ac{Raspi} 1, our kernel modules was "4.4.11+". The other modules in /lib/modules
is for \ac{Raspi} 2.

Thus, to run mkinitramfs, type in the correct kernel version:

% Create an initramfs
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.11+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Although there might be printet a few warnings, this should generate
a new working \texttt{initramfs} that will put a \ac{RAM} filesystem
on top of the root filesystem.

At this point, it is no longer required to be \texttt{chrooted} and the image
can be written to a memory card. The following commands exits the \texttt{chrooted}
environment, unmount all partitions, and detaches the loop device:

% Exit chroot, umount, and write to memory card
\begin{lstlisting}[]
(chroot) # exit
# cd ..
# umount --recursive ${ROOTDIR}
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{(Optional) Persistent subdirectories}

Provided the stacked filesystem was configured, it may be desired to have
some directories where files are not flushed upon rebooting the \ac{Raspi}s.
The following procedure will create an extra partition in the image for
the \ac{Raspi} users home directory that will be made storage persistent.
%Provided that a stacked filesystem was configured to make the root
%filesystem forget chances upon reboots is not always desired.

Expand image according to the desired home directory size. Keep in mind not
to make the image bigger than the target memory card.
% Expand image
\begin{lstlisting}[]
$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create home partition after the root partition. Use \texttt{fdisk} to
find the next available sector in the image. In our case, the new
partition should start at sector 4806656:
\begin{lstlisting}[]
$ fdisk -lu ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create new partition:

\begin{lstlisting}[]
$ (echo n; echo p; echo ; echo 4806656; echo ; echo p; echo w) | fdisk ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create loopback device: 
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
/dev/loop0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Format new partition:
\begin{lstlisting}[]
# mkfs.ext4 ${DEV}p3
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The last thing to remember is to inform Raspbian to mount the home partition
we created during startup. This can be done by adding an entry in \texttt{fstab}:
\begin{lstlisting}[]
# mount ${DEV}p2 ${ROOTDIR}
# sudo sh -c "echo '/dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' >> ${ROOTDIR}/etc/fstab"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The home folder was originally located in the root filesystem. We should
therefore copy the content to the new home partition:
\begin{lstlisting}[]
# mount ${DEV}p3 ${ROOTDIR}/mnt
# mv ${ROOTDIR}/home/* ${ROOTDIR}/mnt/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

That is it. unmount partitions and detach loop device:
\begin{lstlisting}[]
# umount --recursive ${ROOTDIR}
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Write image to memory card}

The final step is to write the image to all the memory cards before they can
be used in the \ac{Raspi}s.

%This is done by first leaving the \texttt{chrooted}
%filesystem and unmounting all the directories we mounted:

% Exit chroot, umount, and write to memory card
%\begin{lstlisting}[]
%# exit
%# cd ..
%# umount --recursive ${ROOTDIR}
%# losetup -d ${DEV}.img
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

%If successfully unmounted, you can 
Insert a memory card in your machine. It will be available as
\texttt{/dev/mmcblkX} or \texttt{/dev/sdX}

%In most systems, this memory card was available under \texttt{/dev/mmcblk0}, but this may be
%different on different machines. For internal memory card readers, it will
%probably be named /dev/mmcblkX where X is some number and if the memory card
%reader is connected through \ac{USB}, then it is likely called /dev/sdX.
Make absolutely sure to write to the correct device as everything will be 
overwritten!
Inserting/Ejecting the memory card a couple of times and using the commands
below may help deducing the correct device:
%The commands below may help deduce which device is the correct:
%Below is a few commands that may help you to deduce which device is correct:

\begin{lstlisting}[]
# lsblk
# df -h
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For us, the device was \texttt{/dev/mmcblk0}. Write the image to a memory card:
\begin{lstlisting}[]
# dd if=${IMAGE}.img of=/dev/mmcblk0 bs=4M && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The memory card can now be started in a \ac{Raspi}.

### 2016-06-23 testbed_http.tex
Major push on http image and image fixes

@@
 
-Mount http root and home directories
+Mount http root and home directories (ADD LOOP OPTION HERE WHEN MOUNTING)
 \begin{lstlisting}[]
@@
 
+#. /scripts/functions # Already loaded in the init script
+
 # Create mount point directory
@@
 # Parse additional command line options
-for x in ${cat /proc/cmdline}; do
+for x in $(cat /proc/cmdline); do
     case $x in
@@
         IMAGEDEV=${IMAGE%:*}
-        IMAGEPATH="/imagedir/${${IMAGE#*:}#/}"
+        IMAGEPATH="${IMAGE#*:}"
+        IMAGEPATH="/imagedir/${IMAGEPATH#/}"
         ;;
     url=*)
-        URL=${${x#url=}%/}  # If url has trailing slash, then remove it
+        URL=${x#url=}
+        URL=${URL%/}  # If url has trailing slash, then remove it
         ;;
+    esac
 done
@@
     mount -r -n -t ext4 -o nodiratime,noatime /dev/loop0 /root
+
+    # Bind /root/home to /home
+    #mount -n -o rbind /home /root/home
 }
@@
 }
-
 \end{lstlisting}
@@
 \begin{lstlisting}[]
-init=http ro ip=dhcp image=/dev/mmcblk0p2:/root.img root=/dev/loop0 url=http://192.168.87.106/raspi/
+boot=http ro ip=dhcp image=/dev/mmcblk0p2:/root.img root=/dev/loop0 url=http://192.168.87.106/raspi/
 \end{lstlisting}
@@
 /dev/loop0      /               ext4    defaults,noatime  0       1
-/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       1
+/dev/mmcblk0p2  /home           ext4    defaults,noatime  0       0
 \end{lstlisting}
@@
 
+Unmount:
+\begin{lstlisting}[]
+# umount ${HTTPROOTIMAGE}/{dev/pts,proc,sys,dev}  
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\subsubsection{Prepare sd-card}
+
 IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?
@@
 
+MAYBE ALSO SET THE BOOT FLAG FOR FIRST PARTITION. ALSO UPDATE SCRIPT IF THIS
+WORKS. THERE MAY HAVE TO BE ECHOED A 1 AFTER echo a FOR SETTING BOOT FLAG
 \begin{lstlisting}[]
@@
     echo t; echo b;                             # Set boot partition type to FAT
+    echo a;                                     # Set boot flag
     echo n; echo p; echo 2; echo ; echo ;       # Create home partition
@@
 
-%Mount image:
-%\begin{lstlisting}[]
-%# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-%# mount ${DEV}p2 ${ROOTDIR}
-%# mount ${DEV}p1 ${ROOTDIR}/boot
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-Copy boot directory:
+Mount the boot partition from the memory card. Copy the content from the
+boot image and unmount it again:
 \begin{lstlisting}[]
-# cp -r ${ROOTDIR}/boot/* ${HTTPBOOTDIR}/ 
+# mount ${HTTPDEV}p1 ${HTTPROOTDIR}/mnt
+# cp -r ${HTTPROOTDIR}/boot/* ${HTTPROOTDIR}/mnt/ && sync
+# umount ${HTTPROOTDIR}/mnt
 \end{lstlisting}
@@
 
-Prepare home directory (Maybe it is better just to copy the content in home from the
-image's root partition):
+Mount the home partition from the memory card. Copy the content from the
+root image and unmount it again:
 \begin{lstlisting}[]
-# install -d -o 1000 -g 1000 -m 700 ${HOMEDIR}/pi && sync
+# mount ${HTTPDEV}p2 ${HTTPROOTDIR}/mnt
+# cp -r ${HTTPROOTDIR}/home/* ${HTTPROOTDIR}/mnt/ && sync
+# umount ${HTTPROOTDIR}/mnt
 \end{lstlisting}
@@
 \begin{lstlisting}[]
-# mount ${HTTPDEV}p1 ${HTTPROOTDIR}/boot
-# mount ${HTTPDEV}p2 ${HTTPROOTDIR}/home
+# umount --recursive ${HTTPROOTDIR}
 \end{lstlisting}
@@
 
-Unmount directories:
+\subsubsection{(Optional) Compressing root image}
+
+PROBABLY A BETTER IDEA TO USE BTRFS TRANSPARENT COMPRESSION
+
 \begin{lstlisting}[]
-# umount ${HTTPBOOTDIR} ${HTTPHOMEDIR}
+# apt-get install squashfs-tools
+# mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR} 
+# mksquashfs ${HTTPROOTDIR} compressed_${HTTPROOTIMAGE}.img       
 \end{lstlisting}
@@
 \vspace{-5mm}
+
+\subsubsection{Try it out}
+
+Upload the root image (\$\{HTTPROOTIMAGE\}.img) to the HTTP server used in the
+configurations above. Then, unplug the memory card from the computer.
+Insert it into a \ac{Raspi} and
+turn it on. It is recommended that the \ac{Raspi} is turned on with a monitor
+connected the first time the newly created Linux installation is tested to
+assure that Raspbian starts up correctly.
+
+PRESENT FILES ON HTTP. E.G. HTTPROOT.IMG AND IMAGES.CSV

### 2016-06-23 setup_scripts/http_sdcard.sh

@@ -3,7 +3,7 @@
 set -e -u
 
 # Target device
-dev=/dev/mmcblk0
+dev="$1"   # E.g. /dev/mmcblk0
 
 # Print error
 echoerr() { cat <<< "$@" 1>&2; }
@@ -41,13 +41,17 @@ fdisk -lu ${dev}
 mkfs.vfat -n BOOT ${dev}p1
 mkfs.ext4 -L HOME ${dev}p2
 
-# SHOULD THE BELOW BE COPIED FROM THE ROOT IMAGE?
+# LETS COPY THE CONTENT IN HOME FROM THE ROOT IMAGE INSTEAD TO MAKE SURE
+# THE CORRECT USER DIRECTORIES ARE CREATED (OTHER USERS MIGHT HAVE BEEN
+# CREATED IN THE ROOT IMAGE)
+
+exit 0 # Skip the rest of the script
 
 # Create temp directory that is erased when script exits
 mnt=$(mktemp -d)
 
-# Create home directory for user: pi (uid=1000) in group pi (gid=1000). (gid=100 if group users are preferred)
-# mode=700 means: drwx------ (i.e. only the user and root has access to his directory)
+# Create home directory for user: pi (uid=1000) in group pi (gid=1000) or group users (gid=100)
+# mode=700 means: drwx------ (i.e. only the user (pi) and root has access to the directory)
 mount ${dev}p2 ${mnt}
 install -d -o 1000 -g 1000 -m 700 ${mnt}/pi && sync
 

### 2016-06-25 testbed.tex
worked on testbed

@@
 # Attempt to fetch new configuration files
-if ! wget -O /tmp/${config_file} ${url%/}/${config_file}; then
+if ! wget -O -q --show-progress /tmp/${config_file} ${url%/}/${config_file}; then
     echo "Warning: Unable to update rasp_config files"
@@
 # Unzip and overwrite configurationn files to root's home directory
-unzip -q -o ${config_file} -d /home/pi/
+unzip -q -o /tmp/${config_file} -d /home/pi/
 \end{lstlisting}
@@
 
-\subsection{Installation and Updating the image}
+\subsection{Installing and updating the image}
 
@@
 $ cd $ROOTDIR
-# cp /etc/resolv.conf etc/resolv.conf
+# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
 \end{lstlisting}
@@
 \begin{lstlisting}[]
-# cp /usr/bin/qemu-arm-static usr/bin
+# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
 \end{lstlisting}
@@
 
-The bootloader should be made aware of the initramfs that we will generate:
-%Raspbian lite uses "u-boot" to boot, so it should be made aware of the
-%initramfs that we will generate.
-
-% Tell uboot to load initramfs
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init.gz" > /boot/config.txt
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Finally, lets generate the image. This is done using "mkinitramfs", that will
-by default look for the kernel modules of the current system. Remember we are
-chrooted, so we need to specify the correct kernel modules to look for. These
-can be found in "/lib/modules".
+%Finally, lets generate the initial \ac{RAM} filesystem. This is done using the command
+%\texttt{mkinitramfs}, that will by default look for the kernel modules of the current system. Remember we are
+%chrooted, so we need to specify the correct kernel modules to look for. These
+%can be found in "/lib/modules".
+Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
+needs a different kernel than \ac{Raspi} version 2 and version 3. Thus, start
+by finding the available kernel modules:
+%Because \texttt{initramfs} is executed within a
+%\texttt{chrooted} environment, it will not be able to detect the
+%kernel modules in the \texttt{chrooted} environment. The kernel modules will
 
@@
 (chroot) # ls /lib/modules/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For \ac{Raspi} 1, our kernel modules was "4.4.11+". The other modules in /lib/modules
-is for \ac{Raspi} 2.
-
-Thus, to run mkinitramfs, type in the correct kernel version:
+4.4.13+  4.4.13-v7+
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Kernel "4.4.13+" is for \ac{Raspi} version 1 and kernel "4.4.13-v7+" works
+for \ac{Raspi} version 2 and version 3.
+
+Lets proceed by generating an initial \ac{RAM} filesystem for both kernels:
+
+%Lets proceed with kernel modules "4.4.13+".
+%Lets create an initial ram filesystem for both kernels.
+
+%Thus, to run mkinitramfs, type in the correct kernel version:
 
@@
 \begin{lstlisting}[]
-(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.11+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Although there might be printet a few warnings, this should generate
-a new working \texttt{initramfs} that will put a \ac{RAM} filesystem
-on top of the root filesystem.
+(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
+(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
+%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+
+
+Although there might be printet a few warnings, the above commands should successfully
+generate working initial \ac{RAM} filesystems.
+
+An initial \ac{RAM} filesystem needs to be called by the bootloader. In Raspbian
+this is done by adding a command to \texttt{config.txt} in the boot partition.
+
+If the system should be run in a \ac{Raspi} version 1, then use \texttt{init.gz}:
+%The bootloader should be made aware of the initramfs that we will generate:
+\begin{lstlisting}[]
+(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+If the system should be run in a \ac{Raspi} version 2 or 3, then use \texttt{init-v7.gz}:
+\begin{lstlisting}[]
+(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # Only for Raspberry Pi version 2 and 3
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
 
@@
 can be written to a memory card. The following commands exits the \texttt{chrooted}
-environment, unmount all partitions, and detaches the loop device:
+environment, unmount all partitions, and detaches the loopback device:
 
@@
 
-\subsection{(Optional) Persistent subdirectories}
+\subsection{(Optional) Persistent subdirectories for distributed Linux image}
+
+EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
+WRITTEN TO THE MEMORY CARDS
 
@@
 \subsection{Write image to memory card}
+
+EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
+WRITTEN TO THE MEMORY CARDS
 

### 2016-06-25 testbed_http.tex
worked on testbed

@@
 
+%This part will explain how to prepare the existing image to be downloaded from
+%a \ac{HTTP} server.
+The Linux image that has been made so far has two partitions. One 
+partition with the boot files and one containing the root file system.
+
+This section will present how the partitions can be extracted from the image,
+and how the boot partition can be configured to download the root filesystem
+from a \ac{HTTP} server during startup.
+
+The advantage of this setup is that multiple \ac{Raspi}s can share the same
+root image. This significantly eases the maintainance task.
+
+%An alternative to the setup explained in this procedure is to use \ac{NFS}
+%instead of \ac{HTTP}.
+
+%The first program called when a \ac{Raspi} starts is the Boot Loader,
+%\texttt{U-Boot}. \texttt{U-Boot} 
+%
+%When Raspbian starts up, the first program called
+%
+%The procedure explained below will customize the initial \ac{RAM} filesystem
+%
+%
+%there exists only one root file
+%
+%The advantage of this is that the root filesystem only one all \ac{Raspi}s can automatically download the 
+%
+%The boot partition will be required in the \ac{Raspi}s, but the root image can
+
+Start by declaring the path and names of the boot- and root-image:. 
 \begin{lstlisting}[]
@@
 
-Extract root partition from image
+The, setup a loopback device of the existing image and use \texttt{dd} to
+extract the boot and root partitions from the image:
 \begin{lstlisting}[]
@@
 # dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
+# losetup -d ${DEV}
 \end{lstlisting}
@@
 
-Mount http root and home directories (ADD LOOP OPTION HERE WHEN MOUNTING)
+Now it is time to configure the boot- and root-image just extracted before
+the content of the boot image is copied to the a memory card and the root
+image is uploaded to a \ac{HTTP} server.
+
+Create a mountpoint and mount the http root and boot images:
+%Mount the http root image and home directories
 \begin{lstlisting}[]
@@
 $ mkdir -p ${HTTPROOTDIR}
-# mount ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
-# mount ${HTTPBOOTIMAGE}.img ${HTTPROOTDIR}/boot
+# mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
+# mount -o loop ${HTTPBOOTIMAGE}.img ${HTTPROOTDIR}/boot
 \end{lstlisting}
@@
 \vspace{-5mm}
+
+In a previous steps, the boot loader was configured to call an initial \ac{RAM}
+filesystem. The coming steps will present how to the initial \ac{RAM}
+filesystem can be extended further to download and cache the root filesystem
+from a \ac{HTTP} server before mounting it.
+
+The initial \ac{RAM} filesystem contains an \texttt{init} script that
+sets up the prepares the yis first
+called. This script initializes the  may be rewritten, but
 
@@
 # Fetch, cache, and mount root image using HTTP
-
-#. /scripts/functions # Already loaded in the init script
 
@@
     # If the image is not cached, then get it
-    if [ ! -f ${img} ] && wget --spider -q "${URL}/${img}"; then
+    if [ ! -f "${IMAGEPATH%/*}/${img}" ] && wget --spider -q "${URL}/${img}"; then
         # Delete old image
@@
 
-Add the below to \$\{HTTPROOTDIR\}/boot/cmdline.txt (CHANGE URL HERE) (remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
-\begin{lstlisting}[]
-boot=http ro ip=dhcp image=/dev/mmcblk0p2:/root.img root=/dev/loop0 url=http://192.168.87.106/raspi/
+Keep the content in \texttt{cmdline.txt}, but add/modify the variables
+\texttt{boot}, \texttt{ip}, \texttt{image}, \texttt{root}, and \texttt{url}.
+
+%Keep also all other variables as they are in
+%\texttt{cmdline.txt}.
+
+%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
+(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
+    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
 \end{lstlisting}
@@
 
-Edit fstab to look like this (exclude the last line if persistent
+Here is how the variables are used in the initramfs:
+\begin{itemize}
+    \item boot=http: tells the init script in initramfs to call the \texttt{http} script presented above
+    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the initial ram filesystem
+    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
+    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
+    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
+\end{itemize}
+
+%The \ac{URL} should be modified to reflect where
+%the \ac{Raspi}s should download the http root image. The \texttt{image} variable informs the
+%\texttt{http} initramfs script to download and cache the http root image in the second partition
+%on its memory card at path \texttt{/httproot.img}.
+
+Edit fstab to look like this (exclude the last line if a persistent
 home storage directory is not desired):
@@
 \begin{lstlisting}[]
-(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.11+
+(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
+(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
 \end{lstlisting}
@@
 \begin{lstlisting}[]
+$ cd ..
 # umount ${HTTPROOTIMAGE}/{dev/pts,proc,sys,dev}  
@@
 # mount ${HTTPDEV}p1 ${HTTPROOTDIR}/mnt
-# cp -r ${HTTPROOTDIR}/boot/* ${HTTPROOTDIR}/mnt/ && sync
+# cp -rp ${HTTPROOTDIR}/boot/* ${HTTPROOTDIR}/mnt/ && sync
 # umount ${HTTPROOTDIR}/mnt
@@
 # mount ${HTTPDEV}p2 ${HTTPROOTDIR}/mnt
-# cp -r ${HTTPROOTDIR}/home/* ${HTTPROOTDIR}/mnt/ && sync
+# cp -rp ${HTTPROOTDIR}/home/* ${HTTPROOTDIR}/mnt/ && sync
 # umount ${HTTPROOTDIR}/mnt
@@
 PRESENT FILES ON HTTP. E.G. HTTPROOT.IMG AND IMAGES.CSV
+
+\begin{itemize}
+    \item Copy http root image (\$\{HTTPROOTIMAGE\}.img) a HTTP server
+    \item Create images.csv
+\end{itemize}
+
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{http://<SERVERIP>/.../images.csv}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+# Ethernet MAC    Root image
+b8:27:eb:5b:da:20 httproot_1.img
+b8:27:eb:7b:c3:91 httproot_1.img
+b8:27:eb:7b:c3:91 httproot_2.img
+...
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Notice that the \ac{Raspi}s can be configured to download different root images.
+This is useful both for having various root images for subsets of the \ac{Raspi}s,
+but also as a basic way inform the \ac{Raspi}s that the root image changed such
+that they can download the new root image at next boot.
+
+As of this writing, there is unfortunately a bug in wget that makes wget not showing
+its progress bar while downloading the root image at startup. Instead, it may appear
+to be unable to connect although it is downloading, so keep patient.
+
+\begin{lstlisting}[]
+Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
+image.csv       100% |*******************************************************| 1287   0:00:00 ETA
+Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
+_
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+REMEMBER TO COPY-PASTE ALL CONTENT FROM SCRIPT INTO LATEX AGAIN IN CASE SOME
+SCRIPT CHANGED, BUT WAS NOT UPDATED IN LATEX

### 2016-06-28 testbed.tex
..

@@
 we summarize all the details regarding on setting the connectivity and
-system files for a set of \ac{Raspi}s in a decentralized fashion.
-Afterwards, we indicate how to cross-compile C++ code and the Kodo
-network coding library in an easy way.
+system files for a set of \ac{Raspi}s in a decentralized and/or centralized fashion
+depending on the testbed admins preferences..
+%Afterwards, we indicate how to cross-compile C++ code and the Kodo
+%network coding library in an easy way.
 %Finally, we provide further information about Kodo itself in terms
@@
 \centering
-\includegraphics[width=0.85\textwidth]{images/testbed_setup2.eps}
+\includegraphics[width=0.45\textwidth]{images/testbed_setup3.eps}
 \caption{Testbed setup}
@@
 
-%This section will present a testbed of thirty \ac{Raspi}s and a \ac{PC}
-The testbed consists of thirty \ac{Raspi}s that will be setup and controlled
-by a testbed administrator from a \ac{PC}.
-%Twenty \ac{Raspi} 1 model B revision 2 and ten \ac{Raspi} 2 model B V1.1.
-The \ac{Raspi}s are each equipped with a memory card and a wired and wireless
-network interface. They are all connected to the
-university network by wire using their ethernet interface (eth0).
-
-To easily discover the \ac{Raspi}s within the university network,
-%by a testbed administrator from a \ac{PC}. The \ac{Raspi}s are each equipped
-%with a memory card and connected to the
-%university network by wire using their ethernet interface (eth0). To easily
-%discover the \ac{Raspi}s within the university network,
-they have all been assigned a unique name under a common domain.
-This eliminates the need to remember their individual \ac{IP} addresses.
+The testbed consists of up to one hundred \ac{Raspi}s of different models.
+More specifically, Raspberry Pi 1 model B rev.~2,
+Raspberry Pi 2 model B V1.1, and Raspberry Pi 3 model B V1.2.
+All \ac{Raspi}s are each equipped with a memory card, a wired, and a wireless
+network interface. They connect to the university network using its wired
+ethernet interface that is named \texttt{eth0} according to the legacy naming
+convention of ethernet interfaces in Linux. \textcolor{red}{MAYBE PUT SOUECE HERE}
+
+All \ac{Raspi}s will be running an \ac{SSH} daemon for easy remote access
+within the university network. We therefore had the university's \ac{IT}
+department configure the university's \ac{DHCP} server to assign each \ac{Raspi} a
+static \ac{IP}. This eliminates the requirements for monitors and keyboards
+for the \ac{Raspi}s for non-graphical applications.
+%This significantly eases the management of the \ac{Raspi}s
+%as \ac{SSH} eliminates the need for monitors and keyboards for the \ac{Raspi}s.
+
 A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
-We have two types of \ac{Raspi}s in our setup.
-Table~\ref{tab:rasp_naming} shows the naming convention.
-
-\begin{table}[ht!]
-  \centering
-  \begin{tabular}{|c|c|}
-    \hline
-      \textbf{Hostname} & \textbf{Model} \\ \hline
-    rasp00 - rasp19 & Raspberry Pi 1 model B rev. 2 \\ \hline
-    rasp20 - rasp29 & Rasperry Pi 2 model B V1.1 \\ \hline
-  \end{tabular}
-  \caption{Raspberry Pi naming convention}
-  \label{tab:rasp_naming}
-\end{table}
-
-%Accessing them is easily done using \ac{SSH}.
-%To address them within the university network,
-%The \ac{Raspi}s
-
-%The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
-%connected by wire to the university's \ac{LAN}. The university network
-%administration was provided
-%the \ac{MAC} addresses of the ethernet interfaces to automatically assign
-%the devices the desired \ac{IP} addresses as shown in the figure. In a personal
-%\ac{LAN}, another option could be to manually provide the devices with static
-%\ac{IP}s.
-
-%This section will be used to present a \ac{Raspi} testbed and the steps to
-%install and configure its devices.
-
-%The testbed is depicted in figure~\ref{fig:testbed_setup}. It consists of twenty
-%Raspberry Pi 1 model B rev 2 (Rasp01 to Rasp20) and
-%ten Raspberry Pi 2 model b V1.1 (Rasp2\_01 to Rasp2\_10).
-
-The testbed is available for students to perform simulations and
-measurements as part of their semester projects. The \ac{Raspi}s should
-be configured identically, but allow the students to alter the
-configurations and how each \ac{Raspi} behaves.
-%It is therefore desired
-%that the \ac{Raspi}s are configured identically, but still allows the
-%students to alter the configurations.
-
-When working on a testbed with many devices, it is not always desired
-to make persistent
-changes to devices. We present how the concept of stacked
-filesystems is used to enable the means of both persistent and
-%The testbed should be usable by students to perform simulations and
-%measurements as part of their semester projects. It is therefore desired
-%that the \ac{Raspi}s are configured identically, but still allows the
-%students to alter the configurations. When working on a testbed, it is
-%not always desired to make persistent
-%changes to the devices. We also present how the concept of stacked
-%filesystems can be used to enable the means of both persistent and
-temporary storage.
-
-%It is therefore
-%mandatory that the \ac{Raspi}s are easily configurable but also
-%desired to
-
-%The testbed should be usable by students that should be able to configure
-%the \ac{Raspi}'s as desired. It is therefore
-
-%The testbed will be used by students that might need to change how the
-%devices are work either temporary or permanently. We will illustrate how this
-%can be accomplished using an overlaying filesystem.
-
-%The goal is to configure all the \ac{Raspi}s identically with the same software
-%and configurations. There are however a few differences between them that we
-%need to take into account. E.g. hostname and network addresses.
-
-%Finally, it should be possible to store data that are generated from
-%simulations on the memory cards that are in the individual \ac{Raspi}.
-%We using a overlaying filesystem that
-%We will illustrate
-%how that can be done. In the case of temporary changes, it is important that
-%all changes can be easily reverted. We put a \ac{RAM} filesystem on top of
-%the the root filesystem to .
-%the configure the
-%devices to behave according to their test specifications. We
-%will also present how that might be done.
-
-%their linkings. That could be to change network config
-
-%The goal is to configure all the devices to be as identical as possible. This
-%means that it is desired to have the same software on every device, but
-
-%The \ac{Raspi}s are all equipped with an ethernet interface (eth0) and
-%connected by wire to the university's \ac{LAN}. The university network
-%administration was provided
-%the \ac{MAC} addresses of the ethernet interfaces to automatically assign
-%the devices the desired \ac{IP} addresses as shown in the figure. In a personal
-%\ac{LAN}, another option could be to manually provide the devices with static
-%\ac{IP}s.
-%that are
-%assigned the \ac{IP} addresses in the figure and connected wired to the
-%university's \ac{LAN}.
-
-%We use standard usb power adapters to power the \ac{Raspi}s.
+
+The testbed should be available for students to perform simulations and
+measurements as part of their semester projects. We intend to configure
+all \ac{Raspi}s identically, while still allowing students to alter
+how each \ac{Raspi} behaves and is configured.
+
+When working on the testbed with this many devices, it may not always be
+desired to make persistent changes to devices. We present how the concept
+of stacked filesystems may be used to enable the means of both persistent
+and temporary storage.
+
+%The testbed should be available for students to perform simulations and
+%measurements as part of their semester projects. The \ac{Raspi}s should
+%be configured identically, but allows the students to alter the
+%configurations and how each \ac{Raspi} behaves.
+%
+%When working on a testbed with many devices, it may not always desired
+%to make persistent changes to devices. We present how the concept of stacked
+%filesystems may be used to enable the means of both persistent and
+%temporary storage.
+
+%A \ac{HTTP} server will be used to store Linux configuration files as well
+%as the root filesystem in the centralized approach.
+
+The following procedure will explain how the testbed may be setup and
+configured both in the traditional distributed fasion where the
+entire Linux filesystem is written to an memory card, but also in a more
+centralized fasion where 
+is written to
+Based on the popular Debian based Raspbian Linux distribution
+We
+first present how a standard Linux distribution for the \ac{Raspi}s
+may be customized
 
@@
 We will be using a \ac{PC} running a Debian-based Linux distribution
-%\ac{Raspi}s. We will use Raspbian Linux~\cite{raspbian}. We will be
-%using a \ac{PC} running a Debian-based Linux distribution
 to setup a customized Raspbian Linux image that can be written to
 a memory card and work in any \ac{Raspi} version 1 or 2.
-%We will use a desktop that has linux installed to download,
-%configure, and write the image to the memory cards.
-%We have decided to use a linux desktop machine to create a
-%common image of Raspbian Linux that can be used in any of the Raspberry Pis.
-
-%for all devices. After the image has been configured it can be written to
-%a memory card and used in any of the Raspberry Pis. This means that each
-%device needs a memory card, but it does not matter which card is put into
-%which device.
-
-%We have decided to go with Raspbian linux~\cite{raspbian} and
-%to install it on memory cards.
-
-%There exists a number of ways to setup a testbed in a centralized manner.
-%E.g. using e.g. \ac{NFS}.
-
-%There exists a number of ways to setup a testbed
-%
-%with multiple devices. We
-%have decided on a very simple approach of simply instal
-
-%-
-%- Download image (wget)
-%- Update/Modify image (chroot)
-%- Burn image
-%- Bang! linux is running
-%-
-
-%Open a terminal, and then...
-
-%The first thing we need to do is to download the linux installation file.
-%For Raspbian, this is available at \url{http://downloads.raspberrypi.org}.
 
@@
 Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
-%connect to another machine through a command shell. In fact it is the successor
-%of \ac{Telnet} that has been designed with network security in mind.
-%
-%Our testbed does not have monitors so we will install Raspbian lite.
-%A desktop environment can be installed using Raspbian's package manager
-%(apt-get) in case a desktop environment will be desired sometime in the future.
-%Below is the procedure we did to setup the \ac{Raspi}s in our testbed. We
 have performed a simple setup where each \ac{Raspi} is equipped with a
@@
 of the \ac{OS} in all \ac{Raspi}s will be identical.
-%with identical settings.
-
-%The reason this can be done easily is because the official Raspbian lite image
-%that we will download simply conntains a complete linux installation that are
-%packaged in an image rather than written to a harddrive, memory card, or usb
-%pen.
 
@@
 user permissions.
-%In the following, the hashmark (\#) and dollar (\$) signs will be used
-%throughout the procedure to indicate if an operation should be run with
-%root permissions or user permissions.
-%\textbf{hashmark (\#) means to perform operation as root, dollar (\$) means
-%to perform operation as user}
 
@@
 
-It is seen that the image is now available as /dev/loop0. Use \texttt{lsblk}
+It is seen that the image is now available as \texttt{/dev/loop0}. Use \texttt{lsblk}
 to view the partitions:

### 2016-06-29 conclusions.tex
.

\label{sec:conclusions}
%Given the need for low-cost, easy to configure, testbeds that enable
%reproducible results, we provide a in-depth description of Aalborg
%University's \ac{Raspi} testbed and how to guarantee replicable
%results and managing scaling of the system.  We also provide a
%performance evaluation for network coding schemes, focusing on
%processing speed and energy behaviour for two \ac{Raspi} models,
%including algorithms that exploit \ac{SIMD} instructions and multicore
%capabilities of the \ac{Raspi} 2. Our measurements show that
%processing speeds of more than 80 Mbps and 800 Mbps are attainable for
%the \ac{Raspi} model 1 and 2, respectively, for a wide range of
%network coding configurations and maintaining a processing energy
%below 1 nJ/bit (or even an order of magnitude lower) in similar
%configurations. Future work in the use of \ac{Raspi} devices will
%focus on expanding the setup and automation of tasks to run the
%testbed, configure specified network topologies (e.g., with specific
%connectivity or packet loss ratios), reserve the use of these
%sub-networks for running tailored experiments, and opening the use of
%the testbed beyond our team at Aalborg University.

### 2016-06-29 introduction.tex
.

%!TEX root = raspi_journal.tex
% At the beginning, there was darkness and then... bang! \ac{NC}
% \cite{ahlswede2000network} appeared to save us from the evilness
% of routing.
%
% General introduction. Introduction to topic addressed in the journal.
% Review of the State of the Art. Specify why our approach has benefits
% and which are they. Indicate contributions.

%The emergence of powerful and inexpensive single-board computers
%running full versions of a standard \ac{OS} from many possible in the last
%years opens exciting possibilities for researchers. This not only
%allows to seamlessly develop implementations that are compatible with
%higher end devices and allows the reuse stable software from the Open
%Source community. It also enables the deployment and testing of large
%scale (hundreds of devices or more) distributed systems at a fraction
%of the cost necessary in previous years. This paper focuses on
%providing the critical steps and mechanisms followed to deploy, setup,
%and maintain a large-scale testbed based on Raspberry Pi devices of
%different models. We also complement this with an in-depth evaluation
%of the Raspberry Pi as a platform to enable a new paradigm in
%networking and storage known as \ac{NC}, which relies on
%processing at intermediate nodes to achieve higher throughput, delay
%and reliability performance than state-of-the-art routing techniques.
%
%
%% INTRODUCTION TO NETWORK CODING
%Introduced by Ahlswede et al.~\cite{ahlswede2000network}, network
%coding constitutes a paradigm shift in the way how researchers and
%industry understand and operate networks, by changing the role of
%intermediate relays in the process of transmission of information.
%Relays are no longer limited to storing and forwarding data, but also
%take part in the coding process, through a process called recoding,
%where the relay generates new linear combinations of incoming coded
%packets without previously decoding the data. Network coding allows
%the increase of throughput, reliability, security and delay
%performance of the networks. 
%
%Compared to other broadly used coding schemes, e.g., Gallager's LDPC
%codes~\cite{gallager1962low} or Reed-Solomon
%codes~\cite{reed1960polynomial}, network coding is a technology that
%has been implemented in real systems since the early years of its
%conception. In 2006, Katabi et al.~\cite{katabi2006practical} published
%the results of the performance of their protocol COPE when implemented
%in a real wireless system, which relied on minimalistic coding to
%provide interesting gains. In 2008, Pedersen et
%al.~\cite{pedersen2008implementation} used commercially available
%Symbian OS enabled mobile phones to implement network coding in a
%\ac{D2D} cooperation-based application. In 2011, researchers at Aalborg
%University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
%coding library with open source code for researchers intended to make
%network coding implementations easily available for both, the research
%community and commercial entities.
%Another case is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
%implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
%IEEE 802.11 multi-hop meshed networks using some of the intuition from
%COPE, but deploying it in real systems. It is currently available as
%open source code in the Linux kernel. Many other implementations have
%been tested on real world systems~
%\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.
%
%Despite this successful deployment in real systems, many of these
%protocols and contributions have been implemented in separate testbeds
%and the experiences are hard to reproduce.  Given the need for
%reproducibility of past and future results among researchers and the
%need of low-cost and easy-to-deploy testbeds, this paper contribution
%is provide detailed instructions on how to set up a low-cost,
%heterogeneous and scalable testbed consisting of Raspberry Pis
%(Raspis), and by providing detailed measurements of goodput and energy
%consumption of such testbed when performing network coding operations
%with different codecs (such as full \ac{RLNC} \cite{ho2006random},
%multicore-enabled \ac{RLNC}, sparse \ac{RLNC} and tunable sparse \ac{RLNC}).
%The heterogeneity of the testbed comes from the fact that old, single core,
%Raspberry Pi 1 model B units are integrated seamlessly in the described
%testbed with newer, multicore versions, like the Raspberry Pi 2 model B.
%Our testbed allows us to obtain a comprehensive set of measurements
%to characterize the goodput and energetic footprint of these devices
%showing that processing speeds of 800 Mbps and energy per bit values
%of 1 nJ or even one order of magnitude less are possible.
%
%Our work is organized as follows. Section~\ref{sec:testbed} introduces the
%testbed system, its setup, files configuration and connectivity.
%Section~\ref{sec:cross_compilation} elaborates on the compilation of
%the Kodo library for the \ac{Raspi}. Section~\ref{sec:schemes} defines the
%coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
%we describe the considered metrics and methodology for performance comparison
%of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
%show the measurements in the \ac{Raspi} models of the mentioned metrics. Final
%conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
%
%%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications should be cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the main conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.

### 2016-06-29 testbed.tex
.

@@
 ethernet interface that is named \texttt{eth0} according to the legacy naming
-convention of ethernet interfaces in Linux. \textcolor{red}{MAYBE PUT SOUECE HERE}
+convention of ethernet interfaces in Linux~\cite{PredictableNetworkInterfaceNames}.
 
@@
 A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
+The testbed is intended for students to perform simulations and measurements
+as part of semester projects. We therefore intend to configure all
+\ac{Raspi}s identically, while still allowing students to alter how each
+\ac{Raspi} behaves and is configured.
+
+The following produre explains a way to setup the testbed to facilitates these
+requirements. The procedure is assumed to be performed by whom we will refer
+to as the testbed administrator in a \ac{PC} running Linux. The exact Linux
+distribution is does not matter, but the procedure will be based on
+Debian Linux.
 
@@
 
-The following procedure will explain how the testbed may be setup and
-configured both in the traditional distributed fasion where the
-entire Linux filesystem is written to an memory card, but also in a more
-centralized fasion where 
-is written to
-Based on the popular Debian based Raspbian Linux distribution
-We
-first present how a standard Linux distribution for the \ac{Raspi}s
-may be customized
+The following procedure explains how the testbed may be setup and configured.
+
+%The following procedure will explain how the testbed may be setup and
+%configured both in the traditional distributed fasion where the
+%entire Linux filesystem is written to an memory card, but also in a more
+%centralized fasion where 
+%is written to
+%Based on the popular Debian based Raspbian Linux distribution
+%We
+%first present how a standard Linux distribution for the \ac{Raspi}s
+%may be customized
 

### 2016-06-29 testbed.tex
.

@@
 
-All \ac{Raspi}s will be running an \ac{SSH} daemon for easy remote access
+All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
 within the university network. We therefore had the university's \ac{IT}
@@
 The testbed is intended for students to perform simulations and measurements
-as part of semester projects. We therefore intend to configure all
+as part of semester projects. We intend to configure all
 \ac{Raspi}s identically, while still allowing students to alter how each
@@
 
-The following produre explains a way to setup the testbed to facilitates these
-requirements. The procedure is assumed to be performed by whom we will refer
-to as the testbed administrator in a \ac{PC} running Linux. The exact Linux
-distribution is does not matter, but the procedure will be based on
-Debian Linux.
-
-The testbed should be available for students to perform simulations and
-measurements as part of their semester projects. We intend to configure
-all \ac{Raspi}s identically, while still allowing students to alter
-how each \ac{Raspi} behaves and is configured.
-
-When working on the testbed with this many devices, it may not always be
-desired to make persistent changes to devices. We present how the concept
-of stacked filesystems may be used to enable the means of both persistent
-and temporary storage.
+The following section propose a procedure to setup the testbed to facilitates these
+requirements. The procedure is performed, by whom we will refer
+to as the testbed administrator, in a regular \ac{PC} running Linux. The exact Linux
+distribution does not matter although the procedure will be based on
+Debian Linux. All the commands in this paper will be performed on
+the testbed administrators \ac{PC}.
+
+The procedure goes as follows. Based on a regular Raspbian Linux image, it will
+be explained how it can be customized to include desired packages and
+configured as desired.
+When working on a testbed with many devices, it may not always be desired to make
+persistent changes to devices. We therefore present how the concept of stacked
+filesystems may be used to enable the meands of both persistent and temporary
+storage. This step of the procedure will be optional if the testbed
+administrator decides to write the whole image to a memory card.
+Writing the whole Linux filesystem to memory cards may complicate the
+maintenance as e.g. updates need to be performed in each \ac{Raspi}.
+\texttt{Python Fabric} may ease this task, but we also propose alternative
+approach where the initial \ac{RAM} filesystem is customized in a way that
+enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP}
+server. This enables the testbed administrator to only update the filesystem
+in one location. The \ac{Raspi}s will then get the changes after a reboot.
+
+%\textcolor{red}{The testbed should be available for students to perform simulations and
+%measurements as part of their semester projects. We intend to configure
+%all \ac{Raspi}s identically, while still allowing students to alter
+%how each \ac{Raspi} behaves and is configured.
+%\\\\
+%When working on the testbed with this many devices, it may not always be
+%desired to make persistent changes to devices. We present how the concept
+%of stacked filesystems may be used to enable the means of both persistent
+%and temporary storage.}
 

### 2016-06-29 introduction.tex
added franks intro

@@
 % and which are they. Indicate contributions.
+
+The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale.
+This will break new ground for new markets such as driverless cars, manufacturing, humanoid robots, smart grids, and many more.
+With 5G the number of wireless devices will increase by two fold up to 500 billion devices.
+It is generally accepted, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connect due to the expected energy consumption.
+Therefore the communication architecture in future communication scenario dominated by mesh technologies.
+
+Mesh technology is known for ages for sensor networks, ad hoc networks, or mobile cloud scenarios, but the technical requirements on 5G mesh communication systems are dramatically increasing.
+Future mesh networks need to support high data rate, low latency, security, network availability, and heterogeneity of devices.
+In state of the art systems those requirements are traded off with each other, but in 5G context we cannot do this anymore.
+
+In our previous works we have shown that random linear network coding [MURIEL] is able to satisfy the aforementioned technical requirements.
+We have actually shown how to increase the throughput [PEYMAN], reduce the delay [WEB1], or support heterogeneity for coding enabled communication nodes [FULCRUM].
+But most of the work was focusing on small mesh networks with a handful of communication nodes.
+Introduced by Ahlswede et al.
+[1], network coding constitutes a paradigm shift in the way how researchers and industry understand and operate networks, by changing the role of intermediate relays in the process of transmission of information.
+Relays are no longer limited to storing and forwarding data, but also take part in the coding process, through a process called recoding, where the relay generates new linear combinations of incoming coded packets without previously decoding the data.
+Network coding allows the increase of throughput, reliability, security and delay performance of the networks.
+
+Based on the need for large scale mesh communication testbeds, in the paper we show a step by step tutorial how to build up a low cost testbed with several communication nodes.
+The emergence of powerful and inexpensive single-board computers running full versions of a standard Operating System (OS) from many possible in the last years opens exciting possibilities for researchers.
+This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from the Open Source community.
+It also enables the deployment and testing of large scale (hundreds of devices or more) distributed systems at a fraction of the cost necessary in previous years.
+This paper focuses on providing the critical steps and mechanisms followed to deploy, setup, and maintain a large-scale testbed based on Raspberry Pi devices of different models as well as including random linear network coding into the testbed.
 
@@
 %and maintain a large-scale testbed based on Raspberry Pi devices of
-%different models. We also complement this with an in-depth evaluation
-%of the Raspberry Pi as a platform to enable a new paradigm in
-%networking and storage known as \ac{NC}, which relies on
-%processing at intermediate nodes to achieve higher throughput, delay
-%and reliability performance than state-of-the-art routing techniques.
-%
+%different models. 
 %
@@
 %performance of the networks.
-%
+
 %Compared to other broadly used coding schemes, e.g., Gallager's LDPC

### 2016-06-29 testbed.tex
added franks intro

@@
 
-The \ac{Raspi}s in our testbed is not connected to monitors, so we
+The \ac{Raspi}s in our testbed are not connected to monitors, so we
 have decided to go with Raspbian lite.
@@
 
-The latest Raspbian lite image can be downloaded at~\cite{raspbian}.
+The latest Raspbian lite bundle can be downloaded from Raspbian's official
+webpage~\cite{raspbian}.
 %To download the latest Raspbian lite, we go to the url
 %\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
-At the time of this writing, the latest bundle was
-2016-05-27-raspbian-jessie-lite.zip.
-
-we want to proceed downloading the image in a shell instead of the browser.
-to do this, first open a command shell on a linux machine. this guide is
-assuming a debian-based distribution, but there should only be minor differences
-if you have another package manager. i.e. the package managers and name of software packages.
-
-After the terminal has been opened, we start by declaring a few variables to reduce
-repeated typing. The first variables we declare is the Raspbian image name.
-Notice that the extension has been omitted. This is because the image has been
-packed into a zip file. The other variable we declare is a working directory.
-This is where we will download the image to and work on it. In other words, it
-will be stored in \texttt{/home/<username>/Raspbian}
+At the time of this writing, the latest available bundle was
+\texttt{2016-05-27-raspbian-jessie-lite.zip}.
+
+To ensure that the content of the bundle does not change, this procedure
+will be based on that particular version of Raspbian lite which we have
+made available at~\cite{tunescode_webpage}. All other files used in this paper
+will also be available there unless specified otherwise.
+
+%We want to proceed downloading the bundle in a shell instead of the browser.
+%To do this, first open a command shell on a linux machine. This guide is
+%assuming a Debian-based distribution, but there should only be minor differences
+%if you have another package manager. i.e. the package managers and name of software packages.
+
+To get started, open a Linux shell (terminal) and declare the variables below.
+The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
+located and \texttt{WORKDIR} specifies a working directory where we will 
+store Raspbian and customize it.
+Do not type in \$ and \# in the shell. Those indicates if a command can be
+executed as a regular user or if root permissions are required.
+Running a command as root can on most systems be done by putting
+\texttt{sudo} in front of the command. This is illustrated in the
+following code block 
+with the command \texttt{whoami} that prints the username. Lines within a code
+block without leading \$ or \# is terminal output or content of a file.
+
+%After the terminal has been opened, we start by declaring a few variables to reduce
+%repeated typing. The first variables we declare is the Raspbian image name.
+%Notice that the extension has been omitted. This is because the image has been
+%packed into a zip file. The other variable we declare is a working directory.
+%This is where we will download the image to and work on it. In other words, it
+%will be stored in \texttt{/home/<username>/Raspbian}
 
@@
 \begin{lstlisting}[]
+$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
 $ export IMAGE="2016-05-27-raspbian-jessie-lite"
@@
 
-A final thing to notice is the way \$ and \# are used in the code blocks. \$
-means that the command can be run with normal user permissions where \# indicates that the
-command needs root permissions.
-Running a command with root permissions can on most systems be done by putting
-the command "sudo" in front of each line. This is illustrated in the code below
-with the command "whoami" that prints the username of the caller. Lines without
-leading \$ or \# is the terminal output.
+%A final thing to notice is the way \$ and \# are used in the code blocks. \$
+%means that the command can be run with normal user permissions where \# indicates that the
+%command needs root permissions.
+%Running a command with root permissions can on most systems be done by putting
+%the command "sudo" in front of each line. This is illustrated in the code below
+%with the command "whoami" that prints the username of the caller. Lines without
+%leading \$ or \# is the terminal output.
 
@@
 
-Next, we make the working directory and change directory (cd) into the working
-directory. We can then use wget to download Raspbian lite. Unpack the image
-when the download is complete.
+Next, make the working directory and used \texttt{cd} to change into the working
+directory. Then \texttt{wget} can be used to download Raspbian lite before
+unpacking the \texttt{zip} file:
 
@@
 $ cd ${WORKDIR}
-$ wget http://kom.aau.dk/project/TuneSCode/raspi/${IMAGE}.zip
+$ wget ${URL%/*}/${IMAGE}.zip
 $ unzip ${IMAGE}.zip
@@
 
-A more recent version of Raspbian lite may be available at~\cite{raspbian}.
+%A more recent version of Raspbian lite may be available at~\cite{raspbian}.
 %\url{http://downloads.raspberrypi.org/raspbian_lite/images}.
@@
 
-After Raspbian lite has been unpacked, there should be an .img file in the directory.
-"fdisk" can be used to display the content of the image. Pass the arguments "-u sectors" to
-display the sizes in sectors and "-l" to display partitions and automatically
-exit fdisk again.
+After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
+\texttt{fdisk} can be used to display the content of the image. Optionally, parse the arguments
+"-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
+image and and automatically exit \texttt{fdisk} again:
 
@@
 
-The image is in total 2709504 sectors (1.3GiB) in size and contains two
+The output provides some useful knowledge about the image. First, the image is in total
+2709504 sectors (1.3GiB) in size and contains two
 partitions. The first partition starts at sector 8192 and

### 2016-06-29 testbed.tex
worked more on testbed description

@@
 After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
-\texttt{fdisk} can be used to display the content of the image. Optionally, parse the arguments
+\texttt{fdisk} can be used to display the content of the image. Parse the arguments
 "-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
@@
 
-The output provides some useful knowledge about the image. First, the image is in total
+The output provides knowledge about the image. The image is in total
 2709504 sectors (1.3GiB) in size and contains two
@@
 the other partition starts at sector 137216. The partitions is of type FAT32
-and Linux and have sizes 63M and 1.2G. This indicates that the first partition
+and Linux and are 63M and 1.2G big. This indicates that the first partition
 is most likely a boot partition and the second partition is a traditional Linux file
@@
 \subsection{Resize image}
-1.2G is not very much to hold the root filesystem given that it is desired to
-in later steps alter the files and install more packages. It may therefore be
-necessary to increase the partition size slightly.
-The following steps illustrates how the image and its root filesystem
+1.2G is not very much to store the root filesystem given that it is desired to
+alter the files and install additional packages. It may therefore be
+necessary to increase the partition size.
+The following illustrates how the image and its root filesystem
 can be expanded by one \ac{GB}.
@@
 
-Use \texttt{fdisk} to check that the image is one \ac{GB} bigger:
+Use \texttt{fdisk} to see that the image is now one \ac{GB} larger:
 \begin{lstlisting}[literate={DOLLAR}{\$}1]
@@
 and then add it again one \ac{GB} bigger.
-We use \texttt{fdisk} to alter the partition table. The echo's within the parenthesis
-will function as
-key-presses in \texttt{fdisk} and hence delete last partition, create a new
+Use \texttt{fdisk} to alter the partition table. The \texttt{echo}'s within the parenthesis
+will be interpreted as
+key-presses in \texttt{fdisk} and hence delete the last partition, create a new
 primary partition, print new partition table, and write new partition table
@@
 
-It is seen that the image is now available as \texttt{/dev/loop0}. Use \texttt{lsblk}
-to view the partitions:
+\texttt{lsblk} may be used to list block devices. It appears that the image is
+available as \texttt{/dev/loop0} which has two partitions \texttt{loop0p1} and
+\texttt{loop0p2}:
+%Use \texttt{lsblk} to view the partitions:
 
@@
 that we will be used as a mountpoint. We name it \texttt{root} and create it in the
-working directory.
-After this, simply mount the partitions:
+working directory before mounting the root filesystem onto the mountpoint.
+%After this, mount the partitions:
 %After this,
@@
 
-After this, we can also mount the boot partition inside the newly mounted
-root filesystem. This is convenient because that partition will be mounted
+The root filesystem already has a boot directory that can be used as
+the mountpoint for the boot partition.
+%After this, we can also mount the boot partition inside the newly mounted
+%root filesystem.
+This is convenient because that partition will be mounted
 on this directory when a \ac{Raspi} starts up with a memory card containing
 the image. %It is therefore the natural place to mount it for later purposes.
-
-Mount boot partition:
+Now, mount boot partition:
 \begin{lstlisting}[]
@@
 
-We can now change all the files in the disk image as desired.
+%We can now change all the files in the disk image as desired.
+It is now possible to change all files within the Raspbian image as desired.
+We will take advantage of this to edit configuration files, append new files,
+and even update and install packages.
 

### 2016-06-29 testbed.tex
.

@@
 The \ac{Raspi}s should be setup as similar as possible. One of those
-few settings that may be different in the devices is their hostname that
-helps the user to distinguish the devices from each other.
+few settings that may be different in the devices is their hostname.
+%The hostname helps the user to distinguish the devices from each other.
 %The hostname helps distinguish them from each other.
-One approach to create unique hostnames would be to rename the hostname
+One approach to create unique hostnames would be to write the hostname
 in each memory card. Another way is to create a start up script to automatically
 configure the hostname. We decided to define the hostnames based on the \ac{MAC}
-addresses of the \ac{Raspi}s wired ethernet interface.
+addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
+addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
+administrator will need to insert the \ac{MAC} addresses and hostnames of
+his/her \ac{Raspi}s.
 
@@
 
-Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
+%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
 
@@
 %We do this using a small Bash script:
-We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
-hostname in \texttt{nodes.csv}.
+%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
+%hostname in \texttt{nodes.csv}.
+A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:
 
@@
 
-Line 1 tells the system to interpret the script using Bash. Line 3 gets the
-path of the script itself. Line 4 gets the MAC address. This script will be
-located in all \ac{Raspi}s and thus return the MAC address of the node itself
-when executed.
-Line 4 searches line by line for this mac address.
-Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
-then it will keep the default hostname of the system.
+The code lines above do the following:
+\begin{itemize}
+    \item Line 1: Tells system to interpret the script using \ac{Bash}
+    \item Line 3-4: Gets the path to the script itself and the list of hostnames
+    \item Line 5: Gets the \ac{MAC} address of the node itself
+    \item Line 6: Gets the current hostname
+    \item Line 7: Gets the new hostname from the hostname list
+    \item Line 10-14: Assigns the new hostname
+\end{itemize}
+
+%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
+%path of the script itself. Line 4 gets the MAC address. This script will be
+%located in all \ac{Raspi}s and thus return the MAC address of the node itself
+%when executed.
+%Line 4 searches line by line for this mac address.
+%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
+%then it will keep the default hostname of the system.
 
@@
 scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
-all \ac{Raspi}s. One way around this is to make the \ac{Raspi}s fetch the
+all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
 configuration scripts during start up before running the configuration scripts.
 
-The below script attempts to fetch the scripts:
+The below script attempts to download/update the scripts:
 
@@
 # Attempt to fetch new configuration files
-if ! wget -O -q --show-progress /tmp/${config_file} ${url%/}/${config_file}; then
+if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
     echo "Warning: Unable to update rasp_config files"
@@
 
+\begin{itemize}
+    \item Line 3-4: Is the url and zip file that should be downloaded
+    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
+        and prints a warning in case of errors
+    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
+\end{itemize}
+
 We have packed the files and uploaded them to a webserver. These can be
-retrieved with \texttt{wget} and unzipped into the Raspbian lite image that
-we are customizing:
+retrieved with \texttt{wget} and unzipped into the Raspbian lite image:
 
@@
 \begin{lstlisting}[]
-$ wget http://kom.aau.dk/project/TuneSCode/raspi/rasp_config.zip
+$ wget ${URL%/*}/rasp_config.zip
 $ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
@@
 
-For the above to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
+For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
 \ac{MAC} address is found in \texttt{nodes.csv}. Thus,
 \texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
-the http server from which the \ac{Raspi} retrieves the files from.
+the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
+The testbed administrator will therefore need to also edit the \ac{URL} in the
+script \texttt{update\_rasp\_config}.
 %"nodes.csv" can be changed using a text editor. For vim, this could be done
@@
 also run the update script before running other scripts.
-Do this with root permissions by inserting the line below before
+Do this with root permissions by inserting the line(s) below before
 \texttt{"exit 0"} using your favorite editor:
@@
 Configuring \texttt{rc.local} makes it really cumbersome for the testbed
-administrator to add new scripts since a new script first needs to be added
-in \texttt{rasp\_config.zip} and then run by adding an additional line of code
-to \texttt{rc.local} in the root image or in all memory cards depending on
+administrator to add new scripts since it first needs to be included
+in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional 
+line added to \texttt{rc.local} in the root image or in all memory cards depending on
 the final testbed setup.
 
-A solution to this is to have yet another script in \texttt{rasp\_config.zip}
+A solution to this is to make yet another script in \texttt{rasp\_config.zip}
 that calls all the other scripts in the desired order. This would look like
@@
 
-Now, to use the above script, it is required to change \texttt{rc.local}
-slightly:
+Now, to use the above script, \texttt{rc.local} needs to look like this:
 \Suppressnumber\begin{lstlisting}[]
@@
 \vspace{-5mm}
-Notice that \texttt{set\_hostname} is now called through the \texttt{main}
+Notice that \texttt{set\_hostname} is now called by the \texttt{main}
 script instead. The update script is still called directly. This ensures that
 all configuration scripts are updated before executed. Changes to the update
-script itself will first take effect at system startup. Had the update script
-been called in \texttt{main} instead would mean that changes to the 
-\texttt{main} script would also first have been available at next startup.
+script itself will first take effect at next system startup.
+%Had the update script
+%been called in \texttt{main} instead would mean that changes to the 
+%\texttt{main} script would also first have been available at next startup.
 
@@
 
-WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
-FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
-
-All other files could be updated as well.
+%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
+%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
+
+%All other files could be updated as well.
+It is not only possible to add and configure files to/within the image. One can
+also install and update the packages within the image which will be illustrated
+in the next section.
 
@@
 Provided that qemu-arm is enabled, we should now be able to change root (chroot)
-into our Raspbian lite image. Change root is a method i linux that enables
-the root (/) to be changed. Thus, enabling a linux installation within a
-linux installation.
-
-There are a few commands to be performed before we can change root.
+into our Raspbian lite image. Change root is a method in Linux that enables
+the root (/) to be changed. Thus, enabling a Linux installation within a
+Linux installation.
+
+There are a few commands to be performed before actually changing root into the
+root partion of the image.
 First, to get internet access from within the Raspbian lite image it is needed
-to copy our resolv.conf into the filesystem. Line 1 steps into Raspbian lite's
+to copy our \texttt{resolv.conf} into the filesystem. Line 1 steps into Raspbian lite's
 root filesystem. Then line 2 copies resolv.conf.

### 2016-06-29 testbed.tex
updated testbed

@@
 First, to get internet access from within the Raspbian lite image it is needed
-to copy our \texttt{resolv.conf} into the filesystem. Line 1 steps into Raspbian lite's
-root filesystem. Then line 2 copies resolv.conf.
+to copy our \texttt{resolv.conf} into the filesystem. Line 1 change directory
+into root filesystem of the image. The second line copies \texttt{resolv.conf}
+from the testbed administrators \ac{PC} to the image.
 
@@
 
-Now, because of the ARM architecture, it is also required to copy
-/usr/bin/qemu-arm-static into the image before we can continue:
+Now, because of the \ac{ARM} architecture,
+\texttt{/usr/bin/qemu-arm-static} needs to be copied into the image before continuing:
 
@@
 
-The final preparation before chaning root is to populate the directories proc,
-sys, and dev:
+The final preparation before changing root is to populate the directories \texttt{proc},
+\texttt{sys}, and \texttt{dev}:
 
@@
 
-Finally, the filesystem is ready for us to change root. This can be done with
-"proot". It may be required to install proot:
+Finally, it is time to change root.
+%Finally, the filesystem is ready for us to change root.
+This can be done with \texttt{proot}. It may be required to install \texttt{proot}
+using "\texttt{apt-get install proot}":
 
@@
 
-Most available online material uses the more known chroot command as written in
-the code block below. This did not work correctly in our machines, but we show
-it as an alternative in case proot may not work on other systems.
+Most available material online uses the more known \texttt{chroot} command as
+written in the code block below instead of \texttt{proot}. This did not work
+correctly in our machines, but we present the command as an alternative to
+\texttt{proot} in case the testbed administrator should be in the opposite
+situration where \texttt{proot} is not working.
 
@@
 
-If everything went well, we should now be standing in the Raspbian lite filesystem
-under its root. Optionally, we can change the prompt title to indicate that we
-are chrooted:
+If everything went well, we should now be \texttt{chrooted} into the Raspbian lite
+filesystem under its root. Optionally, change the prompt title to indicate that it
+is a \texttt{chrooted} environment:
 
@@
 
-The Raspbian lite system should now be possible to use as if it had been booted
-in a \ac{Raspi}. The only difference is that the laptop that we are using is
-significantly faster than a \ac{Raspi}. Thus enabling us to for example update,
-upgrade, and install software packages:
+The Raspbian lite system image should now be possible to use almost as if it
+had been booted in a \ac{Raspi}.
+A major difference is that the testbed administrator's \ac{PC} is likely
+significantly faster than a \ac{Raspi}.
+%The only difference is that the laptop that we are using is
+%significantly faster than a \ac{Raspi}.
+Thus enabling updates, upgrades, and installing new software packages much
+faster than in the \ac{Raspi}s.
+%Thus enabling us to for example update,
+%upgrade, and install software packages:
 
@@
 
-Lets install some useful applications:
+We also install a few packages that we find useful:
+%Lets install some useful applications:
 % Install packages
@@
 
-All the changes that are made here will exists in all \ac{Raspi}s when the
-image is written to a memory card.
+When writing the image to a memory card, all these change that has been made
+to the image so far will exists in all \ac{Raspi}s.
+%All the changes that are made here will exists in all \ac{Raspi}s when the
+%image is written to a memory card.
 

### 2016-06-30 introduction.tex
Add references and massage introduction

--- 2016-06-29 introduction.tex
@@
 % and which are they. Indicate contributions.
+
+The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale.
+This will break new ground for new markets such as driverless cars, manufacturing, humanoid robots, smart grids, and many more.
+With 5G the number of wireless devices will increase by 5$\times$ up to 50 billion devices \cite{cisco2011forecast}.
+It is generally accepted, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connect due to the expected energy consumption.
+Therefore the communication architecture in future communication scenario will be dominated by mesh technologies.
+Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing.
+Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices.
+In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.
+
+Introduced by Ahlswede et al.
+\cite{ahlswede2000network}, network coding constitutes a paradigm shift in the way how researchers and industry understand and operate networks, by changing the role of intermediate relays in the process of transmission of information.
+Relays are no longer limited to storing and forwarding data, but also take part in the coding process, through a process called recoding, where the relay generates new linear combinations of incoming coded packets without previously decoding the data.
+Network coding allows the increase of throughput, reliability, security and delay performance of the networks.
+In our previous works, we have shown that \ac{RLNC} \cite{koetter2003algebraic,ho2006random} is able to satisfy the aforementioned technical requirements.
+We have actually shown how to increase the throughput \cite{pahlevani2013playncool}, reduce the delay \cite{szabo2015towards}, or support heterogeneity for coding enabled communication nodes \cite{lucani2014fulcrum}.
+
+In all these prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities.
+However, most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude.
+Moreover, despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce.
+Furthermore, deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently.
+The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages.
+Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintenable fashion.
+
+The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area.
+This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from widely-used \ac{OS}.
+Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using Raspberry Pi devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
+Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with Raspberry Pi.
+Our work is organized as follows.
+Section~\ref{sec:testbed} introduces the testbed system, its setup, files configuration and connectivity.
+Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}.
+Final conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
 
@@
 %
-%Our work is organized as follows. Section~\ref{sec:testbed} introduces the
-%testbed system, its setup, files configuration and connectivity.
-%Section~\ref{sec:cross_compilation} elaborates on the compilation of
-%the Kodo library for the \ac{Raspi}. Section~\ref{sec:schemes} defines the
-%coding schemes employed in our study. Later, in Section~\ref{sec:metrics}
-%we describe the considered metrics and methodology for performance comparison
-%of the codes deployed in the \ac{Raspi}. In Section~\ref{sec:measurements}, we
-%show the measurements in the \ac{Raspi} models of the mentioned metrics. Final
-%conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
 %

### 2016-06-30 conclusions.tex
New conclusions

@@
 \label{sec:conclusions}
-%Given the need for low-cost, easy to configure, testbeds that enable
-%reproducible results, we provide a in-depth description of Aalborg
-%University's \ac{Raspi} testbed and how to guarantee replicable
-%results and managing scaling of the system.  We also provide a
-%performance evaluation for network coding schemes, focusing on
-%processing speed and energy behaviour for two \ac{Raspi} models,
-%including algorithms that exploit \ac{SIMD} instructions and multicore
-%capabilities of the \ac{Raspi} 2. Our measurements show that
-%processing speeds of more than 80 Mbps and 800 Mbps are attainable for
-%the \ac{Raspi} model 1 and 2, respectively, for a wide range of
-%network coding configurations and maintaining a processing energy
-%below 1 nJ/bit (or even an order of magnitude lower) in similar
-%configurations. Future work in the use of \ac{Raspi} devices will
-%focus on expanding the setup and automation of tasks to run the
-%testbed, configure specified network topologies (e.g., with specific
-%connectivity or packet loss ratios), reserve the use of these
-%sub-networks for running tailored experiments, and opening the use of
-%the testbed beyond our team at Aalborg University.
+OBsersving the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enable reproducible results for mechanisms to address problems in this are, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed and how to guarantee replicability and scaling managing of the system.
+The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges.
+The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and administrable way.
+Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

### 2016-06-30 testbed.tex
Starting to modify the document structure

@@
 
-We proceed to define the testbed environment by each of its conforming
-parts. We later indicate a procedure to set up the testbed. In this procedure,
-we summarize all the details regarding on setting the connectivity and
-system files for a set of \ac{Raspi}s in a decentralized and/or centralized fashion
-depending on the testbed admins preferences..
+% We proceed to define the testbed environment by each of its conforming
+% parts. We later indicate a procedure to set up and configure the testbed
+% for several \ac{Raspi}s. In this procedure, we summarize all the key details
+% regarding on setting the system files, configuration packages, tools and connectivity
+% for the \ac{Raspi}s in a centralized fashion.
+
 %Afterwards, we indicate how to cross-compile C++ code and the Kodo
@@
 
-\subsection{System Overview}
+%\subsection{System Overview and Design Criteria}
+
+A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
+The testbed consists of up to 100 \ac{Raspi}s of different models.
+More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
+Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
+All \ac{Raspi}s are each equipped with a memory card, a wired and wireless
+network interface and a power supply. All the \ac{Raspi} are connected to a common
+\ac{LAN} that provides internal and external connectivity. Without loss of generality,
+in our case they are connected to the university network using their wired Ethernet interface
+that is named \texttt{eth0} according to the legacy naming convention of Ethernet
+interfaces in Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
+university network since we aim the testbed to be used by students and professors
+to perform measurements and experimentation of controlled and reproducible scenarios
+as part of academic research. We intend to configure all \ac{Raspi}s identically,
+while still allowing the end-users to alter how each \ac{Raspi} behaves and is configured.
 
@@
 
-The testbed consists of up to one hundred \ac{Raspi}s of different models.
-More specifically, Raspberry Pi 1 model B rev.~2,
-Raspberry Pi 2 model B V1.1, and Raspberry Pi 3 model B V1.2.
-All \ac{Raspi}s are each equipped with a memory card, a wired, and a wireless
-network interface. They connect to the university network using its wired
-ethernet interface that is named \texttt{eth0} according to the legacy naming
-convention of ethernet interfaces in Linux~\cite{PredictableNetworkInterfaceNames}.
-
-All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
-within the university network. We therefore had the university's \ac{IT}
-department configure the university's \ac{DHCP} server to assign each \ac{Raspi} a
-static \ac{IP}. This eliminates the requirements for monitors and keyboards
-for the \ac{Raspi}s for non-graphical applications.
 %This significantly eases the management of the \ac{Raspi}s
@@
 
-A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
-The testbed is intended for students to perform simulations and measurements
-as part of semester projects. We intend to configure all
-\ac{Raspi}s identically, while still allowing students to alter how each
-\ac{Raspi} behaves and is configured.
-
-The following section propose a procedure to setup the testbed to facilitates these
-requirements. The procedure is performed, by whom we will refer
-to as the testbed administrator, in a regular \ac{PC} running Linux. The exact Linux
-distribution does not matter although the procedure will be based on
-Debian Linux. All the commands in this paper will be performed on
-the testbed administrators \ac{PC}.
-
-The procedure goes as follows. Based on a regular Raspbian Linux image, it will
-be explained how it can be customized to include desired packages and
-configured as desired.
-When working on a testbed with many devices, it may not always be desired to make
-persistent changes to devices. We therefore present how the concept of stacked
-filesystems may be used to enable the meands of both persistent and temporary
+In Fig.~\ref{fig:testbed_setup} and what follows, we will refer
+as the testbed administrator to the person(s) in charge of setting up and configuring
+the testbed with administrator privileges from the \ac{OS} point of view.
+The setting and configuration procedure commands are performed the testbed administrator in a
+regular \ac{PC} running a Linux distribution. Although in principle the administrator
+Linux distribution is not a restriction, we present our procedure in a Debian-based Linux
+distribution. Finally, our design considers to fetch the \ac{Raspi}s
+root filesystem from a \ac{HTTP} server. This enables the testbed administrator to
+make filesystem in one location. The \ac{Raspi}s will then get the changes after a reboot.
+All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
+within the university network. We requested the university \ac{IT} department to
+configure the university \ac{DHCP} server to assign each \ac{Raspi} a static
+\ac{IP}. This eliminates the requirements of monitors and keyboards with the
+\ac{Raspi}s for non-graphical applications.
+
+Based on a regular Raspbian Linux image, we explain how
+to customize it to include desired packages for performing routinary tasks and set it
+in all the \ac{Raspi}s. Also, when working on a testbed with many devices, it may not always be
+desired to make persistent changes to the devices. Thus, we later present how to utilize
+stacked filesystems to enable both persistent and temporary
 storage. This step of the procedure will be optional if the testbed
-administrator decides to write the whole image to a memory card.
-Writing the whole Linux filesystem to memory cards may complicate the
-maintenance as e.g. updates need to be performed in each \ac{Raspi}.
-\texttt{Python Fabric} may ease this task, but we also propose alternative
+administrator decides to write the whole image into a memory card. \todot{Check the
+veracity and correctness of the previous statement. Question: Shoudln't we use
+an overlay FS regardless of how the memory card is configured?}.
+Although being simpler than a stacked filesystem, writing the whole Linux
+filesystem to several memory cards may complicate future
+maintenance of the testbed since updates need to be performed in each \ac{Raspi}.
+Even when tools such as Python Fabric may ease this task, the maintenance might still
+be cumbersome in some cases. Therefore, we propose and encourage the alternative centralized
 approach where the initial \ac{RAM} filesystem is customized in a way that
-enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP}
-server. This enables the testbed administrator to only update the filesystem
-in one location. The \ac{Raspi}s will then get the changes after a reboot.
+enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP} server as
+mentioned previously. The following procedure explains how the testbed may be setup and configured.
 
@@
 %as the root filesystem in the centralized approach.
-
-The following procedure explains how the testbed may be setup and configured.
 

### 2016-07-01 configuration_files.tex
New pieces of the structure

--- 2016-06-29 testbed.tex
@@
-\label{sec:testbed}
-
-We proceed to define the testbed environment by each of its conforming
-parts. We later indicate a procedure to set up the testbed. In this procedure,
-we summarize all the details regarding on setting the connectivity and
-system files for a set of \ac{Raspi}s in a decentralized and/or centralized fashion
-depending on the testbed admins preferences..
-%Afterwards, we indicate how to cross-compile C++ code and the Kodo
-%network coding library in an easy way.
-%Finally, we provide further information about Kodo itself in terms
-%of testing, other platforms supported and source code documentation
-%for further references.
-
-\subsection{System Overview}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=0.45\textwidth]{images/testbed_setup3.eps}
-\caption{Testbed setup}
-\label{fig:testbed_setup}
-\end{figure}
-
-The testbed consists of up to one hundred \ac{Raspi}s of different models.
-More specifically, Raspberry Pi 1 model B rev.~2,
-Raspberry Pi 2 model B V1.1, and Raspberry Pi 3 model B V1.2.
-All \ac{Raspi}s are each equipped with a memory card, a wired, and a wireless
-network interface. They connect to the university network using its wired
-ethernet interface that is named \texttt{eth0} according to the legacy naming
-convention of ethernet interfaces in Linux~\cite{PredictableNetworkInterfaceNames}.
-
-All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
-within the university network. We therefore had the university's \ac{IT}
-department configure the university's \ac{DHCP} server to assign each \ac{Raspi} a
-static \ac{IP}. This eliminates the requirements for monitors and keyboards
-for the \ac{Raspi}s for non-graphical applications.
-%This significantly eases the management of the \ac{Raspi}s
-%as \ac{SSH} eliminates the need for monitors and keyboards for the \ac{Raspi}s.
-
-A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
-The testbed is intended for students to perform simulations and measurements
-as part of semester projects. We intend to configure all
-\ac{Raspi}s identically, while still allowing students to alter how each
-\ac{Raspi} behaves and is configured.
-
-The following section propose a procedure to setup the testbed to facilitates these
-requirements. The procedure is performed, by whom we will refer
-to as the testbed administrator, in a regular \ac{PC} running Linux. The exact Linux
-distribution does not matter although the procedure will be based on
-Debian Linux. All the commands in this paper will be performed on
-the testbed administrators \ac{PC}.
-
-The procedure goes as follows. Based on a regular Raspbian Linux image, it will
-be explained how it can be customized to include desired packages and
-configured as desired.
-When working on a testbed with many devices, it may not always be desired to make
-persistent changes to devices. We therefore present how the concept of stacked
-filesystems may be used to enable the meands of both persistent and temporary
-storage. This step of the procedure will be optional if the testbed
-administrator decides to write the whole image to a memory card.
-Writing the whole Linux filesystem to memory cards may complicate the
-maintenance as e.g. updates need to be performed in each \ac{Raspi}.
-\texttt{Python Fabric} may ease this task, but we also propose alternative
-approach where the initial \ac{RAM} filesystem is customized in a way that
-enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP}
-server. This enables the testbed administrator to only update the filesystem
-in one location. The \ac{Raspi}s will then get the changes after a reboot.
-
-%\textcolor{red}{The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. We intend to configure
-%all \ac{Raspi}s identically, while still allowing students to alter
-%how each \ac{Raspi} behaves and is configured.
-%\\\\
-%When working on the testbed with this many devices, it may not always be
-%desired to make persistent changes to devices. We present how the concept
-%of stacked filesystems may be used to enable the means of both persistent
-%and temporary storage.}
-
-%The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. The \ac{Raspi}s should
-%be configured identically, but allows the students to alter the
-%configurations and how each \ac{Raspi} behaves.
-%
-%When working on a testbed with many devices, it may not always desired
-%to make persistent changes to devices. We present how the concept of stacked
-%filesystems may be used to enable the means of both persistent and
-%temporary storage.
-
-%A \ac{HTTP} server will be used to store Linux configuration files as well
-%as the root filesystem in the centralized approach.
-
-The following procedure explains how the testbed may be setup and configured.
-
-%The following procedure will explain how the testbed may be setup and
-%configured both in the traditional distributed fasion where the
-%entire Linux filesystem is written to an memory card, but also in a more
-%centralized fasion where 
-%is written to
-%Based on the popular Debian based Raspbian Linux distribution
-%We
-%first present how a standard Linux distribution for the \ac{Raspi}s
-%may be customized
-
-\subsubsection{Installing Raspbian}
-
-To get started, we first need to install an \ac{OS} on the
-\ac{Raspi}s. We will use the popular Debian-based Raspbian
-Linux~\cite{raspbian}
-
-We will be using a \ac{PC} running a Debian-based Linux distribution
-to setup a customized Raspbian Linux image that can be written to
-a memory card and work in any \ac{Raspi} version 1 or 2.
-
-Raspbian is made available in two bundles: Raspbian and Raspbian lite.
-The difference between the two is that Raspbian contains a pre-installed desktop
-environment, and Raspbian lite is by default only accessible through a 
-command shell. Both Raspbian and Raspbian lite will both by default be remotely
-accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
-connect to another machine through a command shell.
-%In fact it is the successor
-%of \ac{Telnet} that has been designed with network security in mind.
-
-The \ac{Raspi}s in our testbed are not connected to monitors, so we
-have decided to go with Raspbian lite.
-%Our testbed does not have monitors so we will install Raspbian lite.
-A desktop environment can be installed using the package manager
-(apt-get) in case a desktop environment will be desired at a later
-time.
-
-Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
-have performed a simple setup where each \ac{Raspi} is equipped with a
-memory card that contains a slightly customized Raspbian lite image. The image
-of the \ac{OS} in all \ac{Raspi}s will be identical.
-
-The overall procedure to customize the official Raspbian lite image
-are:
-\begin{enumerate}
-    \item Download Raspbian lite
-    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
-        in the official Raspbian lite image
-    \item Change root, i.e. change root filesystem into the Raspbian lite
-        image to update and install software packages
-    \item Write image to memory cards
-\end{enumerate}
-
-Hashmark (\#) and dollar (\$) will be used throughout the code blocks in
-the paper to indicate if an operation needs to be run with root permissions or
-user permissions.
-
-\subsection{Download Raspbian lite}
-
-The latest Raspbian lite bundle can be downloaded from Raspbian's official
-webpage~\cite{raspbian}.
-%To download the latest Raspbian lite, we go to the url
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
-At the time of this writing, the latest available bundle was
-\texttt{2016-05-27-raspbian-jessie-lite.zip}.
-
-To ensure that the content of the bundle does not change, this procedure
-will be based on that particular version of Raspbian lite which we have
-made available at~\cite{tunescode_webpage}. All other files used in this paper
-will also be available there unless specified otherwise.
-
-%We want to proceed downloading the bundle in a shell instead of the browser.
-%To do this, first open a command shell on a linux machine. This guide is
-%assuming a Debian-based distribution, but there should only be minor differences
-%if you have another package manager. i.e. the package managers and name of software packages.
-
-To get started, open a Linux shell (terminal) and declare the variables below.
-The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
-located and \texttt{WORKDIR} specifies a working directory where we will 
-store Raspbian and customize it.
-Do not type in \$ and \# in the shell. Those indicates if a command can be
-executed as a regular user or if root permissions are required.
-Running a command as root can on most systems be done by putting
-\texttt{sudo} in front of the command. This is illustrated in the
-following code block 
-with the command \texttt{whoami} that prints the username. Lines within a code
-block without leading \$ or \# is terminal output or content of a file.
-
-%After the terminal has been opened, we start by declaring a few variables to reduce
-%repeated typing. The first variables we declare is the Raspbian image name.
-%Notice that the extension has been omitted. This is because the image has been
-%packed into a zip file. The other variable we declare is a working directory.
-%This is where we will download the image to and work on it. In other words, it
-%will be stored in \texttt{/home/<username>/Raspbian}
-
-%To download Raspbian lite, we go to the \url{http://downloads.raspberrypi.org}.
-%There should be a folder called raspbian\lite and
-%to find the download page at which Raspbian lite should be located. Instead of
-%downloading with the browser, we just extract the download url to the latest
-%image. At the time of this writing, that was
-
-% Define variable]
-\begin{lstlisting}[]
-$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
-$ export IMAGE="2016-05-27-raspbian-jessie-lite"
-$ export WORKDIR="${HOME}/Raspbian"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%A final thing to notice is the way \$ and \# are used in the code blocks. \$
-%means that the command can be run with normal user permissions where \# indicates that the
-%command needs root permissions.
-%Running a command with root permissions can on most systems be done by putting
-%the command "sudo" in front of each line. This is illustrated in the code below
-%with the command "whoami" that prints the username of the caller. Lines without
-%leading \$ or \# is the terminal output.
-
-% Root example
-\begin{lstlisting}[]
-$ whoami
-<USERNAME>
-$ sudo whoami
-root
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Next, make the working directory and used \texttt{cd} to change into the working
-directory. Then \texttt{wget} can be used to download Raspbian lite before
-unpacking the \texttt{zip} file:
-
-% Download and unpack image
-\begin{lstlisting}[]
-$ mkdir -p ${WORKDIR}
-$ cd ${WORKDIR}
-$ wget ${URL%/*}/${IMAGE}.zip
-$ unzip ${IMAGE}.zip
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%A more recent version of Raspbian lite may be available at~\cite{raspbian}.
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.
-
-\subsection{The image}
-
-After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
-\texttt{fdisk} can be used to display the content of the image. Parse the arguments
-"-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
-image and and automatically exit \texttt{fdisk} again:
-
-% Check out the image   % The dollar hack was to fix vim syntax
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-
-The output provides knowledge about the image. The image is in total
-2709504 sectors (1.3GiB) in size and contains two
-partitions. The first partition starts at sector 8192 and
-the other partition starts at sector 137216. The partitions is of type FAT32
-and Linux and are 63M and 1.2G big. This indicates that the first partition
-is most likely a boot partition and the second partition is a traditional Linux file
-system. Actually, it is the Linux root filesystem (i.e. /).
-
-\subsection{Resize image}
-1.2G is not very much to store the root filesystem given that it is desired to
-alter the files and install additional packages. It may therefore be
-necessary to increase the partition size.
-The following illustrates how the image and its root filesystem
-can be expanded by one \ac{GB}.
-
-First, expand the image one \ac{GB} and wait for the data to be completely written
-to the disk:
-\begin{lstlisting}[]
-$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Use \texttt{fdisk} to see that the image is now one \ac{GB} larger:
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, to expand the root filesystem it is required to first remove the Linux partition
-and then add it again one \ac{GB} bigger.
-Use \texttt{fdisk} to alter the partition table. The \texttt{echo}'s within the parenthesis
-will be interpreted as
-key-presses in \texttt{fdisk} and hence delete the last partition, create a new
-primary partition, print new partition table, and write new partition table
-to the image:
-
-\begin{lstlisting}[]
-$ (echo d; echo ; echo n; echo p; echo ; echo 137216; echo ; echo p; echo w) | fdisk ${IMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Use a loopback device to make the Raspbian image available as a block device: 
-\begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\texttt{lsblk} may be used to list block devices. It appears that the image is
-available as \texttt{/dev/loop0} which has two partitions \texttt{loop0p1} and
-\texttt{loop0p2}:
-%Use \texttt{lsblk} to view the partitions:
-
-\begin{lstlisting}[]
-# lsblk
-NAME      MAJ:MIN RM  SIZE RO TYPE MOUNTPOINT
-...
-loop0       7:0    0  2.3G  0 loop 
-|-loop0p1 259:2    0   63M  0 loop 
-|-loop0p2 259:3    0  2.2G  0 loop
-...
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Check and resize the root partition:
-\begin{lstlisting}[]
-# e2fsck -f ${DEV}p2
-# resize2fs ${DEV}p2
-\end{lstlisting}
-\FloatBarrier
-
-%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
-%We also notice that each sector is 512 bytes.
-%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"
-%Mount root partition:
-
-\subsection{Mount image}
-
-Browsing and altering the files in the image is easy. Simply mount the partitions.
-%To browser and alter the files in the image, it is possible to mount the partitions.
-
-Lets first mount the root partition. This is done by creating an empty directory
-that we will be used as a mountpoint. We name it \texttt{root} and create it in the
-working directory before mounting the root filesystem onto the mountpoint.
-%After this, mount the partitions:
-%After this,
-%we mount the partition starting at sector 137216. To provide this offset information
-%to mount, we need to find the offset in bytes. From the fdisk command we could also
-%see that each sector is 512 bytes. Thus, we multiply the sector size and the sector
-%start to get the offset in bytes.
-
-\begin{lstlisting}[]
-$ export ROOTDIR="${WORKDIR}/root"
-$ mkdir -p ${ROOTDIR}
-# mount ${DEV}p2 ${ROOTDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%# mount -o loop,offset=$((137216*512)) ${IMAGE}.img ${ROOTDIR}
-
-The root filesystem already has a boot directory that can be used as
-the mountpoint for the boot partition.
-%After this, we can also mount the boot partition inside the newly mounted
-%root filesystem.
-This is convenient because that partition will be mounted
-on this directory when a \ac{Raspi} starts up with a memory card containing
-the image. %It is therefore the natural place to mount it for later purposes.
-Now, mount boot partition:
-\begin{lstlisting}[]
-# mount ${DEV}p1 ${ROOTDIR}/boot
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%# mount -o loop,offset=$((8192*512)) ${IMAGE}.img ${ROOTDIR}/boot
-
-%We can now change all the files in the disk image as desired.
-It is now possible to change all files within the Raspbian image as desired.
-We will take advantage of this to edit configuration files, append new files,
-and even update and install packages.
+\label{sec:configuration_files}
 
@@
 %image is written to a memory card.
-
-\subsection{Installation an overlaying filesystem}
-
-Our goal was to enable the users of the testbed to change however the
-\ac{Raspi} are configured while still being able to revert those changes easily.
-
-We do this by storing a users work in the \ac{Raspi}s in \ac{RAM}. Thus a
-\ac{Raspi}, when a \ac{Raspi} is rebooted, it will forget all changes.
-
-A way to do this is using stacked filesystem. Stacking filesystem enables
-a filesystem to overlay another filesystem. Whenever a file is accessed,
-the upper filesystem will forward the request to the lower filesystem in
-case it does not have it itself. If it has it, it will simpy return the file.
-
-This idea can be used in our setup to mount the root filesystem (i.e. Raspbian
-lite) in the \ac{Raspi}s duing startup as read-only. Thus, the files will not
-be changable. Some programs needs to change files and the testbed admin might
-want to change configurations. To enable this, it is possible to
-overlay a filesystem that is mounted in \ac{RAM} as read/writable on top of
-the regular root. Thus, reading a file may return a file from the lower filesytem,
-but if it is stored, it will be saved in the upper filesystem. Accessing this file
-again will return the stored file from the upper layer.
-
-%We do this using an overlay filesystem. The idea is to let the \ac{Raspi}s
-%mount their root filesystem as 
-%
-%We do this using an overlay filesystem (overlayfs).
-%Overlayfs enables a filesystem to be stacked on top of another filesystem. Thus,
-%if the \ac{Raspi}s mount their root filesystem (/) as read-onlyif we mount our root (/) filesystem as 
-%
-%
-%Overlayfs enables filesystems
-%to be stacked on top of each other. We will be using this to make a lower
-%filesystem and an upper filesystem. The lower filesystem will be our
-%traditional filesystem and the upper filesystem will be mounted in tmpfs
-%(i.e. mounted in \ac{RAM}).
-%
-%Thus, when a file is requsted from stacked filesystem, it will first attempt
-%to fetch the file in the upper filesystem if it does not exists there, it will
-%request it from the lower filessytem.
-
-Assuming that you are still \texttt{chrooted} into the Raspbian lite filesystem, we
-can setup the overlay filesystem (overlayfs).
-
-There already exists implementations overlaying the root filesystem. We will
-use an existing implementation
-available at github: \cite{overlayroot}.
-
-We installed git in a previous step, so lets clone the repository. The below
-code stores it in /tmp which is really mounted in \ac{RAM}. Thus, the files
-will disappear when the machine is rebooted.
-
-% Get files
-\begin{lstlisting}[]
-(chroot) # OVERLAYROOTDIR="/tmp/overlayroot"
-(chroot) # git clone https://github.com/chesty/overlayroot.git $OVERLAYROOTDIR
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-To enable the overlaying filesystem it is required to generate an initial \ac{RAM}
-file system (initramfs). Initramfs is an initial filesystem that gets loaded into
-\ac{RAM} during the startup process of a linux machine to prepare the real
-filesystem.
-
-For this purpose, we need BuxyBox:
-
-% Install required packages
-\begin{lstlisting}[]
-(chroot) # apt-get install busybox
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Add and activate the overlaying filesystem, we need to add the scripts and tell initramfs
-to load the overlaying filesytem:
-
-% Setup initramfs
-\begin{lstlisting}[]
-(chroot) # cp ${OVERLAYROOTDIR}/hooks-overlay /etc/initramfs-tools/hooks/
-(chroot) # cp ${OVERLAYROOTDIR}/init-bottom-overlay /etc/initramfs-tools/scripts/init-bottom/
-(chroot) # echo "overlay" > /etc/initramfs-tools/modules
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%Finally, lets generate the initial \ac{RAM} filesystem. This is done using the command
-%\texttt{mkinitramfs}, that will by default look for the kernel modules of the current system. Remember we are
-%chrooted, so we need to specify the correct kernel modules to look for. These
-%can be found in "/lib/modules".
-Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
-needs a different kernel than \ac{Raspi} version 2 and version 3. Thus, start
-by finding the available kernel modules:
-%Because \texttt{initramfs} is executed within a
-%\texttt{chrooted} environment, it will not be able to detect the
-%kernel modules in the \texttt{chrooted} environment. The kernel modules will
-
-% Create an initramfs
-\begin{lstlisting}[]
-(chroot) # ls /lib/modules/
-4.4.13+  4.4.13-v7+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Kernel "4.4.13+" is for \ac{Raspi} version 1 and kernel "4.4.13-v7+" works
-for \ac{Raspi} version 2 and version 3.
-
-Lets proceed by generating an initial \ac{RAM} filesystem for both kernels:
-
-%Lets proceed with kernel modules "4.4.13+".
-%Lets create an initial ram filesystem for both kernels.
-
-%Thus, to run mkinitramfs, type in the correct kernel version:
-
-% Create an initramfs
-\begin{lstlisting}[]
-(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
-(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
-%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+
-
-Although there might be printet a few warnings, the above commands should successfully
-generate working initial \ac{RAM} filesystems.
-
-An initial \ac{RAM} filesystem needs to be called by the bootloader. In Raspbian
-this is done by adding a command to \texttt{config.txt} in the boot partition.
-
-If the system should be run in a \ac{Raspi} version 1, then use \texttt{init.gz}:
-%The bootloader should be made aware of the initramfs that we will generate:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If the system should be run in a \ac{Raspi} version 2 or 3, then use \texttt{init-v7.gz}:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # Only for Raspberry Pi version 2 and 3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-At this point, it is no longer required to be \texttt{chrooted} and the image
-can be written to a memory card. The following commands exits the \texttt{chrooted}
-environment, unmount all partitions, and detaches the loopback device:
-
-% Exit chroot, umount, and write to memory card
-\begin{lstlisting}[]
-(chroot) # exit
-# cd ..
-# umount --recursive ${ROOTDIR}
-# losetup -d ${DEV}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsection{(Optional) Persistent subdirectories for distributed Linux image}
-
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-Provided the stacked filesystem was configured, it may be desired to have
-some directories where files are not flushed upon rebooting the \ac{Raspi}s.
-The following procedure will create an extra partition in the image for
-the \ac{Raspi} users home directory that will be made storage persistent.
-%Provided that a stacked filesystem was configured to make the root
-%filesystem forget chances upon reboots is not always desired.
-
-Expand image according to the desired home directory size. Keep in mind not
-to make the image bigger than the target memory card.
-% Expand image
-\begin{lstlisting}[]
-$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create home partition after the root partition. Use \texttt{fdisk} to
-find the next available sector in the image. In our case, the new
-partition should start at sector 4806656:
-\begin{lstlisting}[]
-$ fdisk -lu ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create new partition:
-
-\begin{lstlisting}[]
-$ (echo n; echo p; echo ; echo 4806656; echo ; echo p; echo w) | fdisk ${IMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create loopback device: 
-\begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Format new partition:
-\begin{lstlisting}[]
-# mkfs.ext4 ${DEV}p3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The last thing to remember is to inform Raspbian to mount the home partition
-we created during startup. This can be done by adding an entry in \texttt{fstab}:
-\begin{lstlisting}[]
-# mount ${DEV}p2 ${ROOTDIR}
-# sudo sh -c "echo '/dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' >> ${ROOTDIR}/etc/fstab"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The home folder was originally located in the root filesystem. We should
-therefore copy the content to the new home partition:
-\begin{lstlisting}[]
-# mount ${DEV}p3 ${ROOTDIR}/mnt
-# mv ${ROOTDIR}/home/* ${ROOTDIR}/mnt/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-That is it. unmount partitions and detach loop device:
-\begin{lstlisting}[]
-# umount --recursive ${ROOTDIR}
-# losetup -d ${DEV}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsection{Write image to memory card}
-
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-The final step is to write the image to all the memory cards before they can
-be used in the \ac{Raspi}s.
-
-%This is done by first leaving the \texttt{chrooted}
-%filesystem and unmounting all the directories we mounted:
-
-% Exit chroot, umount, and write to memory card
-%\begin{lstlisting}[]
-%# exit
-%# cd ..
-%# umount --recursive ${ROOTDIR}
-%# losetup -d ${DEV}.img
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-%If successfully unmounted, you can 
-Insert a memory card in your machine. It will be available as
-\texttt{/dev/mmcblkX} or \texttt{/dev/sdX}
-
-%In most systems, this memory card was available under \texttt{/dev/mmcblk0}, but this may be
-%different on different machines. For internal memory card readers, it will
-%probably be named /dev/mmcblkX where X is some number and if the memory card
-%reader is connected through \ac{USB}, then it is likely called /dev/sdX.
-Make absolutely sure to write to the correct device as everything will be 
-overwritten!
-Inserting/Ejecting the memory card a couple of times and using the commands
-below may help deducing the correct device:
-%The commands below may help deduce which device is the correct:
-%Below is a few commands that may help you to deduce which device is correct:
-
-\begin{lstlisting}[]
-# lsblk
-# df -h
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For us, the device was \texttt{/dev/mmcblk0}. Write the image to a memory card:
-\begin{lstlisting}[]
-# dd if=${IMAGE}.img of=/dev/mmcblk0 bs=4M && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The memory card can now be started in a \ac{Raspi}.

### 2016-07-01 image_setup.tex
New pieces of the structure

--- 2016-06-29 testbed.tex
@@
-\label{sec:testbed}
+\label{sec:image_setup}
 
-We proceed to define the testbed environment by each of its conforming
-parts. We later indicate a procedure to set up the testbed. In this procedure,
-we summarize all the details regarding on setting the connectivity and
-system files for a set of \ac{Raspi}s in a decentralized and/or centralized fashion
-depending on the testbed admins preferences..
-%Afterwards, we indicate how to cross-compile C++ code and the Kodo
-%network coding library in an easy way.
-%Finally, we provide further information about Kodo itself in terms
-%of testing, other platforms supported and source code documentation
-%for further references.
-
-\subsection{System Overview}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=0.45\textwidth]{images/testbed_setup3.eps}
-\caption{Testbed setup}
-\label{fig:testbed_setup}
-\end{figure}
-
-The testbed consists of up to one hundred \ac{Raspi}s of different models.
-More specifically, Raspberry Pi 1 model B rev.~2,
-Raspberry Pi 2 model B V1.1, and Raspberry Pi 3 model B V1.2.
-All \ac{Raspi}s are each equipped with a memory card, a wired, and a wireless
-network interface. They connect to the university network using its wired
-ethernet interface that is named \texttt{eth0} according to the legacy naming
-convention of ethernet interfaces in Linux~\cite{PredictableNetworkInterfaceNames}.
-
-All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
-within the university network. We therefore had the university's \ac{IT}
-department configure the university's \ac{DHCP} server to assign each \ac{Raspi} a
-static \ac{IP}. This eliminates the requirements for monitors and keyboards
-for the \ac{Raspi}s for non-graphical applications.
-%This significantly eases the management of the \ac{Raspi}s
-%as \ac{SSH} eliminates the need for monitors and keyboards for the \ac{Raspi}s.
-
-A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
-The testbed is intended for students to perform simulations and measurements
-as part of semester projects. We intend to configure all
-\ac{Raspi}s identically, while still allowing students to alter how each
-\ac{Raspi} behaves and is configured.
-
-The following section propose a procedure to setup the testbed to facilitates these
-requirements. The procedure is performed, by whom we will refer
-to as the testbed administrator, in a regular \ac{PC} running Linux. The exact Linux
-distribution does not matter although the procedure will be based on
-Debian Linux. All the commands in this paper will be performed on
-the testbed administrators \ac{PC}.
-
-The procedure goes as follows. Based on a regular Raspbian Linux image, it will
-be explained how it can be customized to include desired packages and
-configured as desired.
-When working on a testbed with many devices, it may not always be desired to make
-persistent changes to devices. We therefore present how the concept of stacked
-filesystems may be used to enable the meands of both persistent and temporary
-storage. This step of the procedure will be optional if the testbed
-administrator decides to write the whole image to a memory card.
-Writing the whole Linux filesystem to memory cards may complicate the
-maintenance as e.g. updates need to be performed in each \ac{Raspi}.
-\texttt{Python Fabric} may ease this task, but we also propose alternative
-approach where the initial \ac{RAM} filesystem is customized in a way that
-enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP}
-server. This enables the testbed administrator to only update the filesystem
-in one location. The \ac{Raspi}s will then get the changes after a reboot.
-
-%\textcolor{red}{The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. We intend to configure
-%all \ac{Raspi}s identically, while still allowing students to alter
-%how each \ac{Raspi} behaves and is configured.
-%\\\\
-%When working on the testbed with this many devices, it may not always be
-%desired to make persistent changes to devices. We present how the concept
-%of stacked filesystems may be used to enable the means of both persistent
-%and temporary storage.}
-
-%The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. The \ac{Raspi}s should
-%be configured identically, but allows the students to alter the
-%configurations and how each \ac{Raspi} behaves.
-%
-%When working on a testbed with many devices, it may not always desired
-%to make persistent changes to devices. We present how the concept of stacked
-%filesystems may be used to enable the means of both persistent and
-%temporary storage.
-
-%A \ac{HTTP} server will be used to store Linux configuration files as well
-%as the root filesystem in the centralized approach.
-
-The following procedure explains how the testbed may be setup and configured.
-
-%The following procedure will explain how the testbed may be setup and
-%configured both in the traditional distributed fasion where the
-%entire Linux filesystem is written to an memory card, but also in a more
-%centralized fasion where 
-%is written to
-%Based on the popular Debian based Raspbian Linux distribution
-%We
-%first present how a standard Linux distribution for the \ac{Raspi}s
-%may be customized
-
-\subsubsection{Installing Raspbian}
+\subsection{Installing Raspbian}
 
@@
 and even update and install packages.
-
-\subsection{Configuring OS files and scripts}
-
-The \ac{Raspi}s should be setup as similar as possible. One of those
-few settings that may be different in the devices is their hostname.
-%The hostname helps the user to distinguish the devices from each other.
-%The hostname helps distinguish them from each other.
-One approach to create unique hostnames would be to write the hostname
-in each memory card. Another way is to create a start up script to automatically
-configure the hostname. We decided to define the hostnames based on the \ac{MAC}
-addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
-addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
-administrator will need to insert the \ac{MAC} addresses and hostnames of
-his/her \ac{Raspi}s.
-
-%The \ac{Raspi}s should be setup to be as similar as possible. One of those
-%things that is preferred to be different on the devices is their hostname.
-%This helps distinguish them from each other. One solution to this in our
-%approach would be to uniquely change the hostname in each memory card.
-%This is however not desired because that would either require modifying
-%the customized Raspbian lite image before writing the content to the
-%memory card or to manually change it in the \ac{Raspi}s after the image has
-%been written to a memory card.
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-%
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames. This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-
-%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
-
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames.
-%This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-% MAC and Hostname file
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-# Ethernet MAC    Hostname
-b8:27:eb:5b:da:20 rasp00
-b8:27:eb:7b:c3:91 rasp01
-b8:27:eb:54:9c:64 rasp02
-b8:27:eb:95:bd:11 rasp03
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
-%the testbed, it still needs to know which name to give itself.
-%We do this using a small Bash script:
-%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
-%hostname in \texttt{nodes.csv}.
-A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:
-
-% Set hostname
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-script_path="$(dirname $(realpath $0))"
-config_file=${script_path}/nodes.csv 
-mac=$(cat /sys/class/net/eth0/address)
-old_hostname=$(hostname)
-new_hostname=$(grep $mac $config_file | cut -f2 -d' ')
-
-# Assign hostname found in nodes.csv
-if [ ! -z ${new_hostname} ]; then
-    echo ${new_hostname} > /etc/hostname
-    hostname ${new_hostname}
-    sed -i.old -e "s:${old_hostname}:${new_hostname}:g" /etc/hosts
-fi
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The code lines above do the following:
-\begin{itemize}
-    \item Line 1: Tells system to interpret the script using \ac{Bash}
-    \item Line 3-4: Gets the path to the script itself and the list of hostnames
-    \item Line 5: Gets the \ac{MAC} address of the node itself
-    \item Line 6: Gets the current hostname
-    \item Line 7: Gets the new hostname from the hostname list
-    \item Line 10-14: Assigns the new hostname
-\end{itemize}
-
-%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
-%path of the script itself. Line 4 gets the MAC address. This script will be
-%located in all \ac{Raspi}s and thus return the MAC address of the node itself
-%when executed.
-%Line 4 searches line by line for this mac address.
-%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
-%then it will keep the default hostname of the system.
-
-It is likely that the testbed administrator needs to either change or add
-scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
-all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
-configuration scripts during start up before running the configuration scripts.
-
-The below script attempts to download/update the scripts:
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-url="http://kom.aau.dk/project/TuneSCode/raspi/"
-config_file="rasp_config.zip"
-
-# Attempt to fetch new configuration files
-if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
-    echo "Warning: Unable to update rasp_config files"
-    exit 1
-fi
-
-# Unzip and overwrite configurationn files to root's home directory
-unzip -q -o /tmp/${config_file} -d /home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\begin{itemize}
-    \item Line 3-4: Is the url and zip file that should be downloaded
-    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
-        and prints a warning in case of errors
-    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
-\end{itemize}
-
-We have packed the files and uploaded them to a webserver. These can be
-retrieved with \texttt{wget} and unzipped into the Raspbian lite image:
-
-% Get the files
-\begin{lstlisting}[]
-$ wget ${URL%/*}/rasp_config.zip
-$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
-\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
-\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
-the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
-The testbed administrator will therefore need to also edit the \ac{URL} in the
-script \texttt{update\_rasp\_config}.
-%"nodes.csv" can be changed using a text editor. For vim, this could be done
-%the following way:
-%\begin{lstlisting}[]
-%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-Finally, to actually make the \ac{Raspi}s change hostname, we have to make
-each \ac{Raspi} call the above script when it starts up. Optionally,
-also run the update script before running other scripts.
-Do this with root permissions by inserting the line(s) below before
-\texttt{"exit 0"} using your favorite editor:
-%"\$\{ROOTDIR\}/etc/rc.local" using
-%your favorite text editor..
-%The file
-%can be opened with your favorite text editor.
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/set_hostname
-...
-exit 0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Configuring \texttt{rc.local} makes it really cumbersome for the testbed
-administrator to add new scripts since it first needs to be included
-in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional 
-line added to \texttt{rc.local} in the root image or in all memory cards depending on
-the final testbed setup.
-
-A solution to this is to make yet another script in \texttt{rasp\_config.zip}
-that calls all the other scripts in the desired order. This would look like
-below:
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-bash /home/pi/rasp_config/set_hostname
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, to use the above script, \texttt{rc.local} needs to look like this:
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/main
-...
-exit 0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-Notice that \texttt{set\_hostname} is now called by the \texttt{main}
-script instead. The update script is still called directly. This ensures that
-all configuration scripts are updated before executed. Changes to the update
-script itself will first take effect at next system startup.
-%Had the update script
-%been called in \texttt{main} instead would mean that changes to the 
-%\texttt{main} script would also first have been available at next startup.
-
-%This can be done by inserting a call to the script
-%set\_hostname in rc.local before it exits.
-%
-%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
-%\begin{lstlisting}[]
-%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
-%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
-%\end{lstlisting}
-%\FloatBarrier
-%$ line="bash /home/pi/.rasp_config/set_hostname"
-%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
-%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>
-
-%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
-%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
-
-%All other files could be updated as well.
-It is not only possible to add and configure files to/within the image. One can
-also install and update the packages within the image which will be illustrated
-in the next section.
-
-\subsection{Installing and updating the image}
-
-It may be desired to pre-install some programs in the image before it is
-written to all the memory cards that goes into the \ac{Raspi}s.
-This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
-(change root).
-%\url{https://wiki.debian.org/QemuUserEmulation}
-%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
-
-Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
-install some additional software:
-%Because \ac{Raspi}s are equipped with an \ac{ARM} processor 
-%%it is not as
-%straightforward as performing the task for two Linux systems of the same
-%architecture.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# apt-get install binfmt-support qemu qemu-user-static
-# update-binfmts --display qemu-arm
-qemu-arm (enabled):
-     package = qemu-user-static
-        type = magic
-      offset = 0
-       magic = \x7fELF\x01\x01\x01\x00\x00\x00\x00\x00\x00\x00\x00\x00\x02\x00\x28\x00
-        mask = \xff\xff\xff\xff\xff\xff\xff\x00\xff\xff\xff\xff\xff\xff\xff\xff\xfe\xff\xff\xff
- interpreter = /usr/bin/qemu-arm-static
-    detector = 
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-% This was neccessary on arch
-%# update-binfmts --importdir /var/lib/binfmts/ --import
-
-Make sure the second command writes "enabled" as the output above. If that 
-is not the case, then try enabling it:
-
-\begin{lstlisting}[]
-# update-binfmts --enable qemu-arm
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Provided that qemu-arm is enabled, we should now be able to change root (chroot)
-into our Raspbian lite image. Change root is a method in Linux that enables
-the root (/) to be changed. Thus, enabling a Linux installation within a
-Linux installation.
-
-There are a few commands to be performed before actually changing root into the
-root partion of the image.
-First, to get internet access from within the Raspbian lite image it is needed
-to copy our \texttt{resolv.conf} into the filesystem. Line 1 steps into Raspbian lite's
-root filesystem. Then line 2 copies resolv.conf.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-$ cd $ROOTDIR
-# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, because of the ARM architecture, it is also required to copy
-/usr/bin/qemu-arm-static into the image before we can continue:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The final preparation before chaning root is to populate the directories proc,
-sys, and dev:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# mount  -t proc proc proc/
-# mount --bind /sys sys/
-# mount --bind /dev dev/
-# mount --bind /dev/pts dev/pts
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Finally, the filesystem is ready for us to change root. This can be done with
-"proot". It may be required to install proot:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# proot -q qemu-arm-static -S ${ROOTDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Most available online material uses the more known chroot command as written in
-the code block below. This did not work correctly in our machines, but we show
-it as an alternative in case proot may not work on other systems.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If everything went well, we should now be standing in the Raspbian lite filesystem
-under its root. Optionally, we can change the prompt title to indicate that we
-are chrooted:
-
-% Optionally, we may create a unique prompt to indicate we have changed root
-\begin{lstlisting}[]
-# export PS1="(chroot) $PS1"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The Raspbian lite system should now be possible to use as if it had been booted
-in a \ac{Raspi}. The only difference is that the laptop that we are using is
-significantly faster than a \ac{Raspi}. Thus enabling us to for example update,
-upgrade, and install software packages:
-
-% Update system
-\begin{lstlisting}[]
-(chroot) # apt-get update
-(chroot) # apt-get upgrade
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Lets install some useful applications:
-% Install packages
-\begin{lstlisting}[]
-(chroot) # apt-get install vim git screen
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-All the changes that are made here will exists in all \ac{Raspi}s when the
-image is written to a memory card.
-
-\subsection{Installation an overlaying filesystem}
-
-Our goal was to enable the users of the testbed to change however the
-\ac{Raspi} are configured while still being able to revert those changes easily.
-
-We do this by storing a users work in the \ac{Raspi}s in \ac{RAM}. Thus a
-\ac{Raspi}, when a \ac{Raspi} is rebooted, it will forget all changes.
-
-A way to do this is using stacked filesystem. Stacking filesystem enables
-a filesystem to overlay another filesystem. Whenever a file is accessed,
-the upper filesystem will forward the request to the lower filesystem in
-case it does not have it itself. If it has it, it will simpy return the file.
-
-This idea can be used in our setup to mount the root filesystem (i.e. Raspbian
-lite) in the \ac{Raspi}s duing startup as read-only. Thus, the files will not
-be changable. Some programs needs to change files and the testbed admin might
-want to change configurations. To enable this, it is possible to
-overlay a filesystem that is mounted in \ac{RAM} as read/writable on top of
-the regular root. Thus, reading a file may return a file from the lower filesytem,
-but if it is stored, it will be saved in the upper filesystem. Accessing this file
-again will return the stored file from the upper layer.
-
-%We do this using an overlay filesystem. The idea is to let the \ac{Raspi}s
-%mount their root filesystem as 
-%
-%We do this using an overlay filesystem (overlayfs).
-%Overlayfs enables a filesystem to be stacked on top of another filesystem. Thus,
-%if the \ac{Raspi}s mount their root filesystem (/) as read-onlyif we mount our root (/) filesystem as 
-%
-%
-%Overlayfs enables filesystems
-%to be stacked on top of each other. We will be using this to make a lower
-%filesystem and an upper filesystem. The lower filesystem will be our
-%traditional filesystem and the upper filesystem will be mounted in tmpfs
-%(i.e. mounted in \ac{RAM}).
-%
-%Thus, when a file is requsted from stacked filesystem, it will first attempt
-%to fetch the file in the upper filesystem if it does not exists there, it will
-%request it from the lower filessytem.
-
-Assuming that you are still \texttt{chrooted} into the Raspbian lite filesystem, we
-can setup the overlay filesystem (overlayfs).
-
-There already exists implementations overlaying the root filesystem. We will
-use an existing implementation
-available at github: \cite{overlayroot}.
-
-We installed git in a previous step, so lets clone the repository. The below
-code stores it in /tmp which is really mounted in \ac{RAM}. Thus, the files
-will disappear when the machine is rebooted.
-
-% Get files
-\begin{lstlisting}[]
-(chroot) # OVERLAYROOTDIR="/tmp/overlayroot"
-(chroot) # git clone https://github.com/chesty/overlayroot.git $OVERLAYROOTDIR
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-To enable the overlaying filesystem it is required to generate an initial \ac{RAM}
-file system (initramfs). Initramfs is an initial filesystem that gets loaded into
-\ac{RAM} during the startup process of a linux machine to prepare the real
-filesystem.
-
-For this purpose, we need BuxyBox:
-
-% Install required packages
-\begin{lstlisting}[]
-(chroot) # apt-get install busybox
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Add and activate the overlaying filesystem, we need to add the scripts and tell initramfs
-to load the overlaying filesytem:
-
-% Setup initramfs
-\begin{lstlisting}[]
-(chroot) # cp ${OVERLAYROOTDIR}/hooks-overlay /etc/initramfs-tools/hooks/
-(chroot) # cp ${OVERLAYROOTDIR}/init-bottom-overlay /etc/initramfs-tools/scripts/init-bottom/
-(chroot) # echo "overlay" > /etc/initramfs-tools/modules
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%Finally, lets generate the initial \ac{RAM} filesystem. This is done using the command
-%\texttt{mkinitramfs}, that will by default look for the kernel modules of the current system. Remember we are
-%chrooted, so we need to specify the correct kernel modules to look for. These
-%can be found in "/lib/modules".
-Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
-needs a different kernel than \ac{Raspi} version 2 and version 3. Thus, start
-by finding the available kernel modules:
-%Because \texttt{initramfs} is executed within a
-%\texttt{chrooted} environment, it will not be able to detect the
-%kernel modules in the \texttt{chrooted} environment. The kernel modules will
-
-% Create an initramfs
-\begin{lstlisting}[]
-(chroot) # ls /lib/modules/
-4.4.13+  4.4.13-v7+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Kernel "4.4.13+" is for \ac{Raspi} version 1 and kernel "4.4.13-v7+" works
-for \ac{Raspi} version 2 and version 3.
-
-Lets proceed by generating an initial \ac{RAM} filesystem for both kernels:
-
-%Lets proceed with kernel modules "4.4.13+".
-%Lets create an initial ram filesystem for both kernels.
-
-%Thus, to run mkinitramfs, type in the correct kernel version:
-
-% Create an initramfs
-\begin{lstlisting}[]
-(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
-(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
-%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+
-
-Although there might be printet a few warnings, the above commands should successfully
-generate working initial \ac{RAM} filesystems.
-
-An initial \ac{RAM} filesystem needs to be called by the bootloader. In Raspbian
-this is done by adding a command to \texttt{config.txt} in the boot partition.
-
-If the system should be run in a \ac{Raspi} version 1, then use \texttt{init.gz}:
-%The bootloader should be made aware of the initramfs that we will generate:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If the system should be run in a \ac{Raspi} version 2 or 3, then use \texttt{init-v7.gz}:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # Only for Raspberry Pi version 2 and 3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-At this point, it is no longer required to be \texttt{chrooted} and the image
-can be written to a memory card. The following commands exits the \texttt{chrooted}
-environment, unmount all partitions, and detaches the loopback device:
-
-% Exit chroot, umount, and write to memory card
-\begin{lstlisting}[]
-(chroot) # exit
-# cd ..
-# umount --recursive ${ROOTDIR}
-# losetup -d ${DEV}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsection{(Optional) Persistent subdirectories for distributed Linux image}
-
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-Provided the stacked filesystem was configured, it may be desired to have
-some directories where files are not flushed upon rebooting the \ac{Raspi}s.
-The following procedure will create an extra partition in the image for
-the \ac{Raspi} users home directory that will be made storage persistent.
-%Provided that a stacked filesystem was configured to make the root
-%filesystem forget chances upon reboots is not always desired.
-
-Expand image according to the desired home directory size. Keep in mind not
-to make the image bigger than the target memory card.
-% Expand image
-\begin{lstlisting}[]
-$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create home partition after the root partition. Use \texttt{fdisk} to
-find the next available sector in the image. In our case, the new
-partition should start at sector 4806656:
-\begin{lstlisting}[]
-$ fdisk -lu ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create new partition:
-
-\begin{lstlisting}[]
-$ (echo n; echo p; echo ; echo 4806656; echo ; echo p; echo w) | fdisk ${IMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create loopback device: 
-\begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Format new partition:
-\begin{lstlisting}[]
-# mkfs.ext4 ${DEV}p3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The last thing to remember is to inform Raspbian to mount the home partition
-we created during startup. This can be done by adding an entry in \texttt{fstab}:
-\begin{lstlisting}[]
-# mount ${DEV}p2 ${ROOTDIR}
-# sudo sh -c "echo '/dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' >> ${ROOTDIR}/etc/fstab"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The home folder was originally located in the root filesystem. We should
-therefore copy the content to the new home partition:
-\begin{lstlisting}[]
-# mount ${DEV}p3 ${ROOTDIR}/mnt
-# mv ${ROOTDIR}/home/* ${ROOTDIR}/mnt/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-That is it. unmount partitions and detach loop device:
-\begin{lstlisting}[]
-# umount --recursive ${ROOTDIR}
-# losetup -d ${DEV}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsection{Write image to memory card}
-
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-The final step is to write the image to all the memory cards before they can
-be used in the \ac{Raspi}s.
-
-%This is done by first leaving the \texttt{chrooted}
-%filesystem and unmounting all the directories we mounted:
-
-% Exit chroot, umount, and write to memory card
-%\begin{lstlisting}[]
-%# exit
-%# cd ..
-%# umount --recursive ${ROOTDIR}
-%# losetup -d ${DEV}.img
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-%If successfully unmounted, you can 
-Insert a memory card in your machine. It will be available as
-\texttt{/dev/mmcblkX} or \texttt{/dev/sdX}
-
-%In most systems, this memory card was available under \texttt{/dev/mmcblk0}, but this may be
-%different on different machines. For internal memory card readers, it will
-%probably be named /dev/mmcblkX where X is some number and if the memory card
-%reader is connected through \ac{USB}, then it is likely called /dev/sdX.
-Make absolutely sure to write to the correct device as everything will be 
-overwritten!
-Inserting/Ejecting the memory card a couple of times and using the commands
-below may help deducing the correct device:
-%The commands below may help deduce which device is the correct:
-%Below is a few commands that may help you to deduce which device is correct:
-
-\begin{lstlisting}[]
-# lsblk
-# df -h
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For us, the device was \texttt{/dev/mmcblk0}. Write the image to a memory card:
-\begin{lstlisting}[]
-# dd if=${IMAGE}.img of=/dev/mmcblk0 bs=4M && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The memory card can now be started in a \ac{Raspi}.

### 2016-07-01 overlay_fs.tex
New pieces of the structure

--- 2016-06-29 testbed.tex
@@
-\label{sec:testbed}
-
-We proceed to define the testbed environment by each of its conforming
-parts. We later indicate a procedure to set up the testbed. In this procedure,
-we summarize all the details regarding on setting the connectivity and
-system files for a set of \ac{Raspi}s in a decentralized and/or centralized fashion
-depending on the testbed admins preferences..
-%Afterwards, we indicate how to cross-compile C++ code and the Kodo
-%network coding library in an easy way.
-%Finally, we provide further information about Kodo itself in terms
-%of testing, other platforms supported and source code documentation
-%for further references.
-
-\subsection{System Overview}
-
-\begin{figure}[ht!]
-\centering
-\includegraphics[width=0.45\textwidth]{images/testbed_setup3.eps}
-\caption{Testbed setup}
-\label{fig:testbed_setup}
-\end{figure}
-
-The testbed consists of up to one hundred \ac{Raspi}s of different models.
-More specifically, Raspberry Pi 1 model B rev.~2,
-Raspberry Pi 2 model B V1.1, and Raspberry Pi 3 model B V1.2.
-All \ac{Raspi}s are each equipped with a memory card, a wired, and a wireless
-network interface. They connect to the university network using its wired
-ethernet interface that is named \texttt{eth0} according to the legacy naming
-convention of ethernet interfaces in Linux~\cite{PredictableNetworkInterfaceNames}.
-
-All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
-within the university network. We therefore had the university's \ac{IT}
-department configure the university's \ac{DHCP} server to assign each \ac{Raspi} a
-static \ac{IP}. This eliminates the requirements for monitors and keyboards
-for the \ac{Raspi}s for non-graphical applications.
-%This significantly eases the management of the \ac{Raspi}s
-%as \ac{SSH} eliminates the need for monitors and keyboards for the \ac{Raspi}s.
-
-A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
-The testbed is intended for students to perform simulations and measurements
-as part of semester projects. We intend to configure all
-\ac{Raspi}s identically, while still allowing students to alter how each
-\ac{Raspi} behaves and is configured.
-
-The following section propose a procedure to setup the testbed to facilitates these
-requirements. The procedure is performed, by whom we will refer
-to as the testbed administrator, in a regular \ac{PC} running Linux. The exact Linux
-distribution does not matter although the procedure will be based on
-Debian Linux. All the commands in this paper will be performed on
-the testbed administrators \ac{PC}.
-
-The procedure goes as follows. Based on a regular Raspbian Linux image, it will
-be explained how it can be customized to include desired packages and
-configured as desired.
-When working on a testbed with many devices, it may not always be desired to make
-persistent changes to devices. We therefore present how the concept of stacked
-filesystems may be used to enable the meands of both persistent and temporary
-storage. This step of the procedure will be optional if the testbed
-administrator decides to write the whole image to a memory card.
-Writing the whole Linux filesystem to memory cards may complicate the
-maintenance as e.g. updates need to be performed in each \ac{Raspi}.
-\texttt{Python Fabric} may ease this task, but we also propose alternative
-approach where the initial \ac{RAM} filesystem is customized in a way that
-enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP}
-server. This enables the testbed administrator to only update the filesystem
-in one location. The \ac{Raspi}s will then get the changes after a reboot.
-
-%\textcolor{red}{The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. We intend to configure
-%all \ac{Raspi}s identically, while still allowing students to alter
-%how each \ac{Raspi} behaves and is configured.
-%\\\\
-%When working on the testbed with this many devices, it may not always be
-%desired to make persistent changes to devices. We present how the concept
-%of stacked filesystems may be used to enable the means of both persistent
-%and temporary storage.}
-
-%The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. The \ac{Raspi}s should
-%be configured identically, but allows the students to alter the
-%configurations and how each \ac{Raspi} behaves.
-%
-%When working on a testbed with many devices, it may not always desired
-%to make persistent changes to devices. We present how the concept of stacked
-%filesystems may be used to enable the means of both persistent and
-%temporary storage.
-
-%A \ac{HTTP} server will be used to store Linux configuration files as well
-%as the root filesystem in the centralized approach.
-
-The following procedure explains how the testbed may be setup and configured.
-
-%The following procedure will explain how the testbed may be setup and
-%configured both in the traditional distributed fasion where the
-%entire Linux filesystem is written to an memory card, but also in a more
-%centralized fasion where 
-%is written to
-%Based on the popular Debian based Raspbian Linux distribution
-%We
-%first present how a standard Linux distribution for the \ac{Raspi}s
-%may be customized
-
-\subsubsection{Installing Raspbian}
-
-To get started, we first need to install an \ac{OS} on the
-\ac{Raspi}s. We will use the popular Debian-based Raspbian
-Linux~\cite{raspbian}
-
-We will be using a \ac{PC} running a Debian-based Linux distribution
-to setup a customized Raspbian Linux image that can be written to
-a memory card and work in any \ac{Raspi} version 1 or 2.
-
-Raspbian is made available in two bundles: Raspbian and Raspbian lite.
-The difference between the two is that Raspbian contains a pre-installed desktop
-environment, and Raspbian lite is by default only accessible through a 
-command shell. Both Raspbian and Raspbian lite will both by default be remotely
-accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
-connect to another machine through a command shell.
-%In fact it is the successor
-%of \ac{Telnet} that has been designed with network security in mind.
-
-The \ac{Raspi}s in our testbed are not connected to monitors, so we
-have decided to go with Raspbian lite.
-%Our testbed does not have monitors so we will install Raspbian lite.
-A desktop environment can be installed using the package manager
-(apt-get) in case a desktop environment will be desired at a later
-time.
-
-Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
-have performed a simple setup where each \ac{Raspi} is equipped with a
-memory card that contains a slightly customized Raspbian lite image. The image
-of the \ac{OS} in all \ac{Raspi}s will be identical.
-
-The overall procedure to customize the official Raspbian lite image
-are:
-\begin{enumerate}
-    \item Download Raspbian lite
-    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
-        in the official Raspbian lite image
-    \item Change root, i.e. change root filesystem into the Raspbian lite
-        image to update and install software packages
-    \item Write image to memory cards
-\end{enumerate}
-
-Hashmark (\#) and dollar (\$) will be used throughout the code blocks in
-the paper to indicate if an operation needs to be run with root permissions or
-user permissions.
-
-\subsection{Download Raspbian lite}
-
-The latest Raspbian lite bundle can be downloaded from Raspbian's official
-webpage~\cite{raspbian}.
-%To download the latest Raspbian lite, we go to the url
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
-At the time of this writing, the latest available bundle was
-\texttt{2016-05-27-raspbian-jessie-lite.zip}.
-
-To ensure that the content of the bundle does not change, this procedure
-will be based on that particular version of Raspbian lite which we have
-made available at~\cite{tunescode_webpage}. All other files used in this paper
-will also be available there unless specified otherwise.
-
-%We want to proceed downloading the bundle in a shell instead of the browser.
-%To do this, first open a command shell on a linux machine. This guide is
-%assuming a Debian-based distribution, but there should only be minor differences
-%if you have another package manager. i.e. the package managers and name of software packages.
-
-To get started, open a Linux shell (terminal) and declare the variables below.
-The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
-located and \texttt{WORKDIR} specifies a working directory where we will 
-store Raspbian and customize it.
-Do not type in \$ and \# in the shell. Those indicates if a command can be
-executed as a regular user or if root permissions are required.
-Running a command as root can on most systems be done by putting
-\texttt{sudo} in front of the command. This is illustrated in the
-following code block 
-with the command \texttt{whoami} that prints the username. Lines within a code
-block without leading \$ or \# is terminal output or content of a file.
-
-%After the terminal has been opened, we start by declaring a few variables to reduce
-%repeated typing. The first variables we declare is the Raspbian image name.
-%Notice that the extension has been omitted. This is because the image has been
-%packed into a zip file. The other variable we declare is a working directory.
-%This is where we will download the image to and work on it. In other words, it
-%will be stored in \texttt{/home/<username>/Raspbian}
-
-%To download Raspbian lite, we go to the \url{http://downloads.raspberrypi.org}.
-%There should be a folder called raspbian\lite and
-%to find the download page at which Raspbian lite should be located. Instead of
-%downloading with the browser, we just extract the download url to the latest
-%image. At the time of this writing, that was
-
-% Define variable]
-\begin{lstlisting}[]
-$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
-$ export IMAGE="2016-05-27-raspbian-jessie-lite"
-$ export WORKDIR="${HOME}/Raspbian"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%A final thing to notice is the way \$ and \# are used in the code blocks. \$
-%means that the command can be run with normal user permissions where \# indicates that the
-%command needs root permissions.
-%Running a command with root permissions can on most systems be done by putting
-%the command "sudo" in front of each line. This is illustrated in the code below
-%with the command "whoami" that prints the username of the caller. Lines without
-%leading \$ or \# is the terminal output.
-
-% Root example
-\begin{lstlisting}[]
-$ whoami
-<USERNAME>
-$ sudo whoami
-root
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Next, make the working directory and used \texttt{cd} to change into the working
-directory. Then \texttt{wget} can be used to download Raspbian lite before
-unpacking the \texttt{zip} file:
-
-% Download and unpack image
-\begin{lstlisting}[]
-$ mkdir -p ${WORKDIR}
-$ cd ${WORKDIR}
-$ wget ${URL%/*}/${IMAGE}.zip
-$ unzip ${IMAGE}.zip
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%A more recent version of Raspbian lite may be available at~\cite{raspbian}.
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.
-
-\subsection{The image}
-
-After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
-\texttt{fdisk} can be used to display the content of the image. Parse the arguments
-"-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
-image and and automatically exit \texttt{fdisk} again:
-
-% Check out the image   % The dollar hack was to fix vim syntax
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-
-The output provides knowledge about the image. The image is in total
-2709504 sectors (1.3GiB) in size and contains two
-partitions. The first partition starts at sector 8192 and
-the other partition starts at sector 137216. The partitions is of type FAT32
-and Linux and are 63M and 1.2G big. This indicates that the first partition
-is most likely a boot partition and the second partition is a traditional Linux file
-system. Actually, it is the Linux root filesystem (i.e. /).
-
-\subsection{Resize image}
-1.2G is not very much to store the root filesystem given that it is desired to
-alter the files and install additional packages. It may therefore be
-necessary to increase the partition size.
-The following illustrates how the image and its root filesystem
-can be expanded by one \ac{GB}.
-
-First, expand the image one \ac{GB} and wait for the data to be completely written
-to the disk:
-\begin{lstlisting}[]
-$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Use \texttt{fdisk} to see that the image is now one \ac{GB} larger:
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, to expand the root filesystem it is required to first remove the Linux partition
-and then add it again one \ac{GB} bigger.
-Use \texttt{fdisk} to alter the partition table. The \texttt{echo}'s within the parenthesis
-will be interpreted as
-key-presses in \texttt{fdisk} and hence delete the last partition, create a new
-primary partition, print new partition table, and write new partition table
-to the image:
-
-\begin{lstlisting}[]
-$ (echo d; echo ; echo n; echo p; echo ; echo 137216; echo ; echo p; echo w) | fdisk ${IMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Use a loopback device to make the Raspbian image available as a block device: 
-\begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\texttt{lsblk} may be used to list block devices. It appears that the image is
-available as \texttt{/dev/loop0} which has two partitions \texttt{loop0p1} and
-\texttt{loop0p2}:
-%Use \texttt{lsblk} to view the partitions:
-
-\begin{lstlisting}[]
-# lsblk
-NAME      MAJ:MIN RM  SIZE RO TYPE MOUNTPOINT
-...
-loop0       7:0    0  2.3G  0 loop 
-|-loop0p1 259:2    0   63M  0 loop 
-|-loop0p2 259:3    0  2.2G  0 loop
-...
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Check and resize the root partition:
-\begin{lstlisting}[]
-# e2fsck -f ${DEV}p2
-# resize2fs ${DEV}p2
-\end{lstlisting}
-\FloatBarrier
-
-%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
-%We also notice that each sector is 512 bytes.
-%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"
-%Mount root partition:
-
-\subsection{Mount image}
-
-Browsing and altering the files in the image is easy. Simply mount the partitions.
-%To browser and alter the files in the image, it is possible to mount the partitions.
-
-Lets first mount the root partition. This is done by creating an empty directory
-that we will be used as a mountpoint. We name it \texttt{root} and create it in the
-working directory before mounting the root filesystem onto the mountpoint.
-%After this, mount the partitions:
-%After this,
-%we mount the partition starting at sector 137216. To provide this offset information
-%to mount, we need to find the offset in bytes. From the fdisk command we could also
-%see that each sector is 512 bytes. Thus, we multiply the sector size and the sector
-%start to get the offset in bytes.
-
-\begin{lstlisting}[]
-$ export ROOTDIR="${WORKDIR}/root"
-$ mkdir -p ${ROOTDIR}
-# mount ${DEV}p2 ${ROOTDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%# mount -o loop,offset=$((137216*512)) ${IMAGE}.img ${ROOTDIR}
-
-The root filesystem already has a boot directory that can be used as
-the mountpoint for the boot partition.
-%After this, we can also mount the boot partition inside the newly mounted
-%root filesystem.
-This is convenient because that partition will be mounted
-on this directory when a \ac{Raspi} starts up with a memory card containing
-the image. %It is therefore the natural place to mount it for later purposes.
-Now, mount boot partition:
-\begin{lstlisting}[]
-# mount ${DEV}p1 ${ROOTDIR}/boot
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%# mount -o loop,offset=$((8192*512)) ${IMAGE}.img ${ROOTDIR}/boot
-
-%We can now change all the files in the disk image as desired.
-It is now possible to change all files within the Raspbian image as desired.
-We will take advantage of this to edit configuration files, append new files,
-and even update and install packages.
-
-\subsection{Configuring OS files and scripts}
-
-The \ac{Raspi}s should be setup as similar as possible. One of those
-few settings that may be different in the devices is their hostname.
-%The hostname helps the user to distinguish the devices from each other.
-%The hostname helps distinguish them from each other.
-One approach to create unique hostnames would be to write the hostname
-in each memory card. Another way is to create a start up script to automatically
-configure the hostname. We decided to define the hostnames based on the \ac{MAC}
-addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
-addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
-administrator will need to insert the \ac{MAC} addresses and hostnames of
-his/her \ac{Raspi}s.
-
-%The \ac{Raspi}s should be setup to be as similar as possible. One of those
-%things that is preferred to be different on the devices is their hostname.
-%This helps distinguish them from each other. One solution to this in our
-%approach would be to uniquely change the hostname in each memory card.
-%This is however not desired because that would either require modifying
-%the customized Raspbian lite image before writing the content to the
-%memory card or to manually change it in the \ac{Raspi}s after the image has
-%been written to a memory card.
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-%
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames. This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-
-%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
-
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames.
-%This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-% MAC and Hostname file
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-# Ethernet MAC    Hostname
-b8:27:eb:5b:da:20 rasp00
-b8:27:eb:7b:c3:91 rasp01
-b8:27:eb:54:9c:64 rasp02
-b8:27:eb:95:bd:11 rasp03
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
-%the testbed, it still needs to know which name to give itself.
-%We do this using a small Bash script:
-%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
-%hostname in \texttt{nodes.csv}.
-A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:
-
-% Set hostname
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-script_path="$(dirname $(realpath $0))"
-config_file=${script_path}/nodes.csv 
-mac=$(cat /sys/class/net/eth0/address)
-old_hostname=$(hostname)
-new_hostname=$(grep $mac $config_file | cut -f2 -d' ')
-
-# Assign hostname found in nodes.csv
-if [ ! -z ${new_hostname} ]; then
-    echo ${new_hostname} > /etc/hostname
-    hostname ${new_hostname}
-    sed -i.old -e "s:${old_hostname}:${new_hostname}:g" /etc/hosts
-fi
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The code lines above do the following:
-\begin{itemize}
-    \item Line 1: Tells system to interpret the script using \ac{Bash}
-    \item Line 3-4: Gets the path to the script itself and the list of hostnames
-    \item Line 5: Gets the \ac{MAC} address of the node itself
-    \item Line 6: Gets the current hostname
-    \item Line 7: Gets the new hostname from the hostname list
-    \item Line 10-14: Assigns the new hostname
-\end{itemize}
-
-%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
-%path of the script itself. Line 4 gets the MAC address. This script will be
-%located in all \ac{Raspi}s and thus return the MAC address of the node itself
-%when executed.
-%Line 4 searches line by line for this mac address.
-%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
-%then it will keep the default hostname of the system.
-
-It is likely that the testbed administrator needs to either change or add
-scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
-all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
-configuration scripts during start up before running the configuration scripts.
-
-The below script attempts to download/update the scripts:
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-url="http://kom.aau.dk/project/TuneSCode/raspi/"
-config_file="rasp_config.zip"
-
-# Attempt to fetch new configuration files
-if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
-    echo "Warning: Unable to update rasp_config files"
-    exit 1
-fi
-
-# Unzip and overwrite configurationn files to root's home directory
-unzip -q -o /tmp/${config_file} -d /home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\begin{itemize}
-    \item Line 3-4: Is the url and zip file that should be downloaded
-    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
-        and prints a warning in case of errors
-    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
-\end{itemize}
-
-We have packed the files and uploaded them to a webserver. These can be
-retrieved with \texttt{wget} and unzipped into the Raspbian lite image:
-
-% Get the files
-\begin{lstlisting}[]
-$ wget ${URL%/*}/rasp_config.zip
-$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
-\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
-\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
-the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
-The testbed administrator will therefore need to also edit the \ac{URL} in the
-script \texttt{update\_rasp\_config}.
-%"nodes.csv" can be changed using a text editor. For vim, this could be done
-%the following way:
-%\begin{lstlisting}[]
-%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-Finally, to actually make the \ac{Raspi}s change hostname, we have to make
-each \ac{Raspi} call the above script when it starts up. Optionally,
-also run the update script before running other scripts.
-Do this with root permissions by inserting the line(s) below before
-\texttt{"exit 0"} using your favorite editor:
-%"\$\{ROOTDIR\}/etc/rc.local" using
-%your favorite text editor..
-%The file
-%can be opened with your favorite text editor.
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/set_hostname
-...
-exit 0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Configuring \texttt{rc.local} makes it really cumbersome for the testbed
-administrator to add new scripts since it first needs to be included
-in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional 
-line added to \texttt{rc.local} in the root image or in all memory cards depending on
-the final testbed setup.
-
-A solution to this is to make yet another script in \texttt{rasp\_config.zip}
-that calls all the other scripts in the desired order. This would look like
-below:
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-bash /home/pi/rasp_config/set_hostname
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, to use the above script, \texttt{rc.local} needs to look like this:
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/main
-...
-exit 0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-Notice that \texttt{set\_hostname} is now called by the \texttt{main}
-script instead. The update script is still called directly. This ensures that
-all configuration scripts are updated before executed. Changes to the update
-script itself will first take effect at next system startup.
-%Had the update script
-%been called in \texttt{main} instead would mean that changes to the 
-%\texttt{main} script would also first have been available at next startup.
-
-%This can be done by inserting a call to the script
-%set\_hostname in rc.local before it exits.
-%
-%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
-%\begin{lstlisting}[]
-%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
-%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
-%\end{lstlisting}
-%\FloatBarrier
-%$ line="bash /home/pi/.rasp_config/set_hostname"
-%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
-%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>
-
-%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
-%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
-
-%All other files could be updated as well.
-It is not only possible to add and configure files to/within the image. One can
-also install and update the packages within the image which will be illustrated
-in the next section.
-
-\subsection{Installing and updating the image}
-
-It may be desired to pre-install some programs in the image before it is
-written to all the memory cards that goes into the \ac{Raspi}s.
-This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
-(change root).
-%\url{https://wiki.debian.org/QemuUserEmulation}
-%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
-
-Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
-install some additional software:
-%Because \ac{Raspi}s are equipped with an \ac{ARM} processor 
-%%it is not as
-%straightforward as performing the task for two Linux systems of the same
-%architecture.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# apt-get install binfmt-support qemu qemu-user-static
-# update-binfmts --display qemu-arm
-qemu-arm (enabled):
-     package = qemu-user-static
-        type = magic
-      offset = 0
-       magic = \x7fELF\x01\x01\x01\x00\x00\x00\x00\x00\x00\x00\x00\x00\x02\x00\x28\x00
-        mask = \xff\xff\xff\xff\xff\xff\xff\x00\xff\xff\xff\xff\xff\xff\xff\xff\xfe\xff\xff\xff
- interpreter = /usr/bin/qemu-arm-static
-    detector = 
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-% This was neccessary on arch
-%# update-binfmts --importdir /var/lib/binfmts/ --import
-
-Make sure the second command writes "enabled" as the output above. If that 
-is not the case, then try enabling it:
-
-\begin{lstlisting}[]
-# update-binfmts --enable qemu-arm
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Provided that qemu-arm is enabled, we should now be able to change root (chroot)
-into our Raspbian lite image. Change root is a method in Linux that enables
-the root (/) to be changed. Thus, enabling a Linux installation within a
-Linux installation.
-
-There are a few commands to be performed before actually changing root into the
-root partion of the image.
-First, to get internet access from within the Raspbian lite image it is needed
-to copy our \texttt{resolv.conf} into the filesystem. Line 1 steps into Raspbian lite's
-root filesystem. Then line 2 copies resolv.conf.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-$ cd $ROOTDIR
-# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, because of the ARM architecture, it is also required to copy
-/usr/bin/qemu-arm-static into the image before we can continue:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The final preparation before chaning root is to populate the directories proc,
-sys, and dev:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# mount  -t proc proc proc/
-# mount --bind /sys sys/
-# mount --bind /dev dev/
-# mount --bind /dev/pts dev/pts
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Finally, the filesystem is ready for us to change root. This can be done with
-"proot". It may be required to install proot:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# proot -q qemu-arm-static -S ${ROOTDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Most available online material uses the more known chroot command as written in
-the code block below. This did not work correctly in our machines, but we show
-it as an alternative in case proot may not work on other systems.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If everything went well, we should now be standing in the Raspbian lite filesystem
-under its root. Optionally, we can change the prompt title to indicate that we
-are chrooted:
-
-% Optionally, we may create a unique prompt to indicate we have changed root
-\begin{lstlisting}[]
-# export PS1="(chroot) $PS1"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The Raspbian lite system should now be possible to use as if it had been booted
-in a \ac{Raspi}. The only difference is that the laptop that we are using is
-significantly faster than a \ac{Raspi}. Thus enabling us to for example update,
-upgrade, and install software packages:
-
-% Update system
-\begin{lstlisting}[]
-(chroot) # apt-get update
-(chroot) # apt-get upgrade
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Lets install some useful applications:
-% Install packages
-\begin{lstlisting}[]
-(chroot) # apt-get install vim git screen
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-All the changes that are made here will exists in all \ac{Raspi}s when the
-image is written to a memory card.
+\label{sec:ovelay_fs}
 

### 2016-07-01 overview.tex
New pieces of the structure

--- 2016-06-30 testbed.tex
@@
-\label{sec:testbed}
+\label{sec:overview}
 
@@
 enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP} server as
-mentioned previously. The following procedure explains how the testbed may be setup and configured.
+mentioned previously.
+
+%The following procedure explains how the testbed may be setup and configured.
 
@@
 %may be customized
-
-\subsubsection{Installing Raspbian}
-
-To get started, we first need to install an \ac{OS} on the
-\ac{Raspi}s. We will use the popular Debian-based Raspbian
-Linux~\cite{raspbian}
-
-We will be using a \ac{PC} running a Debian-based Linux distribution
-to setup a customized Raspbian Linux image that can be written to
-a memory card and work in any \ac{Raspi} version 1 or 2.
-
-Raspbian is made available in two bundles: Raspbian and Raspbian lite.
-The difference between the two is that Raspbian contains a pre-installed desktop
-environment, and Raspbian lite is by default only accessible through a 
-command shell. Both Raspbian and Raspbian lite will both by default be remotely
-accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
-connect to another machine through a command shell.
-%In fact it is the successor
-%of \ac{Telnet} that has been designed with network security in mind.
-
-The \ac{Raspi}s in our testbed are not connected to monitors, so we
-have decided to go with Raspbian lite.
-%Our testbed does not have monitors so we will install Raspbian lite.
-A desktop environment can be installed using the package manager
-(apt-get) in case a desktop environment will be desired at a later
-time.
-
-Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
-have performed a simple setup where each \ac{Raspi} is equipped with a
-memory card that contains a slightly customized Raspbian lite image. The image
-of the \ac{OS} in all \ac{Raspi}s will be identical.
-
-The overall procedure to customize the official Raspbian lite image
-are:
-\begin{enumerate}
-    \item Download Raspbian lite
-    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
-        in the official Raspbian lite image
-    \item Change root, i.e. change root filesystem into the Raspbian lite
-        image to update and install software packages
-    \item Write image to memory cards
-\end{enumerate}
-
-Hashmark (\#) and dollar (\$) will be used throughout the code blocks in
-the paper to indicate if an operation needs to be run with root permissions or
-user permissions.
-
-\subsection{Download Raspbian lite}
-
-The latest Raspbian lite bundle can be downloaded from Raspbian's official
-webpage~\cite{raspbian}.
-%To download the latest Raspbian lite, we go to the url
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
-At the time of this writing, the latest available bundle was
-\texttt{2016-05-27-raspbian-jessie-lite.zip}.
-
-To ensure that the content of the bundle does not change, this procedure
-will be based on that particular version of Raspbian lite which we have
-made available at~\cite{tunescode_webpage}. All other files used in this paper
-will also be available there unless specified otherwise.
-
-%We want to proceed downloading the bundle in a shell instead of the browser.
-%To do this, first open a command shell on a linux machine. This guide is
-%assuming a Debian-based distribution, but there should only be minor differences
-%if you have another package manager. i.e. the package managers and name of software packages.
-
-To get started, open a Linux shell (terminal) and declare the variables below.
-The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
-located and \texttt{WORKDIR} specifies a working directory where we will 
-store Raspbian and customize it.
-Do not type in \$ and \# in the shell. Those indicates if a command can be
-executed as a regular user or if root permissions are required.
-Running a command as root can on most systems be done by putting
-\texttt{sudo} in front of the command. This is illustrated in the
-following code block 
-with the command \texttt{whoami} that prints the username. Lines within a code
-block without leading \$ or \# is terminal output or content of a file.
-
-%After the terminal has been opened, we start by declaring a few variables to reduce
-%repeated typing. The first variables we declare is the Raspbian image name.
-%Notice that the extension has been omitted. This is because the image has been
-%packed into a zip file. The other variable we declare is a working directory.
-%This is where we will download the image to and work on it. In other words, it
-%will be stored in \texttt{/home/<username>/Raspbian}
-
-%To download Raspbian lite, we go to the \url{http://downloads.raspberrypi.org}.
-%There should be a folder called raspbian\lite and
-%to find the download page at which Raspbian lite should be located. Instead of
-%downloading with the browser, we just extract the download url to the latest
-%image. At the time of this writing, that was
-
-% Define variable]
-\begin{lstlisting}[]
-$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
-$ export IMAGE="2016-05-27-raspbian-jessie-lite"
-$ export WORKDIR="${HOME}/Raspbian"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%A final thing to notice is the way \$ and \# are used in the code blocks. \$
-%means that the command can be run with normal user permissions where \# indicates that the
-%command needs root permissions.
-%Running a command with root permissions can on most systems be done by putting
-%the command "sudo" in front of each line. This is illustrated in the code below
-%with the command "whoami" that prints the username of the caller. Lines without
-%leading \$ or \# is the terminal output.
-
-% Root example
-\begin{lstlisting}[]
-$ whoami
-<USERNAME>
-$ sudo whoami
-root
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Next, make the working directory and used \texttt{cd} to change into the working
-directory. Then \texttt{wget} can be used to download Raspbian lite before
-unpacking the \texttt{zip} file:
-
-% Download and unpack image
-\begin{lstlisting}[]
-$ mkdir -p ${WORKDIR}
-$ cd ${WORKDIR}
-$ wget ${URL%/*}/${IMAGE}.zip
-$ unzip ${IMAGE}.zip
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%A more recent version of Raspbian lite may be available at~\cite{raspbian}.
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.
-
-\subsection{The image}
-
-After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
-\texttt{fdisk} can be used to display the content of the image. Parse the arguments
-"-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
-image and and automatically exit \texttt{fdisk} again:
-
-% Check out the image   % The dollar hack was to fix vim syntax
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-
-The output provides knowledge about the image. The image is in total
-2709504 sectors (1.3GiB) in size and contains two
-partitions. The first partition starts at sector 8192 and
-the other partition starts at sector 137216. The partitions is of type FAT32
-and Linux and are 63M and 1.2G big. This indicates that the first partition
-is most likely a boot partition and the second partition is a traditional Linux file
-system. Actually, it is the Linux root filesystem (i.e. /).
-
-\subsection{Resize image}
-1.2G is not very much to store the root filesystem given that it is desired to
-alter the files and install additional packages. It may therefore be
-necessary to increase the partition size.
-The following illustrates how the image and its root filesystem
-can be expanded by one \ac{GB}.
-
-First, expand the image one \ac{GB} and wait for the data to be completely written
-to the disk:
-\begin{lstlisting}[]
-$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Use \texttt{fdisk} to see that the image is now one \ac{GB} larger:
-\begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, to expand the root filesystem it is required to first remove the Linux partition
-and then add it again one \ac{GB} bigger.
-Use \texttt{fdisk} to alter the partition table. The \texttt{echo}'s within the parenthesis
-will be interpreted as
-key-presses in \texttt{fdisk} and hence delete the last partition, create a new
-primary partition, print new partition table, and write new partition table
-to the image:
-
-\begin{lstlisting}[]
-$ (echo d; echo ; echo n; echo p; echo ; echo 137216; echo ; echo p; echo w) | fdisk ${IMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Use a loopback device to make the Raspbian image available as a block device: 
-\begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\texttt{lsblk} may be used to list block devices. It appears that the image is
-available as \texttt{/dev/loop0} which has two partitions \texttt{loop0p1} and
-\texttt{loop0p2}:
-%Use \texttt{lsblk} to view the partitions:
-
-\begin{lstlisting}[]
-# lsblk
-NAME      MAJ:MIN RM  SIZE RO TYPE MOUNTPOINT
-...
-loop0       7:0    0  2.3G  0 loop 
-|-loop0p1 259:2    0   63M  0 loop 
-|-loop0p2 259:3    0  2.2G  0 loop
-...
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Check and resize the root partition:
-\begin{lstlisting}[]
-# e2fsck -f ${DEV}p2
-# resize2fs ${DEV}p2
-\end{lstlisting}
-\FloatBarrier
-
-%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
-%We also notice that each sector is 512 bytes.
-%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"
-%Mount root partition:
-
-\subsection{Mount image}
-
-Browsing and altering the files in the image is easy. Simply mount the partitions.
-%To browser and alter the files in the image, it is possible to mount the partitions.
-
-Lets first mount the root partition. This is done by creating an empty directory
-that we will be used as a mountpoint. We name it \texttt{root} and create it in the
-working directory before mounting the root filesystem onto the mountpoint.
-%After this, mount the partitions:
-%After this,
-%we mount the partition starting at sector 137216. To provide this offset information
-%to mount, we need to find the offset in bytes. From the fdisk command we could also
-%see that each sector is 512 bytes. Thus, we multiply the sector size and the sector
-%start to get the offset in bytes.
-
-\begin{lstlisting}[]
-$ export ROOTDIR="${WORKDIR}/root"
-$ mkdir -p ${ROOTDIR}
-# mount ${DEV}p2 ${ROOTDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%# mount -o loop,offset=$((137216*512)) ${IMAGE}.img ${ROOTDIR}
-
-The root filesystem already has a boot directory that can be used as
-the mountpoint for the boot partition.
-%After this, we can also mount the boot partition inside the newly mounted
-%root filesystem.
-This is convenient because that partition will be mounted
-on this directory when a \ac{Raspi} starts up with a memory card containing
-the image. %It is therefore the natural place to mount it for later purposes.
-Now, mount boot partition:
-\begin{lstlisting}[]
-# mount ${DEV}p1 ${ROOTDIR}/boot
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%# mount -o loop,offset=$((8192*512)) ${IMAGE}.img ${ROOTDIR}/boot
-
-%We can now change all the files in the disk image as desired.
-It is now possible to change all files within the Raspbian image as desired.
-We will take advantage of this to edit configuration files, append new files,
-and even update and install packages.
-
-\subsection{Configuring OS files and scripts}
-
-The \ac{Raspi}s should be setup as similar as possible. One of those
-few settings that may be different in the devices is their hostname.
-%The hostname helps the user to distinguish the devices from each other.
-%The hostname helps distinguish them from each other.
-One approach to create unique hostnames would be to write the hostname
-in each memory card. Another way is to create a start up script to automatically
-configure the hostname. We decided to define the hostnames based on the \ac{MAC}
-addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
-addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
-administrator will need to insert the \ac{MAC} addresses and hostnames of
-his/her \ac{Raspi}s.
-
-%The \ac{Raspi}s should be setup to be as similar as possible. One of those
-%things that is preferred to be different on the devices is their hostname.
-%This helps distinguish them from each other. One solution to this in our
-%approach would be to uniquely change the hostname in each memory card.
-%This is however not desired because that would either require modifying
-%the customized Raspbian lite image before writing the content to the
-%memory card or to manually change it in the \ac{Raspi}s after the image has
-%been written to a memory card.
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-%
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames. This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-
-%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
-
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames.
-%This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-% MAC and Hostname file
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-# Ethernet MAC    Hostname
-b8:27:eb:5b:da:20 rasp00
-b8:27:eb:7b:c3:91 rasp01
-b8:27:eb:54:9c:64 rasp02
-b8:27:eb:95:bd:11 rasp03
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
-%the testbed, it still needs to know which name to give itself.
-%We do this using a small Bash script:
-%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
-%hostname in \texttt{nodes.csv}.
-A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:
-
-% Set hostname
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-script_path="$(dirname $(realpath $0))"
-config_file=${script_path}/nodes.csv 
-mac=$(cat /sys/class/net/eth0/address)
-old_hostname=$(hostname)
-new_hostname=$(grep $mac $config_file | cut -f2 -d' ')
-
-# Assign hostname found in nodes.csv
-if [ ! -z ${new_hostname} ]; then
-    echo ${new_hostname} > /etc/hostname
-    hostname ${new_hostname}
-    sed -i.old -e "s:${old_hostname}:${new_hostname}:g" /etc/hosts
-fi
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The code lines above do the following:
-\begin{itemize}
-    \item Line 1: Tells system to interpret the script using \ac{Bash}
-    \item Line 3-4: Gets the path to the script itself and the list of hostnames
-    \item Line 5: Gets the \ac{MAC} address of the node itself
-    \item Line 6: Gets the current hostname
-    \item Line 7: Gets the new hostname from the hostname list
-    \item Line 10-14: Assigns the new hostname
-\end{itemize}
-
-%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
-%path of the script itself. Line 4 gets the MAC address. This script will be
-%located in all \ac{Raspi}s and thus return the MAC address of the node itself
-%when executed.
-%Line 4 searches line by line for this mac address.
-%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
-%then it will keep the default hostname of the system.
-
-It is likely that the testbed administrator needs to either change or add
-scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
-all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
-configuration scripts during start up before running the configuration scripts.
-
-The below script attempts to download/update the scripts:
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-url="http://kom.aau.dk/project/TuneSCode/raspi/"
-config_file="rasp_config.zip"
-
-# Attempt to fetch new configuration files
-if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
-    echo "Warning: Unable to update rasp_config files"
-    exit 1
-fi
-
-# Unzip and overwrite configurationn files to root's home directory
-unzip -q -o /tmp/${config_file} -d /home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\begin{itemize}
-    \item Line 3-4: Is the url and zip file that should be downloaded
-    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
-        and prints a warning in case of errors
-    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
-\end{itemize}
-
-We have packed the files and uploaded them to a webserver. These can be
-retrieved with \texttt{wget} and unzipped into the Raspbian lite image:
-
-% Get the files
-\begin{lstlisting}[]
-$ wget ${URL%/*}/rasp_config.zip
-$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
-\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
-\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
-the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
-The testbed administrator will therefore need to also edit the \ac{URL} in the
-script \texttt{update\_rasp\_config}.
-%"nodes.csv" can be changed using a text editor. For vim, this could be done
-%the following way:
-%\begin{lstlisting}[]
-%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-Finally, to actually make the \ac{Raspi}s change hostname, we have to make
-each \ac{Raspi} call the above script when it starts up. Optionally,
-also run the update script before running other scripts.
-Do this with root permissions by inserting the line(s) below before
-\texttt{"exit 0"} using your favorite editor:
-%"\$\{ROOTDIR\}/etc/rc.local" using
-%your favorite text editor..
-%The file
-%can be opened with your favorite text editor.
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/set_hostname
-...
-exit 0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Configuring \texttt{rc.local} makes it really cumbersome for the testbed
-administrator to add new scripts since it first needs to be included
-in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional 
-line added to \texttt{rc.local} in the root image or in all memory cards depending on
-the final testbed setup.
-
-A solution to this is to make yet another script in \texttt{rasp\_config.zip}
-that calls all the other scripts in the desired order. This would look like
-below:
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-#!/usr/bin/env bash
-
-bash /home/pi/rasp_config/set_hostname
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, to use the above script, \texttt{rc.local} needs to look like this:
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
------------------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/main
-...
-exit 0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-Notice that \texttt{set\_hostname} is now called by the \texttt{main}
-script instead. The update script is still called directly. This ensures that
-all configuration scripts are updated before executed. Changes to the update
-script itself will first take effect at next system startup.
-%Had the update script
-%been called in \texttt{main} instead would mean that changes to the 
-%\texttt{main} script would also first have been available at next startup.
-
-%This can be done by inserting a call to the script
-%set\_hostname in rc.local before it exits.
-%
-%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
-%\begin{lstlisting}[]
-%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
-%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
-%\end{lstlisting}
-%\FloatBarrier
-%$ line="bash /home/pi/.rasp_config/set_hostname"
-%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
-%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>
-
-%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
-%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
-
-%All other files could be updated as well.
-It is not only possible to add and configure files to/within the image. One can
-also install and update the packages within the image which will be illustrated
-in the next section.
-
-\subsection{Installing and updating the image}
-
-It may be desired to pre-install some programs in the image before it is
-written to all the memory cards that goes into the \ac{Raspi}s.
-This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
-(change root).
-%\url{https://wiki.debian.org/QemuUserEmulation}
-%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
-
-Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
-install some additional software:
-%Because \ac{Raspi}s are equipped with an \ac{ARM} processor 
-%%it is not as
-%straightforward as performing the task for two Linux systems of the same
-%architecture.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# apt-get install binfmt-support qemu qemu-user-static
-# update-binfmts --display qemu-arm
-qemu-arm (enabled):
-     package = qemu-user-static
-        type = magic
-      offset = 0
-       magic = \x7fELF\x01\x01\x01\x00\x00\x00\x00\x00\x00\x00\x00\x00\x02\x00\x28\x00
-        mask = \xff\xff\xff\xff\xff\xff\xff\x00\xff\xff\xff\xff\xff\xff\xff\xff\xfe\xff\xff\xff
- interpreter = /usr/bin/qemu-arm-static
-    detector = 
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-% This was neccessary on arch
-%# update-binfmts --importdir /var/lib/binfmts/ --import
-
-Make sure the second command writes "enabled" as the output above. If that 
-is not the case, then try enabling it:
-
-\begin{lstlisting}[]
-# update-binfmts --enable qemu-arm
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Provided that qemu-arm is enabled, we should now be able to change root (chroot)
-into our Raspbian lite image. Change root is a method in Linux that enables
-the root (/) to be changed. Thus, enabling a Linux installation within a
-Linux installation.
-
-There are a few commands to be performed before actually changing root into the
-root partion of the image.
-First, to get internet access from within the Raspbian lite image it is needed
-to copy our \texttt{resolv.conf} into the filesystem. Line 1 change directory
-into root filesystem of the image. The second line copies \texttt{resolv.conf}
-from the testbed administrators \ac{PC} to the image.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-$ cd $ROOTDIR
-# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, because of the \ac{ARM} architecture,
-\texttt{/usr/bin/qemu-arm-static} needs to be copied into the image before continuing:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The final preparation before changing root is to populate the directories \texttt{proc},
-\texttt{sys}, and \texttt{dev}:
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# mount  -t proc proc proc/
-# mount --bind /sys sys/
-# mount --bind /dev dev/
-# mount --bind /dev/pts dev/pts
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Finally, it is time to change root.
-%Finally, the filesystem is ready for us to change root.
-This can be done with \texttt{proot}. It may be required to install \texttt{proot}
-using "\texttt{apt-get install proot}":
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# proot -q qemu-arm-static -S ${ROOTDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Most available material online uses the more known \texttt{chroot} command as
-written in the code block below instead of \texttt{proot}. This did not work
-correctly in our machines, but we present the command as an alternative to
-\texttt{proot} in case the testbed administrator should be in the opposite
-situration where \texttt{proot} is not working.
-
-% CHROOT to OS image
-\begin{lstlisting}[]
-# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If everything went well, we should now be \texttt{chrooted} into the Raspbian lite
-filesystem under its root. Optionally, change the prompt title to indicate that it
-is a \texttt{chrooted} environment:
-
-% Optionally, we may create a unique prompt to indicate we have changed root
-\begin{lstlisting}[]
-# export PS1="(chroot) $PS1"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The Raspbian lite system image should now be possible to use almost as if it
-had been booted in a \ac{Raspi}.
-A major difference is that the testbed administrator's \ac{PC} is likely
-significantly faster than a \ac{Raspi}.
-%The only difference is that the laptop that we are using is
-%significantly faster than a \ac{Raspi}.
-Thus enabling updates, upgrades, and installing new software packages much
-faster than in the \ac{Raspi}s.
-%Thus enabling us to for example update,
-%upgrade, and install software packages:
-
-% Update system
-\begin{lstlisting}[]
-(chroot) # apt-get update
-(chroot) # apt-get upgrade
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-We also install a few packages that we find useful:
-%Lets install some useful applications:
-% Install packages
-\begin{lstlisting}[]
-(chroot) # apt-get install vim git screen
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-When writing the image to a memory card, all these change that has been made
-to the image so far will exists in all \ac{Raspi}s.
-%All the changes that are made here will exists in all \ac{Raspi}s when the
-%image is written to a memory card.
-
-\subsection{Installation an overlaying filesystem}
-
-Our goal was to enable the users of the testbed to change however the
-\ac{Raspi} are configured while still being able to revert those changes easily.
-
-We do this by storing a users work in the \ac{Raspi}s in \ac{RAM}. Thus a
-\ac{Raspi}, when a \ac{Raspi} is rebooted, it will forget all changes.
-
-A way to do this is using stacked filesystem. Stacking filesystem enables
-a filesystem to overlay another filesystem. Whenever a file is accessed,
-the upper filesystem will forward the request to the lower filesystem in
-case it does not have it itself. If it has it, it will simpy return the file.
-
-This idea can be used in our setup to mount the root filesystem (i.e. Raspbian
-lite) in the \ac{Raspi}s duing startup as read-only. Thus, the files will not
-be changable. Some programs needs to change files and the testbed admin might
-want to change configurations. To enable this, it is possible to
-overlay a filesystem that is mounted in \ac{RAM} as read/writable on top of
-the regular root. Thus, reading a file may return a file from the lower filesytem,
-but if it is stored, it will be saved in the upper filesystem. Accessing this file
-again will return the stored file from the upper layer.
-
-%We do this using an overlay filesystem. The idea is to let the \ac{Raspi}s
-%mount their root filesystem as 
-%
-%We do this using an overlay filesystem (overlayfs).
-%Overlayfs enables a filesystem to be stacked on top of another filesystem. Thus,
-%if the \ac{Raspi}s mount their root filesystem (/) as read-onlyif we mount our root (/) filesystem as 
-%
-%
-%Overlayfs enables filesystems
-%to be stacked on top of each other. We will be using this to make a lower
-%filesystem and an upper filesystem. The lower filesystem will be our
-%traditional filesystem and the upper filesystem will be mounted in tmpfs
-%(i.e. mounted in \ac{RAM}).
-%
-%Thus, when a file is requsted from stacked filesystem, it will first attempt
-%to fetch the file in the upper filesystem if it does not exists there, it will
-%request it from the lower filessytem.
-
-Assuming that you are still \texttt{chrooted} into the Raspbian lite filesystem, we
-can setup the overlay filesystem (overlayfs).
-
-There already exists implementations overlaying the root filesystem. We will
-use an existing implementation
-available at github: \cite{overlayroot}.
-
-We installed git in a previous step, so lets clone the repository. The below
-code stores it in /tmp which is really mounted in \ac{RAM}. Thus, the files
-will disappear when the machine is rebooted.
-
-% Get files
-\begin{lstlisting}[]
-(chroot) # OVERLAYROOTDIR="/tmp/overlayroot"
-(chroot) # git clone https://github.com/chesty/overlayroot.git $OVERLAYROOTDIR
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-To enable the overlaying filesystem it is required to generate an initial \ac{RAM}
-file system (initramfs). Initramfs is an initial filesystem that gets loaded into
-\ac{RAM} during the startup process of a linux machine to prepare the real
-filesystem.
-
-For this purpose, we need BuxyBox:
-
-% Install required packages
-\begin{lstlisting}[]
-(chroot) # apt-get install busybox
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Add and activate the overlaying filesystem, we need to add the scripts and tell initramfs
-to load the overlaying filesytem:
-
-% Setup initramfs
-\begin{lstlisting}[]
-(chroot) # cp ${OVERLAYROOTDIR}/hooks-overlay /etc/initramfs-tools/hooks/
-(chroot) # cp ${OVERLAYROOTDIR}/init-bottom-overlay /etc/initramfs-tools/scripts/init-bottom/
-(chroot) # echo "overlay" > /etc/initramfs-tools/modules
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%Finally, lets generate the initial \ac{RAM} filesystem. This is done using the command
-%\texttt{mkinitramfs}, that will by default look for the kernel modules of the current system. Remember we are
-%chrooted, so we need to specify the correct kernel modules to look for. These
-%can be found in "/lib/modules".
-Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
-needs a different kernel than \ac{Raspi} version 2 and version 3. Thus, start
-by finding the available kernel modules:
-%Because \texttt{initramfs} is executed within a
-%\texttt{chrooted} environment, it will not be able to detect the
-%kernel modules in the \texttt{chrooted} environment. The kernel modules will
-
-% Create an initramfs
-\begin{lstlisting}[]
-(chroot) # ls /lib/modules/
-4.4.13+  4.4.13-v7+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Kernel "4.4.13+" is for \ac{Raspi} version 1 and kernel "4.4.13-v7+" works
-for \ac{Raspi} version 2 and version 3.
-
-Lets proceed by generating an initial \ac{RAM} filesystem for both kernels:
-
-%Lets proceed with kernel modules "4.4.13+".
-%Lets create an initial ram filesystem for both kernels.
-
-%Thus, to run mkinitramfs, type in the correct kernel version:
-
-% Create an initramfs
-\begin{lstlisting}[]
-(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
-(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
-%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+
-
-Although there might be printet a few warnings, the above commands should successfully
-generate working initial \ac{RAM} filesystems.
-
-An initial \ac{RAM} filesystem needs to be called by the bootloader. In Raspbian
-this is done by adding a command to \texttt{config.txt} in the boot partition.
-
-If the system should be run in a \ac{Raspi} version 1, then use \texttt{init.gz}:
-%The bootloader should be made aware of the initramfs that we will generate:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If the system should be run in a \ac{Raspi} version 2 or 3, then use \texttt{init-v7.gz}:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # Only for Raspberry Pi version 2 and 3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-At this point, it is no longer required to be \texttt{chrooted} and the image
-can be written to a memory card. The following commands exits the \texttt{chrooted}
-environment, unmount all partitions, and detaches the loopback device:
-
-% Exit chroot, umount, and write to memory card
-\begin{lstlisting}[]
-(chroot) # exit
-# cd ..
-# umount --recursive ${ROOTDIR}
-# losetup -d ${DEV}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsection{(Optional) Persistent subdirectories for distributed Linux image}
-
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-Provided the stacked filesystem was configured, it may be desired to have
-some directories where files are not flushed upon rebooting the \ac{Raspi}s.
-The following procedure will create an extra partition in the image for
-the \ac{Raspi} users home directory that will be made storage persistent.
-%Provided that a stacked filesystem was configured to make the root
-%filesystem forget chances upon reboots is not always desired.
-
-Expand image according to the desired home directory size. Keep in mind not
-to make the image bigger than the target memory card.
-% Expand image
-\begin{lstlisting}[]
-$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create home partition after the root partition. Use \texttt{fdisk} to
-find the next available sector in the image. In our case, the new
-partition should start at sector 4806656:
-\begin{lstlisting}[]
-$ fdisk -lu ${IMAGE}.img
-Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
-Units: sectors of 1 * 512 = 512 bytes
-Sector size (logical/physical): 512 bytes / 512 bytes
-I/O size (minimum/optimal): 512 bytes / 512 bytes
-Disklabel type: dos
-Disk identifier: 0x6fcf21f3
-
-Device                               Boot  Start     End Sectors  Size Id Type
-2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
-2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create new partition:
-
-\begin{lstlisting}[]
-$ (echo n; echo p; echo ; echo 4806656; echo ; echo p; echo w) | fdisk ${IMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create loopback device: 
-\begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Format new partition:
-\begin{lstlisting}[]
-# mkfs.ext4 ${DEV}p3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The last thing to remember is to inform Raspbian to mount the home partition
-we created during startup. This can be done by adding an entry in \texttt{fstab}:
-\begin{lstlisting}[]
-# mount ${DEV}p2 ${ROOTDIR}
-# sudo sh -c "echo '/dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' >> ${ROOTDIR}/etc/fstab"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The home folder was originally located in the root filesystem. We should
-therefore copy the content to the new home partition:
-\begin{lstlisting}[]
-# mount ${DEV}p3 ${ROOTDIR}/mnt
-# mv ${ROOTDIR}/home/* ${ROOTDIR}/mnt/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-That is it. unmount partitions and detach loop device:
-\begin{lstlisting}[]
-# umount --recursive ${ROOTDIR}
-# losetup -d ${DEV}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsection{Write image to memory card}
-
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-The final step is to write the image to all the memory cards before they can
-be used in the \ac{Raspi}s.
-
-%This is done by first leaving the \texttt{chrooted}
-%filesystem and unmounting all the directories we mounted:
-
-% Exit chroot, umount, and write to memory card
-%\begin{lstlisting}[]
-%# exit
-%# cd ..
-%# umount --recursive ${ROOTDIR}
-%# losetup -d ${DEV}.img
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-%If successfully unmounted, you can 
-Insert a memory card in your machine. It will be available as
-\texttt{/dev/mmcblkX} or \texttt{/dev/sdX}
-
-%In most systems, this memory card was available under \texttt{/dev/mmcblk0}, but this may be
-%different on different machines. For internal memory card readers, it will
-%probably be named /dev/mmcblkX where X is some number and if the memory card
-%reader is connected through \ac{USB}, then it is likely called /dev/sdX.
-Make absolutely sure to write to the correct device as everything will be 
-overwritten!
-Inserting/Ejecting the memory card a couple of times and using the commands
-below may help deducing the correct device:
-%The commands below may help deduce which device is the correct:
-%Below is a few commands that may help you to deduce which device is correct:
-
-\begin{lstlisting}[]
-# lsblk
-# df -h
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For us, the device was \texttt{/dev/mmcblk0}. Write the image to a memory card:
-\begin{lstlisting}[]
-# dd if=${IMAGE}.img of=/dev/mmcblk0 bs=4M && sync
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-The memory card can now be started in a \ac{Raspi}.

### 2016-07-01 image_setup.tex
Disclaimer about proper copied characters rendering in output PDF

@@
 \label{sec:image_setup}
+In this Section, we review the steps to create a common \ac{OS} image for all the
+\ac{Raspi}s. The image setup is composed of three major steps: Select and download the
+\ac{OS} image file, alter the image structure and configure the \ac{OS} files.
+We proceed to detail all these steps providing brief discussions to our setup
+choices when required. To perform these steps, we indicate with command-line
+blocks the required sequential commands to be typed by the testbed
+administrator in its \ac{PC} to obtain the desired setting. In all the
+following sections, the number sign (\#) and dollar sign (\$) will be used in
+the command blocks in the paper to indicate if an operation needs
+to be run with root permissions or common user permissions, respectively. The
+signs will be posted as the terminal prompt in the command blocks.
 
-\subsection{Installing Raspbian}
+\subsection{OS Selection and Download}
 
-To get started, we first need to install an \ac{OS} on the
-\ac{Raspi}s. We will use the popular Debian-based Raspbian
-Linux~\cite{raspbian}
+To get started, we first need to install an \ac{OS} that works properly
+on all the \ac{Raspi} models. We will download and setup the image in the testbed
+administrator \ac{PC} using a Debian-based distribution. We use the popular
+Debian-based Raspbian Linux~\cite{raspbian} given that is the recommended
+and default \ac{OS} for the \ac{Raspi}.
+Raspbian is made available in two bundles: Raspbian and Raspbian lite.
+The difference between the two, is that Raspbian contains a pre-installed desktop
+environment for user interaction and Raspbian lite by default only permits to
+interact through a command shell. Given that the \ac{Raspi}s in our testbed are
+not connected to monitors, we decide to work with Raspbian lite. Still,
+if later desired, a desktop environment can be installed using the package manager
+as it will be shown in the configuration section.
 
-We will be using a \ac{PC} running a Debian-based Linux distribution
-to setup a customized Raspbian Linux image that can be written to
-a memory card and work in any \ac{Raspi} version 1 or 2.
+% The overall procedure to customize the official Raspbian lite image
+% are:
+% \begin{enumerate}
+%     \item Download Raspbian lite
+%     \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
+%         in the official Raspbian lite image
+%     \item Change root, i.e. change root filesystem into the Raspbian lite
+%         image to update and install software packages
+%     \item Write image to memory cards
+% \end{enumerate}
 
-Raspbian is made available in two bundles: Raspbian and Raspbian lite.
-The difference between the two is that Raspbian contains a pre-installed desktop
-environment, and Raspbian lite is by default only accessible through a 
-command shell. Both Raspbian and Raspbian lite will both by default be remotely
-accessible using \ac{SSH}. \ac{SSH} provides the means to remotely
-connect to another machine through a command shell.
-%In fact it is the successor
-%of \ac{Telnet} that has been designed with network security in mind.
-
-The \ac{Raspi}s in our testbed are not connected to monitors, so we
-have decided to go with Raspbian lite.
-%Our testbed does not have monitors so we will install Raspbian lite.
-A desktop environment can be installed using the package manager
-(apt-get) in case a desktop environment will be desired at a later
-time.
-
-Below is the procedure we followed to setup the \ac{Raspi}s in our testbed. We
-have performed a simple setup where each \ac{Raspi} is equipped with a
-memory card that contains a slightly customized Raspbian lite image. The image
-of the \ac{OS} in all \ac{Raspi}s will be identical.
-
-The overall procedure to customize the official Raspbian lite image
-are:
-\begin{enumerate}
-    \item Download Raspbian lite
-    \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
-        in the official Raspbian lite image
-    \item Change root, i.e. change root filesystem into the Raspbian lite
-        image to update and install software packages
-    \item Write image to memory cards
-\end{enumerate}
-
-Hashmark (\#) and dollar (\$) will be used throughout the code blocks in
-the paper to indicate if an operation needs to be run with root permissions or
-user permissions.
-
-\subsection{Download Raspbian lite}
-
-The latest Raspbian lite bundle can be downloaded from Raspbian's official
-webpage~\cite{raspbian}.
+%\subsection{Download Raspbian lite}
 %To download the latest Raspbian lite, we go to the url
 %\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
-At the time of this writing, the latest available bundle was
-\texttt{2016-05-27-raspbian-jessie-lite.zip}.
 
+The latest Raspbian lite bundle can be downloaded from the Raspbian
+official webpage~\cite{raspbian}. At the time of this writing, the latest
+available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
 To ensure that the content of the bundle does not change, this procedure
-will be based on that particular version of Raspbian lite which we have
+is based on that particular version of Raspbian lite which we have
 made available at~\cite{tunescode_webpage}. All other files used in this paper
-will also be available there unless specified otherwise.
-
-%We want to proceed downloading the bundle in a shell instead of the browser.
-%To do this, first open a command shell on a linux machine. This guide is
-%assuming a Debian-based distribution, but there should only be minor differences
-%if you have another package manager. i.e. the package managers and name of software packages.
-
-To get started, open a Linux shell (terminal) and declare the variables below.
-The \texttt{URL} and \texttt{IMAGE} specifies where the Linux bundle will be
-located and \texttt{WORKDIR} specifies a working directory where we will 
-store Raspbian and customize it.
-Do not type in \$ and \# in the shell. Those indicates if a command can be
-executed as a regular user or if root permissions are required.
-Running a command as root can on most systems be done by putting
-\texttt{sudo} in front of the command. This is illustrated in the
-following code block 
-with the command \texttt{whoami} that prints the username. Lines within a code
-block without leading \$ or \# is terminal output or content of a file.
+are available there unless specified otherwise. To get started, open a
+Linux shell (terminal) and declare the environment variables.
+To copy the commands from this procedure to the testbed administrator terminal,
+we strongly recommend to read this document with Adobe Acrobat Reader in Linux.
+Otherwise, copied characters might not render properly.
 
@@
 
-%A final thing to notice is the way \$ and \# are used in the code blocks. \$
-%means that the command can be run with normal user permissions where \# indicates that the
-%command needs root permissions.
-%Running a command with root permissions can on most systems be done by putting
-%the command "sudo" in front of each line. This is illustrated in the code below
-%with the command "whoami" that prints the username of the caller. Lines without
-%leading \$ or \# is the terminal output.
+In the code block, the \texttt{URL} and \texttt{IMAGE} specify where
+the Linux bundle is located and \texttt{WORKDIR} specifies a working
+directory where the Raspbian bundle will be downloaded and customized.
+Notice that even though we use the \$ and \# signs in the shell, in general
+these signs will be particular to the testbed administrator \ac{OS} shell.
+Next, create the working directory and go to it with the \texttt{cd} command.
+To download the image, utilize \texttt{wget} before
+unpacking the \texttt{zip} file as follows:
+
+% Running a command as root can on most systems be done by putting
+% \texttt{sudo} in front of the command. This is illustrated in the
+% following code block 
+% with the command \texttt{whoami} that prints the username. Lines within a code
+% block without leading \$ or \# is terminal output or content of a file.
 
 % Root example
-\begin{lstlisting}[]
-$ whoami
-<USERNAME>
-$ sudo whoami
-root
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Next, make the working directory and used \texttt{cd} to change into the working
-directory. Then \texttt{wget} can be used to download Raspbian lite before
-unpacking the \texttt{zip} file:
+% \begin{lstlisting}[]
+% $ whoami
+% <USERNAME>
+% $ sudo whoami
+% root
+% \end{lstlisting}
+% \FloatBarrier
+% \vspace{-5mm}
 
@@
 
-\subsection{The image}
+\subsection{Image Customization}
 

### 2016-07-01 image_setup.tex
Image resizing edits

@@
 \label{sec:image_setup}
-In this Section, we review the steps to create a common \ac{OS} image for all the
-\ac{Raspi}s. The image setup is composed of three major steps: Select and download the
-\ac{OS} image file, alter the image structure and configure the \ac{OS} files.
-We proceed to detail all these steps providing brief discussions to our setup
-choices when required. To perform these steps, we indicate with command-line
-blocks the required sequential commands to be typed by the testbed
-administrator in its \ac{PC} to obtain the desired setting. In all the
-following sections, the number sign (\#) and dollar sign (\$) will be used in
-the command blocks in the paper to indicate if an operation needs
-to be run with root permissions or common user permissions, respectively. The
-signs will be posted as the terminal prompt in the command blocks.
+In this Section, we review the steps to create a common \ac{OS} image for
+all the \ac{Raspi}s. The image setup is composed of three major steps:
+Select and download the \ac{OS} image file, alter the image structure and
+configure the \ac{OS} files. We proceed to detail all these steps providing
+brief discussions to our setup choices when required. To perform these
+steps, we indicate with command-line blocks the required sequential
+commands to be typed by the testbed administrator in its \ac{PC} to obtain
+the desired setting. In all the following sections, the number sign
+(\#) and dollar sign (\$) will be used in the command blocks in the paper
+to indicate if an operation needs to be run with root permissions or
+common user permissions, respectively. The signs will be posted as the
+terminal prompt in the command blocks.
 
@@
 To get started, we first need to install an \ac{OS} that works properly
-on all the \ac{Raspi} models. We will download and setup the image in the testbed
-administrator \ac{PC} using a Debian-based distribution. We use the popular
-Debian-based Raspbian Linux~\cite{raspbian} given that is the recommended
-and default \ac{OS} for the \ac{Raspi}.
-Raspbian is made available in two bundles: Raspbian and Raspbian lite.
-The difference between the two, is that Raspbian contains a pre-installed desktop
-environment for user interaction and Raspbian lite by default only permits to
-interact through a command shell. Given that the \ac{Raspi}s in our testbed are
-not connected to monitors, we decide to work with Raspbian lite. Still,
-if later desired, a desktop environment can be installed using the package manager
+on all the \ac{Raspi} models. We will download and setup the image in
+the testbed administrator \ac{PC} using a Debian-based distribution. We
+use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
+the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
+available in two bundles: Raspbian and Raspbian lite. The difference
+between the two, is that Raspbian contains a pre-installed desktop environment
+for user interaction and Raspbian lite by default only permits to interact
+through a command shell. Given that the \ac{Raspi}s in our testbed are not
+connected to monitors, we decide to work with Raspbian lite. Still, if later
+desired, a desktop environment can be installed using the package manager
 as it will be shown in the configuration section.
@@
 made available at~\cite{tunescode_webpage}. All other files used in this paper
-are available there unless specified otherwise. To get started, open a
-Linux shell (terminal) and declare the environment variables.
-To copy the commands from this procedure to the testbed administrator terminal,
+are available there unless specified otherwise. To get started, the testbed
+administrator must open a Linux shell (terminal) in and declare the environment
+variables shown in the command block below. We show the procedure by
+performing the role of the testbed administrator. To copy the commands from
+this procedure to the testbed administrator terminal,
 we strongly recommend to read this document with Adobe Acrobat Reader in Linux.
@@
 
-In the code block, the \texttt{URL} and \texttt{IMAGE} specify where
+In this code block, the \texttt{URL} and \texttt{IMAGE} variables specify where
 the Linux bundle is located and \texttt{WORKDIR} specifies a working
@@
 these signs will be particular to the testbed administrator \ac{OS} shell.
-Next, create the working directory and go to it with the \texttt{cd} command.
-To download the image, utilize \texttt{wget} before
+Next, we create the working directory and go to it with the \texttt{cd}
+command. To download the image, we utilize the \texttt{wget} command before
 unpacking the \texttt{zip} file as follows:
@@
 
-After Raspbian lite has been unpacked, there should be an \texttt{img} file in the directory.
-\texttt{fdisk} can be used to display the content of the image. Parse the arguments
-"-u sectors" to display the sizes in sectors and "-l" to display the partitions within the
-image and and automatically exit \texttt{fdisk} again:
+After Raspbian lite has been unpacked, there should be an \texttt{.img}
+file in the \texttt{WORKDIR} directory. Here, \texttt{fdisk} can be used to
+display the content of the image. To do this and obtain administrative
+information, we parse the arguments \texttt{-u sectors} to display the
+sizes in sectors and \texttt{-l} to display the partitions within the
+image. The \texttt{fdisk} command should output to the terminal something
+similar to:
 
-% Check out the image   % The dollar hack was to fix vim syntax
+% Check out the image
+% The dollar hack was to fix vim syntax
 \begin{lstlisting}[literate={DOLLAR}{\$}1]
@@
 
-The output provides knowledge about the image. The image is in total
-2709504 sectors (1.3GiB) in size and contains two
-partitions. The first partition starts at sector 8192 and
-the other partition starts at sector 137216. The partitions is of type FAT32
-and Linux and are 63M and 1.2G big. This indicates that the first partition
-is most likely a boot partition and the second partition is a traditional Linux file
-system. Actually, it is the Linux root filesystem (i.e. /).
+The output provides relevant information about the image. The image is in
+total 2709504 sectors (1.3GiB) in size and contains two partitions. The
+first partition starts at sector 8192 and the other partition starts at
+sector 137216. The first partition type is FAT32 with a size of 63 MB
+and the second partition is a Linux one with a size of 1.2 GB. This
+indicates that the first partition is a boot partition and the second
+one is a traditional Linux root filesystem, i.e. \texttt{/}.
 
-\subsection{Resize image}
-1.2G is not very much to store the root filesystem given that it is desired to
-alter the files and install additional packages. It may therefore be
-necessary to increase the partition size.
-The following illustrates how the image and its root filesystem
-can be expanded by one \ac{GB}.
+\subsection{Image Resizing}
+Given that we want to customize the files stored in the \ac{Raspi}s,
+we need to resize the image file since 1.2GB might not be enough to store
+the root filesystem due to the total size of the additional packages. Thus,
+we need to increase the partition size. The following procedure illustrates
+how the image and its root filesystem can be expanded by one \ac{GB}.
+First, to expand the image one \ac{GB}, we execute:
 
-First, expand the image one \ac{GB} and wait for the data to be completely written
-to the disk:
 \begin{lstlisting}[]
@@
 
-Use \texttt{fdisk} to see that the image is now one \ac{GB} larger:
+Later, we use \texttt{fdisk} with the same arguments as before to see that
+the image is now one \ac{GB} larger:
 \begin{lstlisting}[literate={DOLLAR}{\$}1]
@@
 
-Now, to expand the root filesystem it is required to first remove the Linux partition
-and then add it again one \ac{GB} bigger.
-Use \texttt{fdisk} to alter the partition table. The \texttt{echo}'s within the parenthesis
-will be interpreted as
-key-presses in \texttt{fdisk} and hence delete the last partition, create a new
-primary partition, print new partition table, and write new partition table
-to the image:
+Now, in this command block output, we observe that the change has taken
+effect by noticing the total available image size is 2.3GiB. To expand the
+root filesystem, it is required to first remove the Linux partition
+and then add it again with one \ac{GB} more. To do this, we make use
+again of \texttt{fdisk} in command mode to alter the partition table.
+For this, we pass the arguments of the command mode of \texttt{fdisk}
+through the \texttt{echo} command in Linux and the \texttt{|} operator
+(pipe) as follows:
 
@@
 
-Use a loopback device to make the Raspbian image available as a block device: 
+The \texttt{echo} commands within the parenthesis are interpreted as
+key-presses in the \texttt{fdisk} commmand mode. They (i) delete the last
+partition, (ii) create a new primary partition, (iii) print new partition
+table and (iv) write new partition table to the image file. Use a loopback
+device to make the Raspbian image available as a block device:
+
 \begin{lstlisting}[]

### 2016-07-05 image_setup.tex
Finishing image local mount edits

@@
 \begin{lstlisting}[]
-$ (echo d; echo ; echo n; echo p; echo ; echo 137216; echo ; echo p; echo w) | fdisk ${IMAGE}.img
+$ (echo d; echo 2; echo n; echo ; echo ; echo 137216; echo ; echo w) | fdisk ${IMAGE}.img
 \end{lstlisting}
@@
 The \texttt{echo} commands within the parenthesis are interpreted as
-key-presses in the \texttt{fdisk} commmand mode. They (i) delete the last
-partition, (ii) create a new primary partition, (iii) print new partition
-table and (iv) write new partition table to the image file. Use a loopback
-device to make the Raspbian image available as a block device:
+key-presses in the \texttt{fdisk} commmand mode. They (i) delete partition
+number 2, (ii) create a new primary type partition by default, (iii) set the
+new partition starting point (137216) and (iv) write new partition table to
+the image file. The previously shown command options of \texttt{fdisk}
+should work for both old and new versions of the commands. However, if this
+is not the case, the testbed administrator can do these commands manually.
+In this case, it is only required to delete the old root
+filesystem partition and create the new one starting from the same sector
+number. In our case, if the partitions commands were correct, the output
+should be following:
+
+\begin{lstlisting}[]
+
+Command (m for help): Partition number (1-4):
+Command (m for help): Partition type:
+   p   primary (1 primary, 0 extended, 3 free)
+   e   extended
+Select (default p): Using default response p
+Partition number (1-4, default 2): Using default value 2
+First sector (2048-4806655, default 2048): Last sector, +sectors or
++size{K,M,G} (137216-4806655, default 4806655): Using default value 4806655
+
+Command (m for help): The partition table has been altered!
+
+Syncing disks.
+
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\subsection{Loopback Device Setup}
+After successfully resizing the image file, we use a loopback device to make
+the Raspbian image available as a block device in the filesystem. For this
+command to work, the tested administrator distribution must have the
+\texttt{utils-linux} package with version 2.21 or higher. Otherwise, the
+\texttt{-P} argument of \texttt{losetup} will appear as invalid.
 
@@
 
-\texttt{lsblk} may be used to list block devices. It appears that the image is
-available as \texttt{/dev/loop0} which has two partitions \texttt{loop0p1} and
-\texttt{loop0p2}:
+If the previous command was succesful, the \texttt{lsblk} command can be used
+to list the available block devices in the filesystem as follows:
 %Use \texttt{lsblk} to view the partitions:
@@
 
-Check and resize the root partition:
+The image block device appears as \texttt{/dev/loop0}. This block device has
+two partitions associated to it, e.g. \texttt{loop0p1} and \texttt{loop0p2}.
+Finally, we check the block device status with \texttt{e2fsck} and
+\texttt{resize2fs}, respectively in the command block:
+
 \begin{lstlisting}[]
 # e2fsck -f ${DEV}p2
+e2fsck 1.42.8 (20-Jun-2013)
+Pass 1: Checking inodes, blocks, and sizes
+Pass 2: Checking directory structure
+Pass 3: Checking directory connectivity
+Pass 4: Checking reference counts
+Pass 5: Checking group summary information
+/dev/loop0p2: 35392/80480 files (0.1% non-contiguous), 201968/321536 blocks
 # resize2fs ${DEV}p2
+resize2fs 1.42.8 (20-Jun-2013)
+Resizing the filesystem on /dev/loop0 to 583680 (4k) blocks.
+The filesystem on /dev/loop0 is now 583680 blocks long.
 \end{lstlisting}
@@
 
-%We see that the boot and root partitions starts at sector 8192 and 137216 respectivly.
-%We also notice that each sector is 512 bytes.
-%On newer systems, we can mount the device easier using "losetup --show -f -P IMAGE"
-%Mount root partition:
-
-\subsection{Mount image}
-
-Browsing and altering the files in the image is easy. Simply mount the partitions.
-%To browser and alter the files in the image, it is possible to mount the partitions.
-
-Lets first mount the root partition. This is done by creating an empty directory
-that we will be used as a mountpoint. We name it \texttt{root} and create it in the
-working directory before mounting the root filesystem onto the mountpoint.
-%After this, mount the partitions:
-%After this,
-%we mount the partition starting at sector 137216. To provide this offset information
-%to mount, we need to find the offset in bytes. From the fdisk command we could also
-%see that each sector is 512 bytes. Thus, we multiply the sector size and the sector
-%start to get the offset in bytes.
+\subsection{Block Devices Mounting}
+For browsing and altering the files in the image, we mount the block
+device partitions into a particular path of our \texttt{WORKDIR} in order
+to customize them. We proceed first to mount the block device partition that
+contains the root filesystem and later the boot partition. This is done by
+creating an empty directory that it is used as a mountpoint. We name it
+\texttt{root} and create it in the working directory before mounting the
+root filesystem onto the mountpoint. Thus, we mount the root filesystem
+as follows:
 
@@
 
-The root filesystem already has a boot directory that can be used as
-the mountpoint for the boot partition.
-%After this, we can also mount the boot partition inside the newly mounted
-%root filesystem.
-This is convenient because that partition will be mounted
-on this directory when a \ac{Raspi} starts up with a memory card containing
-the image. %It is therefore the natural place to mount it for later purposes.
-Now, mount boot partition:
+The root filesystem mounted in \texttt{ROOTDIR}, already has a boot
+directory that can be used as the mount point for the boot partition
+in the block device \texttt{loop0p1}. This is convenient because
+the final edited partition from \texttt{ROOTDIR/boot} will be mounted on
+this same directory when a \ac{Raspi} starts up with a memory card
+containing the raw final image. Hence, to mount boot partition we do:
+%It is therefore the natural place to mount it for later purposes.
+
 \begin{lstlisting}[]
@@
 %# mount -o loop,offset=$((8192*512)) ${IMAGE}.img ${ROOTDIR}/boot
+%We can now change all the files in the disk image as desired.
 
-%We can now change all the files in the disk image as desired.
-It is now possible to change all files within the Raspbian image as desired.
-We will take advantage of this to edit configuration files, append new files,
-and even update and install packages.
+In this way, it is now possible to change all files within the Raspbian
+image as desired by editing the files in \texttt{ROOTDIR}. We take
+advantage of this to edit configuration files, append new files and even
+update and install packages.
+
+\subsection{Configuring OS files and scripts}
+
+The \ac{Raspi}s should be setup as similar as possible. One of those
+few settings that may be different in the devices is their hostname.
+%The hostname helps the user to distinguish the devices from each other.
+%The hostname helps distinguish them from each other.
+One approach to create unique hostnames would be to write the hostname
+in each memory card. Another way is to create a start up script to automatically
+configure the hostname. We decided to define the hostnames based on the \ac{MAC}
+addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
+addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
+administrator will need to insert the \ac{MAC} addresses and hostnames of
+his/her \ac{Raspi}s.
+
+%The \ac{Raspi}s should be setup to be as similar as possible. One of those
+%things that is preferred to be different on the devices is their hostname.
+%This helps distinguish them from each other. One solution to this in our
+%approach would be to uniquely change the hostname in each memory card.
+%This is however not desired because that would either require modifying
+%the customized Raspbian lite image before writing the content to the
+%memory card or to manually change it in the \ac{Raspi}s after the image has
+%been written to a memory card.
+%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
+%in our testbed. Those had to be used to put them in a domain name anyway.
+%
+%Below is a part of the file where we keep track of the MAC addresses and assign
+%hostnames. This file will be available online and copied to the Raspbian lite
+%image in a later step under the listed filename. For different \ac{Raspi}s
+%the MAC addresses has to be changed. The MAC address of a network card can be found
+%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
+
+%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
+%in our testbed. Those had to be used to put them in a domain name anyway.
+
+%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
+
+%Below is a part of the file where we keep track of the MAC addresses and assign
+%hostnames.
+%This file will be available online and copied to the Raspbian lite
+%image in a later step under the listed filename. For different \ac{Raspi}s
+%the MAC addresses has to be changed. The MAC address of a network card can be found
+%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
+
+% MAC and Hostname file
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+# Ethernet MAC    Hostname
+b8:27:eb:5b:da:20 rasp00
+b8:27:eb:7b:c3:91 rasp01
+b8:27:eb:54:9c:64 rasp02
+b8:27:eb:95:bd:11 rasp03
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
+%the testbed, it still needs to know which name to give itself.
+%We do this using a small Bash script:
+%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
+%hostname in \texttt{nodes.csv}.
+A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:
+
+% Set hostname
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+#!/usr/bin/env bash
+
+script_path="$(dirname $(realpath $0))"
+config_file=${script_path}/nodes.csv
+mac=$(cat /sys/class/net/eth0/address)
+old_hostname=$(hostname)
+new_hostname=$(grep $mac $config_file | cut -f2 -d' ')
+
+# Assign hostname found in nodes.csv
+if [ ! -z ${new_hostname} ]; then
+    echo ${new_hostname} > /etc/hostname
+    hostname ${new_hostname}
+    sed -i.old -e "s:${old_hostname}:${new_hostname}:g" /etc/hosts
+fi
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+The code lines above do the following:
+\begin{itemize}
+    \item Line 1: Tells system to interpret the script using \ac{Bash}
+    \item Line 3-4: Gets the path to the script itself and the list of hostnames
+    \item Line 5: Gets the \ac{MAC} address of the node itself
+    \item Line 6: Gets the current hostname
+    \item Line 7: Gets the new hostname from the hostname list
+    \item Line 10-14: Assigns the new hostname
+\end{itemize}
+
+%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
+%path of the script itself. Line 4 gets the MAC address. This script will be
+%located in all \ac{Raspi}s and thus return the MAC address of the node itself
+%when executed.
+%Line 4 searches line by line for this mac address.
+%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
+%then it will keep the default hostname of the system.
+
+It is likely that the testbed administrator needs to either change or add
+scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
+all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
+configuration scripts during start up before running the configuration scripts.
+
+The below script attempts to download/update the scripts:
+
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+#!/usr/bin/env bash
+
+url="http://kom.aau.dk/project/TuneSCode/raspi/"
+config_file="rasp_config.zip"
+
+# Attempt to fetch new configuration files
+if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
+    echo "Warning: Unable to update rasp_config files"
+    exit 1
+fi
+
+# Unzip and overwrite configurationn files to root's home directory
+unzip -q -o /tmp/${config_file} -d /home/pi/
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\begin{itemize}
+    \item Line 3-4: Is the url and zip file that should be downloaded
+    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
+        and prints a warning in case of errors
+    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
+\end{itemize}
+
+We have packed the files and uploaded them to a webserver. These can be
+retrieved with \texttt{wget} and unzipped into the Raspbian lite image:
+
+% Get the files
+\begin{lstlisting}[]
+$ wget ${URL%/*}/rasp_config.zip
+$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
+\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
+\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
+the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
+The testbed administrator will therefore need to also edit the \ac{URL} in the
+script \texttt{update\_rasp\_config}.
+%"nodes.csv" can be changed using a text editor. For vim, this could be done
+%the following way:
+%\begin{lstlisting}[]
+%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
+%\end{lstlisting}
+%\FloatBarrier
+%\vspace{-5mm}
+
+Finally, to actually make the \ac{Raspi}s change hostname, we have to make
+each \ac{Raspi} call the above script when it starts up. Optionally,
+also run the update script before running other scripts.
+Do this with root permissions by inserting the line(s) below before
+\texttt{"exit 0"} using your favorite editor:
+%"\$\{ROOTDIR\}/etc/rc.local" using
+%your favorite text editor..
+%The file
+%can be opened with your favorite text editor.
+
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
+<@\textcolor{gray}{
+-----------------------------------------------------}
+\Reactivatenumber @>
+...
+bash /home/pi/rasp_config/update_rasp_config
+bash /home/pi/rasp_config/set_hostname
+...
+exit 0
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Configuring \texttt{rc.local} makes it really cumbersome for the testbed
+administrator to add new scripts since it first needs to be included
+in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional
+line added to \texttt{rc.local} in the root image or in all memory cards depending on
+the final testbed setup.
+
+A solution to this is to make yet another script in \texttt{rasp\_config.zip}
+that calls all the other scripts in the desired order. This would look like
+below:
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
+<@\textcolor{gray}{
+-----------------------------------------------------}
+\Reactivatenumber @>
+#!/usr/bin/env bash
+
+bash /home/pi/rasp_config/set_hostname
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Now, to use the above script, \texttt{rc.local} needs to look like this:
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
+<@\textcolor{gray}{
+-----------------------------------------------------}
+\Reactivatenumber @>
+...
+bash /home/pi/rasp_config/update_rasp_config
+bash /home/pi/rasp_config/main
+...
+exit 0
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+Notice that \texttt{set\_hostname} is now called by the \texttt{main}
+script instead. The update script is still called directly. This ensures that
+all configuration scripts are updated before executed. Changes to the update
+script itself will first take effect at next system startup.
+%Had the update script
+%been called in \texttt{main} instead would mean that changes to the
+%\texttt{main} script would also first have been available at next startup.
+
+%This can be done by inserting a call to the script
+%set\_hostname in rc.local before it exits.
+%
+%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
+%\begin{lstlisting}[]
+%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
+%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
+%\end{lstlisting}
+%\FloatBarrier
+%$ line="bash /home/pi/.rasp_config/set_hostname"
+%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
+%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>
+
+%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
+%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
+
+%All other files could be updated as well.
+It is not only possible to add and configure files to/within the image. One can
+also install and update the packages within the image which will be illustrated
+in the next section.
+
+\subsection{Installing and updating the image}
+
+It may be desired to pre-install some programs in the image before it is
+written to all the memory cards that goes into the \ac{Raspi}s.
+This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
+(change root).
+%\url{https://wiki.debian.org/QemuUserEmulation}
+%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
+
+Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
+install some additional software:
+%Because \ac{Raspi}s are equipped with an \ac{ARM} processor
+%%it is not as
+%straightforward as performing the task for two Linux systems of the same
+%architecture.
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+# apt-get install binfmt-support qemu qemu-user-static
+# update-binfmts --display qemu-arm
+qemu-arm (enabled):
+     package = qemu-user-static
+        type = magic
+      offset = 0
+       magic = \x7fELF\x01\x01\x01\x00\x00\x00\x00\x00\x00\x00\x00\x00\x02\x00\x28\x00
+        mask = \xff\xff\xff\xff\xff\xff\xff\x00\xff\xff\xff\xff\xff\xff\xff\xff\xfe\xff\xff\xff
+ interpreter = /usr/bin/qemu-arm-static
+    detector =
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+% This was neccessary on arch
+%# update-binfmts --importdir /var/lib/binfmts/ --import
+
+Make sure the second command writes "enabled" as the output above. If that
+is not the case, then try enabling it:
+
+\begin{lstlisting}[]
+# update-binfmts --enable qemu-arm
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Provided that qemu-arm is enabled, we should now be able to change root (chroot)
+into our Raspbian lite image. Change root is a method in Linux that enables
+the root (/) to be changed. Thus, enabling a Linux installation within a
+Linux installation.
+
+There are a few commands to be performed before actually changing root into the
+root partion of the image.
+First, to get internet access from within the Raspbian lite image it is needed
+to copy our \texttt{resolv.conf} into the filesystem. Line 1 change directory
+into root filesystem of the image. The second line copies \texttt{resolv.conf}
+from the testbed administrators \ac{PC} to the image.
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+$ cd $ROOTDIR
+# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Now, because of the \ac{ARM} architecture,
+\texttt{/usr/bin/qemu-arm-static} needs to be copied into the image before continuing:
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+The final preparation before changing root is to populate the directories \texttt{proc},
+\texttt{sys}, and \texttt{dev}:
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+# mount  -t proc proc proc/
+# mount --bind /sys sys/
+# mount --bind /dev dev/
+# mount --bind /dev/pts dev/pts
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Finally, it is time to change root.
+%Finally, the filesystem is ready for us to change root.
+This can be done with \texttt{proot}. It may be required to install \texttt{proot}
+using "\texttt{apt-get install proot}":
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+# proot -q qemu-arm-static -S ${ROOTDIR}
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Most available material online uses the more known \texttt{chroot} command as
+written in the code block below instead of \texttt{proot}. This did not work
+correctly in our machines, but we present the command as an alternative to
+\texttt{proot} in case the testbed administrator should be in the opposite
+situration where \texttt{proot} is not working.
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+If everything went well, we should now be \texttt{chrooted} into the Raspbian lite
+filesystem under its root. Optionally, change the prompt title to indicate that it
+is a \texttt{chrooted} environment:
+
+% Optionally, we may create a unique prompt to indicate we have changed root
+\begin{lstlisting}[]
+# export PS1="(chroot) $PS1"
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+The Raspbian lite system image should now be possible to use almost as if it
+had been booted in a \ac{Raspi}.
+A major difference is that the testbed administrator's \ac{PC} is likely
+significantly faster than a \ac{Raspi}.
+%The only difference is that the laptop that we are using is
+%significantly faster than a \ac{Raspi}.
+Thus enabling updates, upgrades, and installing new software packages much
+faster than in the \ac{Raspi}s.
+%Thus enabling us to for example update,
+%upgrade, and install software packages:
+
+% Update system
+\begin{lstlisting}[]
+(chroot) # apt-get update
+(chroot) # apt-get upgrade
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+We also install a few packages that we find useful:
+%Lets install some useful applications:
+% Install packages
+\begin{lstlisting}[]
+(chroot) # apt-get install vim git screen
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+When writing the image to a memory card, all these change that has been made
+to the image so far will exists in all \ac{Raspi}s.
+%All the changes that are made here will exists in all \ac{Raspi}s when the
+%image is written to a memory card.

### 2016-07-05 image_setup.tex
Some edits

@@
 
-The root filesystem mounted in \texttt{ROOTDIR}, already has a boot
+The root filesystem mounted in \texttt{ROOTDIR}, has already a boot
 directory that can be used as the mount point for the boot partition
@@
 
-The \ac{Raspi}s should be setup as similar as possible. One of those
+In general, the \ac{Raspi}s should be setup as similar as possible. However,
+some particularities exist to differentiate the devices in principle
+One of those
 few settings that may be different in the devices is their hostname.
@@
 %Finally, the filesystem is ready for us to change root.
-This can be done with \texttt{proot}. It may be required to install \texttt{proot}
-using "\texttt{apt-get install proot}":
+This can be done with \texttt{proot}. It may be required to install
+\texttt{proot} using \texttt{apt-get install proot}:
 

### 2016-07-06 image_setup.tex
Edits on nodes configuration files section

@@
 
-\subsection{Configuring OS files and scripts}
+\subsection{Default OS Files and Configuration Scripts Setup}
 
 In general, the \ac{Raspi}s should be setup as similar as possible. However,
-some particularities exist to differentiate the devices in principle
-One of those
-few settings that may be different in the devices is their hostname.
-%The hostname helps the user to distinguish the devices from each other.
-%The hostname helps distinguish them from each other.
-One approach to create unique hostnames would be to write the hostname
-in each memory card. Another way is to create a start up script to automatically
-configure the hostname. We decided to define the hostnames based on the \ac{MAC}
-addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
-addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
-administrator will need to insert the \ac{MAC} addresses and hostnames of
-his/her \ac{Raspi}s.
+some particularities exist to differentiate the devices in principle. Also,
+scripts containing further configurations for the \ac{Raspi}s are desirable
+to be distributed as part of the common image. Therefore, here we present
+the steps to setup basic properties of the \ac{Raspi}s and distributing
+configuration scripts to each of them through the image.
 
-%The \ac{Raspi}s should be setup to be as similar as possible. One of those
-%things that is preferred to be different on the devices is their hostname.
-%This helps distinguish them from each other. One solution to this in our
-%approach would be to uniquely change the hostname in each memory card.
-%This is however not desired because that would either require modifying
-%the customized Raspbian lite image before writing the content to the
-%memory card or to manually change it in the \ac{Raspi}s after the image has
-%been written to a memory card.
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-%
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames. This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
-
-%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
-%in our testbed. Those had to be used to put them in a domain name anyway.
-
-%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:
-
-%Below is a part of the file where we keep track of the MAC addresses and assign
-%hostnames.
-%This file will be available online and copied to the Raspbian lite
-%image in a later step under the listed filename. For different \ac{Raspi}s
-%the MAC addresses has to be changed. The MAC address of a network card can be found
-%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.
+Among the settings that may be different in the devices is their hostname.
+The hostname helps the user to physically distinguish the devices from
+each other. Thus, we require the devices in our testbed to have different
+hostnames. Hence, we define the hostnames based on the \ac{MAC} addresses
+of the \ac{Raspi}s wired Ethernet interface. Prior to this stage, the
+\ac{MAC} address of a network card can be found using the command
+\texttt{ifconfig} or \texttt{ip addr} on a given \ac{Raspi}.
+We store the \ac{MAC} addresses and hostnames of the \ac{Raspi}s in the
+configuration file \texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv} as
+shown:
 
@@
 <@\textcolor{gray}{
----------------------------------------------------------------}
+--------------------------------------------------------------------------------}
 \Reactivatenumber @>
@@
 
-%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
-%the testbed, it still needs to know which name to give itself.
-%We do this using a small Bash script:
-%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
-%hostname in \texttt{nodes.csv}.
-A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:
+The testbed administrator has to insert the \ac{MAC} addresses and hostnames
+of his/her \ac{Raspi}s obtained previously in the format shown in the
+configuration file. Once we have defined the hostnames in
+\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}, we use the following
+\ac{Bash} script to have each \ac{Raspi} assign its hostname:
 
@@
 <@\textcolor{gray}{
----------------------------------------------------------------}
+-------------------------------------------------------------------------------------}
 \Reactivatenumber @>
@@
 
-The code lines above do the following:
-\begin{itemize}
-    \item Line 1: Tells system to interpret the script using \ac{Bash}
-    \item Line 3-4: Gets the path to the script itself and the list of hostnames
-    \item Line 5: Gets the \ac{MAC} address of the node itself
-    \item Line 6: Gets the current hostname
-    \item Line 7: Gets the new hostname from the hostname list
-    \item Line 10-14: Assigns the new hostname
-\end{itemize}
-
-%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
-%path of the script itself. Line 4 gets the MAC address. This script will be
-%located in all \ac{Raspi}s and thus return the MAC address of the node itself
-%when executed.
-%Line 4 searches line by line for this mac address.
-%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
-%then it will keep the default hostname of the system.
+The \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} script (in
+script lines): (1) tells the system to interpret the script using \ac{Bash},
+(3-4) gets the path to the script itself and the list of hostnames, (5) gets
+the \ac{MAC} address of the node itself, (6) gets the current hostname, (7)
+gets the new hostname from the hostname list and (10-14) assigns the new
+hostname to the \ac{Raspi} where the script will be executed.
 

### 2016-07-06 image_setup.tex
Most recent changes on image setup file
Image setup

@@
 command to work, the tested administrator distribution must have the
-\texttt{utils-linux} package with version 2.21 or higher. Otherwise, the
+\texttt{util-linux} package with version 2.21 or higher. Otherwise, the
 \texttt{-P} argument of \texttt{losetup} will appear as invalid.
@@
 the steps to setup basic properties of the \ac{Raspi}s and distributing
-configuration scripts to each of them through the image.
+configuration scripts to each of them through the image. First, we indicate
+how to obtain and put our configuration scripts in the image. Later, we
+describe the tasks performed by these configuration scripts. Nevertheless,
+any testbed administrator might modify or include other tasks besides the
+presented in this procedure according to his / her needs.
 
-Among the settings that may be different in the devices is their hostname.
-The hostname helps the user to physically distinguish the devices from
-each other. Thus, we require the devices in our testbed to have different
-hostnames. Hence, we define the hostnames based on the \ac{MAC} addresses
+In our case, we have our configuration scripts are stored in a \texttt{.zip}
+file located in a web server. We first retrieve this compressed file with
+\texttt{wget} and unzip it locally into our Raspbian lite image. These
+commands and the output of the last one are shown as follows:
+
+% Get the configuration files that are compressed in a remote server
+\begin{lstlisting}[]
+$ wget ${URL%/*}/rasp_config.zip
+$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
+Archive:  rasp_config.zip
+   creating: ${ROOTDIR}/home/pi/rasp_config/
+  inflating: ${ROOTDIR}/home/pi/rasp_config/nodes.csv
+  inflating: ${ROOTDIR}/home/pi/rasp_config/set_hostname
+  inflating: ${ROOTDIR}/home/pi/rasp_config/main
+  inflating: ${ROOTDIR}/home/pi/rasp_config/update_rasp_config
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Now,
+
+Among the settings that we want to be different in the devices is
+their hostname. The hostname helps the user to physically distinguish the
+devices from each other. Thus, in our case we require the devices in our
+testbed to have different hostnames. Moreover, besides this standard task,
+other testbed administrators may need to also include other scripts to
+configure their \ac{Raspi}s. Hence, we want to ensure that all the \ac{Raspi}
+configuration scripts of any testbed administrators are obtained in a simple
+way. Therefore, we ease this task by making the \ac{Raspi}s fetch the
+configuration scripts during start up before running them.
+
+Hence, we define the hostnames based on the \ac{MAC} addresses
 of the \ac{Raspi}s wired Ethernet interface. Prior to this stage, the
@@
 \end{itemize}
-
-We have packed the files and uploaded them to a webserver. These can be
-retrieved with \texttt{wget} and unzipped into the Raspbian lite image:
-
-% Get the files
-\begin{lstlisting}[]
-$ wget ${URL%/*}/rasp_config.zip
-$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
 

### 2016-07-07 image_setup.tex
Testing image setup

@@
 
-In our case, we have our configuration scripts are stored in a \texttt{.zip}
+In our case, we have our configuration scripts stored in a \texttt{.zip}
 file located in a web server. We first retrieve this compressed file with
@@
 
-Now,
+The unzippped files are one configuration file and three configuration
+scripts put in the newly created \texttt{\${ROOTDIR}/home/pi/rasp\_config/}
+folder in the image. In what follows, we describe which features do we
+require all the \ac{Raspi}s to have and how are they achieved with
+these configuration scripts.
 
@@
 We store the \ac{MAC} addresses and hostnames of the \ac{Raspi}s in the
-configuration file \texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv} as
-shown:
+configuration file \texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}. A
+sample of our file is shown as follows:
 
@@
 configuration file. Once we have defined the hostnames in
-\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}, we use the following
+\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}, we use the
+\texttt{\${ROOTDIR}/home/pi/rasp\_config/set\_hostname}
 \ac{Bash} script to have each \ac{Raspi} assign its hostname:
@@
 
-It is likely that the testbed administrator needs to either change or add
-scripts to configure the \ac{Raspi}s. This would be tedious to distribute to
-all \ac{Raspi}s at a later stage. We ease this task by making the \ac{Raspi}s fetch the
-configuration scripts during start up before running the configuration scripts.
-
-The below script attempts to download/update the scripts:
+It is possible that the testbed administrator requires to add other
+scripts to configure the \ac{Raspi}s besides the ones introduced here.
+This would be tedious to distribute to all \ac{Raspi}s at a later stage.
+We ease this task by making the \ac{Raspi}s fetch all the configuration
+scripts during start up from a \texttt{.zip} file before running the
+configuration scripts. The below script attempts to download/update the
+scripts:
 
@@
 
-For the above scripts to work in the \ac{Raspi}s, it is required that the \ac{Raspi}
-\ac{MAC} address is found in \texttt{nodes.csv}. Thus,
+For the above scripts to work in the \ac{Raspi}s, it is required that the
+\ac{Raspi} \ac{MAC} address is found in \texttt{nodes.csv}. Thus,
 \texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in

### 2016-07-07 image_setup.tex
Testing image setup. More new info

@@
 
-It is possible that the testbed administrator requires to add other
-scripts to configure the \ac{Raspi}s besides the ones introduced here.
-This would be tedious to distribute to all \ac{Raspi}s at a later stage.
-We ease this task by making the \ac{Raspi}s fetch all the configuration
-scripts during start up from a \texttt{.zip} file before running the
-configuration scripts. The below script attempts to download/update the
+Besides the scripts introduced up to this point of our procedure, it is
+possible that the testbed administrator requires to add other scripts to
+configure its \ac{Raspi}s. We ease this task by making the \ac{Raspi}s
+fetch all the configuration scripts during start up from a \texttt{.zip}
+file before running the configuration scripts. The role of this script
+is the same as theThe below script attempts to download/update the
 scripts:

### 2016-07-07 image_setup.tex
Image OS Files and Configuration Setup Section done

--- 2016-07-07 image_setup.tex
@@
 steps, we indicate with command-line blocks the required sequential
-commands to be typed by the testbed administrator in its \ac{PC} to obtain
-the desired setting. In all the following sections, the number sign
-(\#) and dollar sign (\$) will be used in the command blocks in the paper
-to indicate if an operation needs to be run with root permissions or
+commands to be typed by the testbed administrator in his / her \ac{PC}
+to obtain the desired setting. In all the following sections, the number
+sign (\#) and dollar sign (\$) will be used in the command blocks in the
+paper to indicate if an operation needs to be run with root permissions or
 common user permissions, respectively. The signs will be posted as the
@@
 
-\subsection{Default OS Files and Configuration Scripts Setup}
-
+\subsection{Image OS Files and Configuration Scripts Setup}
 In general, the \ac{Raspi}s should be setup as similar as possible. However,
@@
 the steps to setup basic properties of the \ac{Raspi}s and distributing
-configuration scripts to each of them through the image. First, we indicate
-how to obtain and put our configuration scripts in the image. Later, we
-describe the tasks performed by these configuration scripts. Nevertheless,
-any testbed administrator might modify or include other tasks besides the
-presented in this procedure according to his / her needs.
+configuration scripts to each of them through the image. For this, we first
+indicate how to obtain and put our configuration scripts in the image. Later,
+we describe the tasks performed by these configuration scripts. Finally, we
+indicate how and in which order are the scripts executed to configure all the
+devices. Nevertheless, any testbed administrator might modify or include other
+tasks besides the presented in this procedure according to his / her needs
+as we will show.
 
-In our case, we have our configuration scripts stored in a \texttt{.zip}
-file located in a web server. We first retrieve this compressed file with
-\texttt{wget} and unzip it locally into our Raspbian lite image. These
-commands and the output of the last one are shown as follows:
+\subsubsection{Image Default Configuration Scripts Download}
+\label{sec:configuration_files_download}
+In our case, we have our default configuration scripts stored in a
+\texttt{rasp\_config.zip} file located in the same \ac{URL} where the image
+was retrieved from, i.e. the one in the environment variable \texttt{URL}.
+We first manually download this compressed file with \texttt{wget} and
+extract it locally into our Raspbian lite image. These commands and the
+output of the last one are shown as follows:
 
@@
 
+\subsubsection{Device Hostnames}
 Among the settings that we want to be different in the devices is
@@
 devices from each other. Thus, in our case we require the devices in our
-testbed to have different hostnames. Moreover, besides this standard task,
-other testbed administrators may need to also include other scripts to
-configure their \ac{Raspi}s. Hence, we want to ensure that all the \ac{Raspi}
-configuration scripts of any testbed administrators are obtained in a simple
-way. Therefore, we ease this task by making the \ac{Raspi}s fetch the
-configuration scripts during start up before running them.
+testbed to have different hostnames. Hence, we define the hostnames based
+on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.
 
-Hence, we define the hostnames based on the \ac{MAC} addresses
-of the \ac{Raspi}s wired Ethernet interface. Prior to this stage, the
-\ac{MAC} address of a network card can be found using the command
-\texttt{ifconfig} or \texttt{ip addr} on a given \ac{Raspi}.
-We store the \ac{MAC} addresses and hostnames of the \ac{Raspi}s in the
-configuration file \texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}. A
-sample of our file is shown as follows:
+Prior to this stage, the \ac{MAC} address of a network card can be found
+using the command \texttt{ifconfig} or \texttt{ip addr} on a given
+\ac{Raspi}. We store the \ac{MAC} addresses and hostnames of the
+\ac{Raspi}s in the configuration file
+\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}. A sample of our file
+is shown as follows:
 
@@
 of his/her \ac{Raspi}s obtained previously in the format shown in the
-configuration file. Once we have defined the hostnames in
-\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}, we use the
+configuration file. In the file format, on each line there is a \ac{MAC}
+address and the corresponding hostname for the given \ac{Raspi}. For
+simplicity in the example above, we just include three \ac{Raspi}s.
+This list file will be employed by the
 \texttt{\${ROOTDIR}/home/pi/rasp\_config/set\_hostname}
-\ac{Bash} script to have each \ac{Raspi} assign its hostname:
+\ac{Bash} script to assign the hostname of each \ac{Raspi}.
+The script content is the following:
 
@@
 
-The \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} script (in
-script lines): (1) tells the system to interpret the script using \ac{Bash},
-(3-4) gets the path to the script itself and the list of hostnames, (5) gets
-the \ac{MAC} address of the node itself, (6) gets the current hostname, (7)
-gets the new hostname from the hostname list and (10-14) assigns the new
-hostname to the \ac{Raspi} where the script will be executed.
+The script (in lines): (1) tells the system to interpret the script
+using \ac{Bash}, (3-4) gets the path to the script itself and the list of
+hostnames, (5) gets the \ac{MAC} address of the node itself, (6) gets the
+current hostname, (7) gets the new hostname from the hostname list and
+(10-14) assigns the new hostname to the \ac{Raspi} where the script
+will be executed.
 
-It is possible that the testbed administrator requires to add other
-scripts to configure the \ac{Raspi}s besides the ones introduced here.
-This would be tedious to distribute to all \ac{Raspi}s at a later stage.
-We ease this task by making the \ac{Raspi}s fetch all the configuration
-scripts during start up from a \texttt{.zip} file before running the
-configuration scripts. The below script attempts to download/update the
-scripts:
+\subsubsection{Updating Default Configuration Files and Scripts}
+Besides the single script with its configuration file introduced up
+to this point of our procedure, it is possible that the testbed
+administrator may require to add other scripts to configure
+his / her \ac{Raspi}s. Hence, we want
+to ensure that all the \ac{Raspi} configuration scripts of any testbed
+administrators are obtained in a simple way. We automate this task by
+including the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}
+script in our procedure. The purpose of this script is to make all the
+\ac{Raspi}s fetch all the configuration scripts located with the image
+during a testbed start up.
+
+In our case as the testbed administrator for presenting the procedure, we want
+to fetch all our configuration scripts in
+\texttt{\$\{URL\%/*\}/raspi\_config.zip}. Thus, the update script
+\textit{automatically} downloads all the required configuration files in
+\texttt{rasp\_config.zip} file from a remote location. This is the same that
+we manually did earlier to get our files but this will be made in an
+automated wat after booting up the system. This script contents are:
 
@@
 <@\textcolor{gray}{
----------------------------------------------------------------}
+--------------------------------------------------------------------------------------------------}
 \Reactivatenumber @>
@@
 
-\begin{itemize}
-    \item Line 3-4: Is the url and zip file that should be downloaded
-    \item Line 7-10: Downloads the new configuration files to \texttt{/tmp}
-        and prints a warning in case of errors
-    \item Line 13: Unzips the files to \texttt{/home/pi/} while ignoring existing files/directories
-\end{itemize}
+The update script: (3-4) checks for the \ac{URL} and \texttt{.zip} file that
+should be downloaded, (7-10) downloads the configuration files to \texttt{/tmp}
+folder in the corresponding \ac{Raspi}. It also prints a warning in case of
+errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
+directory, while ignoring existing files and / or directories there.
 
 For the above scripts to work in the \ac{Raspi}s, it is required that the
-\ac{Raspi} \ac{MAC} address is found in \texttt{nodes.csv}. Thus,
-\texttt{rasp\_config.csv} needs to be unzipped, altered, and zipped again in
-the \ac{HTTP} server from which the \ac{Raspi} retrieves the files from.
-The testbed administrator will therefore need to also edit the \ac{URL} in the
-script \texttt{update\_rasp\_config}.
-%"nodes.csv" can be changed using a text editor. For vim, this could be done
-%the following way:
-%\begin{lstlisting}[]
-%$ vim ${ROOTDIR}/home/pi/rasp_config/nodes.csv
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
+\ac{Raspi}s \ac{MAC} addresses are found in \texttt{nodes.csv}. Also, it
+should be noted that for other testbed administrators besides ourselves,
+the \ac{URL} for file fetching and the configuration
+scripts themselves can be modified to fit their requirements. Thus, if
+required for a testbed administrator, the \texttt{rasp\_config.zip}
+will need to be edited to include all the required configuration files
+and scripts. Also, it might be needed to edit the \ac{URL} in the script
+\texttt{update\_rasp\_config} to store and fetch from a different location.
+Nevertheless, both the \ac{URL} and configuration files presented here can be
+used as a starting boilerplate if desired.
 
-Finally, to actually make the \ac{Raspi}s change hostname, we have to make
-each \ac{Raspi} call the above script when it starts up. Optionally,
-also run the update script before running other scripts.
-Do this with root permissions by inserting the line(s) below before
-\texttt{"exit 0"} using your favorite editor:
-%"\$\{ROOTDIR\}/etc/rc.local" using
-%your favorite text editor..
-%The file
-%can be opened with your favorite text editor.
+\subsubsection{Configuration Scripts Execution Order}
+To actually make the \ac{Raspi}s change hostnames and any other considered
+configurations, we have to make each \ac{Raspi} call the above scripts when
+it starts up. After finishing the setup process, all the unzipped files
+presented in Section~\ref{sec:configuration_files_download} should be
+locally available at each \ac{Raspi} after getting the root filesystem.
+Therefore, we first need to run the update script before running
+any other configuration scripts. To do this after boot up, we include
+a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
+file before \texttt{exit 0} using a file editor. To do this, it is required
+to edit this file as root. The resulting file should look like the following:
 
@@
 <@\textcolor{gray}{
------------------------------------------------------}
+---------------------------------------------}
 \Reactivatenumber @>
@@
 bash /home/pi/rasp_config/update_rasp_config
-bash /home/pi/rasp_config/set_hostname
 ...
@@
 
-Configuring \texttt{rc.local} makes it really cumbersome for the testbed
-administrator to add new scripts since it first needs to be included
-in \texttt{rasp\_config.zip} and then each \ac{Raspi} needs to have an additional
-line added to \texttt{rc.local} in the root image or in all memory cards depending on
-the final testbed setup.
+If it is required to have more configuration scripts, adding them
+in the \texttt{rc.local} file difficults its maintenance by the testbed
+administrator since this needs to be both in the image and the
+downloaded \texttt{rasp\_config.zip}. To avoid this problem, we include
+the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/main} script which
+calls all other configuration scripts (besides \texttt{update\_rasp\_config})
+in sequential order. This script content is the following:
 
-A solution to this is to make yet another script in \texttt{rasp\_config.zip}
-that calls all the other scripts in the desired order. This would look like
-below:
 \Suppressnumber\begin{lstlisting}[]
@@
 <@\textcolor{gray}{
------------------------------------------------------}
+---------------------------------------------------------------------}
 \Reactivatenumber @>
@@
 bash /home/pi/rasp_config/set_hostname
+# Any other required configuration scripts...
 \end{lstlisting}
@@
 
-Now, to use the above script, \texttt{rc.local} needs to look like this:
+In this way, the automation process is simplify since we only need to modify
+the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script:
+
 \Suppressnumber\begin{lstlisting}[]
@@
 script itself will first take effect at next system startup.
-%Had the update script
-%been called in \texttt{main} instead would mean that changes to the
-%\texttt{main} script would also first have been available at next startup.
 
-%This can be done by inserting a call to the script
-%set\_hostname in rc.local before it exits.
-%
-%Call set\_hostname script at startup. We insert a line of code to call script in rc.local just after the initial comments (i.e. lines starting with \#).
-%\begin{lstlisting}[]
-%# line_number=$(egrep -n -m1 "(^[^#])|(^$)" ${ROOTDIR}/etc/rc.local | cut -d: -f1)
-%# sed -i "${line_number}c bash /home/pi/rasp_config/set_hostname" ${ROOTDIR}/etc/rc.local
-%\end{lstlisting}
-%\FloatBarrier
-%$ line="bash /home/pi/.rasp_config/set_hostname"
-%# awk -v text="$line" '!/^#/ && !p {print text; p=1} 1' ${ROOTDIR}/etc/rc.local > <@ \Suppressnumber @>
-%    ${ROOTDIR}/etc/rc.local <@ \Reactivatenumber @>
-
-%WE NEED TO ADD A FEW COMMANDS TELLING HOW TO ZIP, MODIFY, AND UNZIP
-%FOR THE READER TO BE ABLE TO UPLOAD A CONSTRUCT HIS/HER OWN BUNDLE.
-
-%All other files could be updated as well.
-It is not only possible to add and configure files to/within the image. One can
-also install and update the packages within the image which will be illustrated
-in the next section.
-
-\subsection{Installing and updating the image}
+\subsection{Image Package Updating}
+Besides adding and configuring files within the image, the testbed
+administrator can also install and update the software packages within
+the image which will be illustrated in this subsection.
 

### 2016-07-07 image_setup.tex
Image Setup Section done!

@@
 \vspace{-5mm}
+
 Notice that \texttt{set\_hostname} is now called by the \texttt{main}
@@
 
+%\url{https://wiki.debian.org/QemuUserEmulation}
+%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
 \subsection{Image Package Updating}
 Besides adding and configuring files within the image, the testbed
-administrator can also install and update the software packages within
-the image which will be illustrated in this subsection.
+administrator may want to install and update the software packages
+within the image before it is written to all the memory cards that goes
+into the \ac{Raspi}s. From any Linux x86 machine as the testbed administrator
+\ac{PC}, this can be done using \texttt{chroot} command in the
+QEMU~\cite{QemuUserEmulation} hypervisor for \ac{ARM} processors.
 
-It may be desired to pre-install some programs in the image before it is
-written to all the memory cards that goes into the \ac{Raspi}s.
-This can be done from any linux x86 machine using QEMU Chroot~\cite{QemuUserEmulation}
-(change root).
-%\url{https://wiki.debian.org/QemuUserEmulation}
-%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
-
-Due to the \ac{ARM} processor that \ac{Raspi}s are using, it is required to
-install some additional software:
-%Because \ac{Raspi}s are equipped with an \ac{ARM} processor
-%%it is not as
-%straightforward as performing the task for two Linux systems of the same
-%architecture.
+\texttt{chroot} is a method in Linux that modifies the apparent root
+filesystem location from \texttt{/} to any other path. Consequently, in
+our case, we can use the Raspbian Lite image root filesystem within the
+testbed administrator Linux distribution. Then, QEMU allows the execution
+of commands for the \ac{Raspi} image (\ac{ARM} instructions) through the ones
+from the testbed administrator \ac{PC} architecture. Due to the \ac{ARM}
+processor that the \ac{Raspi}s employ, t is required to install the QEMU related software first and verigy that QEMU is \ac{ARM} enabled. To do so,
+run the following commands:
 
@@
 
-Make sure the second command writes "enabled" as the output above. If that
-is not the case, then try enabling it:
+In the previous output, the testbed administrator must be sure that the second
+command writes \texttt{qemu-arm (enabled)} as indicated. If that is not the
+case, then it should be possible to enable it by running:
 
@@
 
-Provided that qemu-arm is enabled, we should now be able to change root (chroot)
-into our Raspbian lite image. Change root is a method in Linux that enables
-the root (/) to be changed. Thus, enabling a Linux installation within a
-Linux installation.
+Provided that qemu-arm is enabled, we should now be able to \texttt{chroot}
+into our Raspbian lite image. However, there are a few commands to be
+performed before actually changing root into the root partion of the image.
+First, to get internet access from within the Raspbian lite image it is needed
+to copy the testbed administrator Linux distribution \texttt{resolv.conf} file
+into the image root filesystem. To do this, it is required to run the
+following:
 
-There are a few commands to be performed before actually changing root into the
-root partion of the image.
-First, to get internet access from within the Raspbian lite image it is needed
-to copy our \texttt{resolv.conf} into the filesystem. Line 1 change directory
-into root filesystem of the image. The second line copies \texttt{resolv.conf}
-from the testbed administrators \ac{PC} to the image.
-
-% CHROOT to OS image
 \begin{lstlisting}[]
@@
 
-Now, because of the \ac{ARM} architecture,
-\texttt{/usr/bin/qemu-arm-static} needs to be copied into the image before continuing:
+Now, because of the \ac{ARM} architecture, the
+\texttt{/usr/bin/qemu-arm-static} command needs to be copied into the
+image before continuing by running:
 
@@
 
-The final preparation before changing root is to populate the directories \texttt{proc},
-\texttt{sys}, and \texttt{dev}:
+Before changing root it is necessary to populate the directories
+\texttt{proc}, \texttt{sys} and \texttt{dev} for the image to get control
+as the testbed administrator apparent root filesystem. This is made by the
+following commands:
 
@@
 
-Finally, it is time to change root.
-%Finally, the filesystem is ready for us to change root.
-This can be done with \texttt{proot}. It may be required to install
-\texttt{proot} using \texttt{apt-get install proot}:
+Finally, to change root the following needs to be run:
 
-% CHROOT to OS image
 \begin{lstlisting}[]
-# proot -q qemu-arm-static -S ${ROOTDIR}
+# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash
 \end{lstlisting}
@@
 
-Most available material online uses the more known \texttt{chroot} command as
-written in the code block below instead of \texttt{proot}. This did not work
-correctly in our machines, but we present the command as an alternative to
-\texttt{proot} in case the testbed administrator should be in the opposite
-situration where \texttt{proot} is not working.
+% In some cases, \texttt{chroot} may not work properly. tgus, in this case, an
+% alternative can be \texttt{proot}. It may be required to install it
+% before changing the root with \texttt{apt-get install proot}. Then, it
+% should possible for the testbed administrator to run the following:
 
-% CHROOT to OS image
-\begin{lstlisting}[]
-# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash    (ALTERNATIVE TO proot)
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If everything went well, we should now be \texttt{chrooted} into the Raspbian lite
-filesystem under its root. Optionally, change the prompt title to indicate that it
-is a \texttt{chrooted} environment:
+% % CHROOT to OS image
+% \begin{lstlisting}[]
+% # proot -q qemu-arm-static -S ${ROOTDIR}
+% \end{lstlisting}
+% \FloatBarrier
+% \vspace{-5mm}
+If successfully executed, our terminal should have changed the prompt
+indicating that we are the root user in the Raspbian lite root filesystem as
+the apparent root. To be aware of the mode that we are working now, we change
+the prompt title to indicate that it is a \texttt{chroot} environment as
+follows:
 
@@
 
-The Raspbian lite system image should now be possible to use almost as if it
-had been booted in a \ac{Raspi}.
-A major difference is that the testbed administrator's \ac{PC} is likely
-significantly faster than a \ac{Raspi}.
-%The only difference is that the laptop that we are using is
-%significantly faster than a \ac{Raspi}.
-Thus enabling updates, upgrades, and installing new software packages much
-faster than in the \ac{Raspi}s.
-%Thus enabling us to for example update,
-%upgrade, and install software packages:
+The Raspbian lite image should now be possible to use almost as if it had
+been booted in a \ac{Raspi}. A major difference is that the testbed
+administrator \ac{PC} is likely significantly faster than a \ac{Raspi}.
+Hence, enabling updates, upgrades and installing new software packages
+should be much faster than in a \ac{Raspi}. Still updating and upgrading
+the packages for the \ac{Raspi} might take some amount of time. To get
+the upgraded system packages, it is necessary to run:
 
@@
 
-We also install a few packages that we find useful:
+We further install some packages that we consider useful:
 %Lets install some useful applications:
@@
 
-When writing the image to a memory card, all these change that has been made
-to the image so far will exists in all \ac{Raspi}s.
-%All the changes that are made here will exists in all \ac{Raspi}s when the
-%image is written to a memory card.
+\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
+for managing Git repositories and \texttt{screen} for better handling of
+long-runnning processes. When writing the image either to a memory card
+or storing it remotely \todot{Are we going to store the image
+into a memory card or remotely in a URL? Since this is not completely clear
+yet.}, all these change that has been made to the image so far will exists in
+all \ac{Raspi}s.

### 2016-07-07 appendix.tex
appendix

\label{sec:testbed_http}

\section{Old version of losetup}

\begin{lstlisting}[]
# export DEV="${losetup -f}"
/dev/loop0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\begin{lstlisting}[]
# mknod -m 660 ${DEV}p1 b 7 1
# mknod -m 660 ${DEV}p2 b 7 2
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\begin{lstlisting}[]
# fdisk -lu ${IMAGE}.img
WHAT WILL BE THE OUTPUT?
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

I THINK SIZELMIT IS NOT REQUIRED
\begin{lstlisting}[]
# losetup -o $((8192*512)) --sizelimit $((137215*512)) ${DEV}p1 ${IMAGE}.img
# losetup -o $((137216*512)) --sizelimit $((4806655*512)) ${DEV}p2 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The output from \texttt{lsblk} will look a bit different

### 2016-07-07 image_setup.tex
Document structure changes and new appendix changes

@@
 Now, in this command block output, we observe that the change has taken
-effect by noticing the total available image size is 2.3GiB. To expand the
+effect by noticing the total available image size is 2.3 GiB. To expand the
 root filesystem, it is required to first remove the Linux partition
@@
 \texttt{util-linux} package with version 2.21 or higher. Otherwise, the
-\texttt{-P} argument of \texttt{losetup} will appear as invalid.
+\texttt{-P} argument of \texttt{losetup} will appear as invalid. If the
+version of \texttt{losetup} can not be updated for some reason, an
+alternative option for this part is presented in
+Section~\ref{sec:alternative_losetup} of the Appendix.
 

### 2016-07-07 overlay_fs.tex
Document structure changes and new appendix changes

@@
 \label{sec:ovelay_fs}
+In principle, our procedure modifies the image file only once in
+the testbed administrator \ac{PC} when its setup is made. Also, keeping
+this image in the \ac{Raspi}s provides the same initial system for all
+the devices. This system is later distributed locally to a memory card
+in each device.However, if we do not make any further modifications during
+the image setup, any files created after the initial boot of a \ac{Raspi}
+will remain in the me
+
+to have the same
+configuration files.
+
+Our goal was to enable the users of the testbed to change however the
+
+\ac{Raspi} are configured while still being able to revert those changes easily.
 
 \subsection{Installation an overlaying filesystem}
-
-Our goal was to enable the users of the testbed to change however the
-\ac{Raspi} are configured while still being able to revert those changes easily.
 

### 2016-07-07 overview.tex
Document structure changes and new appendix changes

@@
 \label{sec:overview}
-
-% We proceed to define the testbed environment by each of its conforming
-% parts. We later indicate a procedure to set up and configure the testbed
-% for several \ac{Raspi}s. In this procedure, we summarize all the key details
-% regarding on setting the system files, configuration packages, tools and connectivity
-% for the \ac{Raspi}s in a centralized fashion.
-
-%Afterwards, we indicate how to cross-compile C++ code and the Kodo
-%network coding library in an easy way.
-%Finally, we provide further information about Kodo itself in terms
-%of testing, other platforms supported and source code documentation
-%for further references.
-
-%\subsection{System Overview and Design Criteria}
 
@@
 All \ac{Raspi}s are each equipped with a memory card, a wired and wireless
-network interface and a power supply. All the \ac{Raspi} are connected to a common
-\ac{LAN} that provides internal and external connectivity. Without loss of generality,
-in our case they are connected to the university network using their wired Ethernet interface
-that is named \texttt{eth0} according to the legacy naming convention of Ethernet
-interfaces in Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
-university network since we aim the testbed to be used by students and professors
-to perform measurements and experimentation of controlled and reproducible scenarios
-as part of academic research. We intend to configure all \ac{Raspi}s identically,
-while still allowing the end-users to alter how each \ac{Raspi} behaves and is configured.
+network interface and a power supply. All the \ac{Raspi} are connected to
+a common \ac{LAN} that provides internal and external connectivity. Without
+loss of generality, in our case they are connected to a university network
+using their wired Ethernet interface that is named \texttt{eth0} according
+to the legacy naming convention of Ethernet interfaces in
+Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
+university network since our testbed is used by students and
+professors to perform measurements and experimentation of controlled and
+reproducible scenarios as part of academic research. Nevertheless, the
+testbed description and procedures for setting it up are not restricted
+to the academic scenario. All \ac{Raspi}s will be configured to run
+a \ac{SSH} daemon for easy remote access within the university network.
+We requested the university \ac{IT} department to configure the university
+\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP}. This
+eliminates the requirements of monitors and keyboards with the \ac{Raspi}s
+for non-graphical applications. Finally, our design aims to configure all
+\ac{Raspi}s identically from a customized bootable image in their
+respective memory cards, while still allowing the end-users to store
+files locally in each of the \ac{Raspi}s.
 
@@
 
-In Fig.~\ref{fig:testbed_setup} and what follows, we will refer
-as the testbed administrator to the person(s) in charge of setting up and configuring
-the testbed with administrator privileges from the \ac{OS} point of view.
-The setting and configuration procedure commands are performed the testbed administrator in a
-regular \ac{PC} running a Linux distribution. Although in principle the administrator
-Linux distribution is not a restriction, we present our procedure in a Debian-based Linux
-distribution. Finally, our design considers to fetch the \ac{Raspi}s
-root filesystem from a \ac{HTTP} server. This enables the testbed administrator to
-make filesystem in one location. The \ac{Raspi}s will then get the changes after a reboot.
-All \ac{Raspi}s will be configured to run a \ac{SSH} daemon for easy remote access
-within the university network. We requested the university \ac{IT} department to
-configure the university \ac{DHCP} server to assign each \ac{Raspi} a static
-\ac{IP}. This eliminates the requirements of monitors and keyboards with the
-\ac{Raspi}s for non-graphical applications.
-
-Based on a regular Raspbian Linux image, we explain how
-to customize it to include desired packages for performing routinary tasks and set it
-in all the \ac{Raspi}s. Also, when working on a testbed with many devices, it may not always be
-desired to make persistent changes to the devices. Thus, we later present how to utilize
-stacked filesystems to enable both persistent and temporary
-storage. This step of the procedure will be optional if the testbed
-administrator decides to write the whole image into a memory card. \todot{Check the
-veracity and correctness of the previous statement. Question: Shoudln't we use
-an overlay FS regardless of how the memory card is configured?}.
-Although being simpler than a stacked filesystem, writing the whole Linux
-filesystem to several memory cards may complicate future
-maintenance of the testbed since updates need to be performed in each \ac{Raspi}.
-Even when tools such as Python Fabric may ease this task, the maintenance might still
-be cumbersome in some cases. Therefore, we propose and encourage the alternative centralized
-approach where the initial \ac{RAM} filesystem is customized in a way that
-enables \ac{Raspi}s to fetch the its root filesystem from a \ac{HTTP} server as
-mentioned previously.
-
-%The following procedure explains how the testbed may be setup and configured.
-
-%\textcolor{red}{The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. We intend to configure
-%all \ac{Raspi}s identically, while still allowing students to alter
-%how each \ac{Raspi} behaves and is configured.
-%\\\\
-%When working on the testbed with this many devices, it may not always be
-%desired to make persistent changes to devices. We present how the concept
-%of stacked filesystems may be used to enable the means of both persistent
-%and temporary storage.}
-
-%The testbed should be available for students to perform simulations and
-%measurements as part of their semester projects. The \ac{Raspi}s should
-%be configured identically, but allows the students to alter the
-%configurations and how each \ac{Raspi} behaves.
-%
-%When working on a testbed with many devices, it may not always desired
-%to make persistent changes to devices. We present how the concept of stacked
-%filesystems may be used to enable the means of both persistent and
-%temporary storage.
-
-%A \ac{HTTP} server will be used to store Linux configuration files as well
-%as the root filesystem in the centralized approach.
-
-%The following procedure will explain how the testbed may be setup and
-%configured both in the traditional distributed fasion where the
-%entire Linux filesystem is written to an memory card, but also in a more
-%centralized fasion where 
-%is written to
-%Based on the popular Debian based Raspbian Linux distribution
-%We
-%first present how a standard Linux distribution for the \ac{Raspi}s
-%may be customized
+In Fig.~\ref{fig:testbed_setup} and what follows, we will refer as the
+testbed administrator to the person(s) in charge of setting up and
+configuring the testbed with administrator privileges from the \ac{OS}
+point of view. The setting and configuration procedure commands are
+performed by the testbed administrator in a regular \ac{PC} running a
+Linux distribution. Although in principle the administrator
+Linux distribution is not a restriction, we present our procedure in a
+Debian-based Linux distribution. Our basic design considers to create
+a customized image to later store it in a memory card. Thus, based on a
+regular Raspbian Linux image, we explain how to resize and customize it to
+include desired packages for performing routinary tasks in all the
+\ac{Raspi}s. As another design consideration, when working on a testbed
+with many devices, it may not always be desired to make persistent changes
+on them. One scenario is when different users are interested in
+running experiments on a rebooted testbed. Thus, we later present how
+to utilize stacked filesystems to enable both persistent and
+temporary storage to have this capability. Its purpose is to
+remove non-desired data after a reboot while keeping the original
+customized image structure. This step of the procedure is optional if
+the testbed administrator decides to keep only persistent changes regardless
+of the testbed use. Finally, an advanced step in our design considers to
+fetch the \ac{Raspi}s root filesystem from a \ac{HTTP} server. This enables the
+testbed administrator to put the data files in one location. The
+\ac{Raspi}s will then get these data files after a reboot. We consider
+that this step is particularly useful when there is a large amount
+of devices to be configured and the setup process needs to be automated.
+However, since a more complex scenario that might be required only in special
+cases, we put this step in Section~\ref{sec:http} in the Appendix.

### 2016-07-07 image_setup.tex
Starting with overlay fs

@@
 for managing Git repositories and \texttt{screen} for better handling of
-long-runnning processes. When writing the image either to a memory card
-or storing it remotely \todot{Are we going to store the image
-into a memory card or remotely in a URL? Since this is not completely clear
-yet.}, all these change that has been made to the image so far will exists in
-all \ac{Raspi}s.
+long-runnning processes. When writing the image to a memory card, all the
+changes that have been made to the image so far, will exist in all
+\ac{Raspi}s after fetching it.

### 2016-07-07 overlay_fs.tex
Starting with overlay fs

--- 2016-07-01 overlay_fs.tex
@@
 \label{sec:ovelay_fs}
+In principle, our procedure modifies the image file only once in
+the testbed administrator \ac{PC} when its setup is made. Also, keeping
+this image in the \ac{Raspi}s provides the same initial system for all
+the devices. This system is later distributed locally to a memory card
+in each device. However, if we do not make any further modifications during
+the image setup, any files created after the initial boot of a \ac{Raspi}
+will remain in the memory card. This is cumbersome to maintain since
+the size of the memory card is relatively reduced (8 GB) and there
+might be various users utilizing the testbed. Also, different testbed
+users could be interested in running their experiments in a fresh
+rebooted system with the original customized image. Nevertheless, we
+emphasize that this step is not necessary if the tesbed administrator
+wants to consider only persistent storage for its devices. A use case
+for this scenario could be a single user for the testbed or when a testbed
+administrator only wants to setup a few \ac{Raspi}s.
 
-\subsection{Installation an overlaying filesystem}
+Therefor, if both persistent and non-persistent storage are required for
+the \ac{Raspi}s, here we present the steps to setup an overlay filesystem.
+This type of filesystem enables an \textit{upper} filesystem to overlay
+into a \textit{lower} filesystem. Whenever a file is requested, the upper
+filesystem will forward the request to the lower filesystem in case it
+does not have it itself. If the upper filesystem has the requested file,
+it will simpy return the file. This idea can be used in our setup to mount
+the root filesystem (i.e. Raspbian lite) in the \ac{Raspi}s during startup
+as read-only filesystem. Thus, on the one hand, the image configuration files
+will remain after a reboot and will not be modifiable. On the other hand,
+to enable the possibility of non-persistent changes, we overlay the upper
+filesystem that is mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp}
+as read/writable on top of the lower root filesystem. Thus, reading a file
+may return a file from the lower filesytem, but if it is stored, it will
+be saved in the upper filesystem. Accessing this file again will return
+the stored file from the upper layer. After a reboot, all the stored files
+in the upper filesystem will be erasured.
 
-Our goal was to enable the users of the testbed to change however the
-\ac{Raspi} are configured while still being able to revert those changes easily.
+\subsection{Overlaying Filesystem Installation}
 
-We do this by storing a users work in the \ac{Raspi}s in \ac{RAM}. Thus a
-\ac{Raspi}, when a \ac{Raspi} is rebooted, it will forget all changes.
-
-A way to do this is using stacked filesystem. Stacking filesystem enables
-a filesystem to overlay another filesystem. Whenever a file is accessed,
-the upper filesystem will forward the request to the lower filesystem in
-case it does not have it itself. If it has it, it will simpy return the file.
-
-This idea can be used in our setup to mount the root filesystem (i.e. Raspbian
-lite) in the \ac{Raspi}s duing startup as read-only. Thus, the files will not
-be changable. Some programs needs to change files and the testbed admin might
-want to change configurations. To enable this, it is possible to
-overlay a filesystem that is mounted in \ac{RAM} as read/writable on top of
-the regular root. Thus, reading a file may return a file from the lower filesytem,
-but if it is stored, it will be saved in the upper filesystem. Accessing this file
-again will return the stored file from the upper layer.
-
-%We do this using an overlay filesystem. The idea is to let the \ac{Raspi}s
-%mount their root filesystem as 
-%
-%We do this using an overlay filesystem (overlayfs).
-%Overlayfs enables a filesystem to be stacked on top of another filesystem. Thus,
-%if the \ac{Raspi}s mount their root filesystem (/) as read-onlyif we mount our root (/) filesystem as 
-%
-%
-%Overlayfs enables filesystems
-%to be stacked on top of each other. We will be using this to make a lower
-%filesystem and an upper filesystem. The lower filesystem will be our
-%traditional filesystem and the upper filesystem will be mounted in tmpfs
-%(i.e. mounted in \ac{RAM}).
-%
-%Thus, when a file is requsted from stacked filesystem, it will first attempt
-%to fetch the file in the upper filesystem if it does not exists there, it will
-%request it from the lower filessytem.
-
-Assuming that you are still \texttt{chrooted} into the Raspbian lite filesystem, we
-can setup the overlay filesystem (overlayfs).
-
-There already exists implementations overlaying the root filesystem. We will
-use an existing implementation
-available at github: \cite{overlayroot}.
-
-We installed git in a previous step, so lets clone the repository. The below
-code stores it in /tmp which is really mounted in \ac{RAM}. Thus, the files
-will disappear when the machine is rebooted.
+Assuming that we are still using the \texttt{chroot} commands into the
+Raspbian lite root filesystem for installing packages, we can setup the
+overlay filesystem at this point of the procedure. There already exists
+implementations overlaying the root filesystem. We use an implementation
+available at the Git repository in\cite{overlayroot}. Since we have
+installed git in a previous step, we clone the repository. The
+command block below stores it in \texttt{/tmp} which is really mounted in
+\ac{RAM}. Then, all the files stored here will disappear a \ac{Raspi} is
+rebooted.
 

### 2016-07-07 overlay_fs.tex
Persistent directories

@@
 implementations overlaying the root filesystem. We use an implementation
-available at the Git repository in\cite{overlayroot}. Since we have
+available at the Git repository in \cite{overlayroot}. Since we have
 installed git in a previous step, we clone the repository. The
@@
 
-To enable the overlaying filesystem it is required to generate an initial \ac{RAM}
-file system (initramfs). Initramfs is an initial filesystem that gets loaded into
-\ac{RAM} during the startup process of a linux machine to prepare the real
-filesystem.
-
-For this purpose, we need BuxyBox:
+Before enabling the overlaying filesystem, it is required to generate
+an initial \ac{RAM} file system or \texttt{initramfs}. This is an
+initial filesystem that gets loaded into \ac{RAM} during the startup
+process of a Linux machine to prepare the real filesystem. For this purpose,
+we need the BusyBox package by running:
 
@@
 
-Add and activate the overlaying filesystem, we need to add the scripts and tell initramfs
-to load the overlaying filesytem:
+To create and activate the overlaying filesystem, we need to first
+add the required system scripts to do so. This is done as follows:
 
@@
 
-%Finally, lets generate the initial \ac{RAM} filesystem. This is done using the command
-%\texttt{mkinitramfs}, that will by default look for the kernel modules of the current system. Remember we are
-%chrooted, so we need to specify the correct kernel modules to look for. These
-%can be found in "/lib/modules".
-Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
-needs a different kernel than \ac{Raspi} version 2 and version 3. Thus, start
-by finding the available kernel modules:
-%Because \texttt{initramfs} is executed within a
-%\texttt{chrooted} environment, it will not be able to detect the
-%kernel modules in the \texttt{chrooted} environment. The kernel modules will
+To generate the initial \ac{RAM} filesystem, we have to utilize the
+\texttt{mkinitramfs} command. This searches by default for the available
+kernel modules in the system. Since we are in \texttt{chroot} mode, we need
+to specify the correct kernel modules to search for. The available
+kernel modules are \texttt{/lib/modules}. To see them, we just run:
 
-% Create an initramfs
+% See available kernel modules
 \begin{lstlisting}[]
@@
 \vspace{-5mm}
+%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
+%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+
 
-Kernel "4.4.13+" is for \ac{Raspi} version 1 and kernel "4.4.13-v7+" works
-for \ac{Raspi} version 2 and version 3.
-
-Lets proceed by generating an initial \ac{RAM} filesystem for both kernels:
-
-%Lets proceed with kernel modules "4.4.13+".
-%Lets create an initial ram filesystem for both kernels.
-
-%Thus, to run mkinitramfs, type in the correct kernel version:
+Now, the initial \ac{RAM} filesystem can be generated. \ac{Raspi} version 1
+needs a different kernel than \ac{Raspi} version 2 and version 3. Kernel
+version 4.4.13+ is for \ac{Raspi} version 1 and kernel 4.4.13-v7+ for
+\ac{Raspi} version 2 and version 3. We proceed to generate an initial
+\ac{RAM} filesystem for these kernels by running:
 
@@
 
-%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
-%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+
+Although these commands might be output some warnings, they should
+successfully generate working initial \ac{RAM} filesystems. Later, an
+initial \ac{RAM} filesystem will need to be called by the bootloader.
+In Raspbian this is done by adding a command to \texttt{config.txt} file
+in the boot partition. If the system should be run in a \ac{Raspi} version 1,
+then use \texttt{init.gz}, otherwise use \texttt{init-v7.gz} as follows
+\todot{What happens if I have both a Raspi 1, 2 or 3 in my testbed? Do I have
+to have different images after this point?}:
 
-Although there might be printet a few warnings, the above commands should successfully
-generate working initial \ac{RAM} filesystems.
-
-An initial \ac{RAM} filesystem needs to be called by the bootloader. In Raspbian
-this is done by adding a command to \texttt{config.txt} in the boot partition.
-
-If the system should be run in a \ac{Raspi} version 1, then use \texttt{init.gz}:
 %The bootloader should be made aware of the initramfs that we will generate:
@@
 (chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
+(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # For Raspberry Pi version 2 or 3
 \end{lstlisting}
@@
 
-If the system should be run in a \ac{Raspi} version 2 or 3, then use \texttt{init-v7.gz}:
-\begin{lstlisting}[]
-(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # Only for Raspberry Pi version 2 and 3
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
+After this point, it is no longer required to be in \texttt{chroot} mode.
+The following commands exits the \texttt{chroot} environment, unmount all
+partitions and detaches the loopback devices:
 
-At this point, it is no longer required to be \texttt{chrooted} and the image
-can be written to a memory card. The following commands exits the \texttt{chrooted}
-environment, unmount all partitions, and detaches the loopback device:
-
-% Exit chroot, umount, and write to memory card
+% Exit chroot, unmount and detach loopback devices
 \begin{lstlisting}[]
@@
 
-\subsection{(Optional) Persistent subdirectories for distributed Linux image}
+For the \texttt{--recursive} to work properly, it is necessary that
+the package \texttt{util-linux} version is greater than or equal to 2.22.
+Otherwise, an alternative is to either update the package or follow the
+procedure in Section~\ref{sec:umount} of the Appendix.
 
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
+\subsection{Persistent Image Directories}
+Provided the stacked filesystem is configured, now it is possible to have
+directories where files are not removed upon rebooting the \ac{Raspi}s.
+The following procedure creates an extra partition in the image for
+the \ac{Raspi} user home directory that will be made storage persistent
+using the upper filesystem. We first expand image according to the desired
+home directory size, but avoiding to make the image bigger than the target
+memory card size.
 
-Provided the stacked filesystem was configured, it may be desired to have
-some directories where files are not flushed upon rebooting the \ac{Raspi}s.
-The following procedure will create an extra partition in the image for
-the \ac{Raspi} users home directory that will be made storage persistent.
-%Provided that a stacked filesystem was configured to make the root
-%filesystem forget chances upon reboots is not always desired.
-
-Expand image according to the desired home directory size. Keep in mind not
-to make the image bigger than the target memory card.
 % Expand image

### 2016-07-08 appendix.tex
Document structure changes and new appendix changes
Persistent directories
New appendix and including more description about the persistent directory setup

@@
-\label{sec:testbed_http}
+\section{Advanced Setup: HTTP Image Fetching}
+\label{sec:http}
 
-\section{Old version of losetup}
+The HTTP description
+
+\section{Alternative Commands for Outdated Packages}
+This section describes alternative commands in case the testbed administrator
+is not able to update old packages in its Linux distribution for performing
+the commands, particularly the ones regarding the \texttt{util-linux} package.
+
+\subsection{Losetup for Loopback Devices}
+\label{sec:alternative_losetup}
+
+In case the \texttt{losetup -P} is not available given that a version
+higher or equal than 2.21 for \texttt{util-linux} is neither not available,
+a possible option is to manually set all the loopback devices used during the
+whole procedure with \texttt{mknod}. To do this, the alternative commands are:
 
 \begin{lstlisting}[]
-# export DEV="${losetup -f}"
-/dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\begin{lstlisting}[]
-# mknod -m 660 ${DEV}p1 b 7 1
-# mknod -m 660 ${DEV}p2 b 7 2
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\begin{lstlisting}[]
-# fdisk -lu ${IMAGE}.img
-WHAT WILL BE THE OUTPUT?
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-I THINK SIZELMIT IS NOT REQUIRED
-\begin{lstlisting}[]
+$ export DEV="/dev/loop0"
+# mknod ${DEV}p1 b 7 1
+# mknod ${DEV}p2 b 7 2
 # losetup -o $((8192*512)) --sizelimit $((137215*512)) ${DEV}p1 ${IMAGE}.img
@@
 
-The output from \texttt{lsblk} will look a bit different
+However, there will be a few differences by using the above code as an
+alternative to the \texttt{losetup -P} case. First, the output from
+\texttt{lsblk} will look different:
+
+\begin{lstlisting}[]
+# lsblsk
+  NAME   MAJ:MIN RM   SIZE RO TYPE MOUNTPOINT
+  sda      8:0    0 238.5G  0 disk
+    sda1   8:1    0   512M  0 part /boot
+    sda2   8:2    0   168G  0 part /
+    sda3   8:3    0     2G  0 part [SWAP]
+    sda4   8:4    0    68G  0 part
+  loop1    7:1    0    67M  0 loop
+  loop2    7:2    0   2.2G  0 loop
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Second, detaching the loop devices will also be different:
+
+\begin{lstlisting}[]
+# losetup -d ${DEV}p{1,2}
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Working with these loopback devices will be transparent for the remainding
+procedure.
+
+Also, when creating the persistent home directory partition in
+Section~\ref{sec:persistent_directories}, for recreating the loopback devices
+the commands are:
+
+\begin{lstlisting}[]
+$ export DEV="/dev/loop0"
+# mknod ${DEV}p2 b 7 2
+# mknod ${DEV}p3 b 7 3
+# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512)) ${DEV}p2 ${IMAGE}.img
+# losetup -o $((4806656*512)) --sizelimit $(( (6903807-4806656+1)*512)) ${DEV}p3 ${IMAGE}.img
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Finally, for detaching in this case the command is:
+\begin{lstlisting}[]
+# losetup -d ${DEV}p{2,3}
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+\subsection{Umount after Image Chroot}
+\label{sec:umount}
+
+In case the \texttt{umount --recursive} is not available given that a version
+higher or equal than 2.22 for \texttt{util-linux} is neither not available,
+a possible option is to manually umount all the partitions used during the
+\texttt{chroot} environment. The directories must be unmounted in the reverse
+order as when they were mounted. These commands are:
+
+\begin{lstlisting}[]
+# umount ${ROOTDIR}/dev/pts
+# umount ${ROOTDIR}/dev
+# umount ${ROOTDIR}/sys
+# umount ${ROOTDIR}/proc
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}

### 2016-07-08 overlay_fs.tex
New appendix and including more description about the persistent directory setup

@@
 
-For the \texttt{--recursive} to work properly, it is necessary that
+For the \texttt{----recursive} option to work properly, it is necessary that
 the package \texttt{util-linux} version is greater than or equal to 2.22.
@@
 \subsection{Persistent Image Directories}
+\label{sec:persistent_directories}
 Provided the stacked filesystem is configured, now it is possible to have
@@
 
-Create home partition after the root partition. Use \texttt{fdisk} to
-find the next available sector in the image. In our case, the new
-partition should start at sector 4806656:
+We create a partition for the home directory after the root partition.
+To do this, we again use \texttt{fdisk} to find the next available sector
+in the image. To verify the new available space for the full image and observe
+the next available sector, we run:
 \begin{lstlisting}[]
-$ fdisk -lu ${IMAGE}.img
+$ fdisk -u sectors -l ${IMAGE}.img
 Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
@@
 
-Create new partition:
+We notice that one GB is now available to be used in the partitions. Also, we
+observe the new partition should start at sector 4806656. To create it, we
+show the introduced command and its output:
 
 \begin{lstlisting}[]
-$ (echo n; echo p; echo ; echo 4806656; echo ; echo p; echo w) | fdisk ${IMAGE}.img
+$ (echo n; echo p; echo 3; echo 4806656; echo ; echo w) | fdisk ${IMAGE}.img
+
+Command (m for help): Partition type:
+   p primary (1 primary, 0 extended, 3 free)
+   e extended
+Partition number (1-4, default 3): Using default value 3
+First sector (2048-6903807, default 2048): Last sector, +sectors or
++size{K,M,G} (4806656-6903807, default 6903807): Using default value 6903807
+
+Command (m for help): The partition table has been altered!
+
+Syncing disks.
 \end{lstlisting}
@@
 
-Create loopback device:
+We create again a loopback device and format the new partition, as follows:
 \begin{lstlisting}[]
@@
 /dev/loop0
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Format new partition:
-\begin{lstlisting}[]
 # mkfs.ext4 ${DEV}p3
@@
 
-The last thing to remember is to inform Raspbian to mount the home partition
-we created during startup. This can be done by adding an entry in \texttt{fstab}:
+Again, if for some reason, the \texttt{-P} option is not available for
+\texttt{losetup}, we provide an alternative command line in
+Section~\ref{sec:alternative_losetup}. Finally, if the previous filesystem
+formatting was successful, the filesystem is now available for use. Thus,
+we need to inform the Raspbian root filesystem to mount the home partition
+that we have just created. We will notify to the Raspbian \ac{OS} during
+the startup of the \ac{Raspi}s. This can be done by adding an entry in
+\texttt{fstab} as follows:
+
 \begin{lstlisting}[]

### 2016-07-08 overlay_fs.tex
Structural changes: Moving HTTP to the Appendix and some editing

@@
 formatting was successful, the filesystem is now available for use. Thus,
-we need to inform the Raspbian root filesystem to mount the home partition
-that we have just created. We will notify to the Raspbian \ac{OS} during
-the startup of the \ac{Raspi}s. This can be done by adding an entry in
-\texttt{fstab} as follows:
+we need to inform the Raspbian \ac{OS} to mount the home partition
+that we have just created. We will notify the \ac{OS} while
+booting during the startup of the \ac{Raspi}s. This can be done by adding
+an entry in \texttt{fstab} as follows:
 
@@
 
-That is it. unmount partitions and detach loop device:
+Now, unmount again all the partitions and detach the loop devices as follows:
 \begin{lstlisting}[]
@@
 
-\subsection{Write image to memory card}
+If the steps have been properly made up to this point, the customized image
+is ready to be deployed into the \ac{Raspi}s. In the following section, we
+indicate how to proceed with the writing of the image into various memory
+cards, although a more complex setup with a \ac{HTTP} server is also presented
+in Section~\ref{sec:http} of the Appendix.
 
-EMPHASIZE THAT THIS SETUP IS ONLY FOR THE SETUP WHERE THE ROOT IMAGE ARE
-WRITTEN TO THE MEMORY CARDS
-
-The final step is to write the image to all the memory cards before they can
-be used in the \ac{Raspi}s.
-
-%This is done by first leaving the \texttt{chrooted}
-%filesystem and unmounting all the directories we mounted:
-
-% Exit chroot, umount, and write to memory card
-%\begin{lstlisting}[]
-%# exit
-%# cd ..
-%# umount --recursive ${ROOTDIR}
-%# losetup -d ${DEV}.img
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-%If successfully unmounted, you can
-Insert a memory card in your machine. It will be available as
-\texttt{/dev/mmcblkX} or \texttt{/dev/sdX}
+\subsection{Customized Image Writing to Memory Cards}
+For a basic system setup, the final step is to write the customized image
+to all the memory cards before they can be used in the \ac{Raspi}s. For
+our current considered system, we do this manually for each card. Insert
+a memory card in your machine. It will be available as \texttt{/dev/mmcblkX}
+or \texttt{/dev/sdX}.
 

### 2016-07-08 appendix.tex
Structural changes: Moving HTTP to the Appendix and some editing
Mounting DEVp3 on mnt question

@@
 
-The HTTP description
+\input{testbed_http}
 
@@
 NAME   MAJ:MIN RM   SIZE RO TYPE MOUNTPOINT
-  sda      8:0    0 238.5G  0 disk
-    sda1   8:1    0   512M  0 part /boot
-    sda2   8:2    0   168G  0 part /
-    sda3   8:3    0     2G  0 part [SWAP]
-    sda4   8:4    0    68G  0 part
+...
 loop1    7:1    0    67M  0 loop
 loop2    7:2    0   2.2G  0 loop
+...
 \end{lstlisting}

### 2016-07-08 overlay_fs.tex
Making new partition and detecting and small typographic error when copying
Mounting DEVp3 on mnt question

@@
 # mount ${DEV}p2 ${ROOTDIR}
-# sudo sh -c "echo '/dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' >> ${ROOTDIR}/etc/fstab"
+# sed -i '$a /dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' ${ROOTDIR}/etc/fstab
 \end{lstlisting}
@@
 
-The home folder was originally located in the root filesystem. We should
-therefore copy the content to the new home partition:
+If the last command was executed correctly, the
+\texttt{\${ROOTDIR}/etc/fstab} file should have the new line. However,
+the line will not be properly indented. To verify the correct line
+insertion and fix the indentation, edit the file with
+\texttt{sudo vim \$\{ROOTDIR\}/etc/fstab}. The resulting file should like
+the following:
+
+\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{ROOTDIR\}/etc/fstab}@>
+<@\textcolor{gray}{
+----------------------------------------}
+\Reactivatenumber @>
+proc            /proc            proc    defaults          0       0
+/dev/mmcblk0p1  /boot            vfat    defaults          0       2
+/dev/mmcblk0p2  /                ext4    defaults,noatime  0       1
+/dev/mmcblk0p3  /home            ext4    defaults,noatime  0       2
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Originally, the home folder is located in the root filesystem. Although,
+we have to move its content to the new home partition and store it properly
+
 \begin{lstlisting}[]
@@
 
-If the steps have been properly made up to this point, the customized image
+If the steps are successfully executed up to this point, the customized image
 is ready to be deployed into the \ac{Raspi}s. In the following section, we
@@
 to all the memory cards before they can be used in the \ac{Raspi}s. For
-our current considered system, we do this manually for each card. Insert
-a memory card in your machine. It will be available as \texttt{/dev/mmcblkX}
-or \texttt{/dev/sdX}.
+our current considered system, we do this manually for each card.
+The testbed administrator needs to insert each memory card in its \ac{PC}
+and follow the procedure in this section. A given card will be available as
+\texttt{/dev/mmcblk*} or \texttt{/dev/sd*} with X a given
 
-%In most systems, this memory card was available under \texttt{/dev/mmcblk0}, but this may be
-%different on different machines. For internal memory card readers, it will
-%probably be named /dev/mmcblkX where X is some number and if the memory card
-%reader is connected through \ac{USB}, then it is likely called /dev/sdX.
-Make absolutely sure to write to the correct device as everything will be
-overwritten!
+For internal memory card readers,
+it will probably be named /dev/mmcblkX where X is a given integer number
+and if the memory card reader is connected through \ac{USB}, then it is likely called /dev/sdX.
+It is very important Make absolutely sure to write to the correct device as everything will be overwritten!
 Inserting/Ejecting the memory card a couple of times and using the commands

### 2016-07-08 overlay_fs.tex
Test
Undo Test and change move to copy
minor change
fixed mistake
Trying to write memory card

@@
 Originally, the home folder is located in the root filesystem. Although,
-we have to move its content to the new home partition and store it properly
+we have to move its content to the new home partition and store it properly.
+We do that as follows:
 
@@
 and follow the procedure in this section. A given card will be available as
-\texttt{/dev/mmcblk*} or \texttt{/dev/sd*} with X a given
+\texttt{/dev/mmcblkX} for memory card readers or \texttt{/dev/sdX} USB-through
+readers , where \texttt{X} a natural number. It is \textit{very important}
+to write to the correct device as everything will be overwritten. To avoid
+removing information from the wrong device, a testbed administrator can
+insert / eject a memory card a couple of times and using the
+\texttt{sudo lsblk} and \texttt commands to deduce the correct device.
 
-For internal memory card readers,
-it will probably be named /dev/mmcblkX where X is a given integer number
-and if the memory card reader is connected through \ac{USB}, then it is likely called /dev/sdX.
-It is very important Make absolutely sure to write to the correct device as everything will be overwritten!
-Inserting/Ejecting the memory card a couple of times and using the commands
-below may help deducing the correct device:
-%The commands below may help deduce which device is the correct:
-%Below is a few commands that may help you to deduce which device is correct:
+For our case, the device was \texttt{/dev/mmcblk0}. Once identified, to write
+the image to a memory card, the following command is used:
 
-\begin{lstlisting}[]
-# lsblk
-# df -h
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-For us, the device was \texttt{/dev/mmcblk0}. Write the image to a memory card:
 \begin{lstlisting}[]
@@
 
-The memory card can now be started in a \ac{Raspi}.
+After this is made, the memory card can now be started in a \ac{Raspi}.

### 2016-07-11 overlay_fs.tex
Overlay FS done. Customized image in memory card works!

@@
 If the steps are successfully executed up to this point, the customized image
-is ready to be deployed into the \ac{Raspi}s. In the following section, we
-indicate how to proceed with the writing of the image into various memory
-cards, although a more complex setup with a \ac{HTTP} server is also presented
-in Section~\ref{sec:http} of the Appendix.
+is available in the \texttt{\$\{IMAGE\}.img} file and ready to be deployed
+into the \ac{Raspi}s. In the following section, we indicate how to proceed
+with the writing of the image into various memory cards, although a more
+complex setup with a \ac{HTTP} server is also presented in
+Section~\ref{sec:http} of the Appendix.
 
-\subsection{Customized Image Writing to Memory Cards}
+\subsection{Writing Customized Image to Memory Cards}
 For a basic system setup, the final step is to write the customized image
@@
 and follow the procedure in this section. A given card will be available as
-\texttt{/dev/mmcblkX} for memory card readers or \texttt{/dev/sdX} USB-through
-readers , where \texttt{X} a natural number. It is \textit{very important}
-to write to the correct device as everything will be overwritten. To avoid
-removing information from the wrong device, a testbed administrator can
-insert / eject a memory card a couple of times and using the
+\texttt{/dev/mmcblkX} for memory card readers or as \texttt{/dev/sdbX}
+USB-through readers, where \texttt{X} a natural number.
+
+It is \textit{very
+important} to write to the correct device as everything will be overwritten.
+To avoid removing information from the wrong device, a testbed administrator
+can insert / eject a memory card a couple of times and using the
 \texttt{sudo lsblk} and \texttt commands to deduce the correct device.
-
-For our case, the device was \texttt{/dev/mmcblk0}. Once identified, to write
+For our case, the device was \texttt{/dev/sdb} in an Ubuntu distribution and
+\texttt{/dev/mmcblk0} in an Archlinux distribution. Once identified, to write
 the image to a memory card, the following command is used:
@@
 
-After this is made, the memory card can now be started in a \ac{Raspi}.
+The previous \texttt{dd} and \texttt{sync} commands for copying the image to
+the SD memory card and flushing the remainder in memory to the filesystem
+will take around 15 minutes to be finished. This is an average on a two-core
+1.5 GHz laptop. However, other times might occur depending on the total image
+size, the writing block size and the core capabilities of the testbed
+administrator \ac{PC}. After this is made, it is only required
+to eject the memory card and now it can be started in a \ac{Raspi}.

### 2016-07-12 tools.tex
Changing Chapter 4 structure

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the testbed everyday use, there exists frequent tasks that require
a set of various commands in a given \ac{Raspi}. This could be tedious, prone
to errors and time consuming to realize everytime the task is required to be
made. Therefore, in this section we introduce a set of tools that help to
automate and monitor routinary task execution in the \ac{Raspi}s. Given the
computing capabilities of the \ac{Raspi}, common computational demanding tasks
such as source code compilation and library linking can be challenginng if
not prohibitive in terms of \ac{CPU} or memory usage. Thus, we also include
a description of how to cross-compile source code from the testbed
administrator \ac{PC} to the \ac{Raspi}s. To be able to run all the following
commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
otherwise the commands need to be run locally in a \ac{Raspi} making
necessary to use a keyboard and a monitor.

\subsection{Fabric}
Controlling multiple devices using \ac{SSH} from a single \ac{PC}
often leads to a lot of repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices, (ii) installing applications in multiple
devices and (iii) copying files to and / or from multiple devices. To
simplify the management and avoid execution issues, \texttt{Fabric} provides
a Python library to ease the task of working with multiple devices from
a single \ac{PC}. The code below provides a small script that can perform
the few items above:

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
<@\textcolor{gray}{
-------------------------------------}
\Reactivatenumber @>
from fabric.api import run, env, task
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

env.hosts = ['rasp00.domain.com','rasp01.domain.com','rasp02.domain.com']
env.user = 'pi'
env.password = 'raspberry'

@task
def reboot():
    """ Reboot device """
    sudo('reboot', quiet=True)

@task
def install(program):
    """
    Install a program
    program: program name
    """
    result = sudo('apt-get install -y {}'.format(program), quiet=True)
    print(result)

@task
def push(src,dst):
    """
    Copy file to device
    src: source file path
    dst: destination file path
    """
    put(src, dst)

\end{lstlisting}
\FloatBarrier

The script functions can be called from the shell in the following
manner. \texttt{Fabric} will automatically search the directory for a file
named 'fabfile.py'. The syntax for executing a task and arguments is in the form
'fab <FUNCTION>:arg1,arg2,...:
\begin{lstlisting}[]
$ fab reboot
$ fab install:tmux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
The first function above reboots the \ac{Raspi}s in the lists of hosts, and
the install function installs a program given by an argument. In the above case,
\texttt{tmux} is installed.
%The two functions above reboots the \ac{Raspi}s specified in the list of hosts

Below is a slightly more complicated function call that copies a file from
a \ac{PC} to all \ac{Raspi}s in the list of hosts:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/hello_world",'~/'
[192.168.87.107] Executing task 'push'
[192.168.87.107] put: /home/<USER>/code/hello_world -> /home/pi/hello_world

Done.
Disconnecting from 192.168.87.107... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\texttt{Fabric} has many other useful functionalities that are useful in
controlling a bunch of \ac{Raspi}s.

%execute a script function by calling "fab install:feh"

\subsection{Long running jobs using SSH}

There may be times when a simulation has to run for days in the \ac{Raspi}s.
For this purpose, it may be inconvenient to keep a \ac{SSH} connection
open to the \ac{Raspi}s risking that the connection will be interrupted and
thus leading to the \ac{Raspi} automatically terminating the process.

%There may be times when it is desired to start a simulation that will
%run for days. For that purpose, it may not be desired to keep
%a \ac{SSH} connection open to the \ac{Raspi} risking that the connection
%will be interrupted and thus result in the \ac{Raspi} terminating
%the running simulation.

There exists a few methods to enable the \ac{Raspi}s to continue running
applications although the connection is terminated either on purpose or
unexpectedly.
%There exists a few methods to avoid the risk of simulations terminating
%unexpectedly.
One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a \texttt{screen}
session which never terminates. Users can attach and detach from such a
session as desired.
%a session as it is
%desired. Programs within the screen session just keeps running.

The following procedure presents how to use \texttt{screen} with \ac{SSH}:
\begin{enumerate}
    \item login to a \ac{Raspi}
    \item open a screen session
    \item execute a command
    \item detach from the screen session
    \item terminate SSH connection
    \item login to the \ac{Raspi} again
    \item attach to screen session to see the program still running
\end{enumerate}

Start by establishing SSH connection to a \ac{Raspi} and open a screen session:
\begin{lstlisting}[]
$ SSH pi@<RASP_IP>
$ screen
Screen version 4.02.01 (GNU) 28-Apr-14
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Press space or return to clear the shell.

At this stage, we are in a screen session. Now execute a program that never ends:
\begin{lstlisting}[]
$ top
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

When \texttt{top} is running, press Ctrl+a Crtl+d to detach from
the \texttt{screen} session.

Terminate the SSH connection and login again:
\begin{lstlisting}[]
$ exit
$ SSH pi@<RASP_IP>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now that we are logged in to the \ac{Raspi} again, lets see if a
\texttt{screen} session is still running:
\begin{lstlisting}[]
$ screen -list
There is a screen on:
  1668.pts-0.rasp09 (06/06/16 04:02:15) (Detached)
1 Socket in /var/run/screen/S-pi.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

It is seen that the session is still running and that no user is currently
attached to the session. Attach to the session:
\begin{lstlisting}[]
$ screen -r 1668.pts-0.rasp09
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

After attaching again, \texttt{top} should still be running. \texttt{screen}
has a few more functionalities that is outside the scope of this
paper.

To terminate the \texttt{screen} session. First, terminate \texttt{top}
by pressing 'q', and then type 'exit' two times to first exit the
\texttt{screen} session and then terminate the \ac{SSH} connection:

%Problem when you logout, then applications will terminate. SCREEN is the answer.

\subsection{Cross-compilation: From your PC to the Raspberry Pi}
Some processing expensive tasks such as compiling code libraries can with
advantage be performed in a \ac{PC} rather than a \ac{Raspi} to save time.
This section will present how to setup a toolchain in a \ac{PC} that may
be used to cross compile C++ code to run in \ac{Raspi}s.
There will be given an example of compiling a simple C++ application to run
on \ac{Raspi}, but also how Kodo-cpp can be cross compiled for \ac{Raspi}.

%Some tasks can with advantage be performed in \ac{PC} rather than a \ac{Raspi}
%due to the
%
%Modern \ac{PC}s are significantly better equipped than the \ac{Raspi}s that in
%contrasts to \ac{PC}s target a far more affortable and smaller module size. This
%means that it may be preferred or even mandatory to perform some computational
%expensive tasks in a \ac{PC} rather than in \ac{Raspi}s.
%Example of tasks that are quite computational expensive is to compile software
%packages and large libraries.
%Another reason to use a \ac{PC} may be the hardware limitations of \ac{Raspi}s
%in terms of e.g. memory and disk space. Some software packages might simply be
%too memory demanding to be compiled in a \ac{Raspi}.
%Thus, this section will present how to setup a toolchain in a \ac{PC} that can
%be used to cross compile C++ code to the ARM architecture that the \ac{Raspi}s.
%The toolchain is mandatory on most consumer \ac{PC}s due to the different
%processor architecture.
%
%Furthermore, we will cover how to cross compile a simple C++ example and
%Kodo-cpp as well as how the binaries can be copied from a \ac{PC} to a
%\ac{Raspi} and executed.
%There exists bindings to Kodo in various programming language. We will present
%how to also run a Kodo-python application in a \ac{Raspi}.

%WRITE INTRO OF WHY WE WANT TO CROSS COMPILE. E.G. 1) WE DO NOT HAVE MONITORS
%CONNECTED TO THE RASPIS, 2) A PC IS SIGNIFICANTLY FASTER AND FAR BETTER EQUIPPED (RAM, CPU, HARDDRIVE), 3) MORE CONVINIENT TO DO EVERYTHING IN ONE PC AND DISTRIBUTE IT TO THE RASPIS AFTERWARDS. MORE?

\subsubsection{Setup toolchain on PC}

%Besides the previous description (\textbf{Include compiling Kodo from the
%RasPi from the scratch}), the testbed administrator can compile Kodo in its
%personal workstation and transfer the generated binaries directly to
%a path in the \ac{Raspi}. To achieve this, we get a toolchain that
%contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
%for the \ac{Raspi}. Therefore, we strongly recommend any testbed
%administrator to do the following procedure. In what follows, we provide
%the instructions considering that the NFS server uses the \texttt{\$HOME}
%directory as the working directory. However, the administrator may choose
%some other working directory of its preference if desired.

\begin{enumerate}

\item Create toolchain directory:
\begin{lstlisting}[]
$ TOOLCHAINDIR="${HOME}/toolchains"
$ mkdir -p $TOOLCHAINDIR
$ cd $TOOLCHAINDIR
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Download the \ac{Raspi} toolchain for 64-bit Linux from.
A more recent \ac{Raspi} toolchain may be available at
\texttt{http://buildbot.steinwurf.dk/toolchains/linux/}:

\begin{lstlisting}[]
$ TOOLCHAIN="raspberry-gxx49-arm"
$ wget http://kom.aau.dk/project/TuneSCode/raspi/${TOOLCHAIN}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Extract the downloaded file:
%locally in the NFS server. After
%this operation, there should be a new directory for the toolchain
%in the server. \\
\begin{lstlisting}[]
$ unzip ${TOOLCHAIN}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item (Optional) Delete zip file:
%locally in the NFS server. After
%this operation, there should be a new directory for the toolchain
%in the server. \\
\begin{lstlisting}[]
$ rm ${TOOLCHAIN}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item We can now verify that the ARM cross compiler is working:

\begin{lstlisting}[]
$ ${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++ --version
raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
Copyright (C) 2014 Free Software Foundation, Inc.
This is free software; see the source for copying conditions.  There is NO
warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\begin{lstlisting}[]
%$ echo "${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++"
%/home/<USER>/toolchains/raspberry-gxx49-arm/bin/raspberry-gxx49-arm-g++
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

%    The ARM cross compilers should now be located in

\item Make toolchain binaries available systemwide. Instead of calling
the ARM cross compiler using its full path, we can make it accessible
from the command shell systemwide. One way to do this is by adding the
toolchain binaries directory to the Linux environment variable \texttt{PATH}
when the \ac{OS} starts up:

%\item Add toolchain binaries to \texttt{PATH}. Instead specifying the full
%path to the toolchain binaries we can instead tell the operation system
%where to search for it. This makes the toolchain binaries available
%systemwide.

\begin{lstlisting}[]
$ echo PATH=\"\$PATH:${TOOLCHAINDIR}/${TOOLCHAIN}/bin\" >> ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%$ printf 'PATH="%s/%s/bin"\n' "${TOOLCHAINDIR}" "${TOOLCHAIN}" >> ${HOME}/.profile

%\item Add the \texttt{bin} folder of the toolchain to the \texttt{PATH}
%Linux environment variable of the server. This will help the server OS
%to recognize the location of the compiler command, which will be needed
%later. To do so, edit the \texttt{\$HOME/.profile} to add in a newline:
%\texttt{PATH="\$PATH:\$HOME/raspberry-gxx49-arm/bin"}. Save the
%\texttt{\$HOME/.profile}. \\

\texttt{.profile} should now contain the line we inserted. There may be more
code in you file.
% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$HOME/.profile}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
PATH="$PATH:/home/<USERNAME>/toolchains/raspberry-gxx49-arm/bin"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Update \texttt{PATH}. Source \texttt{.profile} to make the changes
take effect in your system:
\begin{lstlisting}[]
$ source ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item ARM cross compiler should now be available systemwide:

\begin{lstlisting}[]
$ raspberry-gxx49-arm-g++ --version
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\item Restart the server session in order for the changes made in the
%previous step take effect. To verify this, open a new terminal and type:
%\texttt{raspberry-gxx49-arm-g++ --version}. A correct binary installation
%should return an output similar to:
%
%\texttt{raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
%Copyright (C) 2014 Free Software Foundation, Inc.
%This is free software; see the source for copying conditions.  There is NO
%warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.} \\

\end{enumerate}

\subsubsection{Cross compile example}

The following example will provide a very basic example of 1) how to cross compile
a \texttt{hello\_world} c++ example to the \ac{Raspi}'s ARM architecture
and 2) how to copy and execute the binary in a \ac{Raspi} using \ac{SCP} and
\ac{SSH}.

Lets create a directory to hold our code:
\begin{lstlisting}[]
$ CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Create the file \texttt{hello\_world.cpp} with the following content using your
favorite text editor:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\${CODEDIR}/hello\_world.cpp}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#include <iostream>

int main()
{
    std::cout << "Hello World!" << std::endl;
    return 0;
}
\end{lstlisting}
\FloatBarrier

Save \texttt{hello\_world.cpp} and compile it for \ac{Raspi}:

\begin{lstlisting}[]
$ raspberry-gxx49-arm-g++ hello_world.cpp -o hello_world
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This should produce a binary \texttt{hello\_world}. Copy it to a \ac{Raspi}
and test it. We will use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s.

The default username for our Raspbian lite distribution is "pi" and password is "raspberry"

\begin{lstlisting}[]
$ scp main pi:<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If you do not know the \ac{IP} address of a \ac{Raspi} in your network, you can
connect a monitor to it and run the following command after you have logged in:
\begin{lstlisting}[]
pi@rasp01:~ $ ifconfig
eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
          inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
          inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
          UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
          RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
          TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
          collisions:0 txqueuelen:1000
          RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)

\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

After the executable has been copied to a \ac{Raspi}. Then, \ac{SSH} to it:

\begin{lstlisting}[]
$ ssh pi:<RASP_IP>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We can list the directory content after we have logged into the \ac{Raspi} and
see that \texttt{hello\_world} is there:

\begin{lstlisting}[]
pi@rasp07:~ $ ls
hello_world  rasp_config
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, simply execute the \texttt{hello\_world} to confirm that cross compiling
\texttt{hello\_world} worked properly:

\begin{lstlisting}[]
pi@rasp07:~ $ ./hello_world
Hello World!
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsubsection{Cross compile Kodo}

Kodo is a C++11 network coding library. It is available through bindings in
a variety of programming languages.
%There exists a number of bindings for
%Kodo to other popular programming languages.
This procedure will present how to configure the Kodo C++ bindings (Kodo-cpp)
to cross compile applications that can run in \ac{Raspi}. More details are
provided in the code documentation~\cite{kodocppdoc}.

%This procedure will go through
%how to setup Kodo-cpp to cross compile applications that can run in \ac{Raspi}. Kodo-cpp provides a
%simple interface to the underlying C++11 code that exists in the libraries
%Kodo-core and Kodo-rlnc.

To use Kodo, it is required to obtain a research- and
education-friendly license.
It can be be obtained from~\cite{steinwurflicenselink} and is free
for research and educational purposes.

%A more detailed procedure of setting up Kodo-cpp than provided below is available at

\begin{enumerate}

\item Start by creating a directory to hold the code:
\begin{lstlisting}[]
$ CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Clone the Kodo repository and change directory into the repository:
\begin{lstlisting}[]
$ git clone https://github.com/steinwurf/kodo-cpp.git
$ cd kodo-cpp
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Configure Kodo to build executables for the \ac{ARM} architecture using the \ac{Raspi} toolchain:
\begin{lstlisting}[]
$ python waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
...
'configure' finished successfully (0.620s)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Build executables:
\begin{lstlisting}[]
$ python waf build
...
'build' finished successfully (2m22.918s)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Make shared library:
\begin{lstlisting}[]
$ python waf install --install_shared_libs --install_path="./shared_test"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
\begin{lstlisting}[]
$ scp -r shared_test/include shared_test/libkodoc.so pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
\begin{lstlisting}[]
$ scp -r shared_test/encode_decode_simple pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Execute binary:
\begin{lstlisting}[]
$ ssh pi@<RASP_IP> ./encode_decode_simple
Data decoded correctly
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Cross compiling Kodo applications works provided that above command
returned "Data decoded correctly".

%\item Navigate to the repository and configure \texttt{waf} by typing:
%\texttt{python config.py} and select the 16th ``make specification'' file
%for the \ac{Raspi}, e.g. option \texttt{[16]cxx\_raspberry\_gxx49\_arm}
%presented by the file.
%
%This command configures \texttt{waf} to use the proper compiler and its
%required flags to generate the binaries for the \ac{Raspi}. If the
%configuration was correct, the output will indicate:
%\texttt{'configure' finished successfully (X.XXXs)}, where \texttt{X.XXX}
%is total time in seconds for configuring the project in the server. \\
%
%\item Execute \texttt{python waf build}. If the build process was
%successful, the generated binaries for the \ac{Raspi} should be located
%in \texttt{build/cxx\_raspberry\_gxx49\_arm} in the Kodo repository.
%\textbf{Indicate how the binary files should look like}.
%
%Once this procedure is made, the testbed administrator can relocate the
%generated binary files to the \ac{Raspi}s through the network as desired
%by using the \texttt{scp} command during the configuration step.

\end{enumerate}

%\subsection{Kodo Builds for the \ac{Raspi}, Platform Support and Documentation \textcolor{red}{(WHAT IS THE PURPOSE OF THIS SUBSECTION?)}}
%
%You can check the build status of Kodo, Fifi and other relevant projects
%through their respective repository master branch on our buildbot page
%\cite{steinwurf2016buildbot}. Our buildbot displays the status of the builds
%for Raspbian 8 and GCC 4.9 for the ARM architecture which is the relevant one
%for the \ac{Raspi}. At the link, you can check build status and build
%statistics. Also, documentation about Kodo basics with a tutorial is available
%at \cite{kododocs}.

%\subsection{Kodo python}
%
%Often when simplicity and fast prototyping is of higher importance than speed
%it is often desired to use a scripting language rather than the often faster
%compiled language. Kodo-python provides roughly similar bindings to the
%underlying C++11 implementation of Kodo as Kodo-cpp does. This means that
%python calls to Kodo-python are actually executed in the underlying C++11
%compilied code.
%
%Parsing calls and variables between Python and C++ compiled code is
%not free, but this small introduced overhead may in many cases be worth
%it to take advantage of Pythons and other scripting languages many
%benefits.
%
%Below is a short procedure to use Kodo-python in a \ac{Raspi}.
%
%One advantage of Python compared to Kodo-cpp is that only the Kodo-python
%library needs to be cross compiled. The python scripts using the Kodo
%library needs not to be cross compiled but simply copied to the same
%directory as .
%
%Kodo-python is available at \url{https://github.com/steinwurf/kodo-python}.
%
%
%
%\begin{lstlisting}[]
%$ apt-get update
%$ apt-get install python build-essential libpython-dev python-dev
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%  % IN ARCH LINUX, MOD_PYTHON NEEDS TO BE INSTALLED
%
%
%Create a directory to hold our code:
%\begin{lstlisting}[]
%$ CODEDIR="${HOME}/code"
%$ mkdir -p ${CODEDIR}
%$ cd ${CODEDIR}
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%\begin{lstlisting}[]
%$ git clone git@github.com:steinwurf/kodo-python.git
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%\begin{lstlisting}[]
%$ cd kodo-python
%$ python2.7 waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
%$ python waf build
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%KODO-PYTHON IS NOT WORKING!! WAITING FOR HELP ON STEINWURF DEV.
%
%WRITE HOW TO RUN A KODO PYTHON SCRIPT IN THE RASPIS

%\texttt{hello\_world} returns immediately. This is not always the case.
%Sometimes it is desired to run simulations that should run for days.
%A software package called \texttt{screen} can be used for this purpose.
%Simply run ...

### 2016-07-12 conclusions.tex
More changes

@@
 \label{sec:conclusions}
-OBsersving the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enable reproducible results for mechanisms to address problems in this are, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed and how to guarantee replicability and scaling managing of the system.
+Obsersving the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enable reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of the system.
 The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges.
-The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and administrable way.
+The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way.
 Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

### 2016-07-12 tools.tex
More changes

@@
 otherwise the commands need to be run locally in a \ac{Raspi} making
-necessary to use a keyboard and a monitor.
+necessary to use a keyboard and a monitor. To do so, the testbed
+administrator needs to put the memory cards in the \ac{Raspi}s and turn them
+on, for them to be able to boot. The devices should boot properly and have
+connectivity if the image and \ac{DHCP} server were properly set previously.
 
@@
 Controlling multiple devices using \ac{SSH} from a single \ac{PC}
-often leads to a lot of repetitive tasks. Among these, we can mention:
+often leads to many repetitive tasks. Among these, we can mention:
 (i) rebooting a set of devices, (ii) installing applications in multiple
-devices and (iii) copying files to and / or from multiple devices. To
-simplify the management and avoid execution issues, \texttt{Fabric} provides
-a Python library to ease the task of working with multiple devices from
-a single \ac{PC}. The code below provides a small script that can perform
-the few items above:
+devices and (iii) copying files to and / or from multiple devices, etc. To
+simplify the management and avoid execution issues, Fabric provides
+a Python library to ease the major task of working with multiple devices from
+a single \ac{PC}. First, the testbed administrator creates a directory to
+hold the \texttt{Fabric} source code:
+
+\begin{lstlisting}[]
+$ export CODEDIR="${HOME}/code"
+$ mkdir -p ${CODEDIR}
+$ cd ${CODEDIR}
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Then, the \texttt{\$CODEDIR/fabfile.py} file below provides a script with
+some basic functionalities that can perform the few items above (i-iii).
+In general, other administrators may require different functionalities
+but that goes out of the scope of this work. Nevertheless, the following
+file serves as a starting boilerplate:
 
@@
 \Reactivatenumber @>
-from fabric.api import run, env, task
+from fabric.api import env, task, sudo
 # Python Fabric script to run commands on multiple hosts through ssh
@@
 
-The script functions can be called from the shell in the following
-manner. \texttt{Fabric} will automatically search the directory for a file
-named 'fabfile.py'. The syntax for executing a task and arguments is in the form
-'fab <FUNCTION>:arg1,arg2,...:
+The previous \texttt{fabfile} shows three function that perform our
+example tasks. These functions utilize variables and subsequent functions
+from the Fabric \ac{API} such as \texttt{env}, \texttt{task} and \texttt{sudo}
+among others. Each of these \ac{API} functions permit to define environment
+variables, create the administrator tasks through decorators or run the
+mentioned task in \texttt{sudo} mode, respectively.
+When a task is called from the terminal, Fabric searches the directory for
+the \texttt{fabfile.py} file and executes the desired task. The syntax for
+executing a task with arguments is in the form
+\texttt{fab <FUNCTION>:arg1,arg2,...}. Some executions from the terminal
+are shown as follows \todot{Test the following commands later and print the
+outputs}:
+
 \begin{lstlisting}[]
@@
 \vspace{-5mm}
-The first function above reboots the \ac{Raspi}s in the lists of hosts, and
-the install function installs a program given by an argument. In the above case,
-\texttt{tmux} is installed.
-%The two functions above reboots the \ac{Raspi}s specified in the list of hosts
 
-Below is a slightly more complicated function call that copies a file from
-a \ac{PC} to all \ac{Raspi}s in the list of hosts:
+The first function above reboots the \ac{Raspi}s in the lists of hosts and
+the second function installs a program given by an argument. In both cases,
+after a successful \ac{SSH} connection is made, Fabric employs the
+\ac{Raspi}'s \texttt{reboot} and \texttt{apt-get} commands in \texttt{sudo}
+mode to do the required tasks. Below it is shown an example for
+the \texttt{push} task which uses two arguments. Here we copy
+\texttt{my\_file} from the testbed administrator \ac{PC} to all the host
+\ac{Raspi}s:
 
 \begin{lstlisting}[]
-$ fab push:"${CODEDIR}/hello_world",'~/'
+$ fab push:"${CODEDIR}/my_file",'~/'
 [192.168.87.107] Executing task 'push'
-[192.168.87.107] put: /home/<USER>/code/hello_world -> /home/pi/hello_world
+[192.168.87.107] put: /home/<USER>/code/my_file -> /home/pi/my_file
 
@@
 
-\texttt{Fabric} has many other useful functionalities that are useful in
-controlling a bunch of \ac{Raspi}s.
+Fabric has many other functionalities that are useful in
+controlling a large set of \ac{Raspi}s. For example, we may extract files
+or run automated experiments. The included functionalities in
+the \texttt{\$CODEDIR/fabfile.py} file will depend on the requirements
+of the testbed administrator.
 
 %execute a script function by calling "fab install:feh"
+\subsection{Long running jobs using SSH}
+There are times when a task may need to run for several hours or even days
+in the \ac{Raspi}s, particularly when related to simulations or measurement
+campaigns. For this purpose, it might be necessary to keep open a \ac{SSH}
+connection in the \ac{Raspi}s without risking that the connection will
+be interrupted and a given \ac{Raspi} terminating the task.
 
-\subsection{Long running jobs using SSH}
-
-There may be times when a simulation has to run for days in the \ac{Raspi}s.
-For this purpose, it may be inconvenient to keep a \ac{SSH} connection
-open to the \ac{Raspi}s risking that the connection will be interrupted and
-thus leading to the \ac{Raspi} automatically terminating the process.
-
-%There may be times when it is desired to start a simulation that will
-%run for days. For that purpose, it may not be desired to keep
-%a \ac{SSH} connection open to the \ac{Raspi} risking that the connection
-%will be interrupted and thus result in the \ac{Raspi} terminating
-%the running simulation.
-
-There exists a few methods to enable the \ac{Raspi}s to continue running
+There are methods to enable the \ac{Raspi}s to continue running
 applications although the connection is terminated either on purpose or
-unexpectedly.
-%There exists a few methods to avoid the risk of simulations terminating
-%unexpectedly.
-One method is to run programs within a \texttt{screen} session.
-\texttt{screen} enables a user to run applications within a \texttt{screen}
-session which never terminates. Users can attach and detach from such a
-session as desired.
-%a session as it is
-%desired. Programs within the screen session just keeps running.
-
-The following procedure presents how to use \texttt{screen} with \ac{SSH}:
-\begin{enumerate}
-    \item login to a \ac{Raspi}
-    \item open a screen session
-    \item execute a command
-    \item detach from the screen session
-    \item terminate SSH connection
-    \item login to the \ac{Raspi} again
-    \item attach to screen session to see the program still running
-\end{enumerate}
-
-Start by establishing SSH connection to a \ac{Raspi} and open a screen session:
+unexpectedly. One method is to run programs within a \texttt{screen} session.
+\texttt{screen} enables a user to run applications within a shell
+window, a \texttt{screen} session, which does not terminate even with
+connectivity interruptions. Users can attach and detach from such a session
+as desired. The following procedure presents how to use \texttt{screen} with
+\ac{SSH} to: (i) Login to a \ac{Raspi}, (ii) open a screen session, (iii)
+execute an example command, (iv) detach from the screen session, (v)
+terminate the \ac{SSH} connection, (vi) login to the \ac{Raspi} again
+and (vii) attach to screen session to see the program still running.
+From the testbed administrator, we start by establishing \ac{SSH} connection
+to a \ac{Raspi} and open a screen session:
 \begin{lstlisting}[]
-$ SSH pi@<RASP_IP>
+$ ssh pi@<RASP_IP>
 $ screen
@@
 %Problem when you logout, then applications will terminate. SCREEN is the answer.
-
 \subsection{Cross-compilation: From your PC to the Raspberry Pi}
@@
 \ac{SSH}.
-
-Lets create a directory to hold our code:
-\begin{lstlisting}[]
-$ CODEDIR="${HOME}/code"
-$ mkdir -p ${CODEDIR}
-$ cd ${CODEDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
 

### 2016-07-12 cross_compilation.tex
Separate automation tools from cross-compiling

--- 2016-07-12 tools.tex
@@
-%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
-\label{sec:tools}
-Within the testbed everyday use, there exists frequent tasks that require
-a set of various commands in a given \ac{Raspi}. This could be tedious, prone
-to errors and time consuming to realize everytime the task is required to be
-made. Therefore, in this section we introduce a set of tools that help to
-automate and monitor routinary task execution in the \ac{Raspi}s. Given the
-computing capabilities of the \ac{Raspi}, common computational demanding tasks
-such as source code compilation and library linking can be challenginng if
-not prohibitive in terms of \ac{CPU} or memory usage. Thus, we also include
-a description of how to cross-compile source code from the testbed
-administrator \ac{PC} to the \ac{Raspi}s. To be able to run all the following
-commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
-otherwise the commands need to be run locally in a \ac{Raspi} making
-necessary to use a keyboard and a monitor.
+\label{sec:cross_compilation}
+An important case of a computational expensive task is to compile software
+packages and large libraries. Given the computing capabilities of the
+\ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
+\ac{CPU}, memory or space usage and / or compilation time. Thus, in
+this Section we present a procedure of how to cross-compile C++
+source code from the testbed administrator \ac{PC} for the \ac{ARM}
+architecture of the \ac{Raspi}s. By doing this, we take advantage
+of the (typically) much higher computing power of the testbed
+administrator \ac{PC} in order to save time and computational resources.
+Hence, we give an example of compiling a simple C++ program to run
+on a \ac{Raspi}.
 
-\subsection{Fabric}
-Controlling multiple devices using \ac{SSH} from a single \ac{PC}
-often leads to a lot of repetitive tasks. Among these, we can mention:
-(i) rebooting a set of devices, (ii) installing applications in multiple
-devices and (iii) copying files to and / or from multiple devices. To
-simplify the management and avoid execution issues, \texttt{Fabric} provides
-a Python library to ease the task of working with multiple devices from
-a single \ac{PC}. The code below provides a small script that can perform
-the few items above:
+Furthermore, given that our testbed purpose is for
+network coding applications, we also present how to cross-compile Kodo
+\cite{kodo2011pedersen}, a C++11 network coding library to perform
+encoding, decoding and recoding operations. In this way, we aim to present
+a fully configurable and manageable testbed with the capabilities to
+evaluate network coding protocols with several \ac{Raspi}s and locally
+store measurements from different evaluations. Therefore, we also show
+how \texttt{kodo-cpp}, a set of high-level C++ bindings for Kodo, can be
+cross-compiled for applications with the \ac{Raspi}.
 
-% Exit chroot, umount, and write to memory card
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
-<@\textcolor{gray}{
--------------------------------------}
-\Reactivatenumber @>
-from fabric.api import run, env, task
-# Python Fabric script to run commands on multiple hosts through ssh
-#
-# Run script as 'fab <task>', where <task> is one of the scripts functions
-# marked as a tesk. The task marked as 'default' will be run if <task> is not
-# specified
-
-env.hosts = ['rasp00.domain.com','rasp01.domain.com','rasp02.domain.com']
-env.user = 'pi'
-env.password = 'raspberry'
-
-@task
-def reboot():
-    """ Reboot device """
-    sudo('reboot', quiet=True)
-
-@task
-def install(program):
-    """
-    Install a program
-    program: program name
-    """
-    result = sudo('apt-get install -y {}'.format(program), quiet=True)
-    print(result)
-
-@task
-def push(src,dst):
-    """
-    Copy file to device
-    src: source file path
-    dst: destination file path
-    """
-    put(src, dst)
-
-\end{lstlisting}
-\FloatBarrier
-
-The script functions can be called from the shell in the following
-manner. \texttt{Fabric} will automatically search the directory for a file
-named 'fabfile.py'. The syntax for executing a task and arguments is in the form
-'fab <FUNCTION>:arg1,arg2,...:
-\begin{lstlisting}[]
-$ fab reboot
-$ fab install:tmux
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-The first function above reboots the \ac{Raspi}s in the lists of hosts, and
-the install function installs a program given by an argument. In the above case,
-\texttt{tmux} is installed.
-%The two functions above reboots the \ac{Raspi}s specified in the list of hosts
-
-Below is a slightly more complicated function call that copies a file from
-a \ac{PC} to all \ac{Raspi}s in the list of hosts:
-
-\begin{lstlisting}[]
-$ fab push:"${CODEDIR}/hello_world",'~/'
-[192.168.87.107] Executing task 'push'
-[192.168.87.107] put: /home/<USER>/code/hello_world -> /home/pi/hello_world
-
-Done.
-Disconnecting from 192.168.87.107... done.
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\texttt{Fabric} has many other useful functionalities that are useful in
-controlling a bunch of \ac{Raspi}s.
-
-%execute a script function by calling "fab install:feh"
-
-\subsection{Long running jobs using SSH}
-
-There may be times when a simulation has to run for days in the \ac{Raspi}s.
-For this purpose, it may be inconvenient to keep a \ac{SSH} connection
-open to the \ac{Raspi}s risking that the connection will be interrupted and
-thus leading to the \ac{Raspi} automatically terminating the process.
-
-%There may be times when it is desired to start a simulation that will
-%run for days. For that purpose, it may not be desired to keep
-%a \ac{SSH} connection open to the \ac{Raspi} risking that the connection
-%will be interrupted and thus result in the \ac{Raspi} terminating
-%the running simulation.
-
-There exists a few methods to enable the \ac{Raspi}s to continue running
-applications although the connection is terminated either on purpose or
-unexpectedly.
-%There exists a few methods to avoid the risk of simulations terminating
-%unexpectedly.
-One method is to run programs within a \texttt{screen} session.
-\texttt{screen} enables a user to run applications within a \texttt{screen}
-session which never terminates. Users can attach and detach from such a
-session as desired.
-%a session as it is
-%desired. Programs within the screen session just keeps running.
-
-The following procedure presents how to use \texttt{screen} with \ac{SSH}:
-\begin{enumerate}
-    \item login to a \ac{Raspi}
-    \item open a screen session
-    \item execute a command
-    \item detach from the screen session
-    \item terminate SSH connection
-    \item login to the \ac{Raspi} again
-    \item attach to screen session to see the program still running
-\end{enumerate}
-
-Start by establishing SSH connection to a \ac{Raspi} and open a screen session:
-\begin{lstlisting}[]
-$ SSH pi@<RASP_IP>
-$ screen
-Screen version 4.02.01 (GNU) 28-Apr-14
-...
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-Press space or return to clear the shell.
-
-At this stage, we are in a screen session. Now execute a program that never ends:
-\begin{lstlisting}[]
-$ top
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-When \texttt{top} is running, press Ctrl+a Crtl+d to detach from
-the \texttt{screen} session.
-
-Terminate the SSH connection and login again:
-\begin{lstlisting}[]
-$ exit
-$ SSH pi@<RASP_IP>
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now that we are logged in to the \ac{Raspi} again, lets see if a
-\texttt{screen} session is still running:
-\begin{lstlisting}[]
-$ screen -list
-There is a screen on:
-  1668.pts-0.rasp09 (06/06/16 04:02:15) (Detached)
-1 Socket in /var/run/screen/S-pi.
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-It is seen that the session is still running and that no user is currently
-attached to the session. Attach to the session:
-\begin{lstlisting}[]
-$ screen -r 1668.pts-0.rasp09
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-After attaching again, \texttt{top} should still be running. \texttt{screen}
-has a few more functionalities that is outside the scope of this
-paper.
-
-To terminate the \texttt{screen} session. First, terminate \texttt{top}
-by pressing 'q', and then type 'exit' two times to first exit the
-\texttt{screen} session and then terminate the \ac{SSH} connection:
-
-%Problem when you logout, then applications will terminate. SCREEN is the answer.
-
-\subsection{Cross-compilation: From your PC to the Raspberry Pi}
-Some processing expensive tasks such as compiling code libraries can with
-advantage be performed in a \ac{PC} rather than a \ac{Raspi} to save time.
-This section will present how to setup a toolchain in a \ac{PC} that may
-be used to cross compile C++ code to run in \ac{Raspi}s.
-There will be given an example of compiling a simple C++ application to run
-on \ac{Raspi}, but also how Kodo-cpp can be cross compiled for \ac{Raspi}.
-
-%Some tasks can with advantage be performed in \ac{PC} rather than a \ac{Raspi}
-%due to the
-%
-%Modern \ac{PC}s are significantly better equipped than the \ac{Raspi}s that in
-%contrasts to \ac{PC}s target a far more affortable and smaller module size. This
-%means that it may be preferred or even mandatory to perform some computational
-%expensive tasks in a \ac{PC} rather than in \ac{Raspi}s.
-%Example of tasks that are quite computational expensive is to compile software
-%packages and large libraries.
-%Another reason to use a \ac{PC} may be the hardware limitations of \ac{Raspi}s
-%in terms of e.g. memory and disk space. Some software packages might simply be
-%too memory demanding to be compiled in a \ac{Raspi}.
-%Thus, this section will present how to setup a toolchain in a \ac{PC} that can
-%be used to cross compile C++ code to the ARM architecture that the \ac{Raspi}s.
 %The toolchain is mandatory on most consumer \ac{PC}s due to the different
@@
 %\ac{Raspi} and executed.
-%There exists bindings to Kodo in various programming language. We will present
-%how to also run a Kodo-python application in a \ac{Raspi}.
 
-%WRITE INTRO OF WHY WE WANT TO CROSS COMPILE. E.G. 1) WE DO NOT HAVE MONITORS
-%CONNECTED TO THE RASPIS, 2) A PC IS SIGNIFICANTLY FASTER AND FAR BETTER EQUIPPED (RAM, CPU, HARDDRIVE), 3) MORE CONVINIENT TO DO EVERYTHING IN ONE PC AND DISTRIBUTE IT TO THE RASPIS AFTERWARDS. MORE?
-
-\subsubsection{Setup toolchain on PC}
+\subsection{Setup toolchain on PC}
 
@@
 
-\subsubsection{Cross compile example}
+\subsection{Cross compile example}
 
@@
 
-\subsubsection{Cross compile Kodo}
+\subsection{Cross compile Kodo}
 

### 2016-07-12 tools.tex
Separate automation tools from cross-compiling

@@
 made. Therefore, in this section we introduce a set of tools that help to
-automate and monitor routinary task execution in the \ac{Raspi}s. Given the
-computing capabilities of the \ac{Raspi}, common computational demanding tasks
-such as source code compilation and library linking can be challenginng if
-not prohibitive in terms of \ac{CPU} or memory usage. Thus, we also include
-a description of how to cross-compile source code from the testbed
-administrator \ac{PC} to the \ac{Raspi}s. To be able to run all the following
+automate and monitor routinary tasks execution in the \ac{Raspi}s and show
+relevant example commands with them. To be able to run all the following
 commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
@@
 executing a task with arguments is in the form
-\texttt{fab <FUNCTION>:arg1,arg2,...}. Some executions from the terminal
+\texttt{fab <TASK>:arg1,arg2,...}. Some executions from the terminal
 are shown as follows \todot{Test the following commands later and print the
@@
 as desired. The following procedure presents how to use \texttt{screen} with
-\ac{SSH} to: (i) Login to a \ac{Raspi}, (ii) open a screen session, (iii)
+\ac{SSH} to: (i) Login to a generic \ac{Raspi} with \ac{IP} address as
+\texttt{<RASP\_IP>}, (ii) open a screen session, (iii)
 execute an example command, (iv) detach from the screen session, (v)
@@
 to a \ac{Raspi} and open a screen session:
+
 \begin{lstlisting}[]
 $ ssh pi@<RASP_IP>
+...
+pi@<RASP_IP>'s password:
+
+The programs included with the Debian GNU/Linux system are free software;
+the exact distribution terms for each program are described in the
+individual files in /usr/share/doc/*/copyright.
+
+Debian GNU/Linux comes with ABSOLUTELY NO WARRANTY, to the extent
+permitted by applicable law.
+Last login: Tue Jul 12 13:04:31 2016
+
 $ screen
@@
 ...
+[Press space or Return to end.]
 \end{lstlisting}
@@
 \vspace{-5mm}
-Press space or return to clear the shell.
 
-At this stage, we are in a screen session. Now execute a program that never ends:
+To enter in the \texttt{screen} session after the introduction message, we
+have to press either the \texttt{Space} or \texttt{Return} key in the
+keyboard to clear the shell. After doing so, we should be in a
+\texttt{screen} session although its appearance is the same as a regular
+terminal shell. Inside this example session, we execute a program that
+never ends:
+
 \begin{lstlisting}[]
@@
 
-When \texttt{top} is running, press Ctrl+a Crtl+d to detach from
-the \texttt{screen} session.
+The \texttt{top} command simply shows continuously the table of processes
+executed in the \ac{Raspi} like in any Linux distribution. When \texttt{top}
+is running, we first press \texttt{Ctrl+a} and later \texttt{Crtl+d}
+in the keyboard to detach from the \texttt{screen} session. We now
+terminate the \ac{SSH} connection and login again to verify that the
+\texttt{top} command is still running. Without using \texttt{screen},
+the \texttt{top} program should terminate since its hosting shell was
+terminated. To log out, we run:
 
-Terminate the SSH connection and login again:
 \begin{lstlisting}[]
 $ exit
-$ SSH pi@<RASP_IP>
+logout
+Connection to <RASP_IP> closed.
+$ ssh pi@<RASP_IP>
 \end{lstlisting}
@@
 
-Now that we are logged in to the \ac{Raspi} again, lets see if a
-\texttt{screen} session is still running:
+Now that we are logged in to the \ac{Raspi} again, we first check the
+available detached sessions by running:
 \begin{lstlisting}[]
@@
 There is a screen on:
-  1668.pts-0.rasp09 (06/06/16 04:02:15) (Detached)
+  824.pts-0.raspXX (07/12/16 13:17:30) (Detached)
 1 Socket in /var/run/screen/S-pi.
@@
 
-It is seen that the session is still running and that no user is currently
-attached to the session. Attach to the session:
+From the command output, we see that the session is still running in our
+generic \ac{Raspi} number \texttt{XX} and that no user is currently attached
+to the session. To attach to the session, we execute:
+
 \begin{lstlisting}[]
-$ screen -r 1668.pts-0.rasp09
+$ screen -r 824.pts-0.raspXX
 \end{lstlisting}
@@
 
-After attaching again, \texttt{top} should still be running. \texttt{screen}
-has a few more functionalities that is outside the scope of this
-paper.
+After attaching again, we should see the \texttt{top} still running.
+\texttt{screen} has more functionalities that can be used in this
+or other contexts but that is outside the scope of this work.
+To terminate the \texttt{screen} session, first terminate \texttt{top}
+by pressing \texttt{q} in the keyboard. Once out of the \texttt{top}
+output, we need to type \texttt{exit} two times to first exit the
+\texttt{screen} session and then terminate the \ac{SSH} connection.
+An output should be as follows:
 
-To terminate the \texttt{screen} session. First, terminate \texttt{top}
-by pressing 'q', and then type 'exit' two times to first exit the
-\texttt{screen} session and then terminate the \ac{SSH} connection:
-
-%Problem when you logout, then applications will terminate. SCREEN is the answer.
-\subsection{Cross-compilation: From your PC to the Raspberry Pi}
-Some processing expensive tasks such as compiling code libraries can with
-advantage be performed in a \ac{PC} rather than a \ac{Raspi} to save time.
-This section will present how to setup a toolchain in a \ac{PC} that may
-be used to cross compile C++ code to run in \ac{Raspi}s.
-There will be given an example of compiling a simple C++ application to run
-on \ac{Raspi}, but also how Kodo-cpp can be cross compiled for \ac{Raspi}.
-
-%Some tasks can with advantage be performed in \ac{PC} rather than a \ac{Raspi}
-%due to the
-%
-%Modern \ac{PC}s are significantly better equipped than the \ac{Raspi}s that in
-%contrasts to \ac{PC}s target a far more affortable and smaller module size. This
-%means that it may be preferred or even mandatory to perform some computational
-%expensive tasks in a \ac{PC} rather than in \ac{Raspi}s.
-%Example of tasks that are quite computational expensive is to compile software
-%packages and large libraries.
-%Another reason to use a \ac{PC} may be the hardware limitations of \ac{Raspi}s
-%in terms of e.g. memory and disk space. Some software packages might simply be
-%too memory demanding to be compiled in a \ac{Raspi}.
-%Thus, this section will present how to setup a toolchain in a \ac{PC} that can
-%be used to cross compile C++ code to the ARM architecture that the \ac{Raspi}s.
-%The toolchain is mandatory on most consumer \ac{PC}s due to the different
-%processor architecture.
-%
-%Furthermore, we will cover how to cross compile a simple C++ example and
-%Kodo-cpp as well as how the binaries can be copied from a \ac{PC} to a
-%\ac{Raspi} and executed.
-%There exists bindings to Kodo in various programming language. We will present
-%how to also run a Kodo-python application in a \ac{Raspi}.
-
-%WRITE INTRO OF WHY WE WANT TO CROSS COMPILE. E.G. 1) WE DO NOT HAVE MONITORS
-%CONNECTED TO THE RASPIS, 2) A PC IS SIGNIFICANTLY FASTER AND FAR BETTER EQUIPPED (RAM, CPU, HARDDRIVE), 3) MORE CONVINIENT TO DO EVERYTHING IN ONE PC AND DISTRIBUTE IT TO THE RASPIS AFTERWARDS. MORE?
-
-\subsubsection{Setup toolchain on PC}
-
-%Besides the previous description (\textbf{Include compiling Kodo from the
-%RasPi from the scratch}), the testbed administrator can compile Kodo in its
-%personal workstation and transfer the generated binaries directly to
-%a path in the \ac{Raspi}. To achieve this, we get a toolchain that
-%contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
-%for the \ac{Raspi}. Therefore, we strongly recommend any testbed
-%administrator to do the following procedure. In what follows, we provide
-%the instructions considering that the NFS server uses the \texttt{\$HOME}
-%directory as the working directory. However, the administrator may choose
-%some other working directory of its preference if desired.
-
-\begin{enumerate}
-
-\item Create toolchain directory:
 \begin{lstlisting}[]
-$ TOOLCHAINDIR="${HOME}/toolchains"
-$ mkdir -p $TOOLCHAINDIR
-$ cd $TOOLCHAINDIR
+[screen is terminating]
+pi@<RASP_IP>:~ $ exit
+logout
+Connection to <RASP_IP> closed.
+$
 \end{lstlisting}
@@
 \vspace{-5mm}
-
-\item Download the \ac{Raspi} toolchain for 64-bit Linux from.
-A more recent \ac{Raspi} toolchain may be available at
-\texttt{http://buildbot.steinwurf.dk/toolchains/linux/}:
-
-\begin{lstlisting}[]
-$ TOOLCHAIN="raspberry-gxx49-arm"
-$ wget http://kom.aau.dk/project/TuneSCode/raspi/${TOOLCHAIN}.zip
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Extract the downloaded file:
-%locally in the NFS server. After
-%this operation, there should be a new directory for the toolchain
-%in the server. \\
-\begin{lstlisting}[]
-$ unzip ${TOOLCHAIN}.zip
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item (Optional) Delete zip file:
-%locally in the NFS server. After
-%this operation, there should be a new directory for the toolchain
-%in the server. \\
-\begin{lstlisting}[]
-$ rm ${TOOLCHAIN}.zip
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item We can now verify that the ARM cross compiler is working:
-
-\begin{lstlisting}[]
-$ ${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++ --version
-raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
-Copyright (C) 2014 Free Software Foundation, Inc.
-This is free software; see the source for copying conditions.  There is NO
-warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%\begin{lstlisting}[]
-%$ echo "${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++"
-%/home/<USER>/toolchains/raspberry-gxx49-arm/bin/raspberry-gxx49-arm-g++
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-%    The ARM cross compilers should now be located in
-
-\item Make toolchain binaries available systemwide. Instead of calling
-the ARM cross compiler using its full path, we can make it accessible
-from the command shell systemwide. One way to do this is by adding the
-toolchain binaries directory to the Linux environment variable \texttt{PATH}
-when the \ac{OS} starts up:
-
-%\item Add toolchain binaries to \texttt{PATH}. Instead specifying the full
-%path to the toolchain binaries we can instead tell the operation system
-%where to search for it. This makes the toolchain binaries available
-%systemwide.
-
-\begin{lstlisting}[]
-$ echo PATH=\"\$PATH:${TOOLCHAINDIR}/${TOOLCHAIN}/bin\" >> ${HOME}/.profile
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%$ printf 'PATH="%s/%s/bin"\n' "${TOOLCHAINDIR}" "${TOOLCHAIN}" >> ${HOME}/.profile
-
-%\item Add the \texttt{bin} folder of the toolchain to the \texttt{PATH}
-%Linux environment variable of the server. This will help the server OS
-%to recognize the location of the compiler command, which will be needed
-%later. To do so, edit the \texttt{\$HOME/.profile} to add in a newline:
-%\texttt{PATH="\$PATH:\$HOME/raspberry-gxx49-arm/bin"}. Save the
-%\texttt{\$HOME/.profile}. \\
-
-\texttt{.profile} should now contain the line we inserted. There may be more
-code in you file.
-% MAC and Hostname file
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$HOME/.profile}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-PATH="$PATH:/home/<USERNAME>/toolchains/raspberry-gxx49-arm/bin"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Update \texttt{PATH}. Source \texttt{.profile} to make the changes
-take effect in your system:
-\begin{lstlisting}[]
-$ source ${HOME}/.profile
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item ARM cross compiler should now be available systemwide:
-
-\begin{lstlisting}[]
-$ raspberry-gxx49-arm-g++ --version
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%\item Restart the server session in order for the changes made in the
-%previous step take effect. To verify this, open a new terminal and type:
-%\texttt{raspberry-gxx49-arm-g++ --version}. A correct binary installation
-%should return an output similar to:
-%
-%\texttt{raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
-%Copyright (C) 2014 Free Software Foundation, Inc.
-%This is free software; see the source for copying conditions.  There is NO
-%warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.} \\
-
-\end{enumerate}
-
-\subsubsection{Cross compile example}
-
-The following example will provide a very basic example of 1) how to cross compile
-a \texttt{hello\_world} c++ example to the \ac{Raspi}'s ARM architecture
-and 2) how to copy and execute the binary in a \ac{Raspi} using \ac{SCP} and
-\ac{SSH}.
-
-Create the file \texttt{hello\_world.cpp} with the following content using your
-favorite text editor:
-
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\${CODEDIR}/hello\_world.cpp}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-#include <iostream>
-
-int main()
-{
-    std::cout << "Hello World!" << std::endl;
-    return 0;
-}
-\end{lstlisting}
-\FloatBarrier
-
-Save \texttt{hello\_world.cpp} and compile it for \ac{Raspi}:
-
-\begin{lstlisting}[]
-$ raspberry-gxx49-arm-g++ hello_world.cpp -o hello_world
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-This should produce a binary \texttt{hello\_world}. Copy it to a \ac{Raspi}
-and test it. We will use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s.
-
-The default username for our Raspbian lite distribution is "pi" and password is "raspberry"
-
-\begin{lstlisting}[]
-$ scp main pi:<RASP_IP>:~/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-If you do not know the \ac{IP} address of a \ac{Raspi} in your network, you can
-connect a monitor to it and run the following command after you have logged in:
-\begin{lstlisting}[]
-pi@rasp01:~ $ ifconfig
-eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
-          inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
-          inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
-          UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
-          RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
-          TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
-          collisions:0 txqueuelen:1000
-          RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)
-
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-After the executable has been copied to a \ac{Raspi}. Then, \ac{SSH} to it:
-
-\begin{lstlisting}[]
-$ ssh pi:<RASP_IP>
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-We can list the directory content after we have logged into the \ac{Raspi} and
-see that \texttt{hello\_world} is there:
-
-\begin{lstlisting}[]
-pi@rasp07:~ $ ls
-hello_world  rasp_config
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Now, simply execute the \texttt{hello\_world} to confirm that cross compiling
-\texttt{hello\_world} worked properly:
-
-\begin{lstlisting}[]
-pi@rasp07:~ $ ./hello_world
-Hello World!
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\subsubsection{Cross compile Kodo}
-
-Kodo is a C++11 network coding library. It is available through bindings in
-a variety of programming languages.
-%There exists a number of bindings for
-%Kodo to other popular programming languages.
-This procedure will present how to configure the Kodo C++ bindings (Kodo-cpp)
-to cross compile applications that can run in \ac{Raspi}. More details are
-provided in the code documentation~\cite{kodocppdoc}.
-
-%This procedure will go through
-%how to setup Kodo-cpp to cross compile applications that can run in \ac{Raspi}. Kodo-cpp provides a
-%simple interface to the underlying C++11 code that exists in the libraries
-%Kodo-core and Kodo-rlnc.
-
-To use Kodo, it is required to obtain a research- and
-education-friendly license.
-It can be be obtained from~\cite{steinwurflicenselink} and is free
-for research and educational purposes.
-
-%A more detailed procedure of setting up Kodo-cpp than provided below is available at
-
-\begin{enumerate}
-
-\item Start by creating a directory to hold the code:
-\begin{lstlisting}[]
-$ CODEDIR="${HOME}/code"
-$ mkdir -p ${CODEDIR}
-$ cd ${CODEDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Clone the Kodo repository and change directory into the repository:
-\begin{lstlisting}[]
-$ git clone https://github.com/steinwurf/kodo-cpp.git
-$ cd kodo-cpp
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Configure Kodo to build executables for the \ac{ARM} architecture using the \ac{Raspi} toolchain:
-\begin{lstlisting}[]
-$ python waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
-...
-'configure' finished successfully (0.620s)
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Build executables:
-\begin{lstlisting}[]
-$ python waf build
-...
-'build' finished successfully (2m22.918s)
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Make shared library:
-\begin{lstlisting}[]
-$ python waf install --install_shared_libs --install_path="./shared_test"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
-\begin{lstlisting}[]
-$ scp -r shared_test/include shared_test/libkodoc.so pi@<RASP_IP>:~/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
-\begin{lstlisting}[]
-$ scp -r shared_test/encode_decode_simple pi@<RASP_IP>:~/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Execute binary:
-\begin{lstlisting}[]
-$ ssh pi@<RASP_IP> ./encode_decode_simple
-Data decoded correctly
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Cross compiling Kodo applications works provided that above command
-returned "Data decoded correctly".
-
-%\item Navigate to the repository and configure \texttt{waf} by typing:
-%\texttt{python config.py} and select the 16th ``make specification'' file
-%for the \ac{Raspi}, e.g. option \texttt{[16]cxx\_raspberry\_gxx49\_arm}
-%presented by the file.
-%
-%This command configures \texttt{waf} to use the proper compiler and its
-%required flags to generate the binaries for the \ac{Raspi}. If the
-%configuration was correct, the output will indicate:
-%\texttt{'configure' finished successfully (X.XXXs)}, where \texttt{X.XXX}
-%is total time in seconds for configuring the project in the server. \\
-%
-%\item Execute \texttt{python waf build}. If the build process was
-%successful, the generated binaries for the \ac{Raspi} should be located
-%in \texttt{build/cxx\_raspberry\_gxx49\_arm} in the Kodo repository.
-%\textbf{Indicate how the binary files should look like}.
-%
-%Once this procedure is made, the testbed administrator can relocate the
-%generated binary files to the \ac{Raspi}s through the network as desired
-%by using the \texttt{scp} command during the configuration step.
-
-\end{enumerate}
-
-%\subsection{Kodo Builds for the \ac{Raspi}, Platform Support and Documentation \textcolor{red}{(WHAT IS THE PURPOSE OF THIS SUBSECTION?)}}
-%
-%You can check the build status of Kodo, Fifi and other relevant projects
-%through their respective repository master branch on our buildbot page
-%\cite{steinwurf2016buildbot}. Our buildbot displays the status of the builds
-%for Raspbian 8 and GCC 4.9 for the ARM architecture which is the relevant one
-%for the \ac{Raspi}. At the link, you can check build status and build
-%statistics. Also, documentation about Kodo basics with a tutorial is available
-%at \cite{kododocs}.
-
-%\subsection{Kodo python}
-%
-%Often when simplicity and fast prototyping is of higher importance than speed
-%it is often desired to use a scripting language rather than the often faster
-%compiled language. Kodo-python provides roughly similar bindings to the
-%underlying C++11 implementation of Kodo as Kodo-cpp does. This means that
-%python calls to Kodo-python are actually executed in the underlying C++11
-%compilied code.
-%
-%Parsing calls and variables between Python and C++ compiled code is
-%not free, but this small introduced overhead may in many cases be worth
-%it to take advantage of Pythons and other scripting languages many
-%benefits.
-%
-%Below is a short procedure to use Kodo-python in a \ac{Raspi}.
-%
-%One advantage of Python compared to Kodo-cpp is that only the Kodo-python
-%library needs to be cross compiled. The python scripts using the Kodo
-%library needs not to be cross compiled but simply copied to the same
-%directory as .
-%
-%Kodo-python is available at \url{https://github.com/steinwurf/kodo-python}.
-%
-%
-%
-%\begin{lstlisting}[]
-%$ apt-get update
-%$ apt-get install python build-essential libpython-dev python-dev
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-%  % IN ARCH LINUX, MOD_PYTHON NEEDS TO BE INSTALLED
-%
-%
-%Create a directory to hold our code:
-%\begin{lstlisting}[]
-%$ CODEDIR="${HOME}/code"
-%$ mkdir -p ${CODEDIR}
-%$ cd ${CODEDIR}
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-%
-%\begin{lstlisting}[]
-%$ git clone git@github.com:steinwurf/kodo-python.git
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-%
-%\begin{lstlisting}[]
-%$ cd kodo-python
-%$ python2.7 waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
-%$ python waf build
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-%
-%KODO-PYTHON IS NOT WORKING!! WAITING FOR HELP ON STEINWURF DEV.
-%
-%WRITE HOW TO RUN A KODO PYTHON SCRIPT IN THE RASPIS
-
-%\texttt{hello\_world} returns immediately. This is not always the case.
-%Sometimes it is desired to run simulations that should run for days.
-%A software package called \texttt{screen} can be used for this purpose.
-%Simply run ...

### 2016-07-12 cross_compilation.tex
Toolchain description done

@@
 administrator \ac{PC} in order to save time and computational resources.
-Hence, we give an example of compiling a simple C++ program to run
+Hence, we give an example of compiling a simple C++ program and
+copying the generated binaries with \ac{SSH} to run locally
 on a \ac{Raspi}.
@@
 
-%The toolchain is mandatory on most consumer \ac{PC}s due to the different
-%processor architecture.
-%
-%Furthermore, we will cover how to cross compile a simple C++ example and
-%Kodo-cpp as well as how the binaries can be copied from a \ac{PC} to a
-%\ac{Raspi} and executed.
+\subsection{Toolchain Setup}
+To compile in a given architecture that is aimed for a different one,
+the testbed administrator needs to install a toolchain in its \ac{PC}.
+The toolchain is mandatory due to the different processor architectures
+where the source can be compiled from / to.
 
-\subsection{Setup toolchain on PC}
+Given that compiling a toolchaing can be an ardous task, we get the toolchain
+recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
+This toolchain is available from \cite{2016rpitoolchain} and
+it already contains the binaries for different compilers based on
+\texttt{gcc 4.9}. In what follows, we extract the binaries adjusting
+them to our coding style and compiling convention. For this, we use the
+\texttt{\$TOOLCHAIN} directory as the working directory. However,
+the testbed administrator may choose some other working directory of
+its preference if desired. First, we create the toolchain directory:
 
-%Besides the previous description (\textbf{Include compiling Kodo from the
-%RasPi from the scratch}), the testbed administrator can compile Kodo in its
-%personal workstation and transfer the generated binaries directly to
-%a path in the \ac{Raspi}. To achieve this, we get a toolchain that
-%contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
-%for the \ac{Raspi}. Therefore, we strongly recommend any testbed
-%administrator to do the following procedure. In what follows, we provide
-%the instructions considering that the NFS server uses the \texttt{\$HOME}
-%directory as the working directory. However, the administrator may choose
-%some other working directory of its preference if desired.
-
-\begin{enumerate}
-
-\item Create toolchain directory:
 \begin{lstlisting}[]
-$ TOOLCHAINDIR="${HOME}/toolchains"
+$ export TOOLCHAINDIR="${HOME}/toolchains"
 $ mkdir -p $TOOLCHAINDIR
@@
 
-\item Download the \ac{Raspi} toolchain for 64-bit Linux from.
-A more recent \ac{Raspi} toolchain may be available at
-\texttt{http://buildbot.steinwurf.dk/toolchains/linux/}:
+Later, we download the \ac{Raspi} toolchain for 64-bit Linux. However,
+instead of working with the binaries from \cite{2016rpitoolchain},
+we use a zipped version with the original binaries and renamed symbolic
+links to the original files available in
+\cite{2016steinwurfrpitoolchain}. Finally, we unzip the downloaded file.
+This is made as follows:
 
 \begin{lstlisting}[]
-$ TOOLCHAIN="raspberry-gxx49-arm"
-$ wget http://kom.aau.dk/project/TuneSCode/raspi/${TOOLCHAIN}.zip
+$ wget http://files.steinwurf.com/toolchains/linux/raspberry-gcc-4.9.3/raspberry-gxx493-arm.zip
+$ unzip raspberry-gxx493-arm.zip
 \end{lstlisting}
@@
 
-\item Extract the downloaded file:
-%locally in the NFS server. After
-%this operation, there should be a new directory for the toolchain
-%in the server. \\
+Instead of calling the ARM cross compiler using its full path, we make
+the binaries accessible from the command shell systemwide. A way to do
+this is by adding the following commands in the \texttt{\${HOME}/.profile}
+as follows:
+
 \begin{lstlisting}[]
-$ unzip ${TOOLCHAIN}.zip
+$ sed -i '$a export TOOLCHAINDIR=\"$HOME/toolchains\"' ${HOME}/.profile
+$ sed -i '$a export TOOLCHAINBINARY=\"raspberry-gxx49-arm-g++\"' ${HOME}/.profile
+$ sed -i '$a PATH=\"\$PATH:${TOOLCHAINDIR}/arm-rpi-4.9.3-linux-gnueabihf/bin\"' ${HOME}/.profile
 \end{lstlisting}
@@
 
-\item (Optional) Delete zip file:
-%locally in the NFS server. After
-%this operation, there should be a new directory for the toolchain
-%in the server. \\
-\begin{lstlisting}[]
-$ rm ${TOOLCHAIN}.zip
+This helps the \ac{OS} to recognize the location of the compiler
+command everytime a new shell is opened. The \texttt{.profile} should now
+contain the lines we inserted. There might be more other
+code in your file. Leave other parts unmodified.
+% MAC and Hostname file
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$HOME/.profile}@>
+<@\textcolor{gray}{
+---------------------------}
+\Reactivatenumber @>
+...
+export TOOLCHAINDIR="$HOME/toolchains"
+export TOOLCHAINBINARY="raspberry-gxx49-arm-g++"
+PATH="$PATH:${TOOLCHAINDIR}/arm-rpi-4.9.3-linux-gnueabihf/bin"
+...
 \end{lstlisting}
@@
 
-\item We can now verify that the ARM cross compiler is working:
-
-\begin{lstlisting}[]
-$ ${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++ --version
-raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
-Copyright (C) 2014 Free Software Foundation, Inc.
-This is free software; see the source for copying conditions.  There is NO
-warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-%\begin{lstlisting}[]
-%$ echo "${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++"
-%/home/<USER>/toolchains/raspberry-gxx49-arm/bin/raspberry-gxx49-arm-g++
-%\end{lstlisting}
-%\FloatBarrier
-%\vspace{-5mm}
-
-%    The ARM cross compilers should now be located in
-
-\item Make toolchain binaries available systemwide. Instead of calling
-the ARM cross compiler using its full path, we can make it accessible
-from the command shell systemwide. One way to do this is by adding the
-toolchain binaries directory to the Linux environment variable \texttt{PATH}
-when the \ac{OS} starts up:
-
-%\item Add toolchain binaries to \texttt{PATH}. Instead specifying the full
-%path to the toolchain binaries we can instead tell the operation system
-%where to search for it. This makes the toolchain binaries available
-%systemwide.
-
-\begin{lstlisting}[]
-$ echo PATH=\"\$PATH:${TOOLCHAINDIR}/${TOOLCHAIN}/bin\" >> ${HOME}/.profile
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-%$ printf 'PATH="%s/%s/bin"\n' "${TOOLCHAINDIR}" "${TOOLCHAIN}" >> ${HOME}/.profile
-
-%\item Add the \texttt{bin} folder of the toolchain to the \texttt{PATH}
-%Linux environment variable of the server. This will help the server OS
-%to recognize the location of the compiler command, which will be needed
-%later. To do so, edit the \texttt{\$HOME/.profile} to add in a newline:
-%\texttt{PATH="\$PATH:\$HOME/raspberry-gxx49-arm/bin"}. Save the
-%\texttt{\$HOME/.profile}. \\
-
-\texttt{.profile} should now contain the line we inserted. There may be more
-code in you file.
-% MAC and Hostname file
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$HOME/.profile}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-PATH="$PATH:/home/<USERNAME>/toolchains/raspberry-gxx49-arm/bin"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Update \texttt{PATH}. Source \texttt{.profile} to make the changes
-take effect in your system:
+To update the \texttt{PATH} variable and the \texttt{.profile}, we use the
+\texttt{source} command for the changes take effect in the administrator
+system:
 \begin{lstlisting}[]
@@
 
-\item ARM cross compiler should now be available systemwide:
+A working ARM cross-compiler in the testbed administrator \ac{PC} should
+output the following:
 
 \begin{lstlisting}[]
-$ raspberry-gxx49-arm-g++ --version
+$ ${TOOLCHAINBINARY} --version
+raspberry-gxx49-arm-g++ (crosstool-NG crosstool-ng-1.22.0-88-g8460611) 4.9.3
+Copyright (C) 2015 Free Software Foundation, Inc.
+This is free software; see the source for copying conditions.  There is NO
+warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
 \end{lstlisting}
@@
 \vspace{-5mm}
-
-%\item Restart the server session in order for the changes made in the
-%previous step take effect. To verify this, open a new terminal and type:
-%\texttt{raspberry-gxx49-arm-g++ --version}. A correct binary installation
-%should return an output similar to:
-%
-%\texttt{raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
-%Copyright (C) 2014 Free Software Foundation, Inc.
-%This is free software; see the source for copying conditions.  There is NO
-%warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.} \\
-
-\end{enumerate}
 

### 2016-07-12 cross_compilation.tex
Working on kodo cross-compiling

@@
 
-\subsection{Cross compile example}
-
-The following example will provide a very basic example of 1) how to cross compile
-a \texttt{hello\_world} c++ example to the \ac{Raspi}'s ARM architecture
-and 2) how to copy and execute the binary in a \ac{Raspi} using \ac{SCP} and
-\ac{SSH}.
-
-Create the file \texttt{hello\_world.cpp} with the following content using your
-favorite text editor:
+\subsection{Cross-compile Example}
+The following shows: (i) How to cross compile the classic \texttt{hello\_world}
+C++ example for the \ac{Raspi} ARM architecture and (ii) how to copy and
+execute the binary in a \ac{Raspi} using \ac{SCP} and \ac{SSH}. First, we
+create the file \texttt{hello\_world.cpp}. For simplicity, we create it
+in the directory where we stored the \texttt{fabfile.py} file
+with the following content using any text editor:
 
@@
 <@\textcolor{gray}{
----------------------------------------------------------------}
+-----------------------------------------------}
 \Reactivatenumber @>
@@
 \FloatBarrier
+\vspace{-5mm}
 
-Save \texttt{hello\_world.cpp} and compile it for \ac{Raspi}:
+We save the previous file and compile it for \ac{Raspi} in the testbed
+administrator \ac{PC} by doing:
 
 \begin{lstlisting}[]
-$ raspberry-gxx49-arm-g++ hello_world.cpp -o hello_world
+$ ${TOOLCHAINBINARY} hello_world.cpp -o hello_world
 \end{lstlisting}
@@
 
-This should produce a binary \texttt{hello\_world}. Copy it to a \ac{Raspi}
-and test it. We will use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s.
-
-The default username for our Raspbian lite distribution is "pi" and password is "raspberry"
+This should produce a binary \texttt{hello\_world} that should only be
+useful in the \ac{Raspi}. We copy it to a \ac{Raspi} to test it. We will
+use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s. Still, we
+for example might use Fabric if we are interested in deploying a compiled
+binary for many \ac{Raspi}s.
 
 \begin{lstlisting}[]
-$ scp main pi:<RASP_IP>:~/
+$ scp hello_world pi@<RASP_IP>:~/
 \end{lstlisting}
@@
 
-If you do not know the \ac{IP} address of a \ac{Raspi} in your network, you can
-connect a monitor to it and run the following command after you have logged in:
-\begin{lstlisting}[]
-pi@rasp01:~ $ ifconfig
-eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
-          inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
-          inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
-          UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
-          RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
-          TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
-          collisions:0 txqueuelen:1000
-          RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)
+% If you do not know the \ac{IP} address of a \ac{Raspi} in your network, you can
+% connect a monitor to it and run the following command after you have logged in:
+% \begin{lstlisting}[]
+% pi@rasp01:~ $ ifconfig
+% eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
+%           inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
+%           inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
+%           UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
+%           RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
+%           TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
+%           collisions:0 txqueuelen:1000
+%           RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)
 
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
+% \end{lstlisting}
+% \FloatBarrier
+% \vspace{-5mm}
 
-After the executable has been copied to a \ac{Raspi}. Then, \ac{SSH} to it:
+After the executable has been copied to the \ac{Raspi}. We login through
+\ac{SSH} to it:
 
@@
 We can list the directory content after we have logged into the \ac{Raspi} and
-see that \texttt{hello\_world} is there:
+verify that the compiled \texttt{hello\_world} binary is there:
 
 \begin{lstlisting}[]
-pi@rasp07:~ $ ls
+pi@<RASP_IP>:~ $ ls
 hello_world  rasp_config
@@
 
-Now, simply execute the \texttt{hello\_world} to confirm that cross compiling
-\texttt{hello\_world} worked properly:
+Finally, we simply execute the \texttt{hello\_world} to confirm that
+the cross-compiling of \texttt{hello\_world} worked properly:
 
 \begin{lstlisting}[]
-pi@rasp07:~ $ ./hello_world
+pi@<RASP_IP>:~ $ ./hello_world
 Hello World!
@@
 
-\subsection{Cross compile Kodo}
+\subsection{Cross-compile Kodo}
 
-Kodo is a C++11 network coding library. It is available through bindings in
-a variety of programming languages.
-%There exists a number of bindings for
-%Kodo to other popular programming languages.
-This procedure will present how to configure the Kodo C++ bindings (Kodo-cpp)
-to cross compile applications that can run in \ac{Raspi}. More details are
-provided in the code documentation~\cite{kodocppdoc}.
+As we originally mentioned, Kodo is a C++11 network coding library that
+permits to implement network coding functionalities by allowing any network
+protocol designer to use and test the primitive encoding, decoding and
+recoding operations of \ac{RLNC}. In this way, a designer only needs to
+focus on the design and test of a network coding-based protocol. Kodo is
+available through programming bindings for a variety of popular
+programming languages. This procedure will present how to configure the
+Kodo C++ bindings \texttt{kodo-cpp} to cross-compile applications that can
+run in \ac{Raspi}. \texttt{kodo-cpp} provides a simple interface to the
+underlying C++11 code that exists in the libraries \texttt{kodo-core} for
+the object structure and \texttt{kodo-rlnc} for the \ac{RLNC} codec
+implementation. More details about Kodo are provided in the code
+documentation~\cite{kodocppdoc}.
 
-%This procedure will go through
-%how to setup Kodo-cpp to cross compile applications that can run in \ac{Raspi}. Kodo-cpp provides a
-%simple interface to the underlying C++11 code that exists in the libraries
-%Kodo-core and Kodo-rlnc.
+To use Kodo for research, it is required to obtain a research free license.
+To do this, only a request form needs to be filled
+in~\cite{steinwurflicenselink} and wait for grant access. Once
+the access for Kodo has been granted, the source code can be pulled from
+its Git repositories to be compiled. Thus, assuming that testbed
+administrator already has access, we clone the Kodo repository locally
+in \texttt{\${CODEDIR}} and change directory into the repository by doing:
 
-To use Kodo, it is required to obtain a research- and
-education-friendly license.
-It can be be obtained from~\cite{steinwurflicenselink} and is free
-for research and educational purposes.
-
-%A more detailed procedure of setting up Kodo-cpp than provided below is available at
-
-\begin{enumerate}
-
-\item Start by creating a directory to hold the code:
 \begin{lstlisting}[]
-$ CODEDIR="${HOME}/code"
-$ mkdir -p ${CODEDIR}
 $ cd ${CODEDIR}
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-\item Clone the Kodo repository and change directory into the repository:
-\begin{lstlisting}[]
-$ git clone https://github.com/steinwurf/kodo-cpp.git
+$ git clone git@github.com:steinwurf/kodo-cpp.git
 $ cd kodo-cpp
@@
 
-\item Configure Kodo to build executables for the \ac{ARM} architecture using the \ac{Raspi} toolchain:
+Configure Kodo to build executables for the \ac{ARM} architecture using
+the \ac{Raspi} toolchain:
 \begin{lstlisting}[]
@@
 ...
-'configure' finished successfully (0.620s)
+'configure' finished successfully (X.XXXs)
 \end{lstlisting}
@@
 
-\item Build executables:
+Build executables:
 \begin{lstlisting}[]
-$ python waf build
+$ python waf build --run_tests
 ...
@@
 
-\item Make shared library:
+Make shared library:
 \begin{lstlisting}[]
@@
 
-\item Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
+Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
 \begin{lstlisting}[]
@@
 
-\item Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
+Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
 \begin{lstlisting}[]
@@
 
-\item Execute binary:
+Execute binary:
 \begin{lstlisting}[]
@@
 %by using the \texttt{scp} command during the configuration step.
-
-\end{enumerate}
 

### 2016-07-12 conclusions.tex
Only appendix remaining!

@@
 \label{sec:conclusions}
-Obsersving the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enable reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of the system.
+Observing the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enables reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system.
 The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges.

### 2016-07-12 cross_compilation.tex
Only appendix remaining!

@@
 
-Configure Kodo to build executables for the \ac{ARM} architecture using
-the \ac{Raspi} toolchain:
+We first configure \texttt{kodo-cpp} to build executables for the
+\ac{ARM} architecture using the \ac{Raspi} toolchain and later build them
+by running:
 \begin{lstlisting}[]
@@
 'configure' finished successfully (X.XXXs)
+$ python waf build
+...
+'build' finished successfully (XmXX.XXs)
 \end{lstlisting}
@@
 
-Build executables:
+If the configuration and build steps are successful, the binaries should
+have been created. To be able to use them, we need to create a shared library
+that we will use in the \ac{Raspi}. To do this, we run the following command:
+
 \begin{lstlisting}[]
-$ python waf build --run_tests
+$ python waf install --install_shared_libs --install_path="./shared_test"
 ...
-'build' finished successfully (2m22.918s)
+'install' finished successfully (X.XXXs)
 \end{lstlisting}
@@
 
-Make shared library:
-\begin{lstlisting}[]
-$ python waf install --install_shared_libs --install_path="./shared_test"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
+Now, we copy the shared library, binary files and related headers to the
+\ac{Raspi} home directory as follows:
 
-Copy shared library to a \ac{Raspi}'s home directory (Alternatively, Kodo can also generate static libraries):
 \begin{lstlisting}[]
@@
 
-Copy a binary using Kodo shared library to the \ac{Raspi}'s home directory:
+Alternatively and for the testbed administrator reference, Kodo can also
+generate static libraries. We log in to the \ac{Raspi} and execute the unit
+tests and one of the binaries by running:
+
 \begin{lstlisting}[]
-$ scp -r shared_test/encode_decode_simple pi@<RASP_IP>:~/
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Execute binary:
-\begin{lstlisting}[]
-$ ssh pi@<RASP_IP> ./encode_decode_simple
+$ ssh pi@<RASP_IP>
+$ ./kodocpp_tests
+...
+[----------] Global test environment tear-down
+[==========] 17 tests from 15 test cases ran. (XXXX ms total)
+[  PASSED  ] 17 tests.
+$ ./encode_decode_simple
 Data decoded correctly
@@
 
-Cross compiling Kodo applications works provided that above command
-returned "Data decoded correctly".
-
-%\item Navigate to the repository and configure \texttt{waf} by typing:
-%\texttt{python config.py} and select the 16th ``make specification'' file
-%for the \ac{Raspi}, e.g. option \texttt{[16]cxx\_raspberry\_gxx49\_arm}
-%presented by the file.
-%
-%This command configures \texttt{waf} to use the proper compiler and its
-%required flags to generate the binaries for the \ac{Raspi}. If the
-%configuration was correct, the output will indicate:
-%\texttt{'configure' finished successfully (X.XXXs)}, where \texttt{X.XXX}
-%is total time in seconds for configuring the project in the server. \\
-%
-%\item Execute \texttt{python waf build}. If the build process was
-%successful, the generated binaries for the \ac{Raspi} should be located
-%in \texttt{build/cxx\_raspberry\_gxx49\_arm} in the Kodo repository.
-%\textbf{Indicate how the binary files should look like}.
-%
-%Once this procedure is made, the testbed administrator can relocate the
-%generated binary files to the \ac{Raspi}s through the network as desired
-%by using the \texttt{scp} command during the configuration step.
-
-%\subsection{Kodo Builds for the \ac{Raspi}, Platform Support and Documentation \textcolor{red}{(WHAT IS THE PURPOSE OF THIS SUBSECTION?)}}
-%
-%You can check the build status of Kodo, Fifi and other relevant projects
-%through their respective repository master branch on our buildbot page
-%\cite{steinwurf2016buildbot}. Our buildbot displays the status of the builds
-%for Raspbian 8 and GCC 4.9 for the ARM architecture which is the relevant one
-%for the \ac{Raspi}. At the link, you can check build status and build
-%statistics. Also, documentation about Kodo basics with a tutorial is available
-%at \cite{kododocs}.
+If the Kodo cross-compilation worked properly, both the unit tests and
+binary run should provide the shown outputs.
 
@@
 %Kodo-python is available at \url{https://github.com/steinwurf/kodo-python}.
-%
-%
-%
 %\begin{lstlisting}[]

### 2016-07-13 appendix.tex
HTTP ongoing work

@@
-\section{Advanced Setup: HTTP Image Fetching}
+\section{HTTP Image Fetch Setup (Optional)}
 \label{sec:http}
@@
 is not able to update old packages in its Linux distribution for performing
-the commands, particularly the ones regarding the \texttt{util-linux} package.
+the commands, particularly the ones regarding the \texttt{util-linux} package
+or if a command just fails.
 
@@
 
+\subsection{Image Chroot with Proot}
+\label{sec:chroot}
+
+In some cases, \texttt{chroot} may not work properly. In this case, an
+alternative can be \texttt{proot}. It might be required to install it
+before changing the root with \texttt{apt-get install proot}. Then, it
+should be possible for the testbed administrator to run the following:
+
+% CHROOT to OS image
+\begin{lstlisting}[]
+# proot -q qemu-arm-static -S ${ROOTDIR}
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
 \subsection{Umount after Image Chroot}
@@
 \vspace{-5mm}
+
+\subsection{Paramiko Package Update}
+\label{sec:paramiko}
+In case Paramiko version 1.15.1 or higher is not installed, we may observe
+a key exchange algorithm error. In this case, it might be
+necessary to update the Paramiko package and verify this by running:
+
+\begin{lstlisting}[]
+# pip install --upgrade paramiko
+...
+Sucesfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
+# pip show paramiko
+...
+Name: paramiko
+Version: 2.0.1
+...
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+This should allow to make \ac{SSH} connections with Fabric by using the proper
+key exchange algorithms through Open SSH.

### 2016-07-13 conclusions.tex
HTTP ongoing work

@@
 Observing the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enables reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system.
-The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges.
-The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way.
+The description shows how to set up interconnected \ac{Raspi}s with SD memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges.
+We hope this work permits researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way.
 Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

### 2016-07-13 cross_compilation.tex
HTTP ongoing work

@@
 its Git repositories to be compiled. Thus, assuming that testbed
-administrator already has access, we clone the Kodo repository locally
-in \texttt{\${CODEDIR}} and change directory into the repository by doing:
+administrator already has access, we clone the \texttt{kodo-cpp}
+repository locally in \texttt{\${CODEDIR}} and change directory into
+the repository by doing:
 

### 2016-07-13 image_setup.tex
HTTP ongoing work

@@
 The output provides relevant information about the image. The image is in
-total 2709504 sectors (1.3GiB) in size and contains two partitions. The
+total 2709504 sectors (1.3 GiB) in size and contains two partitions. The
 first partition starts at sector 8192 and the other partition starts at
@@
 Given that we want to customize the files stored in the \ac{Raspi}s,
-we need to resize the image file since 1.2GB might not be enough to store
+we need to resize the image file since 1.2 GB might not be enough to store
 the root filesystem due to the total size of the additional packages. Thus,
 we need to increase the partition size. The following procedure illustrates
-how the image and its root filesystem can be expanded by one \ac{GB}.
-First, to expand the image one \ac{GB}, we execute:
+how the image and its root filesystem can be expanded by one GB.
+First, to expand the image one GB, we execute:
 
@@
 Later, we use \texttt{fdisk} with the same arguments as before to see that
-the image is now one \ac{GB} larger:
+the image is now one GB larger:
 \begin{lstlisting}[literate={DOLLAR}{\$}1]
@@
 root filesystem, it is required to first remove the Linux partition
-and then add it again with one \ac{GB} more. To do this, we make use
+and then add it again with one GB more. To do this, we make use
 again of \texttt{fdisk} in command mode to alter the partition table.
@@
 %\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
-\subsection{Image Package Updating}
+\subsection{Image Packages Updating by Changing the Apparent Root Directory}
 Besides adding and configuring files within the image, the testbed
@@
 
-% In some cases, \texttt{chroot} may not work properly. tgus, in this case, an
-% alternative can be \texttt{proot}. It may be required to install it
-% before changing the root with \texttt{apt-get install proot}. Then, it
-% should possible for the testbed administrator to run the following:
-
-% % CHROOT to OS image
-% \begin{lstlisting}[]
-% # proot -q qemu-arm-static -S ${ROOTDIR}
-% \end{lstlisting}
-% \FloatBarrier
-% \vspace{-5mm}
 If successfully executed, our terminal should have changed the prompt
 indicating that we are the root user in the Raspbian lite root filesystem as
-the apparent root. To be aware of the mode that we are working now, we change
+the apparent root. In case the \texttt{chroot} command is not successful,
+we provide an alternative command in Section~\ref{sec:chroot} of the Appendix.
+To be aware of the mode that we are working now, we change
 the prompt title to indicate that it is a \texttt{chroot} environment as

### 2016-07-13 testbed_http.tex
Changing Chapter 4 structure
HTTP ongoing work

@@
 \label{sec:testbed_http}
+Creating manual copies of the customized image for each memory card can
+be also a time-consuming task that is prone to errors and become
+cumbersome. Therefore, we present an optional more advanced \ac{HTTP}
+image fetching setup for the existing image to be downloaded from a
+\ac{HTTP} server. We consider this an optional step since a testbed
+administrator might opt for not including this as part of his / her setup.
+We start our description consideri
 
-\subsection{HTTP}
+At the end of Section~\ref{sec:overlay_fs}, the final customized image was
+proposed to be splitted into three partitions. One for the boot files, one
+containing the root file system and one for the home directory. This is made
+considering an overlay filesystem approach. In this section, instead of
+writing the image direclty to the memory cards, we present how the
+partitions can be extracted from the image, and how the boot partition can
+be configured to download the root filesystem and home directory partitions
+from a \ac{HTTP} server during startup. The advantage of this setup is that
+we completely automate the image fetch which significantly eases its
+maintainance since any changes are required to be made only in the \ac{HTTP}
+server image. We start by declaring the variables for the path names of each
+partition:
 
-%This part will explain how to prepare the existing image to be downloaded from
-%a \ac{HTTP} server.
-The Linux image that has been made so far has two partitions. One 
-partition with the boot files and one containing the root file system.
-
-This section will present how the partitions can be extracted from the image,
-and how the boot partition can be configured to download the root filesystem
-from a \ac{HTTP} server during startup.
-
-The advantage of this setup is that multiple \ac{Raspi}s can share the same
-root image. This significantly eases the maintainance task.
-
-%An alternative to the setup explained in this procedure is to use \ac{NFS}
-%instead of \ac{HTTP}.
-
-%The first program called when a \ac{Raspi} starts is the Boot Loader,
-%\texttt{U-Boot}. \texttt{U-Boot} 
-%
-%When Raspbian starts up, the first program called
-%
-%The procedure explained below will customize the initial \ac{RAM} filesystem
-%
-%
-%there exists only one root file
-%
-%The advantage of this is that the root filesystem only one all \ac{Raspi}s can automatically download the 
-%
-%The boot partition will be required in the \ac{Raspi}s, but the root image can
-
-Start by declaring the path and names of the boot- and root-image:. 
 \begin{lstlisting}[]
@@
 $ export HTTPROOTIMAGE="${WORKDIR}/httproot"
+$ export HTTPHOMEIMAGE="${WORKDIR}/httphome"
 \end{lstlisting}
@@
 
-The, setup a loopback device of the existing image and use \texttt{dd} to
+Then, we setup a loopback device of the existing image and use \texttt{dd} to
 extract the boot and root partitions from the image:
@@
 In a previous steps, the boot loader was configured to call an initial \ac{RAM}
-filesystem. The coming steps will present how to the initial \ac{RAM}
+filesystem. Now, we present how to the initial \ac{RAM}
 filesystem can be extended further to download and cache the root filesystem

### 2016-07-13 tools.tex
HTTP ongoing work

@@
 executing a task with arguments is in the form
-\texttt{fab <TASK>:arg1,arg2,...}. Some executions from the terminal
-are shown as follows \todot{Test the following commands later and print the
-outputs}:
+\texttt{fab <TASK>:arg1,arg2,...}. In what follows, we denote the
+\ac{IP} address of a generic \ac{Raspi} for test as \texttt{<RASP\_IP>}.
+The executions from the terminal of some of these commands are shown as
+follows:
 
@@
 $ fab reboot
+[<RASP_IP>] Executing task 'reboot'
+
+Done.
+Disconnecting from <RASP_IP>... done.
 $ fab install:tmux
+[<RASP_IP>] Executing task 'install'
+...
+The following NEW packages will be installed:
+  tmux
+0 upgraded, 1 newly installed, 0 to remove and 0 not upgraded.
+...
+Preparing to unpack .../archives/tmux_1.9-6_armhf.deb ...
+Unpacking tmux (1.9-6) ...
+Processing triggers for man-db (2.7.0.2-5) ...
+Setting up tmux (1.9-6) ...
+
+Done.
+Disconnecting from <RASP_IP>... done.
 \end{lstlisting}
@@
 The first function above reboots the \ac{Raspi}s in the lists of hosts and
-the second function installs a program given by an argument. In both cases,
-after a successful \ac{SSH} connection is made, Fabric employs the
-\ac{Raspi}'s \texttt{reboot} and \texttt{apt-get} commands in \texttt{sudo}
-mode to do the required tasks. Below it is shown an example for
-the \texttt{push} task which uses two arguments. Here we copy
-\texttt{my\_file} from the testbed administrator \ac{PC} to all the host
-\ac{Raspi}s:
+the second function installs a program given by an argument. For the
+connection to the devices, Fabric calls the Paramiko module from Python
+to make a \ac{SSH} connection. For this to work properly, the Paramiko
+version needs to be higher than or equal to 1.15.1. If not available,
+the \ac{SSH} connections from Fabric may fail. Thus in case of any problems,
+some instructions for updating the Paramiko package are available in
+Section~\ref{sec:paramiko} of the Appendix. This is a standard recommendation
+the Fabric troubleshooting guide~\cite{2016fabricsupport}.
+
+After a successful \ac{SSH}
+connection is made, in the previous two commands towards the \ac{Raspi},
+Fabric employs the \ac{Raspi}'s \texttt{reboot} and \texttt{apt-get}
+commands in \texttt{sudo} mode to do the required tasks. Below it is shown
+an example for the \texttt{push} task which uses two arguments. Here we
+copy \texttt{my\_file} from the testbed administrator \ac{PC} to a test
+host \ac{Raspi}:
 
@@
 $ fab push:"${CODEDIR}/my_file",'~/'
-[192.168.87.107] Executing task 'push'
-[192.168.87.107] put: /home/<USER>/code/my_file -> /home/pi/my_file
+[<RASP_IP>] Executing task 'push'
+[<RASP_IP>] put: /home/<USER>/code/my_file -> /home/pi/my_file
 
 Done.
-Disconnecting from 192.168.87.107... done.
+Disconnecting from <RASP_IP>... done.
 \end{lstlisting}
@@
 
-Fabric has many other functionalities that are useful in
-controlling a large set of \ac{Raspi}s. For example, we may extract files
-or run automated experiments. The included functionalities in
-the \texttt{\$CODEDIR/fabfile.py} file will depend on the requirements
-of the testbed administrator.
+To control a large set of devices, we simply need to include them in the
+\texttt{env.hosts} list in the \texttt{\$CODEDIR/fabfile.py} file. Fabric has
+many other functionalities that are useful in controlling a large set of
+\ac{Raspi}s. For example, we may extract files or run automated
+experiments. The included functionalities in the
+\texttt{\$CODEDIR/fabfile.py} file will depend on the requirements of the
+testbed administrator.
 
@@
 as desired. The following procedure presents how to use \texttt{screen} with
-\ac{SSH} to: (i) Login to a generic \ac{Raspi} with \ac{IP} address as
-\texttt{<RASP\_IP>}, (ii) open a screen session, (iii)
-execute an example command, (iv) detach from the screen session, (v)
+\ac{SSH} to: (i) Login to a generic \ac{Raspi}, (ii) open a screen session,
+(iii) execute an example command, (iv) detach from the screen session, (v)
 terminate the \ac{SSH} connection, (vi) login to the \ac{Raspi} again

### 2016-07-15 testbed_http.tex
provided more text about http method

@@
 $ export HTTPROOTIMAGE="${WORKDIR}/httproot"
-$ export HTTPHOMEIMAGE="${WORKDIR}/httphome"
 \end{lstlisting}
@@
 The initial \ac{RAM} filesystem contains an \texttt{init} script that
-sets up the prepares the yis first
-called. This script initializes the  may be rewritten, but
+is first called to initializes
+\texttt{initramfs}, parse arguments specified in \texttt{cmdline.txt}, and
+call a specialized script that prepares the actual root filesystem. By default,
+Raspbian includes scripts to mount the root filesystem from 1) a block devices
+(i.e. the memory card) and 2) from a \ac{NFS} shared folder. We make a third
+script to mount the root filesystem from an image
+(i.e. the root image \texttt{httproot.img}) that is downloaded from a
+\ac{HTTP} server.
 
-Create the following file (THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT)
+The initramfs script to download and mount the image from a \ac{HTTP} server
+will require a few additional arguments. E.g. to specify from which \ac{HTTP}
+server to download the root filesystem image from.
+%This script that we will create will require a few additional arguments such
+%as which \ac{HTTP} server the image should be downloaded from.
+These arguments
+may be specified in \texttt{cmdline.txt}.
+%and later parsed within the \texttt{initramfs} script.
+
+Keep the content in \texttt{cmdline.txt}, but add/modify the variables
+\texttt{boot}, \texttt{ip}, \texttt{image}, \texttt{root}, and \texttt{url}.
+
+%Keep also all other variables as they are in
+%\texttt{cmdline.txt}.
+
+%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
+\textcolor{red}{(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)}
+\Suppressnumber\begin{lstlisting}[]
+<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
+<@\textcolor{gray}{
+---------------------------------------------------------------}
+\Reactivatenumber @>
+<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
+    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+% ip=dhcp root=/dev/http http-options=hard,intr,ro
+
+Here is how the variables will be used in the \texttt{initramfs}:
+\begin{itemize}
+    \item boot=http: tells the init script in initramfs to call the \texttt{http} script that will be presented below
+    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the \texttt{initramfs}
+    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
+    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
+    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
+\end{itemize}
+
+%The \ac{URL} should be modified to reflect where
+%the \ac{Raspi}s should download the http root image. The \texttt{image} variable informs the
+%\texttt{http} initramfs script to download and cache the http root image in the second partition
+%on its memory card at path \texttt{/httproot.img}.
+
+The specialized script that we made to download and mount the root filesystem image from a \ac{HTTP} server should be created in the root filesystem as presented below:
+%Create the following file
+\textcolor{red}{(THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT. I'M NOT EVEN SURE IT IS POSSIBLE TO MAKE THE IMAGE READ-WRITABLE)}
 \Suppressnumber\begin{lstlisting}[]
@@
     mount -r -n -t ext4 -o nodiratime,noatime /dev/loop0 /root
-
-    # Bind /root/home to /home
-    #mount -n -o rbind /home /root/home
 }
@@
 
-Keep the content in \texttt{cmdline.txt}, but add/modify the variables
-\texttt{boot}, \texttt{ip}, \texttt{image}, \texttt{root}, and \texttt{url}.
+The variable (boot=http) that we specified in \texttt{cmdline.txt} tells the
+\texttt{init} script to call the script above rather than the default scripts
+that mounts the root filesystem from a block device or a \ac{NFS} shared
+folder. The \texttt{init} script will call the function \texttt{http\_mount\_root}
+which in this case can be consired as the \texttt{main} function.
 
-%Keep also all other variables as they are in
-%\texttt{cmdline.txt}.
+As in any other shell scripts, this means that code in the global namespace
+will be executed by the interpretor before the function \texttt{http\_mount\_root}
+is entered.
+%Due to the variable that we specified in \texttt{cmdline.txt}, 
+Hence, line 9 will be the first command executed. This line creates a directory
+in \texttt{initramfs} named
+\texttt{imagedir} that we will use to hold the root image.
 
-%Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
-(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)
-\Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{HTTPROOTDIR\}/boot/cmdline.txt}@>
-<@\textcolor{gray}{
----------------------------------------------------------------}
-\Reactivatenumber @>
-<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
-    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-% ip=dhcp root=/dev/http http-options=hard,intr,ro
+Next, line 12 to line 44 will parse the additional arguments that we provided in
+\texttt{cmdline.txt}.
 
-Here is how the variables are used in the initramfs:
-\begin{itemize}
-    \item boot=http: tells the init script in initramfs to call the \texttt{http} script presented above
-    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the initial ram filesystem
-    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
-    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
-    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
-\end{itemize}
-
-%The \ac{URL} should be modified to reflect where
-%the \ac{Raspi}s should download the http root image. The \texttt{image} variable informs the
-%\texttt{http} initramfs script to download and cache the http root image in the second partition
-%on its memory card at path \texttt{/httproot.img}.
+The \ac{Raspi}s does not have enough \ac{RAM} to hold entire root
+filesystem, so we instead download it to the partition that, as specified in
+\texttt{fstab} later, will later be mounted as the \texttt{HOME} directory
+when the system starts. This will enable the \ac{Raspi}s to cache the root image
+and use the previously downloaded root image again after reboots (see the
+function \texttt{http\_fetch\_image} starting from line 56)\fxnote{is there a better way to reference to a function within the script?
+This reference will be invalid if the script is slightly modified}.
 

### 2016-07-15 cross_compilation.tex
Not updated commits

@@
 The toolchain is mandatory due to the different processor architectures
-where the source can be compiled from / to.
-
+where the source can be compiled from.
 Given that compiling a toolchaing can be an ardous task, we get the toolchain
@@
 command everytime a new shell is opened. The \texttt{.profile} should now
-contain the lines we inserted. There might be more other
-code in your file. Leave other parts unmodified.
+contain the lines we inserted. There might be other
+code in the file of other testbed administrators. We recommend to leave
+other parts unmodified.
 % MAC and Hostname file

### 2016-07-15 overlay_fs.tex
HTTP ongoing work
Not updated commits

@@
-\label{sec:ovelay_fs}
+\label{sec:overlay_fs}
 In principle, our procedure modifies the image file only once in
@@
 this image in the \ac{Raspi}s provides the same initial system for all
-the devices. This system is later distributed locally to a memory card
-in each device. However, if we do not make any further modifications during
+the devices. If we do not make any further modifications during
 the image setup, any files created after the initial boot of a \ac{Raspi}
@@
 
-Therefor, if both persistent and non-persistent storage are required for
+If both persistent and non-persistent storage are required for
 the \ac{Raspi}s, here we present the steps to setup an overlay filesystem.
@@
 it will simpy return the file. This idea can be used in our setup to mount
-the root filesystem (i.e. Raspbian lite) in the \ac{Raspi}s during startup
+the root filesystem (i.e. Raspbian Lite) in the \ac{Raspi}s during startup
 as read-only filesystem. Thus, on the one hand, the image configuration files
@@
 filesystem that is mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp}
-as read/writable on top of the lower root filesystem. Thus, reading a file
+as read / writable on top of the lower root filesystem. Thus, reading a file
 may return a file from the lower filesytem, but if it is stored, it will
@@
 the stored file from the upper layer. After a reboot, all the stored files
-in the upper filesystem will be erasured.
+in the upper filesystem will be retrieved.
 
-\subsection{Overlaying Filesystem Installation}
+\subsection{Filesystem Installation}
 
-Assuming that we are still using the \texttt{chroot} commands into the
-Raspbian lite root filesystem for installing packages, we can setup the
+Assuming that we are still in the \texttt{chroot} environment into the
+Raspbian Lite root filesystem for installing packages, we can setup the
 overlay filesystem at this point of the procedure. There already exists
@@
 available at the Git repository in \cite{overlayroot}. Since we have
-installed git in a previous step, we clone the repository. The
+installed \texttt{git} in a previous step, we clone the repository. The
 command block below stores it in \texttt{/tmp} which is really mounted in
@@
 
-Although these commands might be output some warnings, they should
+Although these commands might output some warnings, they should
 successfully generate working initial \ac{RAM} filesystems. Later, an
@@
 
+Again, if the \texttt{----recursive} option is not available, the
+procedure in Section~\ref{sec:umount} of the Appendix could be followed.
 If the steps are successfully executed up to this point, the customized image
@@
 
-It is \textit{very
-important} to write to the correct device as everything will be overwritten.
-To avoid removing information from the wrong device, a testbed administrator
-can insert / eject a memory card a couple of times and using the
-\texttt{sudo lsblk} and \texttt commands to deduce the correct device.
-For our case, the device was \texttt{/dev/sdb} in an Ubuntu distribution and
-\texttt{/dev/mmcblk0} in an Archlinux distribution. Once identified, to write
+It is \textit{very important} to write to the correct device as
+everything will be overwritten. To avoid removing information from the
+wrong device, a testbed administrator can insert and eject a memory
+card a couple of times and using the \texttt{sudo lsblk} and \texttt{df -h}
+commands to deduce the correct device. For our case, the device was
+\texttt{/dev/sdb} in an Ubuntu distribution and
+\texttt{/dev/mmcblk0} in an Archlinux distribuion. Once identified, to write
 the image to a memory card, the following command is used:

### 2016-07-15 testbed_http.tex
Not updated commits

--- 2016-07-13 testbed_http.tex
@@
 administrator might opt for not including this as part of his / her setup.
-We start our description consideri
 
@@
 considering an overlay filesystem approach. In this section, instead of
-writing the image direclty to the memory cards, we present how the
+writing the image directly to the memory cards, we present how the
 partitions can be extracted from the image, and how the boot partition can
@@
 
-Then, we setup a loopback device of the existing image and use \texttt{dd} to
-extract the boot and root partitions from the image:
+We set loopback devices for each of the image partitions and use \texttt{dd} to
+extract them from the image:
 \begin{lstlisting}[]
@@
 # dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
+# dd if=${DEV}p3 of=${HTTPHOMEIMAGE}.img bs=4M
 # losetup -d ${DEV}
@@
 
-Now it is time to configure the boot- and root-image just extracted before
-the content of the boot image is copied to the a memory card and the root
-image is uploaded to a \ac{HTTP} server.
+Now, we configure each of these partitions. The content of the boot image is
+copied and modified just once into a memory card and the root
+and home images is uploaded to a \ac{HTTP} server. Even though we still
+have to manually configure and set the memory cards, we only have
+to make this one time, since the \ac{Raspi}s will fetch the root filesystem
+from the \ac{HTTP} server every they are turned on.
 
@@
 
-\subsubsection{Prepare sd-card}
+\subsection{Prepare sd-card}
 
@@
 
-\subsubsection{(Optional) Compressing root image}
+\subsection{Compressing root image (Optional)}
 
@@
 
-\subsubsection{Try it out}
+\subsection{Try it out}
 

### 2016-07-15 testbed_http.tex
Few changes

--- 2016-07-15 testbed_http.tex
@@
 considering an overlay filesystem approach. In this section, instead of
-writing the image direclty to the memory cards, we present how the
-partitions can be extracted from the image, and how the boot partition can
-be configured to download the root filesystem and home directory partitions
+working with three partitions and writing the image directly to the memory
+cards, we present how to configure the boot partition to download the root
+filesystem partition and also configure the home directory partitions
 from a \ac{HTTP} server during startup. The advantage of this setup is that

### 2016-07-15 appendix.tex
Preparing http script

@@
 # mknod ${DEV}p2 b 7 2
-# losetup -o $((8192*512)) --sizelimit $((137215*512)) ${DEV}p1 ${IMAGE}.img
-# losetup -o $((137216*512)) --sizelimit $((4806655*512)) ${DEV}p2 ${IMAGE}.img
+# losetup -o $((8192*512)) --sizelimit $(( (137215-8192+1)*512) )) ${DEV}p1 ${IMAGE}.img
+# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) )) ${DEV}p2 ${IMAGE}.img
 \end{lstlisting}
@@
 # mknod ${DEV}p3 b 7 3
-# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512)) ${DEV}p2 ${IMAGE}.img
+# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) ) ${DEV}p2 ${IMAGE}.img
 # losetup -o $((4806656*512)) --sizelimit $(( (6903807-4806656+1)*512)) ${DEV}p3 ${IMAGE}.img

### 2016-07-15 testbed_http.tex
Preparing http script

@@
 cards, we present how to configure the boot partition to download the root
-filesystem partition and also configure the home directory partitions
+filesystem partition and configure the home directory partition
 from a \ac{HTTP} server during startup. The advantage of this setup is that
@@
 # export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
+/dev/loop0
 # dd if=${DEV}p1 of=${HTTPBOOTIMAGE}.img bs=4M
 # dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
-# dd if=${DEV}p3 of=${HTTPHOMEIMAGE}.img bs=4M
 # losetup -d ${DEV}
@@
 copied and modified just once into a memory card and the root
-and home images is uploaded to a \ac{HTTP} server. Even though we still
+image is uploaded to a \ac{HTTP} server. We configuing it, we will indicate
+that the home directory is storage-persistent. Even though we still
 have to manually configure and set the memory cards, we only have
 to make this one time, since the \ac{Raspi}s will fetch the root filesystem
-from the \ac{HTTP} server every they are turned on.
-
-Create a mountpoint and mount the http root and boot images:
+from the \ac{HTTP} server every they are turned on. We create a mountpoint
+and mount the \ac{HTTP} root and boot images:
 %Mount the http root image and home directories
@@
 
-In a previous steps, the boot loader was configured to call an initial \ac{RAM}
-filesystem. Now, we present how to the initial \ac{RAM}
-filesystem can be extended further to download and cache the root filesystem
-from a \ac{HTTP} server before mounting it.
+In Section~\ref{sec:overlay_fs}, the boot loader was configured to call
+the initial \ac{RAM} filesystem. Instead of that, now we present how the
+the initial \ac{RAM} filesystem in the boot partition is configured to
+download and cache the root filesystem from a \ac{HTTP} server before
+mounting it. The initial \ac{RAM} filesystem is initialized by the
+\texttt{\$\{HTTPROOTDIR\}/boot/init.gz} script. This is first called
+when the system is booting up before mounting the root filesystem and
+parses the arguments specified in \texttt{\$\{HTTPROOTDIR\}/boot/cmdline.txt}.
+We modify the arguments in this file to download, cache and mount the
+\ac{HTTP} root image filesystem (i.e. the \texttt{httproot.img} file) with a
+customized script.
 
-The initial \ac{RAM} filesystem contains an \texttt{init} script that
-is first called to initializes
-\texttt{initramfs}, parse arguments specified in \texttt{cmdline.txt}, and
-call a specialized script that prepares the actual root filesystem. By default,
-Raspbian includes scripts to mount the root filesystem from 1) a block devices
-(i.e. the memory card) and 2) from a \ac{NFS} shared folder. We make a third
-script to mount the root filesystem from an image
-(i.e. the root image \texttt{httproot.img}) that is downloaded from a
-\ac{HTTP} server.
-
-The initramfs script to download and mount the image from a \ac{HTTP} server
-will require a few additional arguments. E.g. to specify from which \ac{HTTP}
-server to download the root filesystem image from.
-%This script that we will create will require a few additional arguments such
-%as which \ac{HTTP} server the image should be downloaded from.
-These arguments
-may be specified in \texttt{cmdline.txt}.
-%and later parsed within the \texttt{initramfs} script.
-
-Keep the content in \texttt{cmdline.txt}, but add/modify the variables
-\texttt{boot}, \texttt{ip}, \texttt{image}, \texttt{root}, and \texttt{url}.
-
-%Keep also all other variables as they are in
-%\texttt{cmdline.txt}.
+By default, Raspbian includes scripts to mount the
+root filesystem from block devices (memory cards) or a \ac{NFS} shared folder.
+Thus, our customized \ac{HTTP} initramfs script is an inclusion in the
+\texttt{\$\{HTTPROOTDIR\}/etc/initramfs-tools/scripts} directory to handle
+booting up from a \ac{HTTP} server. This script will require a few
+additional arguments. e.g. to specify \ac{URL} for downloading the root
+filesystem image, among others. These arguments have to be specified in
+\texttt{cmdline.txt} file. Thus, we edit the
+\texttt{\$\{HTTPROOTDIR\}/boot/cmdline.txt} file to: (i) Add / modify
+the variables \texttt{boot}, \texttt{ip}, \texttt{image},
+\texttt{root} and \texttt{url}; and (ii) remove the argument
+\texttt{init=/usr/lib/raspi-config/init\_resize.sh} since we use customized
+partitions. In our case, after these edits, our \texttt{cmdline.txt} file
+is as follows (however it might differ for other testbed administrators):
 
 %Add the below configurations to \$\{HTTPROOTDIR\}/boot/cmdline.txt
-\textcolor{red}{(remove init=/usr/lib/raspi-config/init\_resize.sh earlier in guide)}
 \Suppressnumber\begin{lstlisting}[]
@@
 <@\textcolor{gray}{
----------------------------------------------------------------}
+-------------------------------------------------------------}
 \Reactivatenumber @>
-<@\Suppressnumber @>boot=http ip=dhcp image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0
-    url=http://kom.aau.dk/project/TuneSCode/raspi/<@\Reactivatenumber @>
+<@\Suppressnumber @>dwc_otg.lpm_enable=0 console=serial0,115200 console=tty1 boot=http ip=dhcp
+image=/dev/mmcblk0p2:/httproot.img root=/dev/loop0 url=http://kom.aau.dk/project/TuneSCode/raspi/
+rootfstype=ext4 elevator=deadline fsck.repair=yes rootwait quiet<@\Reactivatenumber @>
 \end{lstlisting}
@@
 
-Here is how the variables will be used in the \texttt{initramfs}:
+The arguments for the \texttt{initramfs} initialization script are employed
+as follows:
 \begin{itemize}
-    \item boot=http: tells the init script in initramfs to call the \texttt{http} script that will be presented below
-    \item ip=dhcp: use \ac{DHCP} to obtain an \ac{IP} address in the \texttt{initramfs}
-    \item image=<device>:<root image path>: is where the \ac{Raspi} will store the downloaded root image.
-    \item root=/dev/loop0: The \texttt{http} initramfs script will make the root filesystem available on /dev/loop0
-    \item url=<URL to root image>: Is the directory on a \ac{HTTP} server where the root image will be iavailable.
+    \item \texttt{boot=http}: Tells the script to call the \texttt{http}
+    script that will be introduced below.
+    \item \texttt{ip=dhcp}: Uses \ac{DHCP} to get an \ac{IP} address in
+    the \texttt{initramfs} for network connectivity.
+    \item \texttt{image=<IMAGEDEV>:<IMAGEPATH>}: Indicates to which device
+    and path will the image be located.
+    \item \texttt{root=/dev/loop0}: Is to make the root image available in
+    the \texttt{/dev/loop0} loopback device.
+    \item \texttt{url=<URL>}: Is the link where the root image must be
+    available.
 \end{itemize}
@@
 %The \ac{URL} should be modified to reflect where
-%the \ac{Raspi}s should download the http root image. The \texttt{image} variable informs the
-%\texttt{http} initramfs script to download and cache the http root image in the second partition
-%on its memory card at path \texttt{/httproot.img}.
+%the \ac{Raspi}s should download the http root image.
 
-The specialized script that we made to download and mount the root filesystem image from a \ac{HTTP} server should be created in the root filesystem as presented below:
+For other testbed administrators, the previous arguments have to be adjusted
+to fit their needs. Then, we show the customized script to download and
+mount the root filesystem image from a \ac{HTTP} as follows:
 %Create the following file
-\textcolor{red}{(THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT. I'M NOT EVEN SURE IT IS POSSIBLE TO MAKE THE IMAGE READ-WRITABLE)}
+
+% \textcolor{red}{(THE READ-ONLY OPTION IS STILL NOT CONFIGURABLE IN THIS SCRIPT. I'M NOT EVEN SURE IT IS POSSIBLE TO MAKE THE IMAGE READ-WRITABLE)}
 \Suppressnumber\begin{lstlisting}[]
@@
 <@\textcolor{gray}{
----------------------------------------------------------------}
+---------------------------------------------------------------------------------------------}
 \Reactivatenumber @>
@@
 # Check if variables are declared
-[ -z "${IMAGE}" ] && echo "Error: Please add 'image' in cmdline.txt" && exit 1
-[ -z "${URL}" ] && echo "Error: Please add 'url' in cmdline.txt" && exit 1
+[ -z "${IMAGE}" ] && echo "Error: Please add 'image' in /boot/cmdline.txt" && exit 1
+[ -z "${URL}" ] && echo "Error: Please add 'url' in /boot/cmdline.txt" && exit 1
 

### 2016-07-15 testbed_http.tex
Almost there, just missing the write image part

@@
 
+\subsection{HTTP Image Fetch Script}
 In Section~\ref{sec:overlay_fs}, the boot loader was configured to call
@@
 
-Make the script executable
+Once the script has been saved, we make the script executable by doing:
 \begin{lstlisting}[]
@@
 
-The variable (boot=http) that we specified in \texttt{cmdline.txt} tells the
-\texttt{init} script to call the script above rather than the default scripts
-that mounts the root filesystem from a block device or a \ac{NFS} shared
-folder. The \texttt{init} script will call the function \texttt{http\_mount\_root}
-which in this case can be consired as the \texttt{main} function.
+When using \texttt{boot=http} in the \texttt{cmdline.txt} file, this
+indicates the \texttt{initramfs} initialization script to call the script
+above instead of other default scripts that mount the root filesystem.
+Once called, the initialization script will first execute the
+\texttt{mount\_root()} function which can be consired as the \texttt{main}
+function for this script. As in any other shell scripts, this means
+that any code in the global namespace will executed by the interpretor
+before the function \texttt{mount\_root()} is entered when booting. Hence,
+line 9 of the \texttt{http} script will be the first command executed.
+This line creates a directory in the \texttt{initramfs} named
+\texttt{imagedir} that we will use to hold the root image. Next, lines 12
+to 44 will parse the additional arguments that we provided in
+the \texttt{cmdline.txt} file.
 
-As in any other shell scripts, this means that code in the global namespace
-will be executed by the interpretor before the function \texttt{http\_mount\_root}
-is entered.
-%Due to the variable that we specified in \texttt{cmdline.txt},
-Hence, line 9 will be the first command executed. This line creates a directory
-in \texttt{initramfs} named
-\texttt{imagedir} that we will use to hold the root image.
+Since the \ac{Raspi}s do not have enough \ac{RAM} to hold the entire
+root filesystem, we download it to the partition that, as specified
+in \texttt{\$\{HTTPROOTDIR\}/etc/fstab} file later, will be mounted as the
+\texttt{\$HOME} directory when the system starts. This will enable the
+\ac{Raspi}s to cache the root image and use the previously downloaded
+root image again after any reboots. For details, see the function
+\texttt{http\_fetch\_image()} just after the variables verification.
+We edit the mentioned \texttt{fstab} file with root privileges to look
+as the following (exclude the last line if a persistent home storage
+directory is not desired):
 
-Next, line 12 to line 44 will parse the additional arguments that we provided in
-\texttt{cmdline.txt}.
-
-The \ac{Raspi}s does not have enough \ac{RAM} to hold entire root
-filesystem, so we instead download it to the partition that, as specified in
-\texttt{fstab} later, will later be mounted as the \texttt{HOME} directory
-when the system starts. This will enable the \ac{Raspi}s to cache the root image
-and use the previously downloaded root image again after reboots (see the
-function \texttt{http\_fetch\_image} starting from line 56)\fxnote{is there a better way to reference to a function within the script?
-This reference will be invalid if the script is slightly modified}.
-
-Edit fstab to look like this (exclude the last line if a persistent
-home storage directory is not desired):
+% \fxnote{is there a better way to reference to a function within the script? This reference will be invalid if the script is slightly modified}.
 \Suppressnumber\begin{lstlisting}[]
@@
 
-Chroot into the image
-% CHROOT to OS image
+\subsection{Initial RAM Filesystem for HTTP Fetch}
+We enter again in \texttt{chroot} mode into the image to make the new
+\texttt{initramfs} files. We note that for other testbed administrator
+kernels, the files might be different. The commands for these tasks are:
+
+% CHROOT to HTTP image
 \begin{lstlisting}[]
@@
 # mount --bind /dev/pts dev/pts
-# proot -q qemu-arm-static -S ${HTTPROOTDIR}
+# chroot ${HTTPROOTDIR} /usr/bin/qemu-arm-static /bin/bash
 # export PS1="(chroot) $PS1"
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Create new initramfs (remember that your kernel might be different)
-\begin{lstlisting}[]
 (chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
@@
 
-Exit chroot
+In case the \texttt{chroot} command does not work, it is possible to use
+the alternative \texttt{proot} command as described in
+Section~\ref{sec:chroot}. If successfully executed, the boot partition for
+the SD card is ready to be written. We exit the \texttt{chroot} environment
+and unmount all the partitions as follows:
+
 \begin{lstlisting}[]
 (chroot) # exit
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
-
-Unmount:
-\begin{lstlisting}[]
 $ cd ..
@@
 
-\subsection{Prepare sd-card}
+\subsection{Writing SD Cards for HTTP Fetch}
 
@@
 
-% Define variable
-% export IMAGE="2016-05-27-raspbian-jessie-lite"
-% export WORKDIR="${HOME}/Raspbian"
 \begin{lstlisting}[]
@@
 
-\subsection{Compressing root image (Optional)}
+% IT WILL BE GOOD TO TRY THE COMPRESSED IMAGE BUT MAYBE LATER:
 
-PROBABLY A BETTER IDEA TO USE BTRFS TRANSPARENT COMPRESSION
+% \subsection{Compressing root image (Optional)}
+% PROBABLY A BETTER IDEA TO USE BTRFS TRANSPARENT COMPRESSION
+% \begin{lstlisting}[]
+% # apt-get install squashfs-tools
+% # mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
+% # mksquashfs ${HTTPROOTDIR} compressed_${HTTPROOTIMAGE}.img
+% \end{lstlisting}
+% \FloatBarrier
+% \vspace{-5mm}
 
-\begin{lstlisting}[]
-# apt-get install squashfs-tools
-# mount -o loop ${HTTPROOTIMAGE}.img ${HTTPROOTDIR}
-# mksquashfs ${HTTPROOTDIR} compressed_${HTTPROOTIMAGE}.img
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
+\subsection{HTTP Images File and Setup Test}
 
-\subsection{Try it out}
-
-Upload the root image (\$\{HTTPROOTIMAGE\}.img) to the HTTP server used in the
-configurations above. Then, unplug the memory card from the computer.
-Insert it into a \ac{Raspi} and
-turn it on. It is recommended that the \ac{Raspi} is turned on with a monitor
-connected the first time the newly created Linux installation is tested to
-assure that Raspbian starts up correctly.
-
-PRESENT FILES ON HTTP. E.G. HTTPROOT.IMG AND IMAGES.CSV
-
-\begin{itemize}
-    \item Copy http root image (\$\{HTTPROOTIMAGE\}.img) a HTTP server
-    \item Create images.csv
-\end{itemize}
+We upload the \texttt{\$\{HTTPROOTIMAGE\}.img} file to the \ac{HTTP}
+server used in the configurations above. Finally, in the \ac{HTTP} server,
+the testbed administrator has to create
+\texttt{http://<SERVERIP>/.../images.csv} file. This file contains the
+information regarding which image should be download for each \ac{Raspi}. A
+sample file is the following:
 
@@
 
-Notice that the \ac{Raspi}s can be configured to download different root images.
-This is useful both for having various root images for subsets of the \ac{Raspi}s,
-but also as a basic way inform the \ac{Raspi}s that the root image changed such
-that they can download the new root image at next boot.
+Notice that the \ac{Raspi}s can be configured to download different
+root images. This is useful both for having various root images for subsets
+of the \ac{Raspi}s, but also as a basic way inform the \ac{Raspi}s that
+the root image changed such that they can download the new root image at
+next boot.
 
-As of this writing, there is unfortunately a bug in wget that makes wget not showing
-its progress bar while downloading the root image at startup. Instead, it may appear
-to be unable to connect although it is downloading, so keep patient.
+Once this file is set in the \ac{HTTP} server, the testbed
+administrator needs to unplug the memory card from its \ac{PC} and insert
+it into a \ac{Raspi} and turn it on. For this part, it is recommended that the
+\ac{Raspi} is turned on with a monitor connected to assure that Raspbian
+starts up correctly. At the time of writing this manuscript, a bug in
+the \texttt{wget} command does not permit to show the image download progress
+bar at startup. Instead, it may appear to be unable to connect.
+Nevertheless, the image should be downloading. On average this process takes
+XX minutes. If all the procedure has been successfully executed,
+the testbed administrator should observe the \ac{Raspi} properly booting up
+after the downloading the root image file. Finally, the testbed adminstrator
+only needs to copy one time the boot partition to all the cards and
+boot the remaining \ac{Raspi}s. This concludes the procedure.
 
-\begin{lstlisting}[]
-Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
-image.csv       100% |*******************************************************| 1287   0:00:00 ETA
-Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
-_
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
+% \begin{lstlisting}[]
+% Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
+% image.csv       100% |*******************************************************| 1287   0:00:00 ETA
+% Connecting to <HTTP SERVER IP> (<HTTP SERVER IP>:<PORT>)
+% \end{lstlisting}
+% \FloatBarrier
+% \vspace{-5mm}
 
-REMEMBER TO COPY-PASTE ALL CONTENT FROM SCRIPT INTO LATEX AGAIN IN CASE SOME
-SCRIPT CHANGED, BUT WAS NOT UPDATED IN LATEX
+% DO WE REALLY NEED TO DO THIS? -->
+% REMEMBER TO COPY-PASTE ALL CONTENT FROM SCRIPT INTO LATEX AGAIN IN CASE SOME
+% SCRIPT CHANGED, BUT WAS NOT UPDATED IN LATEX

### 2016-07-17 appendix.tex
Testing procedure. BusyBox wget command fails when testing remote booting

@@
 In case Paramiko version 1.15.1 or higher is not installed, we may observe
-a key exchange algorithm error. In this case, it might be
-necessary to update the Paramiko package and verify this by running:
+a key exchange algorithm error when trying to log in through \ac{SSH} to a
+Raspi. In this case, it might be necessary to update the Paramiko package
+that Fabric uses for this purpose and verify this by running:
 

### 2016-07-18 appendix.tex
Final submission manuscript

@@
-\section{HTTP Image Fetch Setup (Optional)}
-\label{sec:http}
+% \section{HTTP Image Fetch Setup (Optional)}
+% \label{sec:http}
 
-\input{testbed_http}
+% \input{testbed_http}
 
@@
 Also, when creating the persistent home directory partition in
-Section~\ref{sec:persistent_directories}, for recreating the loopback devices
-the commands are:
+Section~\ref{sec:non_persistent_directories}, for recreating the loopback
+devices the commands are:
 

### 2016-07-18 cross_compilation.tex
Final submission manuscript

@@
 them to our coding style and compiling convention. For this, we use the
-\texttt{\$TOOLCHAIN} directory as the working directory. However,
+\texttt{\$\{TOOLCHAIN\}} directory as the working directory. However,
 the testbed administrator may choose some other working directory of
@@
 $ export TOOLCHAINDIR="${HOME}/toolchains"
-$ mkdir -p $TOOLCHAINDIR
-$ cd $TOOLCHAINDIR
+$ mkdir -p ${TOOLCHAINDIR}
+$ cd ${TOOLCHAINDIR}
 \end{lstlisting}
@@
 the binaries accessible from the command shell systemwide. A way to do
-this is by adding the following commands in the \texttt{\${HOME}/.profile}
+this is by adding the following commands in the \texttt{\$\{HOME\}/.profile}
 as follows:
@@
 
-To update the \texttt{PATH} variable and the \texttt{.profile}, we use the
-\texttt{source} command for the changes take effect in the administrator
-system:
+To update the \texttt{\$\{PATH\}} variable and the \texttt{.profile}, we
+use the \texttt{source} command for the changes take effect in the
+administrator system:
 \begin{lstlisting}[]
@@
 \Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\${CODEDIR}/hello\_world.cpp}@>
+<@\textcolor{gray}{\$\{CODEDIR\}/hello\_world.cpp}@>
 <@\textcolor{gray}{
------------------------------------------------}
+---------------------------------------------------}
 \Reactivatenumber @>

### 2016-07-18 image_setup.tex
Final submission manuscript

@@
 the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
-available in two bundles: Raspbian and Raspbian lite. The difference
+available in two bundles: Raspbian and Raspbian Lite. The difference
 between the two, is that Raspbian contains a pre-installed desktop environment
-for user interaction and Raspbian lite by default only permits to interact
+for user interaction and Raspbian Lite by default only permits to interact
 through a command shell. Given that the \ac{Raspi}s in our testbed are not
-connected to monitors, we decide to work with Raspbian lite. Still, if later
+connected to monitors, we decide to work with Raspbian Lite. Still, if later
 desired, a desktop environment can be installed using the package manager
@@
 
-% The overall procedure to customize the official Raspbian lite image
+% The overall procedure to customize the official Raspbian Lite image
 % are:
 % \begin{enumerate}
-%     \item Download Raspbian lite
-%     \item Alter Raspbian lite. e.g. browse, modify, add, and delete files
-%         in the official Raspbian lite image
-%     \item Change root, i.e. change root filesystem into the Raspbian lite
+%     \item Download Raspbian Lite
+%     \item Alter Raspbian Lite. e.g. browse, modify, add, and delete files
+%         in the official Raspbian Lite image
+%     \item Change root, i.e. change root filesystem into the Raspbian Lite
 %         image to update and install software packages
@@
 
-%\subsection{Download Raspbian lite}
-%To download the latest Raspbian lite, we go to the url
-%\url{http://downloads.raspberrypi.org/raspbian_lite/images/}.
+%\subsection{Download Raspbian Lite}
+%To download the latest Raspbian Lite, we go to the url
+%\url{http://downloads.raspberrypi.org/raspbian_Lite/images/}.
 
-The latest Raspbian lite bundle can be downloaded from the Raspbian
+The latest Raspbian Lite bundle can be downloaded from the Raspbian
 official webpage~\cite{raspbian}. At the time of this writing, the latest
@@
 To ensure that the content of the bundle does not change, this procedure
-is based on that particular version of Raspbian lite which we have
-made available at~\cite{tunescode_webpage}. All other files used in this paper
-are available there unless specified otherwise. To get started, the testbed
-administrator must open a Linux shell (terminal) in and declare the environment
-variables shown in the command block below. We show the procedure by
-performing the role of the testbed administrator. To copy the commands from
-this procedure to the testbed administrator terminal,
-we strongly recommend to read this document with Adobe Acrobat Reader in Linux.
-Otherwise, copied characters might not render properly.
+is based on that particular version of Raspbian Lite which we have
+made available at~\cite{tunescode_webpage} acting as our \ac{HTTP} server.
+All other files used in this paper are available there unless specified
+otherwise. To get started, the testbed administrator must open a Linux
+shell (terminal) in its \ac{PC} and declare the environment variables shown
+in the command block below. We show the whole procedure by performing the role
+of the testbed administrator. To copy the commands from this procedure to the
+testbed administrator terminal, we recommend to read this document
+with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
+render properly when pasted.
 
@@
 
-In this code block, the \texttt{URL} and \texttt{IMAGE} variables specify where
-the Linux bundle is located and \texttt{WORKDIR} specifies a working
-directory where the Raspbian bundle will be downloaded and customized.
-Notice that even though we use the \$ and \# signs in the shell, in general
-these signs will be particular to the testbed administrator \ac{OS} shell.
-Next, we create the working directory and go to it with the \texttt{cd}
-command. To download the image, we utilize the \texttt{wget} command before
-unpacking the \texttt{zip} file as follows:
+In this code block, the \texttt{\$\{URL\}} and \texttt{\$\{IMAGE\}}
+variables specify where the Linux bundle is located and
+\texttt{\$\{WORKDIR\}} specifies a working directory where the Raspbian
+Lite bundle will be downloaded and customized. Notice that even though we
+use the \$ and \# signs in the shell, in general these signs will be
+particular to the testbed administrator \ac{OS} shell. Next, we create
+the working directory and go to it with the \texttt{cd} command. To
+download the image, we utilize the \texttt{wget} command before unpacking
+the \texttt{zip} file as follows:
 
@@
 
-After Raspbian lite has been unpacked, there should be an \texttt{.img}
-file in the \texttt{WORKDIR} directory. Here, \texttt{fdisk} can be used to
-display the content of the image. To do this and obtain administrative
-information, we parse the arguments \texttt{-u sectors} to display the
-sizes in sectors and \texttt{-l} to display the partitions within the
-image. The \texttt{fdisk} command should output to the terminal something
-similar to:
+After Raspbian Lite has been unpacked, there should be an \texttt{.img}
+file in the \texttt{\$\{WORKDIR\}} directory. Here, \texttt{fdisk} can
+be used to display the content of the image. To do this and obtain
+administrative information, we parse the arguments \texttt{-u sectors}
+to display the sizes in sectors and \texttt{-l} to display the partitions
+within the image. The \texttt{fdisk} command should output to the terminal
+something similar to:
 
@@
 For browsing and altering the files in the image, we mount the block
-device partitions into a particular path of our \texttt{WORKDIR} in order
+device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
 to customize them. We proceed first to mount the block device partition that
@@
 
-The root filesystem mounted in \texttt{ROOTDIR}, has already a boot
+The root filesystem mounted in \texttt{\$\{ROOTDIR\}}, has already a boot
 directory that can be used as the mount point for the boot partition
-in the block device \texttt{loop0p1}. This is convenient because
-the final edited partition from \texttt{ROOTDIR/boot} will be mounted on
+in the block device \texttt{/dev/loop0p1}. This is convenient because
+the final edited partition from \texttt{\$\{ROOTDIR\}/boot} will be mounted on
 this same directory when a \ac{Raspi} starts up with a memory card
@@
 In this way, it is now possible to change all files within the Raspbian
-image as desired by editing the files in \texttt{ROOTDIR}. We take
+image as desired by editing the files in \texttt{\$\{ROOTDIR\}}. We take
 advantage of this to edit configuration files, append new files and even
@@
 In our case, we have our default configuration scripts stored in a
-\texttt{rasp\_config.zip} file located in the same \ac{URL} where the image
-was retrieved from, i.e. the one in the environment variable \texttt{URL}.
-We first manually download this compressed file with \texttt{wget} and
-extract it locally into our Raspbian lite image. These commands and the
-output of the last one are shown as follows:
+\texttt{rasp\_config.zip} file located in the same \ac{URL} of the
+\ac{HTTP} server where the image was retrieved from, i.e. the one in
+the environment variable \texttt{\$\{URL\}}. We first manually download
+this compressed file with \texttt{wget} and extract it locally into our
+Raspbian Lite image. These commands and the output of the last one are
+shown as follows:
 
@@
 The unzippped files are one configuration file and three configuration
-scripts put in the newly created \texttt{\${ROOTDIR}/home/pi/rasp\_config/}
+scripts put in the newly created \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/}
 folder in the image. In what follows, we describe which features do we
@@
 \ac{Raspi}s in the configuration file
-\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}. A sample of our file
+\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}. A sample of our file
 is shown as follows:
@@
 The testbed administrator has to insert the \ac{MAC} addresses and hostnames
-of his/her \ac{Raspi}s obtained previously in the format shown in the
+of his / her \ac{Raspi}s obtained previously in the format shown in the
 configuration file. In the file format, on each line there is a \ac{MAC}
@@
 This list file will be employed by the
-\texttt{\${ROOTDIR}/home/pi/rasp\_config/set\_hostname}
+\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}
 \ac{Bash} script to assign the hostname of each \ac{Raspi}.

### 2016-07-18 introduction.tex
Final submission manuscript

@@
 This will break new ground for new markets such as driverless cars, manufacturing, humanoid robots, smart grids, and many more.
-With 5G the number of wireless devices will increase by 5$\times$ up to 50 billion devices \cite{cisco2011forecast}.
+With 5G, the number of wireless devices will increase by 5$\times$ up to 50 billion devices \cite{cisco2011forecast}.
 It is generally accepted, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connect due to the expected energy consumption.
@@
 In our previous works, we have shown that \ac{RLNC} \cite{koetter2003algebraic,ho2006random} is able to satisfy the aforementioned technical requirements.
-We have actually shown how to increase the throughput \cite{pahlevani2013playncool}, reduce the delay \cite{szabo2015towards}, or support heterogeneity for coding enabled communication nodes \cite{lucani2014fulcrum}.
+We have actually shown how to increase the throughput \cite{pahlevani2013playncool}, reduce the delay \cite{szabo2015towards} or support heterogeneity for coding enabled communication nodes \cite{lucani2014fulcrum}.
 
@@
 Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with Raspberry Pi.
-Our work is organized as follows.
-Section~\ref{sec:testbed} introduces the testbed system, its setup, files configuration and connectivity.
+Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system.
+In Section~\ref{sec:image_setup}, we provide details about the testbed setup, scripts, configuration files and connectivity.
+In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step.
+Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed t simplify the execution of routinary and repetitive tasks.
 Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}.
-Final conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
+Conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
+Finally, a set of alternative commands procedures, in case the ones presented in this work might not be executed, are discussed in the Appendix.
 
-%The emergence of powerful and inexpensive single-board computers
-%running full versions of a standard \ac{OS} from many possible in the last
-%years opens exciting possibilities for researchers. This not only
-%allows to seamlessly develop implementations that are compatible with
-%higher end devices and allows the reuse stable software from the Open
-%Source community. It also enables the deployment and testing of large
-%scale (hundreds of devices or more) distributed systems at a fraction
-%of the cost necessary in previous years. This paper focuses on
-%providing the critical steps and mechanisms followed to deploy, setup,
-%and maintain a large-scale testbed based on Raspberry Pi devices of
-%different models. 
-%
-%% INTRODUCTION TO NETWORK CODING
-%Introduced by Ahlswede et al.~\cite{ahlswede2000network}, network
-%coding constitutes a paradigm shift in the way how researchers and
-%industry understand and operate networks, by changing the role of
-%intermediate relays in the process of transmission of information.
-%Relays are no longer limited to storing and forwarding data, but also
-%take part in the coding process, through a process called recoding,
-%where the relay generates new linear combinations of incoming coded
-%packets without previously decoding the data. Network coding allows
-%the increase of throughput, reliability, security and delay
-%performance of the networks.
-
-%Compared to other broadly used coding schemes, e.g., Gallager's LDPC
-%codes~\cite{gallager1962low} or Reed-Solomon
-%codes~\cite{reed1960polynomial}, network coding is a technology that
-%has been implemented in real systems since the early years of its
-%conception. In 2006, Katabi et al.~\cite{katabi2006practical} published
-%the results of the performance of their protocol COPE when implemented
-%in a real wireless system, which relied on minimalistic coding to
-%provide interesting gains. In 2008, Pedersen et
-%al.~\cite{pedersen2008implementation} used commercially available
-%Symbian OS enabled mobile phones to implement network coding in a
-%\ac{D2D} cooperation-based application. In 2011, researchers at Aalborg
-%University developed Kodo~\cite{kodo2011pedersen}, a C++11 network
-%coding library with open source code for researchers intended to make
-%network coding implementations easily available for both, the research
-%community and commercial entities.
-%Another case is CATWOMAN~\cite{hundeboll2012catwoman}, a protocol
-%implemented on top of the BATMAN protocol~\cite{johnson2008simple} for
-%IEEE 802.11 multi-hop meshed networks using some of the intuition from
-%COPE, but deploying it in real systems. It is currently available as
-%open source code in the Linux kernel. Many other implementations have
-%been tested on real world systems~
-%\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.
-%
-%Despite this successful deployment in real systems, many of these
-%protocols and contributions have been implemented in separate testbeds
-%and the experiences are hard to reproduce.  Given the need for
-%reproducibility of past and future results among researchers and the
-%need of low-cost and easy-to-deploy testbeds, this paper contribution
-%is provide detailed instructions on how to set up a low-cost,
-%heterogeneous and scalable testbed consisting of Raspberry Pis
-%(Raspis), and by providing detailed measurements of goodput and energy
-%consumption of such testbed when performing network coding operations
-%with different codecs (such as full \ac{RLNC} \cite{ho2006random},
-%multicore-enabled \ac{RLNC}, sparse \ac{RLNC} and tunable sparse \ac{RLNC}).
-%The heterogeneity of the testbed comes from the fact that old, single core,
-%Raspberry Pi 1 model B units are integrated seamlessly in the described
-%testbed with newer, multicore versions, like the Raspberry Pi 2 model B.
-%Our testbed allows us to obtain a comprehensive set of measurements
-%to characterize the goodput and energetic footprint of these devices
-%showing that processing speeds of 800 Mbps and energy per bit values
-%of 1 nJ or even one order of magnitude less are possible.
-%
-%
 %%The introduction should briefly place the study in a broad context and highlight why it is important.

### 2016-07-18 overlay_fs.tex
Final submission manuscript

@@
 as read-only filesystem. Thus, on the one hand, the image configuration files
-will remain after a reboot and will not be modifiable. On the other hand,
-to enable the possibility of non-persistent changes, we overlay the upper
-filesystem that is mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp}
-as read / writable on top of the lower root filesystem. Thus, reading a file
-may return a file from the lower filesytem, but if it is stored, it will
-be saved in the upper filesystem. Accessing this file again will return
+will remain after a reboot but the local data in these directories will be
+erased after a hypothetic reboot. On the other hand, to enable the
+possibility of persistent changes, we overlay the upper filesystem that is
+mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp} as read / writable
+on top of the lower root filesystem. Thus, reading a file may return a
+file from the lower filesytem, but if it is stored, it will be saved in
+the upper filesystem. Accessing this file again will return
 the stored file from the upper layer. After a reboot, all the stored files
-in the upper filesystem will be retrieved.
+in the upper filesystem will be retrieved, but the ones in the lower
+filesystem that are not part of the original image will be removed.
 
@@
 (chroot) # OVERLAYROOTDIR="/tmp/overlayroot"
-(chroot) # git clone https://github.com/chesty/overlayroot.git $OVERLAYROOTDIR
+(chroot) # git clone https://github.com/chesty/overlayroot.git ${OVERLAYROOTDIR}
 \end{lstlisting}
@@
 in the boot partition. If the system should be run in a \ac{Raspi} version 1,
-then use \texttt{init.gz}, otherwise use \texttt{init-v7.gz} as follows
-\todot{What happens if I have both a Raspi 1, 2 or 3 in my testbed? Do I have
-to have different images after this point?}:
+then use \texttt{init.gz}, otherwise use \texttt{init-v7.gz} as follows:
 
@@
 
-\subsection{Persistent Image Directories}
-\label{sec:persistent_directories}
+\subsection{Persistent and Non-Persistent Image Directories}
+\label{sec:non_persistent_directories}
 Provided the stacked filesystem is configured, now it is possible to have
-directories where files are not removed upon rebooting the \ac{Raspi}s.
+directories where files are removed or not upon rebooting the \ac{Raspi}s.
 The following procedure creates an extra partition in the image for
@@
 into the \ac{Raspi}s. In the following section, we indicate how to proceed
-with the writing of the image into various memory cards, although a more
-complex setup with a \ac{HTTP} server is also presented in
-Section~\ref{sec:http} of the Appendix.
+with the writing of the image into various memory cards.
 
-\subsection{Writing Customized Image to Memory Cards}
+\subsection{Writing Customized Image to SD Memory Cards}
 For a basic system setup, the final step is to write the customized image
@@
 wrong device, a testbed administrator can insert and eject a memory
-card a couple of times and using the \texttt{sudo lsblk} and \texttt{df -h}
+card a couple of times and using the \texttt{sudo lsblk} or \texttt{df -h}
 commands to deduce the correct device. For our case, the device was
 \texttt{/dev/sdb} in an Ubuntu distribution and
-\texttt{/dev/mmcblk0} in an Archlinux distribuion. Once identified, to write
+\texttt{/dev/mmcblk0} in an Archlinux distribution. Once identified, to write
 the image to a memory card, the following command is used:
@@
 administrator \ac{PC}. After this is made, it is only required
-to eject the memory card and now it can be started in a \ac{Raspi}.
+to eject the memory card and now plug it in a \ac{Raspi} so it can boot up.

### 2016-07-18 overview.tex
Starting with overlay fs
Final submission manuscript

@@
 Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
-All \ac{Raspi}s are each equipped with a memory card, a wired and wireless
+All \ac{Raspi}s are each equipped with a 8 GB memory card, a wired and wireless
 network interface and a power supply. All the \ac{Raspi} are connected to
@@
 testbed description and procedures for setting it up are not restricted
-to the academic scenario. All \ac{Raspi}s will be configured to run
+to the academic scenario. All \ac{Raspi}s are configured to run
 a \ac{SSH} daemon for easy remote access within the university network.
@@
 Debian-based Linux distribution. Our basic design considers to create
-a customized image to later store it in a memory card. Thus, based on a
-regular Raspbian Linux image, we explain how to resize and customize it to
-include desired packages for performing routinary tasks in all the
-\ac{Raspi}s. As another design consideration, when working on a testbed
+a customized image to later store it in a memory card for each \ac{Raspi}.
+Once configured, we store the resulting image file in a \ac{HTTP} server
+as backup and in case the testbed administrator requires to make new
+changes to this file. We also put all the required configuration files
+and scripts for the \ac{Raspi}s setup in the \ac{HTTP} server so there is
+a single place where system setup is stored and could be modified.
+This simplifies the system maintenance at a later stage.
+As another design consideration, when working on a testbed
 with many devices, it may not always be desired to make persistent changes
@@
 the testbed administrator decides to keep only persistent changes regardless
-of the testbed use. Finally, an advanced step in our design considers to
-fetch the \ac{Raspi}s root filesystem from a \ac{HTTP} server. This enables the
-testbed administrator to put the data files in one location. The
-\ac{Raspi}s will then get these data files after a reboot. We consider
-that this step is particularly useful when there is a large amount
-of devices to be configured and the setup process needs to be automated.
-However, since a more complex scenario that might be required only in special
-cases, we put this step in Section~\ref{sec:http} in the Appendix.
+of the testbed use. Finally, we include a set of automation, monitoring
+and cross-compilation tools over the top of our system in order to simplify
+the execution of repetitive and long tasks, be able to follow the progress
+of long task processes and compile relevant C++ source code for the testbed
+administrator.
+
+%an advanced step in our design considers to
+% fetch the \ac{Raspi}s root filesystem from a \ac{HTTP} server. This enables the
+% testbed administrator to put the data files in one location. The
+% \ac{Raspi}s will then get these data files after a reboot. We consider
+% that this step is particularly useful when there is a large amount
+% of devices to be configured and the setup process needs to be automated.
+% However, since a more complex scenario that might be required only in special
+% cases, we put this step in Section~\ref{sec:http} in the Appendix.

### 2016-07-18 testbed_http.tex
Keep pushing
Testing procedure. BusyBox wget command fails when testing remote booting
Final submission manuscript

@@
 \vspace{-5mm}
+
+%http://kom.aau.dk/project/TuneSCode/raspi/
 % ip=dhcp root=/dev/http http-options=hard,intr,ro
@@
 
-\subsection{Initial RAM Filesystem for HTTP Fetch}
+\subsection{Initial RAM Filesystem for HTTP Image Fetch}
 We enter again in \texttt{chroot} mode into the image to make the new
@@
 
-\subsection{Writing SD Cards for HTTP Fetch}
+\subsection{Writing SD Cards for HTTP Image Fetch}
 
-IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?
+%IS SETTING UP THE SD-CARD REALLY THE SAME PROCEDURE AS FOR A LOCAL FILESYSTEM?
+%ANSWER FROM NESTOR: YES! :)
+
+We insert a test memory card and verify its associated device in the
+filesystem to export it as a variable name:
 
@@
 
-Format partitions
+Format partitions. Before formating the partitions, take out the SD card
+and plug it again!!!
 \begin{lstlisting}[]
@@
 \begin{lstlisting}[]
-# umount --recursive ${HTTPROOTDIR}
+# umount ${HTTPROOTDIR}/boot
+# umount ${HTTPROOTDIR}
 \end{lstlisting}

### 2016-07-18 tools.tex
Final submission manuscript

@@
 
-Then, the \texttt{\$CODEDIR/fabfile.py} file below provides a script with
+Then, the \texttt{\$\{CODEDIR\}/fabfile.py} file below provides a script with
 some basic functionalities that can perform the few items above (i-iii).
@@
 \Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
+<@\textcolor{gray}{\$\{CODEDIR\}/fabfile.py}@>
 <@\textcolor{gray}{
--------------------------------------}
+-----------------------------------------}
 \Reactivatenumber @>
@@
 To control a large set of devices, we simply need to include them in the
-\texttt{env.hosts} list in the \texttt{\$CODEDIR/fabfile.py} file. Fabric has
-many other functionalities that are useful in controlling a large set of
-\ac{Raspi}s. For example, we may extract files or run automated
-experiments. The included functionalities in the
-\texttt{\$CODEDIR/fabfile.py} file will depend on the requirements of the
+\texttt{env.hosts} list in the \texttt{\$\{CODEDIR\}/fabfile.py} file.
+Fabric has many other functionalities that are useful in controlling a
+large set of \ac{Raspi}s. For example, we may extract files or run
+automated experiments. The included functionalities in the
+\texttt{\$\{CODEDIR\}/fabfile.py} file will depend on the requirements of the
 testbed administrator.

### 2016-08-24 image_setup.tex
page 1 to 10 corrections

@@
 in the command block below. We show the whole procedure by performing the role
-of the testbed administrator. To copy the commands from this procedure to the
-testbed administrator terminal, we recommend to read this document
-with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
-render properly when pasted.
+of the testbed administrator.
+%To copy the commands from this procedure to the
+%testbed administrator terminal, we recommend to read this document
+%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
+%render properly when pasted.
 
@@
 particular to the testbed administrator \ac{OS} shell. Next, we create
-the working directory and go to it with the \texttt{cd} command. To
+the working directory and go to it with the change directory (\texttt{cd}) command. To
 download the image, we utilize the \texttt{wget} command before unpacking
@@
 After Raspbian Lite has been unpacked, there should be an \texttt{.img}
-file in the \texttt{\$\{WORKDIR\}} directory. Here, \texttt{fdisk} can
+file in the working directory \texttt{\$\{WORKDIR\}}. Here, \texttt{fdisk} can
 be used to display the content of the image. To do this and obtain
@@
 sector 137216. The first partition type is FAT32 with a size of 63 MB
-and the second partition is a Linux one with a size of 1.2 GB. This
+and the second partition is of type Linux with a size of 1.2 GB. This
 indicates that the first partition is a boot partition and the second
-one is a traditional Linux root filesystem, i.e. \texttt{/}.
+one is a traditional Linux filesystem. In this case, the root filesystem, i.e. \texttt{/}.
 
 \subsection{Image Resizing}
-Given that we want to customize the files stored in the \ac{Raspi}s,
-we need to resize the image file since 1.2 GB might not be enough to store
-the root filesystem due to the total size of the additional packages. Thus,
+Given that we want to customize the root filesystem in the \ac{Raspi}s,
+we need to expand the image file since 1.2 GB might not be enough to store
+existing root filesystem plus additional files and software packages.
+%Given that we want to customize the files stored in the \ac{Raspi}s,
+%we need to resize the image file since 1.2 GB might not be enough to store
+%the root filesystem with the additional packages that will be installed in this procedure.
+%due to the total size of the additional packages.
+Thus,
 we need to increase the partition size. The following procedure illustrates
@@
 and then add it again with one GB more. To do this, we make use
-again of \texttt{fdisk} in command mode to alter the partition table.
-For this, we pass the arguments of the command mode of \texttt{fdisk}
+of \texttt{fdisk} again in command mode to alter the partition table.
+We pass the arguments of the command mode of \texttt{fdisk}
 through the \texttt{echo} command in Linux and the \texttt{|} operator
@@
 contains the root filesystem and later the boot partition. This is done by
-creating an empty directory that it is used as a mountpoint. We name it
+creating an empty directory that is used as a mountpoint. We name it
 \texttt{root} and create it in the working directory before mounting the
@@
 we describe the tasks performed by these configuration scripts. Finally, we
-indicate how and in which order are the scripts executed to configure all the
+indicate how and in which order the scripts are executed to configure all the
 devices. Nevertheless, any testbed administrator might modify or include other
@@
 In our case, we have our default configuration scripts stored in a
-\texttt{rasp\_config.zip} file located in the same \ac{URL} of the
+file \texttt{rasp\_config.zip} located in the same \ac{URL} of the
 \ac{HTTP} server where the image was retrieved from, i.e. the one in
-the environment variable \texttt{\$\{URL\}}. We first manually download
+the environment variable \texttt{\$\{URL\}}. We first download
 this compressed file with \texttt{wget} and extract it locally into our
@@
 devices from each other. Thus, in our case we require the devices in our
-testbed to have different hostnames. Hence, we define the hostnames based
+testbed to have different hostnames. We define the hostnames based
 on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.
@@
 of his / her \ac{Raspi}s obtained previously in the format shown in the
-configuration file. In the file format, on each line there is a \ac{MAC}
+configuration file. In that file format, on each line there is a \ac{MAC}
 address and the corresponding hostname for the given \ac{Raspi}. For
-simplicity in the example above, we just include three \ac{Raspi}s.
-This list file will be employed by the
+simplicity in the example above, we just include four \ac{Raspi}s.
+This file will be employed by the
 \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}
@@
 we manually did earlier to get our files but this will be made in an
-automated wat after booting up the system. This script contents are:
+automated way after booting up the system. This script content is:
 
@@
 errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
-directory, while ignoring existing files and / or directories there.
+directory. Existing files and directories are simply overwritten.
+%while ignoring existing files and / or directories there.
 
@@
 file before \texttt{exit 0} using a file editor. To do this, it is required
-to edit this file as root. The resulting file should look like the following:
+to edit this file with root permissions. The resulting file should look like the following:
 
@@
 calls all other configuration scripts (besides \texttt{update\_rasp\_config})
-in sequential order. This script content is the following:
+in a sequential order. This script content is:
 
@@
 
-In this way, the automation process is simplify since we only need to modify
+In this way, the automation process is simplified since we only need to modify
 the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script:
@@
 from the testbed administrator \ac{PC} architecture. Due to the \ac{ARM}
-processor that the \ac{Raspi}s employ, t is required to install the QEMU related software first and verigy that QEMU is \ac{ARM} enabled. To do so,
+processor that the \ac{Raspi}s employ, it is required to install the QEMU related software first and verify that QEMU is \ac{ARM} enabled. To do so,
 run the following commands:
@@
 
-Finally, to change root the following needs to be run:
+%Finally, to change root the following needs to be run:
+Finally, run the following command to change root:
 
@@
 Hence, enabling updates, upgrades and installing new software packages
-should be much faster than in a \ac{Raspi}. Still updating and upgrading
+should be faster than in a \ac{Raspi}. Still updating and upgrading
 the packages for the \ac{Raspi} might take some amount of time. To get

### 2016-08-24 introduction.tex
page 1 to 10 corrections

@@
 The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale.
-This will break new ground for new markets such as driverless cars, manufacturing, humanoid robots, smart grids, and many more.
-With 5G, the number of wireless devices will increase by 5$\times$ up to 50 billion devices \cite{cisco2011forecast}.
-It is generally accepted, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connect due to the expected energy consumption.
+This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, smart grids, and many more.
+With 5G, the number of wireless devices will increase by five times up to 50 billion devices \cite{cisco2011forecast}.
+It is generally acceptable, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connected due to the expected energy consumption.
 Therefore the communication architecture in future communication scenario will be dominated by mesh technologies.
@@
 The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages.
-Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintenable fashion.
+Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.
 
@@
 In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step.
-Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed t simplify the execution of routinary and repetitive tasks.
+Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed to simplify the execution of routinary and repetitive tasks.
 Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}.
 Conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
-Finally, a set of alternative commands procedures, in case the ones presented in this work might not be executed, are discussed in the Appendix.
+Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendix.
 

### 2016-08-24 overview.tex
page 1 to 10 corrections

@@
 university network since our testbed is used by students and
-professors to perform measurements and experimentation of controlled and
+employees to perform measurements and experimentation of controlled and
 reproducible scenarios as part of academic research. Nevertheless, the
@@
 We requested the university \ac{IT} department to configure the university
-\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP}. This
-eliminates the requirements of monitors and keyboards with the \ac{Raspi}s
+\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP} address. This
+eliminates the demand for monitors and keyboards with the \ac{Raspi}s
 for non-graphical applications. Finally, our design aims to configure all

### 2016-08-24 cross_compilation.tex
more corrections

@@
 This helps the \ac{OS} to recognize the location of the compiler
-command everytime a new shell is opened. The \texttt{.profile} should now
+command when a new shell is opened. The \texttt{.profile} should now
 contain the lines we inserted. There might be other
@@
 
-This should produce a binary \texttt{hello\_world} that should only be
-useful in the \ac{Raspi}. We copy it to a \ac{Raspi} to test it. We will
-use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s. Still, we
-for example might use Fabric if we are interested in deploying a compiled
-binary for many \ac{Raspi}s.
+This should produce a binary \texttt{hello\_world} that is executable
+in the \ac{Raspi}. We copy it to a \ac{Raspi} to test it. We will
+use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s.
+%Still, we for example might use Fabric if we are interested in deploying
+Still, we might use Fabric instead if we are interested in deploying
+a compiled binary for many \ac{Raspi}s.
 

### 2016-08-24 overlay_fs.tex
more corrections

@@
 will remain in the memory card. This is cumbersome to maintain since
-the size of the memory card is relatively reduced (8 GB) and there
+the size of the memory card is relatively small (8 GB) and there
 might be various users utilizing the testbed. Also, different testbed
@@
 does not have it itself. If the upper filesystem has the requested file,
-it will simpy return the file. This idea can be used in our setup to mount
+it will simply return the file. This idea can be used in our setup to mount
 the root filesystem (i.e. Raspbian Lite) in the \ac{Raspi}s during startup
@@
 command block below stores it in \texttt{/tmp} which is really mounted in
-\ac{RAM}. Then, all the files stored here will disappear a \ac{Raspi} is
+\ac{RAM}. Thus, all the files stored here will disappear when the system is
 rebooted.
@@
 Before enabling the overlaying filesystem, it is required to generate
-an initial \ac{RAM} file system or \texttt{initramfs}. This is an
-initial filesystem that gets loaded into \ac{RAM} during the startup
+an initial \ac{RAM} filesystem or \texttt{initramfs}. This is an
+initial filesystem that is loaded into \ac{RAM} during the startup
 process of a Linux machine to prepare the real filesystem. For this purpose,
@@
 to specify the correct kernel modules to search for. The available
-kernel modules are \texttt{/lib/modules}. To see them, we just run:
+kernel modules are located in \texttt{/lib/modules}.
+To see them, we just run:
 
@@
 in the boot partition. If the system should be run in a \ac{Raspi} version 1,
-then use \texttt{init.gz}, otherwise use \texttt{init-v7.gz} as follows:
+then use \texttt{init.gz} by executing only the first code line below,
+otherwise use \texttt{init-v7.gz} by executing only the second code line:
+%as follows:
 
@@
 \begin{lstlisting}[]
-(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # Only for Raspberry Pi version 1
+(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # For Raspberry Pi version 1
 (chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # For Raspberry Pi version 2 or 3
@@
 The following procedure creates an extra partition in the image for
-the \ac{Raspi} user home directory that will be made storage persistent
-using the upper filesystem. We first expand image according to the desired
+the \ac{Raspi} user home directory that will be made storage persistent.
+%using the upper filesystem.
+We first expand image according to the desired
 home directory size, but avoiding to make the image bigger than the target
@@
 
-We create again a loopback device and format the new partition, as follows:
+We create a loopback device again and format the new partition, as follows:
 \begin{lstlisting}[]
@@
 we need to inform the Raspbian \ac{OS} to mount the home partition
-that we have just created. We will notify the \ac{OS} while
-booting during the startup of the \ac{Raspi}s. This can be done by adding
+that we have just created.
+%We will notify the \ac{OS} while booting during the startup of the
+%\ac{Raspi}s.
+This can be done by adding
 an entry in \texttt{fstab} as follows:

### 2016-08-24 tools.tex
more corrections

@@
 
-The previous \texttt{fabfile} shows three function that perform our
+The previous \texttt{fabfile} shows three functions that perform our
 example tasks. These functions utilize variables and subsequent functions
@@
 window, a \texttt{screen} session, which does not terminate even with
-connectivity interruptions. Users can attach and detach from such a session
+connectivity interruptions. Users can attach and detach from a
+\texttt{screen} session
 as desired. The following procedure presents how to use \texttt{screen} with
@@
 and (vii) attach to screen session to see the program still running.
-From the testbed administrator, we start by establishing \ac{SSH} connection
-to a \ac{Raspi} and open a screen session:
+From the testbed administrator \ac{PC}, we start by establishing a \ac{SSH}
+connection to a \ac{Raspi} and open a screen session:
 
@@
 
-From the command output, we see that the session is still running in our
+From the command output, we can see that the session is still running in our
 generic \ac{Raspi} number \texttt{XX} and that no user is currently attached
@@
 
-After attaching again, we should see the \texttt{top} still running.
+After attaching again, we should see \texttt{top} still running.
 \texttt{screen} has more functionalities that can be used in this
@@
 To terminate the \texttt{screen} session, first terminate \texttt{top}
-by pressing \texttt{q} in the keyboard. Once out of the \texttt{top}
-output, we need to type \texttt{exit} two times to first exit the
+by pressing \texttt{q} in the keyboard.
+Once \texttt{top} is terminated,
+%Once out of the \texttt{top} output,
+we need to type \texttt{exit} two times to first exit the
 \texttt{screen} session and then terminate the \ac{SSH} connection.

### 2016-08-25 appendix.tex
page 1 to 10 corrections
more corrections

@@
 
-In case the \texttt{losetup -P} is not available given that a version
-higher or equal than 2.21 for \texttt{util-linux} is neither not available,
-a possible option is to manually set all the loopback devices used during the
-whole procedure with \texttt{mknod}. To do this, the alternative commands are:
+In case the \texttt{losetup -P} command shows the \texttt{invalid option}
+message, an alternative is to manually set all the loopback devices
+used during the whole procedure with \texttt{mknod}. To do this, the
+alternative commands are:
 
@@
 
-In case the \texttt{umount --recursive} is not available given that a version
-higher or equal than 2.22 for \texttt{util-linux} is neither not available,
-a possible option is to manually umount all the partitions used during the
-\texttt{chroot} environment. The directories must be unmounted in the reverse
-order as when they were mounted. These commands are:
+%The recursive command (\texttt{----recursive}) is not valid for \texttt{umount}
+%is not valid for \texttt{util-linux} version 2.22
+%or later. An alternative option is to manually unmount all the partitions used
+%during the \texttt{chroot} environment. The directories must be unmounted in
+%the reverse order as when they were mounter. These commands are:
+
+%The --recursive parameter is not part of umount until 
+%\texttt{umount ----recursive} is not part of \texttt{util-linux} version 2.22
+
+In case the \texttt{umount ----recursive} command shows the \texttt{invalid option}
+message, an alternative is to manually unmount all the partitions used
+during the \texttt{chroot} environment. The directories must be unmounted in
+the reverse order as when they were mounter. These commands are:
+%In case the \texttt{umount ----recursive} command fails given that a version
+%higher than 2.22 for \texttt{util-linux} is not available,
+%a possible option is to manually umount all the partitions used during the
+%\texttt{chroot} environment. The directories must be unmounted in the reverse
+%order as when they were mounted. These commands are:
 
@@
 ...
-Sucesfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
+Successfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
 # pip show paramiko

### 2016-08-25 conclusions.tex
more corrections

@@
 Observing the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enables reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system.
-The description shows how to set up interconnected \ac{Raspi}s with SD memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges.
+The description shows how to set up interconnected \ac{Raspi}s with memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges.
 We hope this work permits researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way.

### 2016-08-25 cross_compilation.tex
more corrections

@@
 \ac{CPU}, memory or space usage and / or compilation time. Thus, in
-this Section we present a procedure of how to cross-compile C++
+this section we present a procedure of how to cross-compile C++
 source code from the testbed administrator \ac{PC} for the \ac{ARM}
@@
 To do this, only a request form needs to be filled
-in~\cite{steinwurflicenselink} and wait for grant access. Once
+in~\cite{steinwurflicenselink} and wait for it to be processed by the Kodo
+developers. Once
 the access for Kodo has been granted, the source code can be pulled from
-its Git repositories to be compiled. Thus, assuming that testbed
+its Git repositories to be compiled. Thus, assuming that the testbed
 administrator already has access, we clone the \texttt{kodo-cpp}

### 2016-08-25 image_setup.tex
more corrections

@@
 \label{sec:image_setup}
-In this Section, we review the steps to create a common \ac{OS} image for
+In this section, we review the steps to create a common \ac{OS} image for
 all the \ac{Raspi}s. The image setup is composed of three major steps:
@@
 particular to the testbed administrator \ac{OS} shell. Next, we create
-the working directory and go to it with the change directory (\texttt{cd}) command. To
+the working directory and change to it with the \texttt{cd} command. To
 download the image, we utilize the \texttt{wget} command before unpacking
@@
 file in the working directory \texttt{\$\{WORKDIR\}}. Here, \texttt{fdisk} can
-be used to display the content of the image. To do this and obtain
-administrative information, we parse the arguments \texttt{-u sectors}
+be used to display the content of the image. We parse the arguments \texttt{-u sectors}
 to display the sizes in sectors and \texttt{-l} to display the partitions
@@
 \begin{lstlisting}[]
-
 Command (m for help): Partition number (1-4):
@@
 alternative option for this part is presented in
-Section~\ref{sec:alternative_losetup} of the Appendix.
+Section~\ref{sec:alternative_losetup} of the Appendices.
 
@@
 two partitions associated to it, e.g. \texttt{loop0p1} and \texttt{loop0p2}.
-Finally, we check the block device status with \texttt{e2fsck} and
-\texttt{resize2fs}, respectively in the command block:
+Finally, we check the filesystem of the block device with \texttt{e2fsck} and
+resize it with the \texttt{resize2fs} command:%, respectively in the command block:
 
@@
 the apparent root. In case the \texttt{chroot} command is not successful,
-we provide an alternative command in Section~\ref{sec:chroot} of the Appendix.
+we provide an alternative command in Section~\ref{sec:chroot} of the Appendices.
 To be aware of the mode that we are working now, we change

### 2016-08-25 introduction.tex
more corrections

@@
 With 5G, the number of wireless devices will increase by five times up to 50 billion devices \cite{cisco2011forecast}.
-It is generally acceptable, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connected due to the expected energy consumption.
-Therefore the communication architecture in future communication scenario will be dominated by mesh technologies.
+It is generally believed, that those devices will not be connected in the same manner as current devices are connected today.
+Centralized system will collapse in terms of capacity, while distributed systems appear as an alternative.
+Therefore the communication architecture in future systems will be dominated by mesh technologies.
 Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing.
@@
 This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from widely-used \ac{OS}.
-Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using Raspberry Pi devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
-Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with Raspberry Pi.
+Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using \ac{Raspi}~\cite{making_of_pi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
+Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with \ac{Raspi}~\cite{Cox2014,s141224408,s140509290,s16020220}\textcolor{red}{move citations or rewrite a bit here}.
 Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system.
@@
 Conclusions and future work are reviewed in Section~\ref{sec:conclusions}.
-Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendix.
+Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendices.
 

### 2016-08-25 overlay_fs.tex
more corrections

@@
 
-Assuming that we are still in the \texttt{chroot} environment into the
+Assuming that we are still in the \texttt{chroot} environment of the
 Raspbian Lite root filesystem for installing packages, we can setup the
@@
 Otherwise, an alternative is to either update the package or follow the
-procedure in Section~\ref{sec:umount} of the Appendix.
+procedure in Section~\ref{sec:umount} of the Appendices.
 
@@
 If the last command was executed correctly, the
-\texttt{\${ROOTDIR}/etc/fstab} file should have the new line. However,
-the line will not be properly indented. To verify the correct line
-insertion and fix the indentation, edit the file with
-\texttt{sudo vim \$\{ROOTDIR\}/etc/fstab}. The resulting file should like
-the following:
+\texttt{\$\{ROOTDIR\}/etc/fstab} file should have the new line.
+%However,
+%the line will not be properly indented. To verify the correct line
+%insertion and fix the indentation, edit the file with
+%\texttt{sudo vim \$\{ROOTDIR\}/etc/fstab}.
+The resulting file should look like the following:
 
@@
 Again, if the \texttt{----recursive} option is not available, the
-procedure in Section~\ref{sec:umount} of the Appendix could be followed.
+procedure in Section~\ref{sec:umount} of the Appendices could be followed.
 If the steps are successfully executed up to this point, the customized image
@@
 and follow the procedure in this section. A given card will be available as
-\texttt{/dev/mmcblkX} for memory card readers or as \texttt{/dev/sdbX}
-USB-through readers, where \texttt{X} a natural number.
+\texttt{/dev/mmcblkX} or \texttt{/dev/sdYX} where \texttt{X} is a natural number
+and \texttt{Y} is a letter.
 
@@
 everything will be overwritten. To avoid removing information from the
-wrong device, a testbed administrator can insert and eject a memory
-card a couple of times and using the \texttt{sudo lsblk} or \texttt{df -h}
-commands to deduce the correct device. For our case, the device was
-\texttt{/dev/sdb} in an Ubuntu distribution and
-\texttt{/dev/mmcblk0} in an Archlinux distribution. Once identified, to write
+wrong device, a testbed administrator can use the commands \texttt{lsblk}
+and/or \texttt{df -h} before and after inserting the memory card to deduce
+its correct device name.
+For our case, the device was \texttt{/dev/mmcblk0}. Once identified, to write
 the image to a memory card, the following command is used:
@@
 The previous \texttt{dd} and \texttt{sync} commands for copying the image to
-the SD memory card and flushing the remainder in memory to the filesystem
-will take around 15 minutes to be finished. This is an average on a two-core
-1.5 GHz laptop. However, other times might occur depending on the total image
-size, the writing block size and the core capabilities of the testbed
-administrator \ac{PC}. After this is made, it is only required
+the memory card and flushing the remainder in memory to the filesystem
+will take tens of minutes depending on the memory card speed and the size of
+the image. After this is made, it is only required
 to eject the memory card and now plug it in a \ac{Raspi} so it can boot up.
+
+%The previous \texttt{dd} and \texttt{sync} commands for copying the image to
+%the memory card and flushing the remainder in memory to the filesystem
+%will take around 15 minutes to be finished. This is an average on a two-core
+%1.5 GHz laptop. However, other times might occur depending on the total image
+%size, the writing block size and the core capabilities of the testbed
+%administrator \ac{PC}. After this is made, it is only required
+%to eject the memory card and now plug it in a \ac{Raspi} so it can boot up.

### 2016-08-25 overview.tex
more corrections

@@
 Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
-All \ac{Raspi}s are each equipped with a 8 GB memory card, a wired and wireless
+All \ac{Raspi}s are each equipped with a 8 GB \ac{SD} memory card, a wired and wireless
 network interface and a power supply. All the \ac{Raspi} are connected to
@@
 Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
-university network since our testbed is used by students and
-employees to perform measurements and experimentation of controlled and
+university network since our testbed is used by students and academic staff
+to perform measurements and experimentation of controlled and
 reproducible scenarios as part of academic research. Nevertheless, the
@@
 
-In Fig.~\ref{fig:testbed_setup} and what follows, we will refer as the
-testbed administrator to the person(s) in charge of setting up and
+We will refer to the
+testbed administrator as the person(s) in charge of setting up and
 configuring the testbed with administrator privileges from the \ac{OS}
-point of view. The setting and configuration procedure commands are
+point of view. The setting and configuration procedures are
 performed by the testbed administrator in a regular \ac{PC} running a
-Linux distribution. Although in principle the administrator
+Linux distribution as shown in Fig.~\ref{fig:testbed_setup}. Although in principle the administrator
 Linux distribution is not a restriction, we present our procedure in a
 Debian-based Linux distribution. Our basic design considers to create
-a customized image to later store it in a memory card for each \ac{Raspi}.
+a customized image to store it later in a memory card for each \ac{Raspi}.
 Once configured, we store the resulting image file in a \ac{HTTP} server

### 2016-08-25 tools.tex
more corrections

@@
 (i) rebooting a set of devices, (ii) installing applications in multiple
-devices and (iii) copying files to and / or from multiple devices, etc. To
+devices and (iii) copying files to and / or from multiple devices and so on. To
 simplify the management and avoid execution issues, Fabric provides
@@
 
-env.hosts = ['rasp00.domain.com','rasp01.domain.com','rasp02.domain.com']
+env.hosts = ['10.0.0.100','10.0.0.101','10.0.0.102']
 env.user = 'pi'
@@
 from the Fabric \ac{API} such as \texttt{env}, \texttt{task} and \texttt{sudo}
-among others. Each of these \ac{API} functions permit to define environment
+among others. Each of these \ac{API} functions permits to define environment
 variables, create the administrator tasks through decorators or run the
@@
 some instructions for updating the Paramiko package are available in
-Section~\ref{sec:paramiko} of the Appendix. This is a standard recommendation
+Section~\ref{sec:paramiko} of the Appendices. This is a standard recommendation
 the Fabric troubleshooting guide~\cite{2016fabricsupport}.
@@
 %Once out of the \texttt{top} output,
-we need to type \texttt{exit} two times to first exit the
+we need to type \texttt{exit} two times in order to first exit the
 \texttt{screen} session and then terminate the \ac{SSH} connection.

### 2016-08-25 reviewer_reply.txt
added reviewer replies

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for you valuable feedback and detailed comments.
They were very useful. We have entirely rewritten a few of the sentences
indicated, but in most cases made the corrections as suggested.

We like the suggested citations. The first two citations are now used
to improve our introduction, but we do not see how to fit in the last
citation "Data Exfiltration from Internet of Things Devices: iOS
Devices as Case Studies" which we think is related to IoT, but not so
much Raspberry Pi, testbeds, or network coding.

To address the applications for a Raspberry Pi testbed, we have added a few
other citations of practical usecases of the Raspberry Pi. Furthermore,
we will ask the editors to have a citation to another MDPI journal paper of
ours if it is being accepted. This paper evaluates the speed and energy
consumption of three network coding schemes in Raspberry Pi devices from our
testbed.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you for you valuable feedback

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 3:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback and suggested citations.
The additional citations has significantly helped to better address the
usecases for a Raspberry Pi testbed. We think that the
papers are great, but the core idea behind the setups in the first four
seem quite similar. We suggest to add suggested paper 2, 3, 5, and a few
papers indicated by another reviewer.

### 2016-08-25 introduction.tex
Some changes

@@
 The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area.
-This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from widely-used \ac{OS}.
-Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using \ac{Raspi}~\cite{making_of_pi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
-Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with \ac{Raspi}~\cite{Cox2014,s141224408,s140509290,s16020220}\textcolor{red}{move citations or rewrite a bit here}.
+This not only allows to seamlessly develop implementations that are compatible with higher end devices but also allows the reuse stable software from widely-used \ac{OS}.
+For example, the work in \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications.
+Here the authors also include computational, throughput and storage benchmarks to assess its performance.
+However, this work indicates provides only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading.
+Moreover, this work does not consider possible network coding applications.
+Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
+Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with \ac{Raspi}~\cite{leccese2014smartcity,leccese2014newimaging,sapes2016finger}\textcolor{red}{move citations or rewrite a bit here}.
 Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system.

### 2016-08-25 introduction.tex
New introduction

@@
 
-The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale.
+The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on global scale.
 This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, smart grids, and many more.
@@
 Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing.
-Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices.
+Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices to ensure high \ac{QoE} for the final user.
 In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.
@@
 Here the authors also include computational, throughput and storage benchmarks to assess its performance.
-However, this work indicates provides only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading.
+However, this work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading.
 Moreover, this work does not consider possible network coding applications.
-Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
-Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with \ac{Raspi}~\cite{leccese2014smartcity,leccese2014newimaging,sapes2016finger}\textcolor{red}{move citations or rewrite a bit here}.
+Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of a public road isle of lamps and reports it to a monitoring center.
+A use case regarding remote environment surveillance is presented in \cite{leccese2014newimaging}, where the \ac{Raspi} reports air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves.
+This data is sent to a monitoring center to ensure the preservation of the paintings.
+Further, authors in \cite{sapes2016finger} utilize the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints.
+Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi} making their discussed results difficult to reproduce.
+Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users.
+Furthermore, the current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.
+
+Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using potentially several \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
+Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the \ac{Raspi}.
 Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system.

### 2016-08-25 introduction.tex
minor fix

@@
 
-Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using potentially several \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
+Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configure and maintain an inexpensive testbed using potentially several \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
 Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the \ac{Raspi}.

### 2016-08-26 reviewer_reply.txt
Final revised version. Few corrections, typos and removing an extrablank page

@@
 Dear reviewer,
+
 Thank you very much for you valuable feedback and detailed comments.
-They were very useful. We have entirely rewritten a few of the sentences
-indicated, but in most cases made the corrections as suggested.
+They have been very useful. To address your comments, we list
+the ones that require our modification below and provide our answer.
 
-We like the suggested citations. The first two citations are now used
-to improve our introduction, but we do not see how to fit in the last
-citation "Data Exfiltration from Internet of Things Devices: iOS
-Devices as Case Studies" which we think is related to IoT, but not so
-much Raspberry Pi, testbeds, or network coding.
+- "To improve the article and considering that already others authors have faced problems similar even if different, the reviewer suggests to enrich the bibliography with others articles.
+For instance, in the introduction could be possible to cite these:
 
-To address the applications for a Raspberry Pi testbed, we have added a few
-other citations of practical usecases of the Raspberry Pi. Furthermore,
-we will ask the editors to have a citation to another MDPI journal paper of
-ours if it is being accepted. This paper evaluates the speed and energy
-consumption of three network coding schemes in Raspberry Pi devices from our
-testbed.
+1) ...'An Intelligent and High Efficiency Street Lighting System Isle Based on...' ...
+2) ...'A Fully Controlled Street Lighting System Isle Based on...' ...
+3)  ...'A New Acquisition and Imaging System for Environmental...' ...
+4) ...'A New Remote and Automated Control System for the Vineyard...'
+5) ...'FingerScanner: Embedding a Fingerprint Scanner in a Raspberry Pi' "
+
+We consider relevant the suggested references. Still, we include different use
+cases and journal versions of articles with similar core ideas behind. Thus, after
+reading them, references 2-3) and 5) are now used as references 11-13)
+in the new revised manuscript to improve our introduction. Besides these, in the
+revised manuscript, references 9-10) have been included to enhance
+the quality of it.
+
+- "As minor revision please check and correct the following rows: 62, 171, 546, 705."
+
+We have corrected all the typos by proof-reading the manuscript.
+Also, we have entirely rewritten a few of the sentences
+indicated as misunderstanding, but in most cases made the corrections as suggested.
 
@@
 Dear reviewer,
-Thank you for you valuable feedback
+
+Thank you for your time with the revision.
 
@@
 Dear reviewer,
+
 Thank you very much for your valuable feedback and suggested citations.
-The additional citations has significantly helped to better address the
-usecases for a Raspberry Pi testbed. We think that the
-papers are great, but the core idea behind the setups in the first four
-seem quite similar. We suggest to add suggested paper 2, 3, 5, and a few
-papers indicated by another reviewer.
+To address your comments, we list the ones that require our modification
+below and provide our answer. For simplicity, we list them by type
+of comment
+
+- Comments of the form: "The suggestions by :
+
+...
+Page XX,
+Line YY
+Line ZZ
+...
+"
+
+We have corrected all the typos by proof-reading the manuscript.
+Also, we have entirely rewritten a few of the sentences
+indicated as misunderstanding, but in most cases made the corrections as suggested.
+
+- Comment of the form: "Suggest to add:
+
+1) Cox, S. ...'Iridis-pi: a low-cost, compact demonstration cluster'...
+2) Raspberry Pi Foundation ...'The Making of Pi,'...
+3) D'Orazio C ...'Data Exfiltration from Internet of Things Devices: iOS Devices as Case Studies'...
+
+"
+
+These citations have helped to better address the
+prior work for a Raspberry Pi testbed. We think 1) and 2) are relevant
+references and are used in the new revised manuscript as references 9-10).
+Nevertheless, but we do not see how to fit the reference
+"Data Exfiltration from Internet of Things Devices: iOS
+Devices as Case Studies" which we think is related to IoT, but not very much
+to Raspberry Pi, testbeds or network coding. Besides these, references 11-13)
+in the new revised manuscript, have been included to enhance the quality of it.

### 2016-08-26 appendix.tex
Final revised version. Few corrections, typos and removing an extrablank page

@@
 message, an alternative is to manually unmount all the partitions used
-during the \texttt{chroot} environment. The directories must be unmounted in
-the reverse order as when they were mounter. These commands are:
+during the \texttt{chroot} environment in the reverse order which they were mounted.
+These commands are:
 %In case the \texttt{umount ----recursive} command fails given that a version
@@
 a key exchange algorithm error when trying to log in through \ac{SSH} to a
-Raspi. In this case, it might be necessary to update the Paramiko package
-that Fabric uses for this purpose and verify this by running:
+Raspi with Open SSH. In this case, it might be necessary to update the Paramiko
+package that Fabric uses to remove this error by running:
 
@@
 \vspace{-5mm}
-
-This should allow to make \ac{SSH} connections with Fabric by using the proper
-key exchange algorithms through Open SSH.

### 2016-08-26 introduction.tex
Final revised version. Few corrections, typos and removing an extrablank page

@@
 Network coding allows the increase of throughput, reliability, security and delay performance of the networks.
-In our previous works, we have shown that \ac{RLNC} \cite{koetter2003algebraic,ho2006random} is able to satisfy the aforementioned technical requirements.
+In previous works, we have shown that \ac{RLNC} \cite{koetter2003algebraic,ho2006random} is able to satisfy the aforementioned technical requirements.
 We have actually shown how to increase the throughput \cite{pahlevani2013playncool}, reduce the delay \cite{szabo2015towards} or support heterogeneity for coding enabled communication nodes \cite{lucani2014fulcrum}.
 
-In all these prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities.
+In our prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities.
 However, most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude.
@@
 The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area.
-This not only allows to seamlessly develop implementations that are compatible with higher end devices but also allows the reuse stable software from widely-used \ac{OS}.
+This not only allows to seamlessly develop implementations that are compatible with higher end devices, but also permits the reuse of stable software from widely-used \ac{OS}.
 For example, the work in \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications.
-Here the authors also include computational, throughput and storage benchmarks to assess its performance.
+Here the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance.
 However, this work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading.
@@
 Further, authors in \cite{sapes2016finger} utilize the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints.
-Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi} making their discussed results difficult to reproduce.
+Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi}.
 Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users.

### 2016-09-14 cross_compilation.tex
first round of corrections

--- 2016-08-24 cross_compilation.tex
@@
 \ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
-\ac{CPU}, memory or space usage and / or compilation time. Thus, in
-this Section we present a procedure of how to cross-compile C++
+\ac{CPU}, memory or space usage and / or compilation time. In
+this section we present a procedure of how to cross-compile C++
 source code from the testbed administrator \ac{PC} for the \ac{ARM}
@@
 them to our coding style and compiling convention. For this, we use the
-\texttt{\$\{TOOLCHAIN\}} directory as the working directory. However,
-the testbed administrator may choose some other working directory of
+\texttt{\$\{TOOLCHAIN\}} directory as the working directory.
+The testbed administrator may choose some other working directory of
 its preference if desired. First, we create the toolchain directory:
@@
 
-Later, we download the \ac{Raspi} toolchain for 64-bit Linux. However,
-instead of working with the binaries from \cite{2016rpitoolchain},
+Later, we download the \ac{Raspi} toolchain for 64-bit Linux.
+Instead of working with the binaries from \cite{2016rpitoolchain},
 we use a zipped version with the original binaries and renamed symbolic
@@
 To use Kodo for research, it is required to obtain a research free license.
-To do this, only a request form needs to be filled
-in~\cite{steinwurflicenselink} and wait for grant access. Once
+To do this, a request form needs to be filled
+in~\cite{steinwurflicenselink} and wait for it to be processed by the Kodo
+developers. Once
 the access for Kodo has been granted, the source code can be pulled from
-its Git repositories to be compiled. Thus, assuming that testbed
+its Git repositories to be compiled. Assuming that the testbed
 administrator already has access, we clone the \texttt{kodo-cpp}
@@
 ...
-[----------] Global test environment tear-down
-[==========] 17 tests from 15 test cases ran. (XXXX ms total)
-[  PASSED  ] 17 tests.
+[  PASSED  ] X tests.
 $ ./encode_decode_simple

### 2016-09-14 image_setup.tex
first round of corrections

@@
 through a command shell. Given that the \ac{Raspi}s in our testbed are not
-connected to monitors, we decide to work with Raspbian Lite. Still, if later
-desired, a desktop environment can be installed using the package manager
+connected to monitors, we decide to work with Raspbian Lite.
+A desktop environment can be installed using the package manager
 as it will be shown in the configuration section.
@@
 After Raspbian Lite has been unpacked, there should be an \texttt{.img}
-file in the working directory \texttt{\$\{WORKDIR\}}. Here, \texttt{fdisk} can
+file in the working directory \texttt{\$\{WORKDIR\}}. \texttt{fdisk} can
 be used to display the content of the image. We parse the arguments \texttt{-u sectors}
@@
 device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
-to customize them. We proceed first to mount the block device partition that
+to customize them. We mount the block device partition that
 contains the root filesystem and later the boot partition. This is done by
@@
 \texttt{root} and create it in the working directory before mounting the
-root filesystem onto the mountpoint. Thus, we mount the root filesystem
+root filesystem onto the mountpoint. We mount the root filesystem
 as follows:
@@
 indicate how and in which order the scripts are executed to configure all the
-devices. Nevertheless, any testbed administrator might modify or include other
+devices. Any testbed administrator might modify or include other
 tasks besides the presented in this procedure according to his / her needs
@@
 b8:27:eb:95:bd:11 rasp03
+...
 \end{lstlisting}
@@
 administrator may require to add other scripts to configure
-his / her \ac{Raspi}s. Hence, we want
+his / her \ac{Raspi}s. We want
 to ensure that all the \ac{Raspi} configuration scripts of any testbed
@@
 to fetch all our configuration scripts in
-\texttt{\$\{URL\%/*\}/raspi\_config.zip}. Thus, the update script
+\texttt{\$\{URL\%/*\}/raspi\_config.zip}. The update script
 \textit{automatically} downloads all the required configuration files in
@@
 
-The update script: (3-4) checks for the \ac{URL} and \texttt{.zip} file that
-should be downloaded, (7-10) downloads the configuration files to \texttt{/tmp}
+The update script lines (3-4) checks for the \ac{URL} and \texttt{.zip} file that
+should be downloaded. Lines (7-10) downloads the configuration files to \texttt{/tmp}
 folder in the corresponding \ac{Raspi}. It also prints a warning in case of
-errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
+errors and line (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
 directory. Existing files and directories are simply overwritten.
@@
 the \ac{URL} for file fetching and the configuration
-scripts themselves can be modified to fit their requirements. Thus, if
+scripts themselves can be modified to fit their requirements. If
 required for a testbed administrator, the \texttt{rasp\_config.zip}
@@
 locally available at each \ac{Raspi} after getting the root filesystem.
-Therefore, we first need to run the update script before running
+We first need to run the update script before running
 any other configuration scripts. To do this after boot up, we include
@@
 Provided that qemu-arm is enabled, we should now be able to \texttt{chroot}
-into our Raspbian lite image. However, there are a few commands to be
+into our Raspbian lite image. There are a few commands to be
 performed before actually changing root into the root partion of the image.
@@
 \texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
-for managing Git repositories and \texttt{screen} for better handling of
+for managing Git repositories and \texttt{screen}~\cite{gnu_screen} for better handling of
 long-runnning processes. When writing the image to a memory card, all the

### 2016-09-14 introduction.tex
first round of corrections

@@
 
-The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on global scale.
-This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, smart grids, and many more.
-With 5G, the number of wireless devices will increase by five times up to 50 billion devices \cite{cisco2011forecast}.
+The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on a global scale.
+This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, and smart grids.
+The number of wireless devices is expected to increase by five times up to 50 billion devices \cite{cisco2011forecast}.
 It is generally believed, that those devices will not be connected in the same manner as current devices are connected today.
@@
 In our prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities.
-However, most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude.
-Moreover, despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce.
-Furthermore, deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently.
+Most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude.
+Despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce.
+Deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently.
 The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages.
-Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.
+There is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.
 
@@
 Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users.
-Furthermore, the current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.
+The current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.
 
 Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configure and maintain an inexpensive testbed using potentially several \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo.
-Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the \ac{Raspi}.
+The architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the \ac{Raspi}.
 Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system.

### 2016-09-14 overlay_fs.tex
first round of corrections

@@
 users could be interested in running their experiments in a fresh
-rebooted system with the original customized image. Nevertheless, we
+rebooted system with the original customized image. We
 emphasize that this step is not necessary if the tesbed administrator
@@
 the root filesystem (i.e. Raspbian Lite) in the \ac{Raspi}s during startup
-as read-only filesystem. Thus, on the one hand, the image configuration files
+as read-only filesystem. On the one hand, the image configuration files
 will remain after a reboot but the local data in these directories will be
-erased after a hypothetic reboot. On the other hand, to enable the
+erased after a reboot. To enable the
 possibility of persistent changes, we overlay the upper filesystem that is
-mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp} as read / writable
-on top of the lower root filesystem. Thus, reading a file may return a
+mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp} as rewritable
+on top of the lower root filesystem. Reading a file may return a
 file from the lower filesytem, but if it is stored, it will be saved in
@@
 command block below stores it in \texttt{/tmp} which is really mounted in
-\ac{RAM}. Thus, all the files stored here will disappear when the system is
+\ac{RAM}. All the files stored here will disappear when the system is
 rebooted.
@@
 
-Again, if for some reason, the \texttt{-P} option is not available for
+If the \texttt{-P} option is not available for
 \texttt{losetup}, we provide an alternative command line in
 Section~\ref{sec:alternative_losetup}. Finally, if the previous filesystem
-formatting was successful, the filesystem is now available for use. Thus,
-we need to inform the Raspbian \ac{OS} to mount the home partition
+formatting was successful, the filesystem is now available for use.
+We need to inform the Raspbian \ac{OS} to mount the home partition
 that we have just created.
@@
 
-Again, if the \texttt{----recursive} option is not available, the
-procedure in Section~\ref{sec:umount} of the Appendices could be followed.
+If the \texttt{----recursive} option is not available, then follow the
+procedure in Section~\ref{sec:umount} of the Appendices.
 If the steps are successfully executed up to this point, the customized image

### 2016-09-14 overview.tex
first round of corrections

@@
 to perform measurements and experimentation of controlled and
-reproducible scenarios as part of academic research. Nevertheless, the
+reproducible scenarios as part of academic research. The
 testbed description and procedures for setting it up are not restricted
@@
 a single place where system setup is stored and could be modified.
-This simplifies the system maintenance at a later stage.
-As another design consideration, when working on a testbed
-with many devices, it may not always be desired to make persistent changes
-on them. One scenario is when different users are interested in
-running experiments on a rebooted testbed. Thus, we later present how
+This simplifies the system maintenance.
+As
+%another design consideration, when working on a testbed with many devices,
+it may not always be desirable to make persistent changes
+on the \ac{Raspi}s. For example when different users are interested in
+running experiments on a rebooted testbed. We later present how
 to utilize stacked filesystems to enable both persistent and

### 2016-09-14 tools.tex
first round of corrections

@@
 otherwise the commands need to be run locally in a \ac{Raspi} making
-necessary to use a keyboard and a monitor. To do so, the testbed
+necessary to use a keyboard and a monitor. The testbed
 administrator needs to put the memory cards in the \ac{Raspi}s and turn them
@@
 devices and (iii) copying files to and / or from multiple devices and so on. To
-simplify the management and avoid execution issues, Fabric provides
+simplify the management and avoid execution issues, Fabric~\cite{python_fabric} provides
 a Python library to ease the major task of working with multiple devices from
@@
 In general, other administrators may require different functionalities
-but that goes out of the scope of this work. Nevertheless, the following
+but that goes out of the scope of this work. The following
 file serves as a starting boilerplate:
@@
 the second function installs a program given by an argument. For the
-connection to the devices, Fabric calls the Paramiko module from Python
+connection to the devices, Fabric calls the Paramiko~\cite{python_paramiko} module from Python
 to make a \ac{SSH} connection. For this to work properly, the Paramiko
 version needs to be higher than or equal to 1.15.1. If not available,
-the \ac{SSH} connections from Fabric may fail. Thus in case of any problems,
+the \ac{SSH} connections from Fabric may fail. In case of any problems,
 some instructions for updating the Paramiko package are available in
@@
 The \texttt{top} command simply shows continuously the table of processes
-executed in the \ac{Raspi} like in any Linux distribution. When \texttt{top}
+executed on the \ac{Raspi} like in any Linux distribution. When \texttt{top}
 is running, we first press \texttt{Ctrl+a} and later \texttt{Crtl+d}

### 2016-09-14 introduction.tex
Corrections up to page 3

@@
 It is generally believed, that those devices will not be connected in the same manner as current devices are connected today.
-Centralized system will collapse in terms of capacity, while distributed systems appear as an alternative.
-Therefore the communication architecture in future systems will be dominated by mesh technologies.
+Centralized systems will collapse in terms of capacity, while distributed systems appear as an alternative.
+Therefore, we believe mesh technologies will play a major role in the communication architecture in future systems.
 Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing.
@@
 
-The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area.
-This not only allows to seamlessly develop implementations that are compatible with higher end devices, but also permits the reuse of stable software from widely-used \ac{OS}.
-For example, the work in \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications.
+The emergence of powerful and inexpensive single-board computers opens new possibilities in this area.
+By running a standard \ac{OS}, they allow implementations that are compatible with higher end devices.
+Also, they utilize stable software supported by their communities.
+For example, the Iridis-pi platform \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications.
 Here the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance.
-However, this work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading.
+This work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading.
 Moreover, this work does not consider possible network coding applications.
-Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of a public road isle of lamps and reports it to a monitoring center.
-A use case regarding remote environment surveillance is presented in \cite{leccese2014newimaging}, where the \ac{Raspi} reports air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves.
+Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of an isle of lamps on a public road and reports it to a monitoring center.
+A use case regarding remote environment surveillance using the Raspi and the Arduino \cite{arduino} tehnologies is presented in \cite{leccese2014newimaging}.
+Here both devices report air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves.
 This data is sent to a monitoring center to ensure the preservation of the paintings.
-Further, authors in \cite{sapes2016finger} utilize the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints.
+Further, authors in \cite{sapes2016finger} consider FingerScanner, a technology that utilizes the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints.
 Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi}.
-Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users.
+These applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users.
 The current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.

### 2016-09-14 overview.tex
Corrections up to page 3

@@
 testbed description and procedures for setting it up are not restricted
-to the academic scenario. All \ac{Raspi}s are configured to run
+to this academic scenario. All \ac{Raspi}s are configured to run
 a \ac{SSH} daemon for easy remote access within the university network.
@@
 point of view. The setting and configuration procedures are
-performed by the testbed administrator in a regular \ac{PC} running a
+performed by the testbed administrator in a \ac{PC} running a
 Linux distribution as shown in Fig.~\ref{fig:testbed_setup}. Although in principle the administrator
@@
 Debian-based Linux distribution. Our basic design considers to create
-a customized image to store it later in a memory card for each \ac{Raspi}.
+a customized image to store it later on a memory card for each \ac{Raspi}.
 Once configured, we store the resulting image file in a \ac{HTTP} server
 as backup and in case the testbed administrator requires to make new
-changes to this file. We also put all the required configuration files
+changes to this file. For our case, we use the university \ac{HTTP} server but any other servers can be used. We also put all the required configuration files
 and scripts for the \ac{Raspi}s setup in the \ac{HTTP} server so there is

### 2016-09-16 image_setup.tex
Today changes

@@
 steps, we indicate with command-line blocks the required sequential
-commands to be typed by the testbed administrator in his / her \ac{PC}
-to obtain the desired setting. In all the following sections, the number
-sign (\#) and dollar sign (\$) will be used in the command blocks in the
-paper to indicate if an operation needs to be run with root permissions or
-common user permissions, respectively. The signs will be posted as the
-terminal prompt in the command blocks.
+commands to be typed by the testbed administrator on his / her \ac{PC}
+to obtain the desired setting.
+In all the command blocks in the paper, we indicate if a command needs to be run with root permissions (\#) or common user permissions (\$).
+These signs will prefix the commands.
 

### 2016-09-16 overview.tex
Today changes

@@
 as backup and in case the testbed administrator requires to make new
-changes to this file. For our case, we use the university \ac{HTTP} server but any other servers can be used. We also put all the required configuration files
-and scripts for the \ac{Raspi}s setup in the \ac{HTTP} server so there is
-a single place where system setup is stored and could be modified.
-This simplifies the system maintenance.
-As
-%another design consideration, when working on a testbed with many devices,
-it may not always be desirable to make persistent changes
+changes to this file. In our case, we use the university \ac{HTTP} server but
+any others can be used. We also put all the required
+configuration files and scripts for the \ac{Raspi}s setup in the \ac{HTTP}
+server so there is a single place where system setup is stored and could be
+modified. This simplifies the system maintenance.
+As it may not always be desirable to make persistent changes
 on the \ac{Raspi}s. For example when different users are interested in

### 2016-09-16 image_setup.tex
Setting DOI for configuration scripts and raw image

@@
 is based on that particular version of Raspbian Lite which we have
-made available at~\cite{tunescode_webpage} acting as our \ac{HTTP} server.
+made available at~\cite{soerensen_chres_wiant_2016_154143} acting as our
+\ac{HTTP} server.
 All other files used in this paper are available there unless specified

### 2016-09-16 overlay_fs.tex

@@
 observe the new partition should start at sector 4806656. To create it, we
-show the introduced command and its output:
+show the introduced command:% and its output:
 
+%\begin{lstlisting}[]
+%$ (echo n; echo p; echo 3; echo 4806656; echo ; echo w) | fdisk ${IMAGE}.img
+%
+%Command (m for help): Partition type:
+%   p primary (1 primary, 0 extended, 3 free)
+%   e extended
+%Partition number (1-4, default 3): Using default value 3
+%First sector (2048-6903807, default 2048): Last sector, +sectors or
+%+size{K,M,G} (4806656-6903807, default 6903807): Using default value 6903807
+%
+%Command (m for help): The partition table has been altered!
+%
+%Syncing disks.
+%\end{lstlisting}
 \begin{lstlisting}[]
-$ (echo n; echo p; echo 3; echo 4806656; echo ; echo w) | fdisk ${IMAGE}.img
+$ fdisk ${IMAGE}.img <<EOF
+n
+p
+3
+4806656
 
-Command (m for help): Partition type:
-   p primary (1 primary, 0 extended, 3 free)
-   e extended
-Partition number (1-4, default 3): Using default value 3
-First sector (2048-6903807, default 2048): Last sector, +sectors or
-+size{K,M,G} (4806656-6903807, default 6903807): Using default value 6903807
-
-Command (m for help): The partition table has been altered!
-
-Syncing disks.
+w
+EOF
 \end{lstlisting}
@@
 \vspace{-5mm}
+
+%\textcolor{red}{SHOW NEW PARTITION TABLE HERE}
 

### 2016-09-16 image_setup.tex
changed fdisks
Corrections

--- 2016-09-16 image_setup.tex
@@
 is based on that particular version of Raspbian Lite which we have
-made available at~\cite{tunescode_webpage} acting as our \ac{HTTP} server.
-All other files used in this paper are available there unless specified
-otherwise. To get started, the testbed administrator must open a Linux
+made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are available there. The testbed
+administrator has to move these files to his / her \ac{HTTP} server.
+To get started, the testbed administrator must open a Linux
 shell (terminal) in its \ac{PC} and declare the environment variables shown
@@
 
-Now, in this command block output, we observe that the change has taken
+Now, in the above command block output, we observe that the change has taken
 effect by noticing the total available image size is 2.3 GiB. To expand the
-root filesystem, it is required to first remove the Linux partition
-and then add it again with one GB more. To do this, we make use
-of \texttt{fdisk} again in command mode to alter the partition table.
-We pass the arguments of the command mode of \texttt{fdisk}
-through the \texttt{echo} command in Linux and the \texttt{|} operator
-(pipe) as follows:
+root filesystem, we replace the Linux partition with a new partition one GB
+larger. The starting point of this new partition should be the same as the old
+one. We make use of \texttt{fdisk} to alter the partition table
+in the commands below. They (i) delete partition number 2, (ii) create a new
+primary partition and (iii) set the new partition starting point. The new
+partition starting point value is 137216 in our case. Finally, we (iv) write
+the new partition table to the image file. This is made as follows:
 
+%\begin{lstlisting}[]
+%$ (echo d; echo 2; echo n; echo ; echo ; echo 137216; echo ; echo w) | fdisk ${IMAGE}.img
+%\end{lstlisting}
 \begin{lstlisting}[]
-$ (echo d; echo 2; echo n; echo ; echo ; echo 137216; echo ; echo w) | fdisk ${IMAGE}.img
+$ fdisk ${IMAGE}.img << EOF
+d
+2
+n
+p
+2
+137216
+
+w
+EOF
 \end{lstlisting}
@@
 
-The \texttt{echo} commands within the parenthesis are interpreted as
-key-presses in the \texttt{fdisk} commmand mode. They (i) delete partition
-number 2, (ii) create a new primary type partition by default, (iii) set the
-new partition starting point (137216) and (iv) write new partition table to
-the image file. The previously shown command options of \texttt{fdisk}
-should work for both old and new versions of the commands. However, if this
-is not the case, the testbed administrator can do these commands manually.
-In this case, it is only required to delete the old root
-filesystem partition and create the new one starting from the same sector
-number. In our case, if the partitions commands were correct, the output
-should be following:
+If the partitions commands were correct, the partition table should now
+look like the following: \todot{PRINT PARTITION TABLE}
 
-\begin{lstlisting}[]
-Command (m for help): Partition number (1-4):
-Command (m for help): Partition type:
-   p   primary (1 primary, 0 extended, 3 free)
-   e   extended
-Select (default p): Using default response p
-Partition number (1-4, default 2): Using default value 2
-First sector (2048-4806655, default 2048): Last sector, +sectors or
-+size{K,M,G} (137216-4806655, default 4806655): Using default value 4806655
+% \begin{lstlisting}[]
+% Command (m for help): Partition number (1-4):
+% Command (m for help): Partition type:
+%    p   primary (1 primary, 0 extended, 3 free)
+%    e   extended
+% Select (default p): Using default response p
+% Partition number (1-4, default 2): Using default value 2
+% First sector (2048-4806655, default 2048): Last sector, +sectors or
+% +size{K,M,G} (137216-4806655, default 4806655): Using default value 4806655
 
-Command (m for help): The partition table has been altered!
+% Command (m for help): The partition table has been altered!
 
-Syncing disks.
+% Syncing disks.
 
-\end{lstlisting}
-\FloatBarrier
-\vspace{-5mm}
+% \end{lstlisting}
+% \FloatBarrier
+% \vspace{-5mm}
 
@@
 devices. Any testbed administrator might modify or include other
-tasks besides the presented in this procedure according to his / her needs
-as we will show.
+tasks according to his / her needs as we will show.
 
@@
 The unzippped files are one configuration file and three configuration
-scripts put in the newly created \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/}
+scripts in the newly created \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/}
 folder in the image. In what follows, we describe which features do we
@@
 \subsubsection{Device Hostnames}
-Among the settings that we want to be different in the devices is
-their hostname. The hostname helps the user to physically distinguish the
-devices from each other. Thus, in our case we require the devices in our
+The hostname helps the user to physically distinguish the
+devices from each other. In our case we require the devices in our
 testbed to have different hostnames. We define the hostnames based
@@
 of his / her \ac{Raspi}s obtained previously in the format shown in the
-configuration file. In that file format, on each line there is a \ac{MAC}
-address and the corresponding hostname for the given \ac{Raspi}. For
-simplicity in the example above, we just include four \ac{Raspi}s.
-This file will be employed by the
-\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}
-\ac{Bash} script to assign the hostname of each \ac{Raspi}.
-The script content is the following:
+configuration file. For each given \ac{Raspi}, there is a \ac{MAC} address and
+the corresponding hostname. This file will be employed by the
+\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} \ac{Bash} script to
+assign the hostname of each \ac{Raspi}. The script content is the following:
 

### 2016-09-16 tools.tex
Corrections

@@
 commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
-otherwise the commands need to be run locally in a \ac{Raspi} making
+otherwise the commands need to be run locally on a \ac{Raspi} making
 necessary to use a keyboard and a monitor. The testbed
 administrator needs to put the memory cards in the \ac{Raspi}s and turn them
-on, for them to be able to boot. The devices should boot properly and have
-connectivity if the image and \ac{DHCP} server were properly set previously.
+on, for them to be able to boot. The devices should now be bootable.
 
@@
 (i) rebooting a set of devices, (ii) installing applications in multiple
-devices and (iii) copying files to and / or from multiple devices and so on. To
-simplify the management and avoid execution issues, Fabric~\cite{python_fabric} provides
-a Python library to ease the major task of working with multiple devices from
-a single \ac{PC}. First, the testbed administrator creates a directory to
-hold the \texttt{Fabric} source code:
+devices and (iii) copying files to / from multiple devices.
+Fabric~\cite{python_fabric} provides a Python library that simplifies the
+management of working with many devices from a single \ac{PC}. First, the
+testbed administrator creates a directory to hold the \texttt{Fabric} source
+code:
 

### 2016-09-17 conclusions.tex
Some more corrections up to the conclusions

@@
 \label{sec:conclusions}
-Observing the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enables reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system.
+Observing the expectation of the \ac{IoT} and lack for a low-cost, easy-to-configure testbed in this area for reproducible research, we provide an in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system.
 The description shows how to set up interconnected \ac{Raspi}s with memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges.
+Using the presented procedure permits to setup a Raspbian Lite for the \ac{Raspi}s, but this restrics the testbed administrators to specifically using this \ac{OS}.
+Other Linux distributions might be setup with Yocto \todot{Include this as a pro/con?
+Nestor: Here I would say that a good idea to address this comment is to compare our approach with another one :)} to create an image from the scratch, but that goes beyond the scope of this procedure.
 We hope this work permits researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way.
 Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.
+Future work in this area will consider to make the testbed fetch the image through the \ac{HTTP} server.
+This is expected to simplify the maintenance of the memory cards.

### 2016-09-17 cross_compilation.tex
Corrections
Some more corrections up to the conclusions

@@
 where the source can be compiled from.
-Given that compiling a toolchaing can be an ardous task, we get the toolchain
+Given that compiling a toolchaing can be an arduous task, we get the toolchain
 recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
-This toolchain is available from \cite{2016rpitoolchain} and
+This toolchain is available from \cite{2016steinwurfrpitoolchain} and
 it already contains the binaries for different compilers based on
@@
 
-Later, we download the \ac{Raspi} toolchain for 64-bit Linux.
-Instead of working with the binaries from \cite{2016rpitoolchain},
-we use a zipped version with the original binaries and renamed symbolic
-links to the original files available in
-\cite{2016steinwurfrpitoolchain}. Finally, we unzip the downloaded file.
-This is made as follows:
+Later, we download a \ac{Raspi} toolchain with the binaries for a 64-bit Linux
+distribution available in \cite{2016steinwurfrpitoolchain}. Finally, we unzip the
+downloaded file. This is made as follows:
 
@@
 This should produce a binary \texttt{hello\_world} that is executable
-in the \ac{Raspi}. We copy it to a \ac{Raspi} to test it. We will
-use \ac{SCP} to copy the binary file to one of the \ac{Raspi}s.
-%Still, we for example might use Fabric if we are interested in deploying
-Still, we might use Fabric instead if we are interested in deploying
-a compiled binary for many \ac{Raspi}s.
+on the \ac{Raspi}. We copy it to a \ac{Raspi} using \ac{SCP} and using
+Fabric instead if we are interested in deploying a compiled binary for
+many \ac{Raspi}s.
 

### 2016-09-18 image_setup.tex
changes

@@
 on all the \ac{Raspi} models. We will download and setup the image in
-the testbed administrator \ac{PC} using a Debian-based distribution. We
+the testbed administrator \ac{PC} using a Debian-based distribution.
+\todot{We could present the two methods here (Yocto vs our method).}
+We
 use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
@@
 connected to monitors, we decide to work with Raspbian Lite.
-A desktop environment can be installed using the package manager
-as it will be shown in the configuration section.
+If required, a desktop environment can be installed using the package manager later.
+%\todot{remove the rest of this line} as it will be shown in the configuration section.
 
@@
 is based on that particular version of Raspbian Lite which we have
-made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are available there. The testbed
-administrator has to move these files to his / her \ac{HTTP} server.
+made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are also available there. The testbed
+administrator has to move these files to his / her own \ac{HTTP} server.
 To get started, the testbed administrator must open a Linux
@@
 \vspace{-5mm}
+\todot{change to zenodo link above}
 
@@
 \begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
+DOLLAR fdisk -u=sectors -l ${IMAGE}.img
 Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
@@
 we need to expand the image file since 1.2 GB might not be enough to store
-existing root filesystem plus additional files and software packages.
+the existing root filesystem plus additional files and software packages.
 %Given that we want to customize the files stored in the \ac{Raspi}s,
@@
 \begin{lstlisting}[literate={DOLLAR}{\$}1]
-DOLLAR fdisk -u sectors -l ${IMAGE}.img
+DOLLAR fdisk -u=sectors -l ${IMAGE}.img
 Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
@@
 If the partitions commands were correct, the partition table should now
-look like the following: \todot{PRINT PARTITION TABLE}
+look like the following:
+\begin{lstlisting}[literate={DOLLAR}{\$}1]
+DOLLAR fdisk -u=sectors -l ${IMAGE}.img
+Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
+Units: sectors of 1 * 512 = 512 bytes
+Sector size (logical/physical): 512 bytes / 512 bytes
+I/O size (minimum/optimal): 512 bytes / 512 bytes
+Disklabel type: dos
+Disk identifier: 0x6fcf21f3
+
+Device                               Boot  Start     End Sectors  Size Id Type
+2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
+2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
 
@@
 the Raspbian image available as a block device in the filesystem. For this
-command to work, the tested administrator distribution must have the
+command to work, the testbed administrator distribution must have the
 \texttt{util-linux} package with version 2.21 or higher. Otherwise, the
@@
 \begin{lstlisting}[]
-# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
+$ export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
 /dev/loop0
@@
 any other configuration scripts. To do this after boot up, we include
-a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
-file before \texttt{exit 0} using a file editor. To do this, it is required
-to edit this file with root permissions. The resulting file should look like the following:
+a call for the update script in \texttt{\$\{ROOTDIR\}/etc/rc.local} before
+\texttt{exit 0} in the file:
+
+%before \texttt{exit 0} using a file editor. To do this, it is required
+%to edit this file with root permissions. The resulting file should look like the following:
 
 \Suppressnumber\begin{lstlisting}[]
-<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
-<@\textcolor{gray}{
----------------------------------------------}
-\Reactivatenumber @>
-...
-bash /home/pi/rasp_config/update_rasp_config
-...
-exit 0
+# sed -i '/^exit 0/i bash /home/pi/rasp_config/update_rasp_config' ${ROOTDIR}/etc/rc.local
 \end{lstlisting}
@@
 \vspace{-5mm}
+
+%\Suppressnumber\begin{lstlisting}[]
+%<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
+%<@\textcolor{gray}{
+%---------------------------------------------}
+%\Reactivatenumber @>
+%...
+%bash /home/pi/rasp_config/update_rasp_config
+%...
+%exit 0
+%\end{lstlisting}
+%\FloatBarrier
+%\vspace{-5mm}
 
@@
 
+\todot{insert sed command below instead}
 \Suppressnumber\begin{lstlisting}[]

### 2016-09-18 overview.tex
changes

@@
 as backup and in case the testbed administrator requires to make new
-changes to this file. In our case, we use the university \ac{HTTP} server but
-any others can be used. We also put all the required
+changes to this file.
+%In our case, we use the university \ac{HTTP} server but any others can be used.
+In our case, we store all files at Zenodo~\cite{soerensen_chres_wiant_2016_154143}, but the testbed administrator should copy the our files to his / her own \ac{HTTP} server to get read/write permissions.
+We also put all the required
 configuration files and scripts for the \ac{Raspi}s setup in the \ac{HTTP}
 server so there is a single place where system setup is stored and could be
-modified. This simplifies the system maintenance.
-As it may not always be desirable to make persistent changes
+modified. This simplifies the system maintenance, as it may not always be desirable to make persistent changes
 on the \ac{Raspi}s. For example when different users are interested in

### 2016-09-18 image_setup.tex
minor changes

@@
 \begin{lstlisting}[]
-$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
+$ export URL="https://zenodo.org/record/154328/files/"
 $ export IMAGE="2016-05-27-raspbian-jessie-lite"
@@
 $ cd ${WORKDIR}
-$ wget ${URL%/*}/${IMAGE}.zip
+$ wget ${URL%/}/${IMAGE}.zip
 $ unzip ${IMAGE}.zip
@@
 \begin{lstlisting}[]
-$ wget ${URL%/*}/rasp_config.zip
+$ wget ${URL%/}/rasp_config.zip
 $ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
@@
 to fetch all our configuration scripts in
-\texttt{\$\{URL\%/*\}/raspi\_config.zip}. The update script
+\texttt{\$\{URL\%/\}/raspi\_config.zip}. The update script
 \textit{automatically} downloads all the required configuration files in

### 2016-09-18 image_setup.tex
Include Yocto reference

@@
 the testbed administrator \ac{PC} using a Debian-based distribution.
-\todot{We could present the two methods here (Yocto vs our method).}
-We
-use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
-the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
-available in two bundles: Raspbian and Raspbian Lite. The difference
+An alternative to this method, is to create a tailoared Linux distribution
+for the \ac{Raspi} platform using the Yocto Project \cite{yocto}. However, this
+process would require to assemble and compile all the software for the
+\ac{Raspi} platform from scratch which goes beyond of the scope of our
+work. We use the popular Debian-based Raspbian Linux~\cite{raspbian} given
+that is the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is
+made available in two bundles: Raspbian and Raspbian Lite. The difference
 between the two, is that Raspbian contains a pre-installed desktop environment
@@
 \vspace{-5mm}
-\todot{change to zenodo link above}
 
@@
 \texttt{\$\{WORKDIR\}} specifies a working directory where the Raspbian
-Lite bundle will be downloaded and customized. Notice that even though we
+Lite bundle will be downloaded and customized. If the testbed administrator
+allocates his / her files into another location, then it will be required to
+change the \texttt{\$\{URL\}} environment variable. Notice that even though we
 use the \$ and \# signs in the shell, in general these signs will be
@@
 
-url="http://kom.aau.dk/project/TuneSCode/raspi/"
+url="https://zenodo.org/record/154328/files/"
 config_file="rasp_config.zip"

### 2016-09-18 image_setup.tex
updated image setup

@@
 
-In this way, the automation process is simplified since we only need to modify
-the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script:
+In this way, the automation process is simplified since we do not need to
+modity \texttt{\$\{ROOTDIR\}/etc/rc.local} again after the image has been
+written to the memory cards.
+%In this way, the automation process is simplified since we only need to modify
+%the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script.
+Now, insert a call to the \texttt{main} script in \texttt{\$\{ROOTDIR\}/etc/rc.local}:
 
-\todot{insert sed command below instead}
+%\todot{insert sed command below instead}
+\Suppressnumber\begin{lstlisting}[]
+# sed -i '/^exit 0/i bash /home/pi/rasp_config/main' ${ROOTDIR}/etc/rc.local
+\end{lstlisting}
+\FloatBarrier
+\vspace{-5mm}
+
+Finally, \texttt{\$\{ROOTDIR\}/etc/rc.local} should look like the following:
+
 \Suppressnumber\begin{lstlisting}[]
@@
 bash /home/pi/rasp_config/main
-...
 exit 0

### 2016-09-18 conclusions.tex

@@
 The description shows how to set up interconnected \ac{Raspi}s with memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges.
-Using the presented procedure permits to setup a Raspbian Lite for the \ac{Raspi}s, but this restrics the testbed administrators to specifically using this \ac{OS}.
-Other Linux distributions might be setup with Yocto \todot{Include this as a pro/con?
-Nestor: Here I would say that a good idea to address this comment is to compare our approach with another one :)} to create an image from the scratch, but that goes beyond the scope of this procedure.
+Using the presented procedure permits to setup a Raspbian Lite for the \ac{Raspi}s, but this restrics the testbed administrators to specifically using this \ac{OS} regardless of the \ac{Raspi} model.
+Other Linux distributions might be setup with Yocto to create an image from the scratch that are tailored for the \ac{Raspi}, even to specific models.
+However, that would require to compile the software for the \ac{Raspi} from scratch which can be a tedious angd long task.
+Still, this method could be adequate for an expert user.
 We hope this work permits researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way.

### 2016-09-18 conclusions.tex
Corrections for the conclusions

@@
 The description shows how to set up interconnected \ac{Raspi}s with memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges.
-Using the presented procedure permits to setup a Raspbian Lite for the \ac{Raspi}s, but this restrics the testbed administrators to specifically using this \ac{OS} regardless of the \ac{Raspi} model.
-Other Linux distributions might be setup with Yocto to create an image from the scratch that are tailored for the \ac{Raspi}, even to specific models.
-However, that would require to compile the software for the \ac{Raspi} from scratch which can be a tedious angd long task.
+Using the presented procedure permits to setup a Raspbian Lite image for the \ac{Raspi}s.
+A tailored Linux distribution might be created from the scratch using the Yocto project.
+However, to assemble and compile the software for the \ac{Raspi} can be a tedious and time-consuming task.
 Still, this method could be adequate for an expert user.

### 2016-09-18 image_setup.tex
Corrections for the conclusions

@@
 In this way, the automation process is simplified since we do not need to
-modity \texttt{\$\{ROOTDIR\}/etc/rc.local} again after the image has been
+modify \texttt{\$\{ROOTDIR\}/etc/rc.local} again after the image has been
 written to the memory cards.
@@
 %the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script.
-Now, insert a call to the \texttt{main} script in \texttt{\$\{ROOTDIR\}/etc/rc.local}:
+Now, we insert a call to the \texttt{main} script in \texttt{\$\{ROOTDIR\}/etc/rc.local} as follows:
 

### 2016-09-18 appendix.tex
changed toolchain link

@@
 This section describes alternative commands in case the testbed administrator
-is not able to update old packages in its Linux distribution for performing
+is not able to update old packages on his / her Linux distribution for performing
 the commands, particularly the ones regarding the \texttt{util-linux} package

### 2016-09-18 cross_compilation.tex
changed toolchain link

@@
 To compile in a given architecture that is aimed for a different one,
-the testbed administrator needs to install a toolchain in its \ac{PC}.
+the testbed administrator needs to install a toolchain on his / her \ac{PC}.
 The toolchain is mandatory due to the different processor architectures
 where the source can be compiled from.
-Given that compiling a toolchaing can be an arduous task, we get the toolchain
+Given that compiling a toolchain can be an arduous task, we get the toolchain
 recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
-This toolchain is available from \cite{2016steinwurfrpitoolchain} and
+This toolchain is available from \cite{soerensen_chres_wiant_2016_154143} and
 it already contains the binaries for different compilers based on
@@
 Later, we download a \ac{Raspi} toolchain with the binaries for a 64-bit Linux
-distribution available in \cite{2016steinwurfrpitoolchain}. Finally, we unzip the
+distribution available in \cite{soerensen_chres_wiant_2016_154143}. Finally, we unzip the
 downloaded file. This is made as follows:
@@
 \begin{lstlisting}[]
-$ wget http://files.steinwurf.com/toolchains/linux/raspberry-gcc-4.9.3/raspberry-gxx493-arm.zip
+$ wget https://zenodo.org/record/154328/files/raspberry-gxx493-arm.zip
 $ unzip raspberry-gxx493-arm.zip

### 2016-09-18 image_setup.tex
changed toolchain link

@@
 To get started, the testbed administrator must open a Linux
-shell (terminal) in its \ac{PC} and declare the environment variables shown
+shell (terminal) on his / her \ac{PC} and declare the environment variables shown
 in the command block below. We show the whole procedure by performing the role

### 2016-09-18 overlay_fs.tex
Corrections
changed toolchain link

@@
 \begin{lstlisting}[]
-$ fdisk -u sectors -l ${IMAGE}.img
+$ fdisk -u=sectors -l ${IMAGE}.img
 Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
@@
 observe the new partition should start at sector 4806656. To create it, we
-show the introduced command:% and its output:
+use \texttt{fdisk} as follows:% and its output:
 
@@
 \begin{lstlisting}[]
-$ fdisk ${IMAGE}.img <<EOF
+$ fdisk ${IMAGE}.img << EOF
 n
@@
 
-%\textcolor{red}{SHOW NEW PARTITION TABLE HERE}
+\textcolor{red}{SHOW NEW PARTITION TABLE HERE}
 
@@
 our current considered system, we do this manually for each card.
-The testbed administrator needs to insert each memory card in its \ac{PC}
+The testbed administrator needs to insert each memory card in his / her \ac{PC}
 and follow the procedure in this section. A given card will be available as

### 2016-09-18 image_setup.tex
minor
draf

@@
 
-The update script lines (3-4) checks for the \ac{URL} and \texttt{.zip} file that
+The update script lines (3-4) specifies the \ac{URL} and \texttt{.zip} file that
 should be downloaded. Lines (7-10) downloads the configuration files to \texttt{/tmp}
@@
 should be faster than in a \ac{Raspi}. Still updating and upgrading
-the packages for the \ac{Raspi} might take some amount of time. To get
-the upgraded system packages, it is necessary to run:
+the packages for the \ac{Raspi} might take some amount of time.
+%To get the upgraded system packages, it is necessary to run:
+To update the system package list, run the following command:
 
@@
 (chroot) # apt-get update
-(chroot) # apt-get upgrade
 \end{lstlisting}

### 2016-09-20 image_setup.tex
minor change
fixed line numbers

--- 2016-09-18 image_setup.tex
@@
 
-\Suppressnumber\begin{lstlisting}[]
+\begin{lstlisting}[]
 # sed -i '/^exit 0/i bash /home/pi/rasp_config/update_rasp_config' ${ROOTDIR}/etc/rc.local
@@
 %\todot{insert sed command below instead}
-\Suppressnumber\begin{lstlisting}[]
+\begin{lstlisting}[]
 # sed -i '/^exit 0/i bash /home/pi/rasp_config/main' ${ROOTDIR}/etc/rc.local
@@
 should be faster than in a \ac{Raspi}. Still updating and upgrading
-the packages for the \ac{Raspi} might take some amount of time. To get
-the upgraded system packages, it is necessary to run:
+the packages for the \ac{Raspi} might take some amount of time.
+To get the upgraded system packages, it is necessary to run:
+%To update the system package list, run the following command:
 

### 2016-09-20 overlay_fs.tex
draf
fixed line numbers again

@@
 
-\textcolor{red}{SHOW NEW PARTITION TABLE HERE}
+%\textcolor{red}{SHOW NEW PARTITION TABLE HERE}
 
@@
 
-\begin{lstlisting}[]
+\Suppressnumber\begin{lstlisting}[]
 <@\textcolor{gray}{\$\{ROOTDIR\}/etc/fstab}@>

### 2016-10-10 electronics-144007-proof/electronics-144007-english.tex
proof read

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[electronics,article,accept,moreauthors,pdftex,10pt,a4paper]{mdpi}
\synctex=1
\usepackage{fixme}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}             % Include the listings-package
\usepackage{placeins}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{textcomp}  % To get the correct quotes in lstlistings (')
\usepackage{appendix}
\usepackage{soul}

\newcommand\floor[1]{\left\lfloor#1\right\rfloor}
\newcommand\ceil[1]{\left\lceil#1\right\rceil}

\newcommand{\todot}[1]{%
  \textbf{\textcolor{red}{TO DO: #1}}
}

%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, atmosphere, atoms, axioms, batteries, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, condensedmatter, cosmetics, cryptography, crystals, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, fishes, fluids, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, inventions, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, lubricants, machines, magnetochemistry, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nutrients, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, universe, urbansci, vaccines, vetsci, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by:
% addendum, article, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimage, letter, newbookreceived, opinion, obituary, projectreport, reply, retraction, review, sciprints, shortnote, supfile, technicalnote
% supfile = supplementary materials
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. In addition, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figure are used, remove the option pdftex and use LaTeX and dvi2pdf.

% *** ACRONYM PACKAGE ***
\usepackage[nolist]{acronym}
%\input{acronym} % Name of acronyms file

% \acf{GS} writes Ground Station
% \ac{GS} writes GS exept for the first time
%syntax: \acro{<acronym>}[<short name>]{<full name>}
\begin{acronym}[]
\acro{3GPP}{3rd Generation Partnership Project}
\acro{ACK}{Acknowledgment}
\acro{AMR}{Adaptive Multi-Rate}
\acro{AP}{Access Point}
\acro{ARM}{Advanced RISC Machine}
\acro{ARQ}{Automatic Repeat-reQuest}
\acro{ARF}{Auto Rate Fallback}
\acro{API}{Application Programming Interface}
\acro{BER}{Bit Error Rate}
\acro{BLAS}{Basic Linear Algebra Subprograms}
\acro{BS}{Base Station}
\acro{BSS}{Basic Service Set}
\acro{cdf}{Cummulative Density Function}
\acro{CFP}{Contention Freeze Periode}
\acro{CPU}{Central Processing Unit}
\acro{CSMA/CA}{Carrier Sense Multiple Access with Collision Avoidance}
\acro{CW}{Contention Window}
\acro{CWN}{Cooperative Wireless Network}
\acro{CTS}{Clear to Send}
\acro{D2D}{Device to Device}
\acro{DAG}{Direct Acyclic Graph}
\acro{DC}{Direct Current}
\acro{DCF}{Distributed Coordination Function}
\acro{DIFS}{Distributed Interframe Space}
\acro{ERP}{Extended Rate PHY}
\acro{FEC}{Forward Error Correction}
\acro{FCS}{Frame Check Sequence}
\acro{FF}{Finite Field}
\acro{GF}{Galois Field}
\acro{GPS}{Global Positioning System}
\acro{GPU}{Graphic Processing Unit}
\acro{HTTP}{Hyper Text Transfer Protocol}
\acro{IBSS}{Independent BSS}
\acro{IC}{Interference Cancellation}
\acro{ICST}{Institute for Computer Sciences, Social-Informatics and Telecommunications Engineering}
\acro{IFS}{Interframe Space}
\acro{IoT}{Internet of Things}
\acro{IP}{Internet Protocol}
\acro{IPTV}{Internet Protocol TeleVision}
\acro{JCN}{Journal of Communications and Networks}
\acro{KL}{Kullback-Leibler}
\acro{LAN}{Local Area Network}
\acro{l.d.}{linearly dependent}
\acro{l.i.}{linearly independent}
\acro{LNC}{Linear Network Coding}
\acro{LOS}{Line Of Sight}
\acro{LTE-A}{Long Term Evolution Advanced}
\acro{MAC}{Medium Access Control}
\acro{MANET}{Mobile Ad hoc NETwork}
\acro{MIMO}{Multiple Input Multiple Output}
\acro{MIT}{Massachusetts Institute of Technology}
\acro{MTBF}{Mean Time Between Failure}
\acro{M2M}{Machine-to-Machine}
\acro{NAV}{Network Allocation Vector}
\acro{NEP}{Nokia Energy Profiler}
\acro{NC}{Network Coding}
\acro{NLOS}{Non Line Of Sight}
\acro{NIC}{Network Interface Card}
\acro{OSI}{Open Systems Interconnect}
\acro{PC}{Personal Computer}
\acro{PDA}{Personal Digital Assistant}
\acro{PEP}{Packet Error Probability}
\acro{pgf}{Probability Generating Function}
\acro{PHY}{Physical Layer}
\acro{PLCP}{Physical Layer Convergence Procedure}
\acro{pmf}{Probability Mass Function}
\acro{PNC}{Physical layer Network Coding}
\acro{PPDU}{PLCP Protocol Data Unit}
\acro{QEMU}{Quick Emulator}
\acro{QoE}{Quality of Experience}
\acro{QoS}{Quality of Service}
\acro{RLNC}{Random Linear Network Coding}
\acro{ROC}{Region of Convergence}
\acro{SRLNC}{Sparse Random Linear Network Coding}
\acro{SIMD}{Single Instruction Multiple Data}
\acro{SAIC}{Single Antenna Interference Cancellation}
\acro{SD}{Secure Digital}
\acro{SIFS}{Short Interframe Space}
\acro{SIMD}{Single Instruction Multiple Data}
\acro{SINR}{Signal-to-Interference-plus-Noise Ratio}
\acro{SMP}{Symmetric Multiprocessor}
\acro{SNMP}{Simple Network Management Protocol}
\acro{SNR}{Signal-to-Noise Ratio}
\acro{SOC}{System on Chip}
\acro{SSE}{Streaming SIMD Extensions}
\acro{SSID}{Service Set Identifier}
\acro{TSNC}{Tunable Sparse Network Coding}
\acro{UDP}{User Datagram Protocol}
\acro{UI}{User Interface}
\acro{UML}{Unified Modeling Language}
\acro{URL}{Uniform Resource Locator}
\acro{VANET}{Vehicular Ad-hoc Network}
\acro{VoIP}{Voice over Internet Protocol}
\acro{WBN}{Wireless Broadcast Network}
\acro{WiFi}{Wireless Fidelity}
\acro{WLAN}{Wireless Local Area Network}
\acro{WMN}{Wireless Mesh Network}
\acro{pmf}{Probability Mass Function}
\acro{TC}{Telescopic Codes}
\acro{TBTT}{Target Beacon Transmision Times}
\acro{TCP}{Transmission Control Protocol}
\acro{dof}{degrees of freedom}
\acro{IP}{Internet Protocol}
\acro{OS}{Operating System}
\acro{NFS}{Network File System}
\acro{RAM}{Random Access Memory}
\acro{Raspi}{Raspberry Pi}
\acro{SSH}{Secure Shell}
\acro{USB}{Universal Serial Bus}
\acro{SCP}{Secure Copy}
\acro{Telnet}{Telnet}
\acro{GB}{Gigabyte}
\acro{Bash}{Bourne Again SHell}
\acro{DHCP}{Dynamic Host Configuration Protocol}
\acro{IT}{Information Technology}
\end{acronym}

% *** ALGORITHM PACKAGE ***
\usepackage[]{algorithm2e}

% *** PDF PACKAGE *** ..to extract a page from a PDF with multiple pages.
\usepackage{pdfpages}

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2016}
\copyrightyear{2016}
\externaleditor{Academic Editors: Steven Johnston and Simon J. Cox}
\history{Received: 19 July 2016; Accepted: 28 September 2016; Published: date}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed

%=================================================================
%% Please use the following mathematics environments:
 \theoremstyle{mdpi}
 \newcounter{thm}
 \setcounter{thm}{0}
 \newcounter{ex}
 \setcounter{ex}{0}
 \newcounter{re}
 \setcounter{re}{0}

% \newtheorem{Theorem}[thm]{Theorem}
% \newtheorem{Lemma}[thm]{Lemma}
% \newtheorem{Corollary}[thm]{Corollary}
% \newtheorem{Proposition}[thm]{Proposition}

 \theoremstyle{mdpidefinition}
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Easy as Pi: A Network Coding Raspberry Pi Testbed}

% Authors, for the paper (add full first names)
\Author{Chres W. S{\o}rensen $^{1,}$*${^,}$${^\dagger}$, N\'estor J. Hern\'andez Marcano $^{1,2,\dagger}$, Juan A. Cabrera Guerrero $^{3,4}$, Simon~Wunderlich $^{3}$, Daniel E. Lucani $^{1,\dagger}$ and Frank H. P. Fitzek $^{3}$}

% please provide the whole name.
% Authors, for metadata in PDF
\AuthorNames{Chres W. Soerensen, Nestor J. Hernandez M., Juan Cabrera, Simon Wunderlich, Daniel E. Lucani and Frank H. P. Fitzek}
% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
    $^{1}$ \quad  Department of Electronic Systems, Aalborg University, Aalborg 9220, Denmark; del@es.aau.dk (D.E.L.)\\
$^{2}$ \quad Steinwurf ApS, Aalborg 9220, Denmark; nestor@steinwurf.com (N.J.H.M.)\\
$^{3}$ \quad Deutsche Telekom Chair of Communication Networks, Technische Universit\"at Dresden, Dresden 01069, Germany; juan.cabrera@tu-dresden.de (J.A.C.G.); simon.wunderlich@mailbox.tu-dresden.de (S.W.); frank.fitzek@tu-dresden.de (F.H.P.F.)\\
$^{4}$ \quad SFB 912---Collaborative Research Center HAEC, Dresden 01069, Germany}
% the zip code in affiliations 1--3 is newly added, the third is made up of five numbers please confirm, and please add the zip code for affiliation 4. finally, please check the e-mail in affiliation 2.

% Contact information of the corresponding author
\corres{Correspondence: cws@es.aau.dk; Tel.: +45-99-408-723}
%\corres{Correspondence: nestor@steinwurf.com, nh@es.aau.dk; Tel.: +45 51 20 03 49}

% Current address and/or shared authorship
\firstnote{Fredrik Bajers Vej 7A, Room A3-110, Aalborg 9220, Denmark}
% please add the zip code.

% Simple summary
%\simplesumm{}

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{In the near future, upcoming communications and storage networks are expected to tolerate major difficulties produced by huge amounts of data being generated from the Internet of Things (IoT). For these types of networks, strategies and mechanisms based on network coding have appeared as an alternative to overcome these difficulties in a holistic manner, e.g., without sacrificing the benefit of a given network metric when improving another. There has been recurrent issues on: (i) making large-scale deployments akin to the Internet of Things; (ii) assessing and (iii) replicating the obtained results in preliminary studies. Therefore, finding testbeds that can deal with large-scale deployments and not lose historic data in order to evaluate these mechanisms are greatly needed and desirable from a research perspective. However, this can be hard to manage, not only due to the inherent costs of the hardware, but also due to maintenance challenges. In this paper, we present the required key steps to design, setup and maintain an inexpensive testbed using Raspberry Pi devices for communications and storage networks with network coding capabilities. This testbed can be utilized for any applications requiring results replicability.}

% A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article: it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.

% Keywords
\keyword{Linux; network coding; Raspberry Pi; testbed; C++}

% keyword 1; keyword 2; keyword 3. List three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\S$ in the end of the title of this article.
%\conference{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:

%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize\ttfamily,
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  %commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  %escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  escapeinside={<@}{@>},
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  %keywordstyle=\color{blue},       % keyword style
  %language=Bash,                 % the language of the code
  otherkeywords={*,...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  %stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname,                  % show the filename of files included with \lstinputlisting; also try caption instead of title
  xleftmargin=5mm,
  xrightmargin=5mm,
  columns=fullflexible,
  upquote=true
}

% Used to suppress line numbers in lstlistings
\let\origthelstnumber\thelstnumber
\makeatletter
\newcommand*\Suppressnumber{%
  \lst@AddToHook{OnNewLine}{%
    \let\thelstnumber\relax%
     \advance\c@lstnumber-\@ne\relax%
    }%
}

% Used to reactivate line numbers in lstlistrings
\newcommand*\Reactivatenumber{%
  \lst@AddToHook{OnNewLine}{%
   \let\thelstnumber\origthelstnumber%
   \advance\c@lstnumber\@ne\relax}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Sections that are not mandatory are listed as such. The section titles given are for Articles. Review papers and other article types have a more flexible structure.

%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \setcounter{section}{-1} %% Remove this when starting to work on the template.
% \section{How to Use this Template}

% The template details the sections that can be used in a manuscript. Sections that are not mandatory are listed as such. The section titles given are for Articles. Review papers and other article types have a more flexible structure. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.

\section{Introduction}
%\input{introduction}

%!TEX root = raspi_journal.tex
% At the beginning, there was darkness and then... bang! \ac{NC}
% \cite{ahlswede2000network} appeared to save us from the evilness
% of routing.
%
% General introduction. Introduction to topic addressed in the journal.
% Review of the State of the Art. Specify why our approach has benefits
% and which are they. Indicate contributions.

Upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on a global scale. This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, and smart grids. The number of wireless devices is expected to increase by five times to up to 50 billion devices \cite{cisco2011forecast}. It is generally believed that those devices will not be connected in the same manner as current devices are connected today. Centralized systems will collapse in terms of capacity, while distributed systems appear as an alternative. Therefore, we believe mesh technologies will play a major role in the communication architecture in future systems. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices to ensure high Quality of Experience (QoE) for the final user. In state-of-the-art systems, those requirements are traded-off with each other, but in the 5G context, we cannot do this anymore.

Introduced by Ahlswede et al. \cite{ahlswede2000network}, network coding constitutes a paradigm shift in the way that researchers and industry understand and operate networks, by changing the role of intermediate relays in the process of transmission of information. Relays are no longer limited to storing and forwarding data, but also take part in the coding process, through a process called recoding, \mbox{where the relay} generates new linear combinations of incoming coded packets without previously decoding the data. Network coding allows the increase of throughput, reliability, security and delay performance of the networks. In previous works, we have shown that Random Linear Network Coding (RLNC)
\cite{koetter2003algebraic,ho2006random} is able to satisfy the aforementioned technical requirements. We have actually shown how to increase the throughput \cite{pahlevani2013playncool}, reduce the delay \cite{szabo2015towards} or support heterogeneity for coding enabled communication nodes \cite{lucani2014fulcrum}.

In our prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic RLNC functionalities. Most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude. Despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce. Deploying a large-scale and configurable testbed for networking and storage can be challenging, not~only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently. The latter requires not only the devices to run the same Operating System (OS), but also have exactly the same configurations and software packages. \mbox{There is a need} to evaluate large-scale network deployments of low-cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.

The emergence of powerful and inexpensive single-board computers opens new possibilities in this area. By running a standard OS, they allow implementations that are compatible with higher end devices. In addition, they utilize stable software supported by their communities. For example, the~Iridis-pi platform \cite{cox2014iridis} provides a detailed description of a Raspberry Pi (Raspi)~\cite{making_of_pi} testbed ideal for educational applications. Here, the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance. This work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-consuming. Moreover, this work does not consider possible network coding applications. Different studies of IoT applications consider using the Raspi for data processing: In \cite{leccese2014smartcity}, the Raspi is the processing unit that coordinates and controls the activity of an isle of lamps on a public road and reports it to a monitoring center. A use case regarding remote environment surveillance using the Raspi and the Arduino \cite{arduino} technologies is presented in~\cite{leccese2014newimaging}. Here,~both devices report air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves. This data is sent to a monitoring center to ensure the preservation of the paintings. Furthermore, authors in \cite{sapes2016finger} consider FingerScanner, a~technology that utilizes the Raspi to act as the data server in a finger scanning application that collects the fingerprints. Even though all these applications consider the use of the Raspi as a core block, they~provide few to no descriptions of their procedures to configure the Raspi. These applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users. The current way that the data is sent in the considered networks for these IoT applications will not be feasible in future 5G systems as mentioned previously.

Given this set of specific needs, in this work, we present the design, key step-by-step instructions and mechanisms to setup, configure and maintain an inexpensive testbed using potentially several Raspi devices for networking (wireless or wired) and storage applications including RLNC functionalities into the testbed through Kodo. The architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the Raspi. Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system. In Section~\ref{sec:image_setup}, we provide details about the testbed setup, scripts, configuration files and connectivity. In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step. Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed to simplify the execution of routinary and repetitive tasks. Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the Raspi. Conclusions and future work are reviewed in Section~\ref{sec:conclusions}. Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendices.

%%The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications should be cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the main conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.

\section{Testbed Overview and Design Criteria}
%\input{overview}
\label{sec:overview}

A sketch of the testbed is depicted in Figure~\ref{fig:testbed_setup}.
The testbed consists of up to 100 Raspis of different models.
More specifically, in our design, we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All Raspis are each equipped with a 8 GB \ac{SD} memory card, a wired and wireless
network interface and a power supply. All the Raspi are connected to
a common Local Area Network (LAN) that provides internal and external connectivity. Without
loss of generality, in our case, they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and academic staff
to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. The
testbed description and procedures for setting it up are not restricted
to this academic scenario. All Raspis are configured to run
a Secure Shell (SSH) daemon for easy remote access within the university network.
We requested the university Information Technology (IT) department to configure the university Dynamic Host Configuration Protocol (DHCP) server to assign each Raspi a static Internet Protocol (IP) address. This~eliminates the demand for monitors and keyboards with the Raspis
for non-graphical applications. Finally, our~design aims to configure all
Raspis identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the Raspis.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.45\textwidth]{testbed_setup3.eps}
\caption{Testbed setup.}
\label{fig:testbed_setup}
\end{figure}

%This significantly eases the management of the Raspis
%as SSH eliminates the need for monitors and keyboards for the Raspis.

% i can not find the picture in folder, so I canceled the image of the insert, please provide.

We will refer to the
testbed administrator as the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the OS
point of view. The setting and configuration procedures are
performed by the testbed administrator in a PC running a
Linux distribution as shown in Figure~\ref{fig:testbed_setup}. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to store it later on a memory card for each Raspi.
Once~configured, we store the resulting image file in a Hyper Text Transfer Protocol (HTTP) server
as backup and in case the testbed administrator requires the making of new
changes to this file.
%In our case, we use the university HTTP server but any others can be used.
In our case, we store all files at Zenodo~\cite{soerensen_chres_wiant_2016_154143}, but the testbed administrator should copy the our files to his/her own HTTP server to get read/write permissions.
We also put all the required
configuration files and scripts for the Raspis setup in the HTTP
server so there is a single place where system setup is stored and could be
modified. This simplifies the system maintenance, as it may not always be desirable to make persistent changes
on the Raspis---for example, when different users are interested in
running experiments on a rebooted testbed. We later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, we include a set of automation, monitoring
and cross-compilation tools over the top of our system in order to simplify
the execution of repetitive and long tasks, be able to follow the progress
of long task processes and compile relevant C++ source code for the testbed
administrator.

%an advanced step in our design considers to
% fetch the Raspis root filesystem from a HTTP server. This enables the
% testbed administrator to put the data files in one location. The
% Raspis will then get these data files after a reboot. We consider
% that this step is particularly useful when there is a large amount
% of devices to be configured and the setup process needs to be automated.
% However, since a more complex scenario that might be required only in special
% cases, we put this step in Section~\ref{sec:http} in the Appendix.

\section{OS Image Setup}
%\input{image_setup}

\label{sec:image_setup}
In this section, we review the steps to create a common OS image for
all the Raspis. The image setup is composed of three major steps:
select and download the OS image file, alter the image structure and
configure the OS files. We proceed to detail all these steps providing
brief discussions to our setup choices when required. To perform these
steps, we indicate with command-line blocks the required sequential
commands to be typed by the testbed administrator on his/her PC
to obtain the desired setting. In all the command blocks in the paper, we indicate if a command needs to be run with root permissions (\#) or common user permissions (\$). These signs will prefix the commands.

\subsection{OS Selection and Download}

To get started, we first need to install an OS that works properly
on all the Raspi models. We will download and setup the image in
the testbed administrator PC using a Debian-based distribution.
An alternative to this method, is to create a tailoared Linux distribution
for the Raspi platform using the Yocto Project \cite{yocto}. However, this
process would require assembly and compilation of all the software for the
Raspi platform from scratch, which goes beyond the scope of our
work. We use the popular Debian-based Raspbian Linux~\cite{raspbian} given
that is the recommended and default OS for the Raspi. Raspbian is
made available in two bundles: Raspbian and Raspbian Lite. The difference
between the two is that Raspbian contains a pre-installed desktop environment
for user interaction, and Raspbian Lite by default only permits interaction
through a command shell. Given that the Raspis in our testbed are not
connected to monitors, we decide to work with Raspbian Lite.
If required, a~desktop environment can be installed using the package manager later.
%\todot{remove the rest of this line} as it will be shown in the configuration section.

% The overall procedure to customize the official Raspbian Lite image
% are:
% \begin{enumerate}
%     \item Download Raspbian Lite
%     \item Alter Raspbian Lite. e.g. browse, modify, add, and delete files
%         in the official Raspbian Lite image
%     \item Change root, i.e. change root filesystem into the Raspbian Lite
%         image to update and install software packages
%     \item Write image to memory cards
% \end{enumerate}

%\subsection{Download Raspbian Lite}
%To download the latest Raspbian Lite, we go to the url
%\url{http://downloads.raspberrypi.org/raspbian_Lite/images/}.

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite, which we have
made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are also available there. The testbed
administrator has to move these files to his/her own HTTP server.
To~get started, the testbed administrator must open a Linux
shell (terminal) on his/her PC and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

%After the terminal has been opened, we start by declaring a few variables to reduce
%repeated typing. The first variables we declare is the Raspbian image name.
%Notice that the extension has been omitted. This is because the image has been
%packed into a zip file. The other variable we declare is a working directory.
%This is where we will download the image to and work on it. In other words, it
%will be stored in \texttt{/home/<username>/Raspbian}

%To download Raspbian lite, we go to the \url{http://downloads.raspberrypi.org}.
%There should be a folder called raspbian\lite and
%to find the download page at which Raspbian lite should be located. Instead of
%downloading with the browser, we just extract the download url to the latest
%image. At the time of this writing, that was

% Define variable]
\begin{lstlisting}[]
$ export URL="https://zenodo.org/record/154328/files/"
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

In this code block, the \texttt{\$\{URL\}} and \texttt{\$\{IMAGE\}}
variables specify where the Linux bundle is located and
\texttt{\$\{WORKDIR\}} specifies a working directory where the Raspbian
Lite bundle will be downloaded and customized. If the testbed administrator
allocates his / her files into another location, then it will be required to
change the \texttt{\$\{URL\}} environment variable. Notice that even though we
use the \$ and \# signs in the shell, in general, these signs will be
particular to the testbed administrator OS shell. Next, we create
the working directory and change to it with the \texttt{cd} command. To
download the image, we utilize the \texttt{wget} command before unpacking
the \texttt{zip} file as follows:

% Running a command as root can on most systems be done by putting
% \texttt{sudo} in front of the command. This is illustrated in the
% following code block
% with the command \texttt{whoami} that prints the username. Lines within a code
% block without leading \$ or \# is terminal output or content of a file.

% Root example
% \begin{lstlisting}[]
% $ whoami
% <USERNAME>
% $ sudo whoami
% root
% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

% Download and unpack image
\begin{lstlisting}[]
$ mkdir -p ${WORKDIR}
$ cd ${WORKDIR}
$ wget ${URL%/}/${IMAGE}.zip
$ unzip ${IMAGE}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-6mm}

%A more recent version of Raspbian lite may be available at~\cite{raspbian}.
%\url{http://downloads.raspberrypi.org/raspbian_lite/images}.

\subsection{Image Customization}

After Raspbian Lite has been unpacked, there should be an \texttt{.img}
file in the working directory \texttt{\$\{WORKDIR\}}. \texttt{fdisk} can
be used to display the content of the image. We parse the arguments \texttt{-u sectors}
to display the sizes in sectors and \texttt{-l} to display the partitions
within the image. The \texttt{fdisk} command should output to the terminal
something similar to:

% Check out the image
% The dollar hack was to fix vim syntax
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The output provides relevant information about the image. The image is in
total 2,709,504 sectors (1.3 GiB) in size and contains two partitions. The
first partition starts at sector 8192 and the other partition starts at
\hl{sector} 137,216. The first partition type is FAT32 with a size of 63 MB
and the second partition is of type Linux with a size of 1.2 GB. This
indicates that the first partition is a boot partition, and the second
one is a traditional Linux filesystem. In this case, the root filesystem, i.e., \texttt{/}.
% i add a comma here, please confirm if it is right.
\subsection{Image Resizing}
Given that we want to customize the root filesystem in the Raspis,
we need to expand the image file since 1.2 GB might not be enough to store
the existing root filesystem plus additional files and software packages.
%Given that we want to customize the files stored in the Raspis,
%we need to resize the image file since 1.2 GB might not be enough to store
%the root filesystem with the additional packages that will be installed in this procedure.
%due to the total size of the additional packages.
Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one GB.
First, to expand the image one GB, we execute:

\begin{lstlisting}[]
$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Later, we use \texttt{fdisk} with the same arguments as before to see that
the image is now one GB~larger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 2709503 2572288  1.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, in the above command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, we replace the Linux partition with a new partition one GB
larger. The starting point of this new partition should be the same as the old
one. We make use of \texttt{fdisk} to alter the partition table
in the commands below. They~(i)~delete partition number 2; (ii) create a new
primary partition and (iii) set the new partition starting point. The new
partition starting point value is 137,216 in our case; Finally, we (iv) write
the new partition table to the image file. This is made as follows:
\newpage
%\begin{lstlisting}[]
%$ (echo d; echo 2; echo n; echo ; echo ; echo 137216; echo ; echo w) | fdisk ${IMAGE}.img
%\end{lstlisting}
\begin{lstlisting}[]
$ fdisk ${IMAGE}.img << EOF
d
2
n
p
2
137216

w
EOF
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the partitions commands were correct, the partition table should now
look like the following:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

% \begin{lstlisting}[]
% Command (m for help): Partition number (1-4):
% Command (m for help): Partition type:
%    p   primary (1 primary, 0 extended, 3 free)
%    e   extended
% Select (default p): Using default response p
% Partition number (1-4, default 2): Using default value 2
% First sector (2048-4806655, default 2048): Last sector, +sectors or
% +size{K,M,G} (137216-4806655, default 4806655): Using default value 4806655

% Command (m for help): The partition table has been altered!

% Syncing disks.

% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

\subsection{Loopback Device Setup}
After successfully resizing the image file, we use a loopback device to make
the Raspbian image available as a block device in the filesystem. For this
command to work, the testbed administrator distribution must have the
\texttt{util-linux} package with version 2.21 or higher. Otherwise, the
\texttt{-P} argument of \texttt{losetup} will appear as invalid. If the
version of \texttt{losetup} can not be updated for some reason, an
alternative option for this part is presented in
Section~\ref{sec:alternative_losetup} of the Appendices.

\begin{lstlisting}[]
$ export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
/dev/loop0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the previous command was succesful, the \texttt{lsblk} command can be used
to list the available block devices in the filesystem as follows:
%Use \texttt{lsblk} to view the partitions:

\begin{lstlisting}[]
# lsblk
NAME      MAJ:MIN RM  SIZE RO TYPE MOUNTPOINT
...
loop0       7:0    0  2.3G  0 loop
|-loop0p1 259:2    0   63M  0 loop
|-loop0p2 259:3    0  2.2G  0 loop
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The image block device appears as \texttt{/dev/loop0}. This block device has
two partitions associated with it, e.g., \texttt{loop0p1} and \texttt{loop0p2}.
Finally, we check the filesystem of the block device with \texttt{e2fsck} and
resize it with the \texttt{resize2fs} command:%, respectively in the command block:
\newpage
\begin{lstlisting}[]
# e2fsck -f ${DEV}p2
e2fsck 1.42.8 (20-Jun-2013)
Pass 1: Checking inodes, blocks, and sizes
Pass 2: Checking directory structure
Pass 3: Checking directory connectivity
Pass 4: Checking reference counts
Pass 5: Checking group summary information
/dev/loop0p2: 35392/80480 files (0.1% non-contiguous), 201968/321536 blocks
# resize2fs ${DEV}p2
resize2fs 1.42.8 (20-Jun-2013)
Resizing the filesystem on /dev/loop0 to 583680 (4k) blocks.
The filesystem on /dev/loop0 is now 583680 blocks long.
\end{lstlisting}
\FloatBarrier

\subsection{Block Device Mounting}
For browsing and altering the files in the image, we mount the block
device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
to customize them. We mount the block device partition that
contains the root filesystem and later the boot partition. This is done by
creating an empty directory that is used as a mountpoint. We name it
\texttt{root} and create it in the working directory before mounting the
root filesystem onto the mountpoint. We mount the root filesystem
as follows:

\begin{lstlisting}[]
$ export ROOTDIR="${WORKDIR}/root"
$ mkdir -p ${ROOTDIR}
# mount ${DEV}p2 ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%# mount -o loop,offset=$((137216*512)) ${IMAGE}.img ${ROOTDIR}

The root filesystem mounted in \texttt{\$\{ROOTDIR\}} already has a boot
directory that can be used as the mount point for the boot partition
in the block device \texttt{/dev/loop0p1}. This is convenient because
the final edited partition from \texttt{\$\{ROOTDIR\}/boot} will be mounted on
this same directory when a Raspi starts up with a memory card
containing the raw final image. Hence, to mount boot partition we do:
%It is therefore the natural place to mount it for later purposes.

\begin{lstlisting}[]
# mount ${DEV}p1 ${ROOTDIR}/boot
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%# mount -o loop,offset=$((8192*512)) ${IMAGE}.img ${ROOTDIR}/boot
%We can now change all the files in the disk image as desired.

In this way, it is now possible to change all files within the Raspbian
image as desired by editing the files in \texttt{\$\{ROOTDIR\}}. We take
advantage of this to edit configuration files, append new files and even
update and install packages.

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the Raspis should be setup as similarly as possible. However,
some particularities exist to differentiate the devices in principle. In addition,
scripts containing further configurations for the Raspis are desirable
to be distributed as part of the common image. Therefore, we present here
the steps to setup basic properties of the Raspis and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order the scripts are executed to configure all the
devices. Any testbed administrator might modify or include other
tasks according to his / her needs as we will show.

\subsubsection{Image Default Configuration Scripts Download}
\label{sec:configuration_files_download}
In our case, we have our default configuration scripts stored in a
file \texttt{rasp\_config.zip} located in the same URL of the
HTTP server where the image was retrieved from, i.e., the one in
the environment variable \texttt{\$\{URL\}}. We first download
this compressed file with \texttt{wget} and extract it locally into our
Raspbian Lite image. These commands and the output of the last one are
shown as follows:

% Get the configuration files that are compressed in a remote server
\begin{lstlisting}[]
$ wget ${URL%/}/rasp_config.zip
$ unzip rasp_config.zip -d ${ROOTDIR}/home/pi/
Archive:  rasp_config.zip
   creating: ${ROOTDIR}/home/pi/rasp_config/
  inflating: ${ROOTDIR}/home/pi/rasp_config/nodes.csv
  inflating: ${ROOTDIR}/home/pi/rasp_config/set_hostname
  inflating: ${ROOTDIR}/home/pi/rasp_config/main
  inflating: ${ROOTDIR}/home/pi/rasp_config/update_rasp_config
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The unzippped files are one configuration file and three configuration
scripts in the newly created \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/}
folder in the image. We describe which features that we
require all the Raspis to have and how are they achieved with
these configuration~scripts.

\subsubsection{Device Hostnames}
The hostname helps the user to physically distinguish the
devices from each other. In our case, we require the devices in our
testbed to have different hostnames. We define the hostnames based
on the Medium Access Control (MAC) addresses of the Raspis wired Ethernet interface.

Prior to this stage, the MAC address of a network card can be found
using the command \texttt{ifconfig} or \texttt{ip addr} on a given
Raspi. We store the MAC addresses and hostnames of the
Raspis in the configuration file
\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}. A sample of our file
is shown as follows:

% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
<@\textcolor{gray}{
--------------------------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Hostname
b8:27:eb:5b:da:20 rasp00
b8:27:eb:7b:c3:91 rasp01
b8:27:eb:54:9c:64 rasp02
b8:27:eb:95:bd:11 rasp03
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The testbed administrator has to insert the MAC addresses and hostnames
of his/her Raspis obtained previously in the format shown in the
configuration file. For each given Raspi, there is a MAC address and
the corresponding hostname. This file will be employed by the
\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} \ac{Bash} script to
assign the hostname of each Raspi. The script content is the following:

% Set hostname
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
<@\textcolor{gray}{
-------------------------------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

script_path="$(dirname $(realpath $0))"
config_file=${script_path}/nodes.csv
mac=$(cat /sys/class/net/eth0/address)
old_hostname=$(hostname)
new_hostname=$(grep $mac $config_file | cut -f2 -d' ')

# Assign hostname found in nodes.csv
if [ ! -z ${new_hostname} ]; then
    echo ${new_hostname} > /etc/hostname
    hostname ${new_hostname}
    sed -i.old -e "s:${old_hostname}:${new_hostname}:g" /etc/hosts
fi
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The script (in lines): (1) tells the system to interpret the script
using \ac{Bash}; (3--4) gets the path to the script itself and the list of
hostnames; (5) gets the MAC address of the node itself; (6) gets the
current hostname; (7) gets the new hostname from the hostname list; and
(10--14) assigns the new hostname to the Raspi where the script
will be executed.

\subsubsection{Updating Default Configuration Files and Scripts}

Besides the single script with its configuration file introduced up
to this point of our procedure, it is possible that the testbed~administrator may require to add other scripts to configure
his/her Raspis. We want
to ensure that all the Raspi configuration scripts of any~testbed~administrators are obtained in a simple way. We automate this task by
including the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}
script in our procedure. The purpose of this script is to make all the
Raspis fetch all the configuration scripts located with the image
during a testbed start up.

In our case, as the testbed administrator for presenting the procedure, we want
to fetch all our configuration scripts in
\texttt{\$\{URL\%/\}/raspi\_config.zip}. The update script
\textit{automatically} downloads all the required configuration files in
\texttt{rasp\_config.zip} file from a remote location. This is the same that
we manually did earlier to get our files, but this will be made in an
automated way after booting up the system. This script content is:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}@>
<@\textcolor{gray}{
--------------------------------------------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

url="https://zenodo.org/record/154328/files/"
config_file="rasp_config.zip"

# Attempt to fetch new configuration files
if ! wget -q --show-progress -O /tmp/${config_file} ${url%/}/${config_file}; then
    echo "Warning: Unable to update rasp_config files"
    exit 1
fi

# Unzip and overwrite configurationn files to root's home directory
unzip -q -o /tmp/${config_file} -d /home/pi/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The update script lines (3--4) specify the URL and \texttt{.zip} file that
should be downloaded. Lines~(7--10) download the configuration files to \texttt{/tmp}
folder in the corresponding Raspi. It also prints a warning in case of
errors and line (13) unzips the files to \texttt{/home/pi/}, the Raspi home
directory. Existing files and directories are simply overwritten.
%while ignoring existing files and / or directories there.

For the above scripts to work in the Raspis, it is required that the
Raspis MAC addresses are found in \texttt{nodes.csv}. In addition, it
should be noted that for other testbed administrators besides ourselves,
the URL for file fetching and the configuration
scripts themselves can be modified to fit their requirements. If
required for a testbed administrator, the \texttt{rasp\_config.zip}
will need to be edited to include all the required configuration files
and scripts. In addition, it might be necessary to edit the URL in the script
\texttt{update\_rasp\_config} to store and fetch from a different location.
Nevertheless, both~the URL and configuration files presented here can be
used as a starting boilerplate if desired.

\subsubsection{Configuration Scripts Execution Order}
To actually make the Raspis change hostnames and any other considered
configurations, we~have to make each Raspi call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each Raspi after getting the root filesystem.
We first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in \texttt{\$\{ROOTDIR\}/etc/rc.local} before
\texttt{exit 0} in the file:
\newpage
%before \texttt{exit 0} using a file editor. To do this, it is required
%to edit this file with root permissions. The resulting file should look like the following:

\Suppressnumber\begin{lstlisting}[]
# sed -i '/^exit 0/i bash /home/pi/rasp_config/update_rasp_config' ${ROOTDIR}/etc/rc.local
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\Suppressnumber\begin{lstlisting}[]
%<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
%<@\textcolor{gray}{
%---------------------------------------------}
%\Reactivatenumber @>
%...
%bash /home/pi/rasp_config/update_rasp_config
%...
%exit 0
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

If it is required to have more configuration scripts, adding them
in the \texttt{rc.local} file makes maintenance by the testbed administrator difficult since this needs to be both in the image and the
downloaded \texttt{rasp\_config.zip}. To avoid this problem, we include
the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/main} script that
calls all other configuration scripts (besides \texttt{update\_rasp\_config})
in a sequential order. This script content is:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/main}@>
<@\textcolor{gray}{
---------------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

bash /home/pi/rasp_config/set_hostname
# Any other required configuration scripts...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

In this way, the automation process is simplified since we do not need to
modify \texttt{\$\{ROOTDIR\}/etc/rc.local} again after the image has been
written to the memory cards.
%In this way, the automation process is simplified since we only need to modify
%the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script.
Now, we~insert a call to the \texttt{main} script in \texttt{\$\{ROOTDIR\}/etc/rc.local} as follows:

%\todot{insert sed command below instead}
\Suppressnumber\begin{lstlisting}[]
# sed -i '/^exit 0/i bash /home/pi/rasp_config/main' ${ROOTDIR}/etc/rc.local
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, \texttt{\$\{ROOTDIR\}/etc/rc.local} should look like the following:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
---------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Notice that \texttt{set\_hostname} is now called by the \texttt{main}
script instead. The update script is still called directly. This ensures that
all configuration scripts are updated before executed. Changes to the update
script itself will first take effect at the next system startup.

%\url{https://wiki.debian.org/QemuUserEmulation}
%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
\subsection{Image Package Updating by Changing the Apparent Root Directory}
Besides adding and configuring files within the image, the testbed
administrator may want to install and update the software packages
within the image before it is written to all the memory cards that goes
into the Raspis. From any Linux x86 machine as the testbed administrator
PC, \mbox{this can be done using} \texttt{chroot} command in the
\ac{QEMU}~\cite{QemuUserEmulation} hypervisor for Advanced RISC Machine (ARM) processors.
%Does this need to be defined?

\texttt{chroot} is a method in Linux that modifies the apparent root
filesystem location from \texttt{/} to any other path. Consequently, in
our case, we can use the Raspbian Lite image root filesystem within the
testbed administrator Linux distribution. Then, \ac{QEMU} allows the execution
of commands for the Raspi image (ARM instructions) through the ones
from the testbed administrator PC architecture. Due to the ARM
processor that the Raspis employ, installation of the \ac{QEMU} related software is required first and verification that \ac{QEMU} is ARM enabled. To do so,
run the following commands:
\newpage

% CHROOT to OS image
\begin{lstlisting}[]
# apt-get install binfmt-support qemu qemu-user-static
# update-binfmts --display qemu-arm
qemu-arm (enabled):
     package = qemu-user-static
        type = magic
      offset = 0
       magic = \x7fELF\x01\x01\x01\x00\x00\x00\x00\x00\x00\x00\x00\x00\x02\x00\x28\x00
        mask = \xff\xff\xff\xff\xff\xff\xff\x00\xff\xff\xff\xff\xff\xff\xff\xff\xfe\xff\xff\xff
 interpreter = /usr/bin/qemu-arm-static
    detector =
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% This was neccessary on arch
%# update-binfmts --importdir /var/lib/binfmts/ --import

In the previous output, the testbed administrator must be sure that the second
command writes \texttt{qemu-arm (enabled)} as indicated. If that is not the
case, then it should be possible to enable it by~running:

\begin{lstlisting}[]
# update-binfmts --enable qemu-arm
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Provided that \texttt{qemu-arm} is enabled, we should now be able to \texttt{chroot}
into our Raspbian lite image. There are a few commands to be
performed before actually changing root into the root partion of the image.
First, to get internet access from within the Raspbian lite image, copying the testbed administrator Linux distribution \texttt{resolv.conf} file
into the image root filesystem is required. To do this, it is necessary to run the
following:
\begin{lstlisting}[]
$ cd $ROOTDIR
# cp /etc/resolv.conf ${ROOTDIR}/etc/resolv.conf
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, because of the ARM architecture, the
\texttt{/usr/bin/qemu-arm-static} command needs to be copied into the
image before continuing by running:

% CHROOT to OS image
\begin{lstlisting}[]
# cp /usr/bin/qemu-arm-static ${ROOTDIR}/usr/bin
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Before changing the root, it is necessary to populate the directories
\texttt{proc}, \texttt{sys} and \texttt{dev} for the image to get control
as the testbed administrator apparent root filesystem. This is made by the
following~commands:

% CHROOT to OS image
\begin{lstlisting}[]
# mount  -t proc proc proc/
# mount --bind /sys sys/
# mount --bind /dev dev/
# mount --bind /dev/pts dev/pts
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%Finally, to change root the following needs to be run:
Finally, run the following command to change root:

\begin{lstlisting}[]
# chroot ${ROOTDIR} /usr/bin/qemu-arm-static /bin/bash
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If successfully executed, our terminal should have changed the prompt,
indicating that we are the root user in the Raspbian lite root filesystem as
the apparent root. In case the \texttt{chroot} command is not successful,
we provide an alternative command in Section~\ref{sec:chroot} of the Appendices.
To be aware of the mode that we are working now, we change
the prompt title to indicate that it is a \texttt{chroot} environment as
follows:

% Optionally, we may create a unique prompt to indicate we have changed root
\begin{lstlisting}[]
# export PS1="(chroot) $PS1"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The Raspbian lite image should now be possible to use almost as if it had
been booted in a Raspi. A major difference is that the testbed
administrator PC is likely significantly faster than a Raspi.
Hence, enabling updates, upgrades and installing new software packages
should be faster than in a Raspi. Still updating and upgrading
the packages for the Raspi might take some amount of time.
%To get the upgraded system packages, it is necessary to run:
To~update the system package list, run the following command:

% Update system
\begin{lstlisting}[]
(chroot) # apt-get update
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We further install some packages that we consider useful:
%Lets install some useful applications:
% Install packages
\begin{lstlisting}[]
(chroot) # apt-get install vim git screen
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
for managing Git repositories and \texttt{screen}~\cite{gnu_screen} for better handling of
long-runnning processes. When writing the image to a memory card, all the
changes that have been made to the image so far will exist in all
Raspis after fetching it.

\section{Overlay File System}
%\input{overlay_fs}

\label{sec:overlay_fs}
In principle, our procedure modifies the image file only once in
the testbed administrator PC when its setup is made. In addition, keeping
this image in the Raspis provides the same initial system for all
the devices. If we do not make any further modifications during
the image setup, any files created after the initial boot of a Raspi
will remain in the memory card. This is cumbersome to maintain since
the size of the memory card is relatively small (8 GB), and there
might be various users utilizing the testbed. In addition, different testbed
users could be interested in running their experiments in a fresh
rebooted system with the original customized image. We
emphasize that this step is not necessary if the tesbed administrator
wants to consider only persistent storage for its devices. A use case
for this scenario could be a single user for the testbed or when a testbed
administrator only wants to setup a few Raspis.

If both persistent and non-persistent storage are required for
the Raspis, we present here the steps to setup an overlay filesystem.
This type of filesystem enables an \textit{upper} filesystem to overlay
into a \textit{lower} filesystem. Whenever a file is requested, the upper
filesystem will forward the request to the lower filesystem in case it
does not have it itself. If the upper filesystem has the requested file,
it will simply return the file. This idea can be used in our setup to mount
the root filesystem (i.e., Raspbian Lite) in the Raspis during startup
as read-only filesystem. On the one hand, the image configuration files
will remain after a reboot but the local data in these directories will be
erased after a reboot. To enable the
possibility of persistent changes, we overlay the upper filesystem that is
mounted in the Raspi Random Access Memory (RAM), i.e., \texttt{/tmp} as rewritable
on top of the lower root filesystem. Reading a file may return a
file from the lower filesytem, but if it is stored, it will be saved in
the upper filesystem. Accessing this file again will return
the stored file from the upper layer. After a reboot, all the stored files
in the upper filesystem will be retrieved, but the ones in the lower
filesystem that are not part of the original image will be removed.

\subsection{Filesystem Installation}

Assuming that we are still in the \texttt{chroot} environment of the
Raspbian Lite root filesystem for installing packages, we can setup the
overlay filesystem at this point of the procedure. There already exists
implementations overlaying the root filesystem. We use an implementation
available at the Git repository in \cite{overlayroot}. Since we have
installed \texttt{git} in a previous step, we clone the repository. The~command block below stores it in \texttt{/tmp} which is really mounted in
RAM. All the files stored here will disappear when the system is
rebooted.

% Get files
\begin{lstlisting}[]
(chroot) # OVERLAYROOTDIR="/tmp/overlayroot"
(chroot) # git clone https://github.com/chesty/overlayroot.git ${OVERLAYROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Before enabling the overlaying filesystem, it is necessary to generate
an initial RAM filesystem or \texttt{initramfs}. This is an
initial filesystem that is loaded into RAM during the startup
process of a Linux machine to prepare the real filesystem. For this purpose,
we need the BusyBox package by running:

% Install required packages
\begin{lstlisting}[]
(chroot) # apt-get install busybox
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To create and activate the overlaying filesystem, we need to first
add the required system scripts to do so. This is done as follows:

% Setup initramfs
\begin{lstlisting}[]
(chroot) # cp ${OVERLAYROOTDIR}/hooks-overlay /etc/initramfs-tools/hooks/
(chroot) # cp ${OVERLAYROOTDIR}/init-bottom-overlay /etc/initramfs-tools/scripts/init-bottom/
(chroot) # echo "overlay" > /etc/initramfs-tools/modules
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To generate the initial RAM filesystem, we have to utilize the
\texttt{mkinitramfs} command. This~searches by default for the available
kernel modules in the system. Since we are in \texttt{chroot} mode, we need
to specify the correct kernel modules to search for. The available
kernel modules are located in \texttt{/lib/modules}.
To see them, we just run:

% See available kernel modules
\begin{lstlisting}[]
(chroot) # ls /lib/modules/
4.4.13+  4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%IT SEEM THAT THE RASPIS DEMANDS DIFFERENT KERNEL MODULES AFTER ALL.
%RASP-VERSION1=4.4.13+, RASP-VERSION2=4.4.13-V7+, RASP-VERSION3=4.4.13-V7+

Now, the initial RAM filesystem can be generated. Raspi version 1
needs a different kernel than Raspi \hl{version} 2 and version 3. Kernel
version 4.4.13+ is for Raspi version 1 and kernel 4.4.13-v7+ for
Raspi version 2 and version 3. We proceed to generate an initial
RAM filesystem for these kernels by~running:
% Create an initramfs
\begin{lstlisting}[]
(chroot) # mkinitramfs -o /boot/init.gz -k 4.4.13+
(chroot) # mkinitramfs -o /boot/init-v7.gz -k 4.4.13-v7+
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
% please confirm if  version 2 and version 3 should be versions 2 and  3.

Although these commands might output some warnings, they should
successfully generate working initial RAM filesystems. Later, an
initial RAM filesystem will need to be called by the bootloader.
In Raspbian, this is done by adding a command to \texttt{config.txt} file
in the boot partition. If~the system should be run in a Raspi version 1,
then use \texttt{init.gz} by executing only the first code line below;
otherwise, use \texttt{init-v7.gz} by executing only the second code line:
%as follows:

%The bootloader should be made aware of the initramfs that we will generate:
\begin{lstlisting}[]
(chroot) # echo "initramfs init.gz" >> /boot/config.txt     # For Raspberry Pi version 1
(chroot) # echo "initramfs init-v7.gz" >> /boot/config.txt  # For Raspberry Pi version 2 or 3
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

After this point, it is no longer required to be in \texttt{chroot} mode.
The following commands exit the \texttt{chroot} environment, unmount all
partitions and detach the loopback devices:

% Exit chroot, unmount and detach loopback devices
\begin{lstlisting}[]
(chroot) # exit
# cd ..
# umount --recursive ${ROOTDIR}
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

For the \texttt{----recursive} option to work properly, it is necessary that
the package \texttt{util-linux} version is greater than or equal to 2.22.
Otherwise, an alternative is to either update the package or follow the
procedure in Section~\ref{sec:umount} of the Appendices.

\subsection{Persistent and Non-Persistent Image Directories}
\label{sec:non_persistent_directories}
Provided the stacked filesystem is configured, it is now possible to have
directories where files are removed or not upon rebooting the Raspis.
The following procedure creates an extra partition in the image for
the Raspi user home directory that will be made storage persistent.
%using the upper filesystem.
We first expand image according to the desired
home directory size, but avoid to making the image bigger than the target
memory card size.

% Expand image
\begin{lstlisting}[]
$ dd if=/dev/zero bs=1M count=1024 >> ${IMAGE}.img && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We create a partition for the home directory after the root partition.
To do this, we again use \texttt{fdisk} to find the next available sector
in the image. To verify the new available space for the full image and observe
the next available sector, we run:
\begin{lstlisting}[]
$ fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 3.3 GiB, 3534749696 bytes, 6903808 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We notice that one GB is now available to be used in the partitions. In addition, we
observe the new partition should start at sector 4806656. To create it, we
use \texttt{fdisk} as follows:% and its output:

%\begin{lstlisting}[]
%$ (echo n; echo p; echo 3; echo 4806656; echo ; echo w) | fdisk ${IMAGE}.img
%
%Command (m for help): Partition type:
%   p primary (1 primary, 0 extended, 3 free)
%   e extended
%Partition number (1-4, default 3): Using default value 3
%First sector (2048-6903807, default 2048): Last sector, +sectors or
%+size{K,M,G} (4806656-6903807, default 6903807): Using default value 6903807
%
%Command (m for help): The partition table has been altered!
%
%Syncing disks.
%\end{lstlisting}
\begin{lstlisting}[]
$ fdisk ${IMAGE}.img << EOF
n
p
3
4806656

w
EOF
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\textcolor{red}{SHOW NEW PARTITION TABLE HERE}

We create a loopback device again and format the new partition, as follows:
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
/dev/loop0
# mkfs.ext4 ${DEV}p3
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the \texttt{-P} option is not available for
\texttt{losetup}, we provide an alternative command line in
Section~\ref{sec:alternative_losetup}. Finally, if the previous filesystem
formatting was successful, the filesystem is now available for use.
We need to inform the Raspbian OS to mount the home partition
that we have just created.
%We will notify the OS while booting during the startup of the
%Raspis.
This can be done by adding
an entry in \texttt{fstab} as follows:

\begin{lstlisting}[]
# mount ${DEV}p2 ${ROOTDIR}
# sed -i '$a /dev/mmcblk0p3 /home ext4 defaults,noatime 0 2' ${ROOTDIR}/etc/fstab
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the last command was executed correctly, the
\texttt{\$\{ROOTDIR\}/etc/fstab} file should have the new line.
%However,
%the line will not be properly indented. To verify the correct line
%insertion and fix the indentation, edit the file with
%\texttt{sudo vim \$\{ROOTDIR\}/etc/fstab}.
The resulting file should look like the following:

\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/fstab}@>
<@\textcolor{gray}{
----------------------------------------}
\Reactivatenumber @>
proc            /proc            proc    defaults          0       0
/dev/mmcblk0p1  /boot            vfat    defaults          0       2
/dev/mmcblk0p2  /                ext4    defaults,noatime  0       1
/dev/mmcblk0p3  /home            ext4    defaults,noatime  0       2
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Originally, the home folder is located in the root filesystem. However,
we have to move its content to the new home partition and store it properly.
We do that as follows:

\begin{lstlisting}[]
# mount ${DEV}p3 ${ROOTDIR}/mnt
# mv ${ROOTDIR}/home/* ${ROOTDIR}/mnt/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, unmount again all the partitions and detach the loop devices as follows:
\begin{lstlisting}[]
# umount --recursive ${ROOTDIR}
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the \texttt{----recursive} option is not available, then follow the
procedure in Section~\ref{sec:umount} of the Appendices.
If the steps are successfully executed up to this point, the customized image
is available in the \texttt{\$\{IMAGE\}.img} file and is ready to be deployed
into the Raspis. In the following section, we~indicate how to proceed
with the writing of the image into various memory cards.

\subsection{Writing Customized Image to SD Memory Cards}
For a basic system setup, the final step is to write the customized image
to all the memory cards before they can be used in the Raspis. For
our current considered system, we do this manually for each card.
The testbed administrator needs to insert each memory card in his/her PC
and follow the procedure in this section. A given card will be available as
\texttt{/dev/mmcblkX} or \texttt{/dev/sdYX} where \texttt{X} is a natural number
and \texttt{Y} is a letter.

It is \textit{very important} to write to the correct device as
everything will be overwritten. To avoid removing information from the
wrong device, a testbed administrator can use the commands \texttt{lsblk}
and/or \texttt{df -h} before and after inserting the memory card to deduce
its correct device name.
For~our case, the device was \texttt{/dev/mmcblk0}. Once identified, to write
the image to a memory card, the~following command is used:

\begin{lstlisting}[]
# dd if=${IMAGE}.img of=/dev/mmcblk0 bs=4M && sync
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The previous \texttt{dd} and \texttt{sync} commands for copying the image to
the memory card and flushing the remainder in memory to the filesystem
will take tens of minutes depending on the memory card speed and the size of
the image. After this is made, it is only necessary
to eject the memory card and now plug it in a Raspi so it can boot up.

%The previous \texttt{dd} and \texttt{sync} commands for copying the image to
%the memory card and flushing the remainder in memory to the filesystem
%will take around 15 minutes to be finished. This is an average on a two-core
%1.5 GHz laptop. However, other times might occur depending on the total image
%size, the writing block size and the core capabilities of the testbed
%administrator PC. After this is made, it is only required
%to eject the memory card and now plug it in a Raspi so it can boot up.

\section{Automation and Monitoring Tools}
%\input{tools}

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the daily testbed use, there exists frequent tasks that require
a set of various commands in a given Raspi. This could be tedious, prone
to errors and time-consuming to realize every time the task is required to be
made. Therefore, in this section, we introduce a set of tools that help to
automate and monitor routinary task execution in the Raspis and show
relevant example commands with them. To be able to run all the following
commands, it is necessary to have SSH connectivity with the Raspis;
otherwise, the commands need to be run locally on a Raspi making
necessary to use a keyboard and a monitor. The testbed
administrator needs to put the memory cards in the Raspis and turn them
on for them to be able to boot. The devices should now be bootable.

\subsection{Fabric}
Controlling multiple devices using SSH from a single PC
often leads to many repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices; (ii) installing applications in multiple
devices; and (iii) copying files to/from multiple devices.
Fabric~\cite{python_fabric} provides a Python library that simplifies the
management of working with many devices from a single PC. First, the
testbed administrator creates a directory to hold the \texttt{Fabric} source
code:

\begin{lstlisting}[]
$ export CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Then, the \texttt{\$\{CODEDIR\}/fabfile.py} file below provides a script with
some basic functionalities that can perform the few items above (i--iii).
In general, other administrators may require different functionalities,
but this is out of the scope of this work. The following
file serves as\mbox{ a starting boilerplate}:
\newpage

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{CODEDIR\}/fabfile.py}@>
<@\textcolor{gray}{
-----------------------------------------}
\Reactivatenumber @>
from fabric.api import env, task, sudo
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

env.hosts = ['10.0.0.100','10.0.0.101','10.0.0.102']
env.user = 'pi'
env.password = 'raspberry'

@task
def reboot():
    """ Reboot device """
    sudo('reboot', quiet=True)

@task
def install(program):
    """
    Install a program
    program: program name
    """
    result = sudo('apt-get install -y {}'.format(program), quiet=True)
    print(result)

@task
def push(src,dst):
    """
    Copy file to device
    src: source file path
    dst: destination file path
    """
    put(src, dst)

\end{lstlisting}
\FloatBarrier
\vspace{-14pt}

The previous \texttt{fabfile} shows three functions that perform our
example tasks. These functions utilize variables and subsequent functions
from the Fabric Application Programming Interface (API) such as \texttt{env}, \texttt{task} and \texttt{sudo}
among others. Each of these API functions permits defining environment
variables, creating the administrator tasks through decorators or running the
mentioned task in \texttt{sudo} mode, respectively.
When a task is called from the terminal, Fabric searches the directory for
the \texttt{fabfile.py} file and executes the desired task. The syntax for
executing a task with arguments is in the form
\texttt{fab <TASK>:arg1,arg2,...} We denote the
%Pls check the format with the ... here
IP address of a generic Raspi for test as \texttt{<RASP\_IP>}.
The executions from the terminal of some of these commands are shown as
follows:

\begin{lstlisting}[]
$ fab reboot
[<RASP_IP>] Executing task 'reboot'

Done.
Disconnecting from <RASP_IP>... done.
$ fab install:tmux
[<RASP_IP>] Executing task 'install'
...
The following NEW packages will be installed:
  tmux
0 upgraded, 1 newly installed, 0 to remove and 0 not upgraded.
...
Preparing to unpack .../archives/tmux_1.9-6_armhf.deb ...
Unpacking tmux (1.9-6) ...
Processing triggers for man-db (2.7.0.2-5) ...
Setting up tmux (1.9-6) ...

Done.
Disconnecting from <RASP_IP>... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The first function above reboots the Raspis in the lists of hosts and
the second function installs a program given by an argument. For the
connection to the devices, Fabric calls the Paramiko~\cite{python_paramiko} module from Python
to make an SSH connection. For this to work properly, the Paramiko
version needs to be higher than or equal to 1.15.1. If not available,
the SSH connections from Fabric may fail. In case of any problems,
some instructions for updating the Paramiko package are available in
Section~\ref{sec:paramiko} of the Appendices. This is a standard recommendation
from the Fabric troubleshooting guide~\cite{2016fabricsupport}.
%Pls confirm - this is not a complete sentence - Do you mean "according to the Fabric troubleshooting guide?

After a successful SSH
connection is made, in the previous two commands towards the Raspi,
Fabric employs the Raspi's \texttt{reboot} and \texttt{apt-get}
commands in \texttt{sudo} mode to do the required tasks. Below,
an example is shown for the \texttt{push} task which uses two arguments. Here, we
copy \texttt{my\_file} from the testbed administrator PC to a test
host Raspi:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/my_file",'~/'
[<RASP_IP>] Executing task 'push'
[<RASP_IP>] put: /home/<USER>/code/my_file -> /home/pi/my_file

Done.
Disconnecting from <RASP_IP>... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To control a large set of devices, we simply need to include them in the
\texttt{env.hosts} list in the \texttt{\$\{CODEDIR\}/fabfile.py} file.
Fabric has many other functionalities that are useful in controlling a
large set of Raspis. For example, we may extract files or run
automated experiments. The~included functionalities in the
\texttt{\$\{CODEDIR\}/fabfile.py} file will depend on the requirements of the
testbed~administrator.

%execute a script function by calling "fab install:feh"
\subsection{Long-Running Jobs Using SSH}
There are times when a task may need to run for several hours or even days
on the Raspis, particularly when related to simulations or measurement
campaigns. For this purpose, it might be necessary to keep open an SSH
connection on the Raspis without risking that the connection will
be interrupted and a given Raspi will terminate the task.

There are methods to enable the Raspis to continue running
applications although the connection is terminated either on purpose or
unexpectedly. One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a shell
window, a \texttt{screen} session, which~does not terminate even with
connectivity interruptions. Users can attach and detach from a
\texttt{screen} session
as desired. The following procedure presents how to use \texttt{screen} with
SSH to: (i)~login to a generic Raspi; (ii) open a screen session;
(iii) execute an example command; (iv) detach from the screen session; (v)
terminate the SSH connection; (vi) login to the Raspi again;
and (vii) attach to \texttt{screen} session to see the program still running.
From the testbed administrator PC, we start by establishing an SSH
connection to a Raspi and open a screen session:

\begin{lstlisting}[]
$ ssh pi@<RASP_IP>
...
pi@<RASP_IP>'s password:

The programs included with the Debian GNU/Linux system are free software;
the exact distribution terms for each program are described in the
individual files in /usr/share/doc/*/copyright.

Debian GNU/Linux comes with ABSOLUTELY NO WARRANTY, to the extent
permitted by applicable law.
Last login: Tue Jul 12 13:04:31 2016

$ screen
Screen version 4.02.01 (GNU) 28-Apr-14
...
[Press space or Return to end.]
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To enter in the \texttt{screen} session after the introduction message, we
have to press either the \texttt{Space} or \texttt{Return} key in the
keyboard to clear the shell. After doing so, we should be in a
\texttt{screen} session although its appearance is the same as a regular
terminal shell. Inside this example session, we~execute a program that
never ends:

\begin{lstlisting}[]
$ top
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The \texttt{top} command simply continuously shows the table of processes
executed on the Raspi like in any Linux distribution. When \texttt{top}
is running, we first press \texttt{Ctrl+a} and later \texttt{Crtl+d}
in the keyboard to detach from the \texttt{screen} session. We now
terminate the SSH connection and login again to verify that the
\texttt{top} command is still running. Without using \texttt{screen},
the \texttt{top} program should terminate since its hosting shell was
terminated. To log out, we run:

\begin{lstlisting}[]
$ exit
logout
Connection to <RASP_IP> closed.
$ ssh pi@<RASP_IP>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now that we are logged in to the Raspi again, we first check the
available detached sessions by~running:
\begin{lstlisting}[]
$ screen -list
There is a screen on:
  824.pts-0.raspXX (07/12/16 13:17:30) (Detached)
1 Socket in /var/run/screen/S-pi.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

From the command output, we can see that the session is still running in our
generic Raspi number \texttt{XX} and that no user is currently attached
to the session. To attach to the session, we execute:

\begin{lstlisting}[]
$ screen -r 824.pts-0.raspXX
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

After attaching again, we should see \texttt{top} still running.
\texttt{screen} has more functionalities that can be used in this
or other contexts, but this is outside the scope of this work.
To terminate the \texttt{screen} session, first terminate \texttt{top}
by pressing \texttt{q} in the keyboard.
Once \texttt{top} is terminated,
%Once out of the \texttt{top} output,
we need to type \texttt{exit} two times in order to first exit the
\texttt{screen} session and then terminate the SSH connection.
An~output should be as follows:

\begin{lstlisting}[]
[screen is terminating]
pi@<RASP_IP>:~ $ exit
logout
Connection to <RASP_IP> closed.
$
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\section{Cross-Compilation: From the PC to the Raspberry Pi}
%\input{cross_compilation}

\label{sec:cross_compilation}
An important case of a computational expensive task is to compile software
packages and large libraries. Given the computing capabilities of the
Raspi, such tasks can be challenging if not prohibitive in terms of
\ac{CPU}, memory or space usage and/or compilation time. In
this section, we present a procedure of how to cross-compile C++
source code from the testbed administrator PC for the ARM
architecture of the Raspis. By doing this, we take advantage
of the (typically) much higher computing power of the testbed
administrator PC in order to save time and computational resources.
Hence, we give an example of compiling a simple C++ program and
copying the generated binaries with SSH to run locally
on a Raspi.

Furthermore, given that our testbed purpose is for
network coding applications, we also present how to cross-compile Kodo
\cite{kodo2011pedersen}, a C++11 network coding library to perform
encoding, decoding and recoding operations. In this way, we aim to present
a fully configurable and manageable testbed with the capabilities to
evaluate network coding protocols with several Raspis and locally
store measurements from different evaluations. Therefore, we also show
how \texttt{kodo-cpp}, a set of high-level C++ bindings for Kodo, can be
cross-compiled for applications with the Raspi.

\subsection{Toolchain Setup}
To compile in a given architecture that is aimed for a different one,
the testbed administrator needs to install a toolchain on his/her PC.
The toolchain is mandatory due to the different processor architectures
where the source can be compiled from.
Given that compiling a toolchain can be an arduous task, we get the toolchain
recommended for the ARM architecture of the Raspis.
This~toolchain is available from \cite{soerensen_chres_wiant_2016_154143} and
it already contains the binaries for different compilers based on
\texttt{gcc 4.9}. We extract the binaries adjusting
them to our coding style and compiling convention. For this, we use the
\texttt{\$\{TOOLCHAIN\}} directory as the working directory.
The testbed administrator may choose some other working directory of
its preference if desired. First, we create the toolchain directory:

\begin{lstlisting}[]
$ export TOOLCHAINDIR="${HOME}/toolchains"
$ mkdir -p ${TOOLCHAINDIR}
$ cd ${TOOLCHAINDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Later, we download a Raspi toolchain with the binaries for a 64-bit Linux
distribution available in \cite{soerensen_chres_wiant_2016_154143}. Finally, we unzip the
downloaded file. This is made as follows:

\begin{lstlisting}[]
$ wget https://zenodo.org/record/154328/files/raspberry-gxx493-arm.zip
$ unzip raspberry-gxx493-arm.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Instead of calling the ARM cross compiler using its full path, we make
the binaries accessible from the command shell systemwide. A way to do
this is by adding the following commands in the \texttt{\$\{HOME\}/.profile}
as follows:

\begin{lstlisting}[]
$ sed -i '$a export TOOLCHAINDIR=\"$HOME/toolchains\"' ${HOME}/.profile
$ sed -i '$a export TOOLCHAINBINARY=\"raspberry-gxx49-arm-g++\"' ${HOME}/.profile
$ sed -i '$a PATH=\"\$PATH:${TOOLCHAINDIR}/arm-rpi-4.9.3-linux-gnueabihf/bin\"' ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This helps the OS to recognize the location of the compiler
command when a new shell is opened. The \texttt{.profile} should now
contain the lines we inserted. There might be other
code in the file of other testbed administrators. We recommend to leave
other parts unmodified.
% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$HOME/.profile}@>
<@\textcolor{gray}{
---------------------------}
\Reactivatenumber @>
...
export TOOLCHAINDIR="$HOME/toolchains"
export TOOLCHAINBINARY="raspberry-gxx49-arm-g++"
PATH="$PATH:${TOOLCHAINDIR}/arm-rpi-4.9.3-linux-gnueabihf/bin"
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To update the \texttt{\$\{PATH\}} variable and the \texttt{.profile}, we
use the \texttt{source} command for the changes take effect in the
administrator system:
\begin{lstlisting}[]
$ source ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

A working ARM cross-compiler in the testbed administrator PC should
output the following:

\begin{lstlisting}[]
$ ${TOOLCHAINBINARY} --version
raspberry-gxx49-arm-g++ (crosstool-NG crosstool-ng-1.22.0-88-g8460611) 4.9.3
Copyright (C) 2015 Free Software Foundation, Inc.
This is free software; see the source for copying conditions.  There is NO
warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Cross-Compile Example}
The following shows: (i) how to cross compile the classic \texttt{hello\_world}
C++ example for the Raspi ARM architecture and (ii) how to copy and
execute the binary in a Raspi using Secure Copy (SCP) and SSH. First, we
create the file \texttt{hello\_world.cpp}. For simplicity, we create it
in the directory where we stored the \texttt{fabfile.py} file
with the following content using any text editor:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{CODEDIR\}/hello\_world.cpp}@>
<@\textcolor{gray}{
---------------------------------------------------}
\Reactivatenumber @>
#include <iostream>

int main()
{
    std::cout << "Hello World!" << std::endl;
    return 0;
}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We save the previous file and compile it for Raspi in the testbed
administrator PC by doing:

\begin{lstlisting}[]
$ ${TOOLCHAINBINARY} hello_world.cpp -o hello_world
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This should produce a binary \texttt{hello\_world} that is executable
on the Raspi. We copy it to a Raspi using SCP and using
Fabric instead if we are interested in deploying a compiled binary \mbox{for
many Raspis.}

\begin{lstlisting}[]
$ scp hello_world pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

% If you do not know the IP address of a Raspi in your network, you can
% connect a monitor to it and run the following command after you have logged in:
% \begin{lstlisting}[]
% pi@rasp01:~ $ ifconfig
% eth0      Link encap:Ethernet  HWaddr b8:27:eb:72:77:54
%           inet addr:192.168.87.107  Bcast:192.168.87.255  Mask:255.255.255.0
%           inet6 addr: fe80::e0a5:38f3:6f82:bc79/64 Scope:Link
%           UP BROADCAST RUNNING MULTICAST  MTU:1500  Metric:1
%           RX packets:1537 errors:0 dropped:0 overruns:0 frame:0
%           TX packets:445 errors:0 dropped:0 overruns:0 carrier:0
%           collisions:0 txqueuelen:1000
%           RX bytes:259117 (253.0 KiB)  TX bytes:52551 (51.3 KiB)

% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

After the executable has been copied to the Raspi, we login through
SSH to it:

\begin{lstlisting}[]
$ ssh pi:<RASP_IP>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We can list the directory content after we have logged into the Raspi and
verify that the compiled \texttt{hello\_world} binary is there:

\begin{lstlisting}[]
pi@<RASP_IP>:~ $ ls
hello_world  rasp_config
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, we simply execute the \texttt{hello\_world} to confirm that
the cross-compiling of \texttt{hello\_world} worked properly:

\begin{lstlisting}[]
pi@<RASP_IP>:~ $ ./hello_world
Hello World!
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Cross-Compile Kodo}

As we originally mentioned, Kodo is a C++11 network coding library that
permits implementation of network coding functionalities by allowing any network
protocol designer to use and test the primitive encoding, decoding and
recoding operations of RLNC. In this way, a designer only needs to
focus on the design and test of a network coding-based protocol. Kodo is
available through programming bindings for a variety of popular
programming languages. This procedure will present how to configure the
Kodo C++ bindings \texttt{kodo-cpp} to cross-compile applications that can
run in Raspi. \texttt{kodo-cpp} provides a simple interface to the
underlying C++11 code that exists in the libraries \texttt{kodo-core} for
the object structure and \texttt{kodo-rlnc} for the RLNC codec
implementation. More details about Kodo are provided in the code
documentation~\cite{kodocppdoc}.

To use Kodo for research, it is necessary to obtain a research free license.
To do this, a request form needs to be filled
in~\cite{steinwurflicenselink} and wait for it to be processed by the Kodo
developers. Once
the access for Kodo has been granted, the source code can be pulled from
its Git repositories to be compiled. Assuming that the testbed
administrator already has access, we clone the \texttt{kodo-cpp}
repository locally in \texttt{\${CODEDIR}} and change directory into
the repository by doing:

\begin{lstlisting}[]
$ cd ${CODEDIR}
$ git clone git@github.com:steinwurf/kodo-cpp.git
$ cd kodo-cpp
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

We first configure \texttt{kodo-cpp} to build executables for the
ARM architecture using the Raspi toolchain and later build them
by running:
\begin{lstlisting}[]
$ python waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
...
'configure' finished successfully (X.XXXs)
$ python waf build
...
'build' finished successfully (XmXX.XXs)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the configuration and build steps are successful, the binaries should
have been created. To be able to use them, we need to create a shared library
that we will use in the Raspi. To do this, we run the following command:

\begin{lstlisting}[]
$ python waf install --install_shared_libs --install_path="./shared_test"
...
'install' finished successfully (X.XXXs)
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now, we copy the shared library, binary files and related headers to the
Raspi home directory as~follows:

\begin{lstlisting}[]
$ scp -r shared_test/include shared_test/libkodoc.so pi@<RASP_IP>:~/
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Alternatively, and for the testbed administrator reference, Kodo can also
generate static libraries. We log in to the Raspi and execute the unit
tests and one of the binaries by running:

\begin{lstlisting}[]
$ ssh pi@<RASP_IP>
$ ./kodocpp_tests
...
[  PASSED  ] X tests.
$ ./encode_decode_simple
Data decoded correctly
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the Kodo cross-compilation worked properly, both the unit tests and
binaries run should provide the shown outputs.

%\subsection{Kodo python}
%
%Often when simplicity and fast prototyping is of higher importance than speed
%it is often desired to use a scripting language rather than the often faster
%compiled language. Kodo-python provides roughly similar bindings to the
%underlying C++11 implementation of Kodo as Kodo-cpp does. This means that
%python calls to Kodo-python are actually executed in the underlying C++11
%compilied code.
%
%Parsing calls and variables between Python and C++ compiled code is
%not free, but this small introduced overhead may in many cases be worth
%it to take advantage of Pythons and other scripting languages many
%benefits.
%
%Below is a short procedure to use Kodo-python in a Raspi.
%
%One advantage of Python compared to Kodo-cpp is that only the Kodo-python
%library needs to be cross compiled. The python scripts using the Kodo
%library needs not to be cross compiled but simply copied to the same
%directory as .
%
%Kodo-python is available at \url{https://github.com/steinwurf/kodo-python}.
%\begin{lstlisting}[]
%$ apt-get update
%$ apt-get install python build-essential libpython-dev python-dev
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%  % IN ARCH LINUX, MOD_PYTHON NEEDS TO BE INSTALLED
%
%
%Create a directory to hold our code:
%\begin{lstlisting}[]
%$ CODEDIR="${HOME}/code"
%$ mkdir -p ${CODEDIR}
%$ cd ${CODEDIR}
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%\begin{lstlisting}[]
%$ git clone git@github.com:steinwurf/kodo-python.git
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%\begin{lstlisting}[]
%$ cd kodo-python
%$ python2.7 waf configure --cxx_mkspec=cxx_raspberry_gxx49_arm
%$ python waf build
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}
%
%KODO-PYTHON IS NOT WORKING!! WAITING FOR HELP ON STEINWURF DEV.
%
%WRITE HOW TO RUN A KODO PYTHON SCRIPT IN THE RASPIS

%\texttt{hello\_world} returns immediately. This is not always the case.
%Sometimes it is desired to run simulations that should run for days.
%A software package called \texttt{screen} can be used for this purpose.
%Simply run ...

\section{Conclusions}
%\input{conclusions}

\label{sec:conclusions}
Observing the expectation of the IoT and lack for a low-cost, easy-to-configure testbed in this area for reproducible research, we provide an in-depth description of the new Aalborg University's Raspi testbed for network coding evaluation and how to guarantee replicability and scaling management of this system. The description shows how to set up interconnected Raspis with memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges. Using the presented procedure permits setting up a Raspbian Lite image for the Raspis. A tailored Linux distribution might be created from the scratch using the Yocto project. However, to assemble and compile the software for the Raspi can be a tedious and time-consuming task. However, this method could be adequate for an expert user. We hope this work permits researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way. Future work in the use of Raspi devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and open the use of the testbed beyond our team at Aalborg University. Future work in this area will consider making the testbed fetch the image through the HTTP server. This is expected to simplify the maintenance of the memory cards.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Subsection}
% \subsubsection{Subsubsection}

% Bulleted lists look like this:
% \begin{itemize}[leftmargin=*,labelsep=4mm]
% \item	First bullet
% \item	Second bullet
% \item	Third bullet
% \end{itemize}

% Numbered lists can be added as follows:
% \begin{enumerate}[leftmargin=*,labelsep=3mm]
% \item	First item
% \item	Second item
% \item	Third item
% \end{enumerate}

% The text continues here.

% \subsection{Figures, Tables and Schemes}

% All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

% \begin{figure}[H]
% \centering
% %\includegraphics[width=3cm]{logo-mdpi}
% \caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
% \end{figure}

% \begin{table}[H]
% \caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
% \small % Font size can be changed to match table content. Recommend 10 pt.
% \centering
% \begin{tabular}{ccc}
% \toprule
% \textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
% \midrule
% entry 1		& data			& data\\
% entry 2		& data			& data\\
% \bottomrule
% \end{tabular}
% \end{table}

% \subsection{Formatting of Mathematical Components}

% This is an example of an equation:

% \begin{equation}
% \mathbb{S}
% \end{equation}

%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
% Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
% \begin{Theorem}
% Example text of a theorem.
% \end{Theorem}
% The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
% \begin{proof}[Proof of Theorem 1]
% Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
% \end{proof}
% The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Discussion}

%This section may be divided by subheadings. Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Materials and Methods}

%This section should be divided by subheadings. Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

%Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusions}

%This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{2pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This research has been partially financed by the
Marie Curie Initial Training Network (ITN) CROSSFIRE project
(Grant No. EU-FP7-CROSSFIRE-317126) from the European Comission FP7
framework, the~Green Mobile Cloud project (Grant No. DFF-0602-01372B),
the TuneSCode project (Grant No. DFF-1335-00125) both granted by the
Danish Council for Independent Research (Det Frie Forskningsr\r{a}d), and~by the German Research Foundation (DFG) within the Collaborative Research
Center SFB 912---HAEC.}

%All sources of funding of the study should be disclosed. Please clearly indicate grants that you have received in support of your research work. Clearly state if you received funds for covering the costs to publish in open access.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{The authors contributed equally to this work.
%\hl{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``X.X. and Y.Y. conceived and designed the experiments; X.X. performed the experiments; X.X. and Y.Y. analyzed the data; W.W. contributed reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictofinterests{The authors declare no conflict of interest. The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results.}

%Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:
% Use these two commands to get used acronyms
%used=$(grep -ohR '\\ac{[^}]*' *.tex | cut -f2 -d'{' | sort | uniq)
%for ac in ${used[@]}; do echo "$(grep -R "acro{${ac}}" acronym.tex)"; done | cut -f2-4 -d'{'
% Use text editor to replace '}{' --> ': ' and '}' --> '\\'
\vspace{12pt}

\noindent\hspace{-0.85em}
\begin{tabular}{ll}
API & Application Programming Interface\\
ARM & Advanced RISC Machine\\
Bash & Bourne Again SHell\\
CPU & Central Processing Unit\\
DHCP & Dynamic Host Configuration Protocol\\
HTTP & Hyper Text Transfer Protocol\\
IoT & Internet of Things\\
IP & Internet Protocol\\
\end{tabular}
% it is the same as above.

\noindent\hspace{-0.85em}
\begin{tabular}{ll}
IT & Information Technology\\
LAN & Local Area Network\\
MAC & Medium Access Control\\
NC & Network Coding\\
NFS & Network File System\\
OS & Operating System\\
PC & Personal Computer\\
QEMU & Quick Emulator\\
RAM & Random Access Memory\\
Raspi & Raspberry Pi\\
RLNC & Random Linear Network Coding\\
SCP & Secure Copy\\
SD & Secure Digital\\
SSH & Secure Shell\\
URL & Uniform Resource Locator\\
\end{tabular}}

\appendixtitles{yes}
\appendixsections{multiple}
\appendix
%\input{appendix}

% \section{HTTP Image Fetch Setup (Optional)}
% \label{sec:http}

% \input{testbed_http}

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages on his/her Linux distribution for performing
the commands, particularly the~ones regarding the \texttt{util-linux} package
or if a command just fails.

\subsection{Losetup for Loopback Devices}
%Pls confirm this word "Losetup" here******************
\label{sec:alternative_losetup}

If the \texttt{losetup -P} command shows the \texttt{invalid option}
message, an alternative is to manually set all the loopback devices
used during the whole procedure with \texttt{mknod}. To do this, the~alternative commands are:

\begin{lstlisting}[]
$ export DEV="/dev/loop0"
# mknod ${DEV}p1 b 7 1
# mknod ${DEV}p2 b 7 2
# losetup -o $((8192*512)) --sizelimit $(( (137215-8192+1)*512) )) ${DEV}p1 ${IMAGE}.img
# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) )) ${DEV}p2 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

However, there will be a few differences by using the above code as an
alternative to the \texttt{losetup -P} case. First, the output from
\texttt{lsblk} will look different:

\begin{lstlisting}[]
# lsblsk
NAME   MAJ:MIN RM   SIZE RO TYPE MOUNTPOINT
...
loop1    7:1    0    67M  0 loop
loop2    7:2    0   2.2G  0 loop
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Second, detaching the loop devices will also be different:

\begin{lstlisting}[]
# losetup -d ${DEV}p{1,2}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Working with these loopback devices will be transparent for the remainding
procedure.

In addition, when creating the persistent home directory partition in
Section~\ref{sec:non_persistent_directories}, for recreating the loopback
devices, the commands are:

\begin{lstlisting}[]
$ export DEV="/dev/loop0"
# mknod ${DEV}p2 b 7 2
# mknod ${DEV}p3 b 7 3
# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) ) ${DEV}p2 ${IMAGE}.img
# losetup -o $((4806656*512)) --sizelimit $(( (6903807-4806656+1)*512)) ${DEV}p3 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, for detaching in this case, the command is:
\begin{lstlisting}[]
# losetup -d ${DEV}p{2,3}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Image Chroot with Proot}
\label{sec:chroot}

In some cases, \texttt{chroot} may not work properly. In this case, an
alternative can be \texttt{proot}. It might be required to install it
before changing the root with \texttt{apt-get install proot}. Then, it
should be possible for the testbed administrator to run the following:

% CHROOT to OS image
\begin{lstlisting}[]
# proot -q qemu-arm-static -S ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Umount after Image Chroot}
\label{sec:umount}

%The recursive command (\texttt{----recursive}) is not valid for \texttt{umount}
%is not valid for \texttt{util-linux} version 2.22
%or later. An alternative option is to manually unmount all the partitions used
%during the \texttt{chroot} environment. The directories must be unmounted in
%the reverse order as when they were mounter. These commands are:

%The --recursive parameter is not part of umount until 
%\texttt{umount ----recursive} is not part of \texttt{util-linux} version 2.22

In case the \texttt{umount ----recursive} command shows the \texttt{invalid option}
message, an alternative is to manually unmount all the partitions used
during the \texttt{chroot} environment in the reverse order from which they were mounted.
These commands are:
%In case the \texttt{umount ----recursive} command fails given that a version
%higher than 2.22 for \texttt{util-linux} is not available,
%a possible option is to manually umount all the partitions used during the
%\texttt{chroot} environment. The directories must be unmounted in the reverse
%order as when they were mounted. These commands are:

\begin{lstlisting}[]
# umount ${ROOTDIR}/dev/pts
# umount ${ROOTDIR}/dev
# umount ${ROOTDIR}/sys
# umount ${ROOTDIR}/proc
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Paramiko Package Update}
\label{sec:paramiko}
In case Paramiko version 1.15.1 or higher is not installed, we may observe
a key exchange algorithm error when trying to log in through SSH to a
Raspi with Open SSH. In this case, it might be necessary to update the Paramiko
package that Fabric uses to remove this error by running:

\begin{lstlisting}[]
# pip install --upgrade paramiko
...
Successfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
# pip show paramiko
...
Name: paramiko
Version: 2.0.1
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
% \appendix
% \section{}
% The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathemtaical proofs of results not central to the paper can be added as an appendix.

% \section{}
% All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{mdpi}

%=====================================
% References, variant A: internal bibliography
%=====================================
 \renewcommand\bibname{References}
 \begin{thebibliography}{999}
% % Reference 1
% \bibitem{ref-journal}
% Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
% % Reference 2
% \bibitem{ref-book}
% Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
% \end{thebibliography}
%\begin{thebibliography}{-------}
\providecommand{\natexlab}[1]{#1}

\bibitem[Evans(2011)]{cisco2011forecast}
Evans, D.
\newblock The Internet of Things: How the next evolution of the Internet is
  changing everything.
\newblock {\em CISCO. White Papers} {\bf 2011}, {\em 1},~1--11.

\bibitem[Ahlswede \em{et~al.}(2000)Ahlswede, Cai, Li, and
  Yeung]{ahlswede2000network}
Ahlswede, R.; Cai, N.; Li, S.Y.; Yeung, R.W.
\newblock Network Information Flow.
\newblock {\em IEEE Trans. Inf. Theory } {\bf 2000}, {\em
  46},~1204--1216.

\bibitem[Koetter and M\'edard(2003)]{koetter2003algebraic}
Koetter, R.; M\'edard, M.
\newblock An Algebraic Approach to Network Coding.
\newblock {\em IEEE/ACM Trans. Netw.} {\bf 2003}, {\em 11},~782--795.

\bibitem[Ho \em{et~al.}(2006)Ho, M{\'e}dard, Koetter, Karger, Effros, Shi, and
  Leong]{ho2006random}
Ho, T.; M{\'e}dard, M.; Koetter, R.; Karger, D.R.; Effros, M.; Shi, J.; Leong,
  B.
\newblock A Random Linear Network Coding Approach to Multicast.
\newblock {\em IEEE Trans. Inf. Theory} {\bf 2006}, {\em
  52},~4413--4430.

\bibitem[Pahlevani \em{et~al.}(2013)Pahlevani, Lucani, Pedersen, and
  Fitzek]{pahlevani2013playncool}
Pahlevani, P.; Lucani, D.E.; Pedersen, M.V.; Fitzek, F.H.
\newblock Playncool: Opportunistic network coding for local optimization of
  routing in wireless mesh networks.
\newblock In Proceedings of the 2013 IEEE Globecom Workshops (GC Wkshps); pp.
  812--817.
% newly added information, please confirm.
\bibitem[Szabo \em{et~al.}(2015)Szabo, Gulyas, Fitzek, Fitzek, and
  Lucani]{szabo2015towards}
Szabo, D.; Gulyas, A.; Fitzek, F.H.P.; Fitzek, F.H.P.; Lucani, D.E.
\newblock Towards the Tactile Internet: Decreasing Communication Latency with
  Network Coding and Software Defined Networking.
\newblock  In Proceedings of the 21th European Wireless Conference European Wireless 2015; pp. 1--6.

\bibitem[Lucani \em{et~al.}(2014)Lucani, Pedersen, Heide, and
  Fitzek]{lucani2014fulcrum}
Lucani, D.E.; Pedersen, M.V.; Heide, J.; Fitzek, F.H.P.
\newblock Fulcrum Network Codes: {A} Code for Fluid Allocation of Complexity.
\newblock {\em CoRR} {\bf 2014}, DOI: {\em abs/1404.6620}.
% please confrim this ref.'s type.
\bibitem[Pedersen \em{et~al.}(2011)Pedersen, Heide, and
  Fitzek]{kodo2011pedersen}
Pedersen, M.; Heide, J.; Fitzek, F.
\newblock Kodo: An Open and Research Oriented Network Coding Library. In {\em
  Networking 2011 Workshops}; Lecture   Notes in Computer ScienceSpringer: Berlin/Heidelberg, Germany,  2011; Volume 6827, pp. 145--152.

\bibitem[Cox \em{et~al.}(2014)Cox, Cox, Boardman, Johnston, Scott, and
  O'Brien]{cox2014iridis}
Cox, S.J.; Cox, J.T.; Boardman, R.P.; Johnston, S.J.; Scott, M.; O'Brien, N.S.
\newblock Iridis-pi: A low-cost, compact demonstration cluster.
\newblock {\em Clust. Comput.} {\bf 2014}, {\em 17},~349--358.

\bibitem[mak(2016)]{making_of_pi}
Raspberry Pi Foundation. The Making of Pi.
\newblock Available online: https://www.raspberrypi.org/about (accessed on October 10, 2016).

\bibitem[Leccese \em{et~al.}(2014)Leccese, Cagnetti, and
  Trinca]{leccese2014smartcity}
Leccese, F.; Cagnetti, M.; Trinca, D.
\newblock A Smart City Application: A Fully Controlled Street Lighting Isle
  Based on Raspberry-Pi Card, a ZigBee Sensor Network and WiMAX.
\newblock {\em Sensors} {\bf 2014}, {\em 14},~24408--24424.

\bibitem[ard()]{arduino}
ARDUINO Corp.
\newblock Available online: http://arduino.cc (accessed on October 10, 2016).
% plaese add  day month year.
\bibitem[Leccese \em{et~al.}(2014)Leccese, Cagnetti, Calogero, Trinca,
  Pasquale, Giarnetti, and Cozzella]{leccese2014newimaging}
Leccese, F.; Cagnetti, M.; Calogero, A.; Trinca, D.; Pasquale, S.d.; Giarnetti,
  S.; Cozzella, L.
\newblock A New Acquisition and Imaging System for Environmental Measurements:
  An Experience on the Italian Cultural Heritage.
\newblock {\em Sensors} {\bf 2014}, {\em 14},~9290--9312.

\bibitem[Sapes and Solsona(2016)]{sapes2016finger}
Sapes, J.; Solsona, F.
\newblock FingerScanner: Embedding a Fingerprint Scanner in a Raspberry Pi.
\newblock {\em Sensors} {\bf 2016}, {\em 16},~220.

\bibitem[Pre()]{PredictableNetworkInterfaceNames}
Predictable Network Interface Names. 
%\newblock \\
  Available online:  https://www.freedesktop.org/wiki/Software/\linebreak systemd/PredictableNetworkInterfaceNames.

%\bibitem[Soerensen \em{et~al.}(2016)Soerensen, Hernandez-Marcano, Cabrera,
%  Wunderlich, Lucani, and Fitzek]{soerensen_chres_wiant_2016_154143}
%S{\o}rensen, C.W.; Hernandez-Marcano, N.J.; Cabrera, J.A.; Wunderlich, S.;
%  Lucani, D.E.; Fitzek, F.H.P.
%\newblock \hl{Easy as Pi}: A Network Coding Raspberry Pi Testbed (Configuration
%  Files and Scripts), DOI: 10.5281/zenodo.154328,  2016.

\bibitem[Soerensen \em{et~al.}(2016)Soerensen, Hernandez-Marcano, Cabrera,
Wunderlich, Lucani, and Fitzek]{soerensen_chres_wiant_2016_154143}
  Configuration Files and Scripts.
  \newblock DOI: 10.5281/zenodo.154328.
  \newblock\\ Available online:  https://doi.org/10.5281/zenodo.154328

\bibitem[yoc()]{yocto}
The Yocto Project.
\newblock  Available online: https://www.yoctoproject.org (accessed on October 10, 2016).

\bibitem[ras()]{raspbian}
The Raspbian Distro.
\newblock  Available online: https://www.raspbian.org (accessed on October 10, 2016).

\bibitem[Qem()]{QemuUserEmulation}
Qemu User Emulation.
\newblock  Available online: https://wiki.debian.org/QemuUserEmulation  (accessed on October 10, 2016).

\bibitem[gnu()]{gnu_screen}
GNU Screen.
\newblock  Available online: https://www.gnu.org/software/screen  (accessed on October 10, 2016).

\bibitem[ove()]{overlayroot}
Overlayroot.
\newblock  Available online: https://github.com/chesty/overlayroot  (accessed on October 10, 2016).

\bibitem[pyt({\natexlab{a}})]{python_fabric}
Fabric Documentation.
\newblock  Available online: http://www.fabfile.org (accessed on October 10, 2016).

\bibitem[pyt({\natexlab{b}})]{python_paramiko}
Python Paramiko.
\newblock  Available online: http://www.paramiko.org (accessed on October 10, 2016).

\bibitem[201()]{2016fabricsupport}
Fabric Troubleshooting Guide.
\newblock  Available online: http://www.fabfile.org/troubleshooting.html (accessed on October 10, 2016).

\bibitem[kod()]{kodocppdoc}
Kodo-cpp documentation.
\newblock  Available online: http://docs.steinwurf.com/kodo/kodo-cpp/index.html (accessed on October 10, 2016).

\bibitem[ste()]{steinwurflicenselink}
Steinwurf Research License webpage.
\newblock  Available online: http://steinwurf.com/license (accessed on October 10, 2016).

\end{thebibliography}

%=====================================
% References, variant B: external bibliography
%=====================================
%\bibliography{raspi_journal}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\sampleavailability{\hl{The testbed and} measurements in this publication are both available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
